### 2019-04-08 paper/paper.Rnw
template modified for our paper

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\journal{Forensic Science International}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{A Convolutional Neural Network for Outsole Recognition}

%% Group authors per affiliation:
% \author{Miranda Tilton and Susan Vanderplas\fnref{csafe}}
% \address{Iowa State University Statistics Department}
% \fntext[csafe]{195 Durham Center, 613 Morill Rd., Ames, IA 50011}

% or include affiliations in footnotes:
\author[mymainaddress]{Miranda Tilton}

\author[mymainaddress]{Susan Vanderplas\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead[url]{https://forensicstats.org}
\ead{srvander@iastate.edu}

\address[mymainaddress]{Statistics Department, Iowa State University}

\begin{abstract}
This template helps you to create a properly formatted \LaTeX\ manuscript.
\end{abstract}

\begin{keyword}
footwear,class characteristics,computer vision, neural networks
\end{keyword}

\end{frontmatter}

\linenumbers

\section{The Elsevier article class}

\paragraph{Installation} If the document class \emph{elsarticle} is not available on your computer, you can download and install the system package \emph{texlive-publishers} (Linux) or install the \LaTeX\ package \emph{elsarticle} using the package manager of your \TeX\ installation, which is typically \TeX\ Live or Mik\TeX.

\paragraph{Usage} Once the package is properly installed, you can use the document class \emph{elsarticle} to create a manuscript. Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal. It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.

\paragraph{Functionality} The Elsevier article class is based on the standard article class and supports almost all of the functionality of that class. In addition, it features commands and options to format the
\begin{itemize}
\item document style
\item baselineskip
\item front matter
\item keywords and MSC codes
\item theorems, definitions and proofs
\item lables of enumerations
\item citation style and labeling.
\end{itemize}

\section{Front matter}

The author names and affiliations could be formatted in two ways:
\begin{enumerate}[(1)]
\item Group the authors per affiliation.
\item Use footnotes to indicate the affiliations.
\end{enumerate}
See the front matter of this document for examples. You are recommended to conform your choice to the journal you are submitting to.

\section{Bibliography styles}

There are various bibliography styles available. You can select the style of your choice in the preamble of this document. These styles are Elsevier styles based on standard styles like Harvard and Vancouver. Please use Bib\TeX\ to generate your bibliography and include DOIs whenever available.

Here are two sample references: \cite{Feynman1963118,Dirac1953888}.

\section*{References}

\bibliography{mybibfile}

\end{document}

### 2019-04-08 paper/paper.Rnw
Add setup chunk, fix author formatting

@@
 
+<<setup, include = F>>=
+knitr::opts_chunk$set(echo = F, dev = "pdf", dpi = 1000)
+@
+
 \begin{document}
@@
 
-\title{A Convolutional Neural Network for Outsole Recognition}
+\title{A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}}
+\tnotetext[t1]{This document is the results of the research
+project funded ....}
 
@@
 % or include affiliations in footnotes:
-\author[mymainaddress]{Miranda Tilton}
+\author[1]{Miranda Tilton\fnref{isu}}
+\ead{tiltonm@iastate.edu}
 
-\author[mymainaddress]{Susan Vanderplas\corref{mycorrespondingauthor}}
-\cortext[mycorrespondingauthor]{Corresponding author}
-\ead[url]{https://forensicstats.org}
+\author[1]{Susan Vanderplas\corref{cor1}\fnref{isu}}
+\cortext[cor1]{Corresponding author}
 \ead{srvander@iastate.edu}
 
-\address[mymainaddress]{Statistics Department, Iowa State University}
+\address[1]{195 Durham Center, 613 Morill Rd, Ames, IA 50010}
+\fntext[isu]{Center for Statistical Applications in Forensic Evidence, Iowa State University}
 

### 2019-04-08 paper/paper.Rnw
modify to add sections

@@
 
-\section{The Elsevier article class}
+\section{Introduction}
 
-\paragraph{Installation} If the document class \emph{elsarticle} is not available on your computer, you can download and install the system package \emph{texlive-publishers} (Linux) or install the \LaTeX\ package \emph{elsarticle} using the package manager of your \TeX\ installation, which is typically \TeX\ Live or Mik\TeX.
+\section{Materials and Methods}
 
-\paragraph{Usage} Once the package is properly installed, you can use the document class \emph{elsarticle} to create a manuscript.
-Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal.
-It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.
+\section{Results}
 
-\paragraph{Functionality} The Elsevier article class is based on the standard article class and supports almost all of the functionality of that class.
-In addition, it features commands and options to format the
-\begin{itemize}
-\item document style
-\item baselineskip
-\item front matter
-\item keywords and MSC codes
-\item theorems, definitions and proofs
-\item lables of enumerations
-\item citation style and labeling.
-\end{itemize}
-
-\section{Front matter}
-
-The author names and affiliations could be formatted in two ways:
-\begin{enumerate}[(1)]
-\item Group the authors per affiliation.
-\item Use footnotes to indicate the affiliations.
-\end{enumerate}
-See the front matter of this document for examples. You are recommended to conform your choice to the journal you are submitting to.
-
-\section{Bibliography styles}
-
-There are various bibliography styles available.
-You can select the style of your choice in the preamble of this document.
-These styles are Elsevier styles based on standard styles like Harvard and Vancouver.
-Please use Bib\TeX\ to generate your bibliography and include DOIs whenever available.
-
-Here are two sample references: \cite{Feynman1963118,Dirac1953888}.
+\section{Conclusions}
 

### 2019-04-08 paper/paper.Rnw
fixing makefile, references

@@
 \section{Introduction}
+Most forensic examinations of footwear pattern evidence are limited to class characteristic based matches; that is, matches based on the positioning of geometric features within the tread pattern.
+Some of these features may include brand logos and size markings, while others are more generic geometric shapes, such as circles, triangles, stars, and chevrons.
+Class characteristic matches can be used for elimination, but are not sufficient for identification, as the characteristics are shared by all shoes with the same size, make, and model \citep{grossVariabilitySignificanceClass2013,bodziakfootwear2000}.
+Several attempts have been made to automatically identify features in both prints and shoe tread images, ranging from the
 

### 2019-04-08 paper/paper.Rnw
Paper outline

@@
 Class characteristic matches can be used for elimination, but are not sufficient for identification, as the characteristics are shared by all shoes with the same size, make, and model \citep{grossVariabilitySignificanceClass2013,bodziakfootwear2000}.
-Several attempts have been made to automatically identify features in both prints and shoe tread images, ranging from the
+Several attempts have been made to automatically identify features in both prints and shoe tread images, using a range of low-level techniques including fractal decomposition \citep{alexanderAutomaticClassificationRecognition1999}, scale-invariant feature recognition \citep{pavlouAutomaticExtractionClassification2006}, Fourier-Mellin transformation \citep{guehamAutomaticRecognitionShoeprints2008}, and other classical image analysis methods.
+Vast advances have been made in the field of computer vision since 2012 \citep{krizhevskyImageNetClassificationDeep2012}, when convolutional neural networks (CNNs) were introduced.
+These networks are designed to mimic the process of human vision, and typically involve application of sets of filters, some of which mimic the filters used in the low-level techniques used by early automatic classification attempts.
+Recently, raw output from CNN models have been used to identify or match shoes to prints \citep{kongCrossDomainImageMatching2019,kongCrossDomainForensicShoeprint2017,zhangAdaptingConvolutionalNeural2017}; these methods work relatively well, but the features used for matching are not generally informative for humans, who typically classify shoes by patterns and spatial relationships \citep{davisIntelligenceApproachFootwear1981,grossVariabilitySignificanceClass2013}.
+In this paper, we discuss a hybrid approach, using features from a pre-trained CNN to develop a classifier which automatically identifies geometric patterns in images of shoe treads.
+Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain.
+
+% This is a really really high level overview - may need to be a bit deeper
 
 \section{Materials and Methods}
+\paragraph{VGG16}
+
+\paragraph{LabelMe Shoe Annotations}
+
+\paragraph{Transfer Learning}
 
 \section{Results}
+\paragraph{Overall Accuracy} % Model-wise and class-by-class, w/ confusion matrix
+
+\paragraph{Model Diagnostics}
 
 \section{Conclusions}
+
+\section{Future Work}
 

### 2019-04-12 paper/paper.Rnw
update introduction - finish out comparison of cv methods

--- 2019-04-08 paper/paper.Rnw
@@
 Class characteristic matches can be used for elimination, but are not sufficient for identification, as the characteristics are shared by all shoes with the same size, make, and model \citep{grossVariabilitySignificanceClass2013,bodziakfootwear2000}.
-Several attempts have been made to automatically identify features in both prints and shoe tread images, ranging from the
+Several attempts have been made to automatically identify features in both prints and shoe tread images, with low-level image analysis methods \citep{guehamAutomaticRecognitionShoeprints2008,chazalAutomatedProcessingShoeprint2005,cervelliTranslationalRotationalInvariant} such as Fourier transforms and keypoint detection methods \citep{luostarinenMeasuringAccuracyAutomatic2014a}.
+Some of these low-level methods perform relatively well in good conditions \citep{luostarinenMeasuringAccuracyAutomatic2014a} but have degraded performance under conditions which are suboptimal, including those commonly found at crimescenes.
+In the wider field of computer vision, attention has turned to more robust methods for image recognition, such as convolutional neural networks (CNNs), which are capable of achieving near-human accuracy under even degraded image conditions.
+CNNs have been applied in footwear forensics, making use of more general neural networks optimized to detect objects found in natural scenes, such as trees, animals, buildings, and cars; the pre-trained network's features are then used to compare shoes or prints to determine how well they match \citep{kongCrossDomainImageMatching2019a,kongCrossDomainForensicShoeprint2017,zhangAdaptingConvolutionalNeural2017}.
+Neural network methods appear to be more successful than traditional image analysis techniques, but work with features that are not directly useful to forensic examiners - a sequence of 500+ numbers with little contextual meaning.
+In this paper, we discuss an alternate approach that uses an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images.
+The output from this model is along the same dimensions as those used by forensics examiners, bridging the gap between theoretical machine learning and practical applications of computer vision techniques to forensics.
 

### 2019-04-12 paper/paper.Rnw

--- 2019-04-08 paper/paper.Rnw
@@
 Class characteristic matches can be used for elimination, but are not sufficient for identification, as the characteristics are shared by all shoes with the same size, make, and model \citep{grossVariabilitySignificanceClass2013,bodziakfootwear2000}.
-Several attempts have been made to automatically identify features in both prints and shoe tread images, ranging from the
+Several attempts have been made to automatically identify features in both prints and shoe tread images, with low-level image analysis methods including fractal decomposition \citep{alexanderAutomaticClassificationRecognition1999}, scale-invariant feature recognition \citep{pavlouAutomaticExtractionClassification2006}, Fourier-Mellin transformation \citep{guehamAutomaticRecognitionShoeprints2008}, and other classical image analysis methods.
+Some of these low-level methods perform relatively well in good conditions \citep{luostarinenMeasuringAccuracyAutomatic2014a} but have degraded performance under conditions which are suboptimal, including those commonly found at crimescenes.
+
+In the wider field of computer vision, attention has turned to more robust methods for image recognition, such as convolutional neural networks (CNNs), which are capable of achieving near-human accuracy under even degraded image conditions\citep{krizhevskyImageNetClassificationDeep2012}.
+These networks are designed to mimic the process of human vision, and typically involve application of sets of filters, some of which mimic the filters used in the low-level techniques used by early automatic classification attempts.
+CNNs have been applied in footwear forensics, making use of more general neural networks optimized to detect objects found in natural scenes, such as trees, animals, buildings, and cars; the pre-trained network's features are then used to compare shoes or prints to determine how well they match \citep{kongCrossDomainImageMatching2019a,kongCrossDomainForensicShoeprint2017,zhangAdaptingConvolutionalNeural2017}.
+Neural network methods appear to be more successful than traditional image analysis techniques, but the features used for matching are not generally informative for humans, who typically classify shoes by patterns and spatial relationships \citep{davisIntelligenceApproachFootwear1981,grossVariabilitySignificanceClass2013}.
+
+In this paper, we discuss an alternate approach that uses an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images.
+Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain.
 
 \section{Materials and Methods}
+\paragraph{VGG16}
+
+\paragraph{LabelMe Shoe Annotations}
+
+\paragraph{Transfer Learning}
 
 \section{Results}
+\paragraph{Overall Accuracy} % Model-wise and class-by-class, w/ confusion matrix
+
+\paragraph{Model Diagnostics}
 
 \section{Conclusions}
+
+\section{Future Work}
 

### 2019-04-22 paper/paper.Rnw
Take a stab at methods

@@
 \section{Materials and Methods}
+\paragraph{Transfer Learning}
+New content: Convolutional neural networks consist of a model base, comprised of feature detectors which are used to map the characteristics of an image, and a classifier (or model head) that uses weights to meaningfully connect image features to classification labels.
+A relatively simple CNN may contain more than 20 million parameters that must be trained on human-labelled data.
+Training a CNN from scratch for a new image classification task requires millions of training images and a substantial amount of computational power.
+(XXX Features are relatively flexible, and new objects are likely some combination of those features.) Thus, although a model base may be trained for one task, it is often easily applied to a new task.
+
+Stolen from CC: Transfer learning is the process of using layers from a CNN (or other classification model) trained on a general image recognition task when fitting a model intended for a more specific purpose.
+The layers which are deemed to identify broadly generalizable patterns are used with their pre-trained weights; new layers are added to customize the model to the specific task at hand, and typically, only these new weights are updated when the model is fit.
+In many cases, the entire model base is used, and fitted with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with diffierent sets of image data.
+Transfer learning allows CNNs to be applied to smaller datasets of several thousand images and also reduces the amount of computational time required to fit the model.
+Transfer learning leverages the modularity of neural networks|the pretrained base of the model can be separated from the full model and a new model head can be trained to connect that base to the output classes.
+
 \paragraph{VGG16}
+Developed by Oxford's Visual Graphics Group, VGG16 is a CNN with 16 ``\functional" (i.e., convolutional and densely connected) layers and 5 ``\structural" max pooling layers.
+In contrast to other popular networks, like ResNet, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy.
+The simplicity of this structure provides the ability to peer into the inner workings of the network for diagnostic purposes, providing a distinct advantage over more complicated network structures with slightly higher accuracy ratings.
+
+Also mention ImageNet somewhere...
 
 \paragraph{LabelMe Shoe Annotations}
-
-\paragraph{Transfer Learning}
+Nine geometric class characteristics were chosen: Circle, ..., and Text.
+Thousands of outsole images were scraped online shoe retail sites.
+These images were then uploaded for use in a tool called LabelMe (Russell et al., 2008), a labeling/annotating interface which allows users to easily select and label regions of an image.
+To date, 4371 shoes have been labeled, yielding 27135 multi-label images.
 

### 2019-04-22 paper/paper.Rnw
transfer learning refs

--- 2019-04-12 paper/paper.Rnw
@@
 \section{Materials and Methods}
+\paragraph{Transfer Learning}
+Neural networks are composed of a set of layers, called the model, base which contain feature detectors, and classifier, the model head, which connect meaningful features to classification labels.
+A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power.
+Assembling data to train a network from scratch is often a nearly impossible task; the process of \emph{transfer learning} is a natural solution to this common problem.
+Transfer learning leverages the modularity of neural networks, that is, that the model base can be separated from the classifier which produces predictions.
+
+Many convolutional neural networks have been trained on a set of images known as ImageNet; these images are hierarchically labeled, so the same image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity.
+CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously.
+As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks.
+
+Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated\citep{oquabLearningTransferringMidlevel2014}.
+In many cases, the entire model base is used, and fitted with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with diffierent sets of image data.
+Transfer learning allows CNNs to be applied to smaller datasets of several thousand images, reduces the amount of computational time required to fit the model, and provides boosted performance compared to training a new model from scratch \citep{yosinskiHowTransferableAre2014}.
+It has been successfully applied in automatic classification of medical images \citep{shinDeepConvolutionalNeural2016} as well as in various applications of shoe forensics \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017}.
+
+In this paper, we will demonstrate the use of transfer learning to automatically identify features in shoe treads which are used by forensic examiners, using a general-purpose CNN with relatively simple structure, VGG16, and a database of labeled shoe tread images we have assembled for this project.
+This approach differs from other approaches \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017} in automatic footwear identification which use the output from the model base directly and do not attempt to add human-friendly contextual information with an additional classifier.
+
 \paragraph{VGG16}
+Developed by Oxford's Visual Graphics Group, VGG16 is a CNN with 16 ``\functional" (i.e., convolutional and densely connected) layers and 5 ``\structural" max pooling layers\citep{krizhevskyImageNetClassificationDeep2012}.
+In contrast to other popular networks, like ResNet, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy.
+The simplicity of this structure provides the ability to peer into the inner workings of the network for diagnostic purposes, providing a distinct advantage over more complicated network structures with slightly higher accuracy ratings.
+
+Also mention ImageNet somewhere...
 
 \paragraph{LabelMe Shoe Annotations}
-
-\paragraph{Transfer Learning}
+Nine geometric class characteristics were chosen: Circle, ..., and Text.
+Thousands of outsole images were scraped online shoe retail sites.
+These images were then uploaded for use in a tool called LabelMe (Russell et al., 2008), a labeling/annotating interface which allows users to easily select and label regions of an image.
+To date, 4371 shoes have been labeled, yielding 27135 multi-label images.
 

### 2019-04-22 paper/paper.Rnw
add commenting, fix errors, add 'todos'

@@
 %%%%%%%%%%%%%%%%%%%%%%%
+
+%%% Commenting %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage[table]{xcolor}
+\newcommand{\mt}[1]{{\color{magenta} #1}}
+\newcommand{\svp}[1]{{\color{teal} #1}}
+\newcommand{\fix}[1]{{\color{orange} #1}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-Many convolutional neural networks have been trained on a set of images known as ImageNet; these images are hierarchically labeled, so the same image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity.
+Many convolutional neural networks have been trained on a set of images known as ImageNet; these images are hierarchically labeled, so the same image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity\fix{cite imagenet here}.
 CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously.
 As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks.
+\fix{add a few VGG16 filters here to demonstrate the generic filters in layers 1-4}
 
@@
 \paragraph{VGG16}
-Developed by Oxford's Visual Graphics Group, VGG16 is a CNN with 16 ``\functional" (i.e., convolutional and densely connected) layers and 5 ``\structural" max pooling layers\citep{krizhevskyImageNetClassificationDeep2012}.
+Developed by Oxford's Visual Graphics Group, VGG16 is a CNN with 16 ``functional" (i.e., convolutional and densely connected) layers and 5 ``structural" max pooling layers\citep{krizhevskyImageNetClassificationDeep2012}.
 In contrast to other popular networks, like ResNet, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy.

### 2019-04-22 paper/paper.Rnw
add images to makefile, add test image to paper
update comments on transfer learning, modify image a bit, and fix bib file

@@
 \paragraph{Transfer Learning}
-Neural networks are composed of a set of layers, called the model, base which contain feature detectors, and classifier, the model head, which connect meaningful features to classification labels.
+\svp{Neural networks are composed of a set of layers, called the model, base which contain feature detectors, and classifier, the model head, which connect meaningful features to classification labels.
 A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power.
 Assembling data to train a network from scratch is often a nearly impossible task; the process of \emph{transfer learning} is a natural solution to this common problem.
-Transfer learning leverages the modularity of neural networks, that is, that the model base can be separated from the classifier which produces predictions.
+Transfer learning leverages the modularity of neural networks, that is, that the model base can be separated from the classifier which produces predictions.}
 
-Many convolutional neural networks have been trained on a set of images known as ImageNet; these images are hierarchically labeled, so the same image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity\fix{cite imagenet here}.
+\svp{Many convolutional neural networks have been trained on a set of images known as ImageNet; these images are hierarchically labeled, so the same image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity\citep{dengImageNetLargeScaleHierarchical}.
 CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously.
-As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks.
-\fix{add a few VGG16 filters here to demonstrate the generic filters in layers 1-4}
+As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks.} \fix{add a few VGG16 filters here to demonstrate the generic filters in layers 1-4}
 
-Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated\citep{oquabLearningTransferringMidlevel2014}.
-In many cases, the entire model base is used, and fitted with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with diffierent sets of image data.
+\svp{Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated\citep{oquabLearningTransferringMidlevel2014}.} In many cases, the entire model base is used, and fitted with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with diffierent sets of image data.
 Transfer learning allows CNNs to be applied to smaller datasets of several thousand images, reduces the amount of computational time required to fit the model, and provides boosted performance compared to training a new model from scratch \citep{yosinskiHowTransferableAre2014}.
@@
 
-In this paper, we will demonstrate the use of transfer learning to automatically identify features in shoe treads which are used by forensic examiners, using a general-purpose CNN with relatively simple structure, VGG16, and a database of labeled shoe tread images we have assembled for this project.
-This approach differs from other approaches \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017} in automatic footwear identification which use the output from the model base directly and do not attempt to add human-friendly contextual information with an additional classifier.
+\svp{In this paper, we will demonstrate the use of transfer learning to automatically identify features in shoe treads which are used by forensic examiners, using a general-purpose CNN with relatively simple structure, VGG16, and a database of labeled shoe tread images we have assembled for this project, which are both described below.
+This approach differs from other approaches \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017} in automatic footwear identification which use the output from the model base directly and do not attempt to add human-friendly contextual information with an additional classifier.}
 
@@
 The simplicity of this structure provides the ability to peer into the inner workings of the network for diagnostic purposes, providing a distinct advantage over more complicated network structures with slightly higher accuracy ratings.
+VGG16 is a common choice for transfer learning because of this structural simplicity\svp{; it has been used for detection of text in natural images \citep{zhongDeepTextUnifiedFramework2016}, medical imaging classification \citep{oquabLearningTransferringMidlevel2014}, classification of weld defects \citep{liuWeldDefectImages2018}, and many other domain-specific image recognition tasks that are more specific than the ImageNet data on which it was trained.
+\autoref{fig:VGG16-structure} shows the architecture of VGG16; with transfer learning, the VGG16 model head is replaced with a trained model head specific to recognition of shoeprint class characteristics.
+}
 
-Also mention ImageNet somewhere...
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{images/vgg16-shoe-nolabel}
+\caption[VGG16 structure]{VGG16 consists of five convolutional blocks that make up the model base.
+Each convolutional block contains an increasing number of increasingly complex features.
+After the convolutional blocks, the fully connected layers of the model head are used to make global connections between separate features.}\label{fig:VGG16-structure}
+\end{figure}
 
-\paragraph{LabelMe Shoe Annotations}
+\svp{In order to train a new head for the VGG16 model base, human-labeled images are necessary. The next section describes how these images were produced and used to fit our custom classifier. }
+
+\paragraph{Annotated Training Data}
+
+\svp{Start with classification scheme. Mention that it was developed in consultation with experts as well as based on past papers and classification schemes.}
+
+\fix{Add classification images}
+
+\svp{Discuss LabelMe and the annotation process. Multiple students working over the course of 6 months... etc.}
+
+\svp{It might be good to add in some (new) statistics relating brand information to common geometric features in tread - show, for instance, the top N brands with bowties vs.
+the top N brands with polygons.
+}
+
 Nine geometric class characteristics were chosen: Circle, ..., and Text.

### 2019-04-27 paper/paper.Rnw
fix zip

--- 2019-04-22 paper/paper.Rnw
@@
 
-\address[1]{195 Durham Center, 613 Morill Rd, Ames, IA 50010}
+\address[1]{195 Durham Center, 613 Morill Rd, Ames, IA 50011}
 \fntext[isu]{Center for Statistical Applications in Forensic Evidence, Iowa State University}
@@
 Class characteristic matches can be used for elimination, but are not sufficient for identification, as the characteristics are shared by all shoes with the same size, make, and model \citep{grossVariabilitySignificanceClass2013,bodziakfootwear2000}.
-Several attempts have been made to automatically identify features in both prints and shoe tread images, with low-level image analysis methods including fractal decomposition \citep{alexanderAutomaticClassificationRecognition1999}, scale-invariant feature recognition \citep{pavlouAutomaticExtractionClassification2006}, Fourier-Mellin transformation \citep{guehamAutomaticRecognitionShoeprints2008}, and other classical image analysis methods.
-Some of these low-level methods perform relatively well in good conditions \citep{luostarinenMeasuringAccuracyAutomatic2014a} but have degraded performance under conditions which are suboptimal, including those commonly found at crimescenes.
-
-In the wider field of computer vision, attention has turned to more robust methods for image recognition, such as convolutional neural networks (CNNs), which are capable of achieving near-human accuracy under even degraded image conditions\citep{krizhevskyImageNetClassificationDeep2012}.
+Several attempts have been made to automatically identify features in both prints and shoe tread images, using a range of low-level techniques including fractal decomposition \citep{alexanderAutomaticClassificationRecognition1999}, scale-invariant feature recognition \citep{pavlouAutomaticExtractionClassification2006}, Fourier-Mellin transformation \citep{guehamAutomaticRecognitionShoeprints2008}, and other classical image analysis methods.
+Vast advances have been made in the field of computer vision since 2012 \citep{krizhevskyImageNetClassificationDeep2012}, when convolutional neural networks (CNNs) were introduced.
 These networks are designed to mimic the process of human vision, and typically involve application of sets of filters, some of which mimic the filters used in the low-level techniques used by early automatic classification attempts.
-CNNs have been applied in footwear forensics, making use of more general neural networks optimized to detect objects found in natural scenes, such as trees, animals, buildings, and cars; the pre-trained network's features are then used to compare shoes or prints to determine how well they match \citep{kongCrossDomainImageMatching2019a,kongCrossDomainForensicShoeprint2017,zhangAdaptingConvolutionalNeural2017}.
-Neural network methods appear to be more successful than traditional image analysis techniques, but the features used for matching are not generally informative for humans, who typically classify shoes by patterns and spatial relationships \citep{davisIntelligenceApproachFootwear1981,grossVariabilitySignificanceClass2013}.
-
-In this paper, we discuss an alternate approach that uses an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images.
+Recently, raw output from CNN models have been used to identify or match shoes to prints \citep{kongCrossDomainImageMatching2019,kongCrossDomainForensicShoeprint2017,zhangAdaptingConvolutionalNeural2017}; these methods work relatively well, but the features used for matching are not generally informative for humans, who typically classify shoes by patterns and spatial relationships \citep{davisIntelligenceApproachFootwear1981,grossVariabilitySignificanceClass2013}.
+In this paper, we discuss a hybrid approach, using features from a pre-trained CNN to develop a classifier which automatically identifies geometric patterns in images of shoe treads.
 Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain.
 
+% This is a really really high level overview - may need to be a bit deeper
+
 \section{Materials and Methods}
-\paragraph{Transfer Learning}
-Neural networks are composed of a set of layers, called the model, base which contain feature detectors, and classifier, the model head, which connect meaningful features to classification labels.
-A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power.
-Assembling data to train a network from scratch is often a nearly impossible task; the process of \emph{transfer learning} is a natural solution to this common problem.
-Transfer learning leverages the modularity of neural networks, that is, that the model base can be separated from the classifier which produces predictions.
-
-Many convolutional neural networks have been trained on a set of images known as ImageNet; these images are hierarchically labeled, so the same image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity.
-CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously.
-As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks.
-
-Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated\citep{oquabLearningTransferringMidlevel2014}.
-In many cases, the entire model base is used, and fitted with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with diffierent sets of image data.
-Transfer learning allows CNNs to be applied to smaller datasets of several thousand images, reduces the amount of computational time required to fit the model, and provides boosted performance compared to training a new model from scratch \citep{yosinskiHowTransferableAre2014}.
-It has been successfully applied in automatic classification of medical images \citep{shinDeepConvolutionalNeural2016} as well as in various applications of shoe forensics \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017}.
-
-In this paper, we will demonstrate the use of transfer learning to automatically identify features in shoe treads which are used by forensic examiners, using a general-purpose CNN with relatively simple structure, VGG16, and a database of labeled shoe tread images we have assembled for this project.
-This approach differs from other approaches \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017} in automatic footwear identification which use the output from the model base directly and do not attempt to add human-friendly contextual information with an additional classifier.
-
 \paragraph{VGG16}
-Developed by Oxford's Visual Graphics Group, VGG16 is a CNN with 16 ``\functional" (i.e., convolutional and densely connected) layers and 5 ``\structural" max pooling layers\citep{krizhevskyImageNetClassificationDeep2012}.
-In contrast to other popular networks, like ResNet, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy.
-The simplicity of this structure provides the ability to peer into the inner workings of the network for diagnostic purposes, providing a distinct advantage over more complicated network structures with slightly higher accuracy ratings.
-
-Also mention ImageNet somewhere...
 
 \paragraph{LabelMe Shoe Annotations}
-Nine geometric class characteristics were chosen: Circle, ..., and Text.
-Thousands of outsole images were scraped online shoe retail sites.
-These images were then uploaded for use in a tool called LabelMe (Russell et al., 2008), a labeling/annotating interface which allows users to easily select and label regions of an image.
-To date, 4371 shoes have been labeled, yielding 27135 multi-label images.
+
+\paragraph{Transfer Learning}
 

### 2019-05-01 paper/paper.Rnw
Try to add class images- no longer compiles :(

--- 2019-04-22 paper/paper.Rnw
@@
 
-In the wider field of computer vision, attention has turned to more robust methods for image recognition, such as convolutional neural networks (CNNs), which are capable of achieving near-human accuracy under even degraded image conditions\citep{krizhevskyImageNetClassificationDeep2012}.
+In the wider field of computer vision, attention has turned to more robust methods for image recognition, such as convolutional neural networks (CNNs), which are capable of achieving near-human accuracy under even degraded image conditions \citep{krizhevskyImageNetClassificationDeep2012}.
 These networks are designed to mimic the process of human vision, and typically involve application of sets of filters, some of which mimic the filters used in the low-level techniques used by early automatic classification attempts.
@@
 \paragraph{Transfer Learning}
-Neural networks are composed of a set of layers, called the model, base which contain feature detectors, and classifier, the model head, which connect meaningful features to classification labels.
-A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power.
+\mt{Neural networks are composed of sets of layers; the early layers, which are called the model base, contain feature detectors, and the final layers make up the classifier, or model head, which connects meaningful features to classification labels.} A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power.
 Assembling data to train a network from scratch is often a nearly impossible task; the process of \emph{transfer learning} is a natural solution to this common problem.
@@
 
-Many convolutional neural networks have been trained on a set of images known as ImageNet; these images are hierarchically labeled, so the same image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity\fix{cite imagenet here}.
+Many convolutional neural networks have been trained on a set of images known as ImageNet; these images are hierarchically labeled, so the same image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity \citep{dengImageNetLargeScaleHierarchical}.
 CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously.
 As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks.
-\fix{add a few VGG16 filters here to demonstrate the generic filters in layers 1-4}
+\fix{add a few VGG16 filters here to demonstrate the generic filters in layers 1-4} \mt{All of them?
+Four per block?}
 
-Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated\citep{oquabLearningTransferringMidlevel2014}.
+Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated \citep{oquabLearningTransferringMidlevel2014}.
 In many cases, the entire model base is used, and fitted with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with diffierent sets of image data.
@@
 
-In this paper, we will demonstrate the use of transfer learning to automatically identify features in shoe treads which are used by forensic examiners, using a general-purpose CNN with relatively simple structure, VGG16, and a database of labeled shoe tread images we have assembled for this project.
+In this paper, we will demonstrate the use of transfer learning to automatically identify features in shoe treads which are used by forensic examiners, using a general-purpose CNN with relatively simple structure, VGG16, and a database of labeled shoe tread images we have assembled for this project, which are both described below.
 This approach differs from other approaches \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017} in automatic footwear identification which use the output from the model base directly and do not attempt to add human-friendly contextual information with an additional classifier.
@@
 The simplicity of this structure provides the ability to peer into the inner workings of the network for diagnostic purposes, providing a distinct advantage over more complicated network structures with slightly higher accuracy ratings.
+VGG16 is a common choice for transfer learning because of this structural simplicity; it has been used for detection of text in natural images \citep{zhongDeepTextUnifiedFramework2016}, medical imaging classification \citep{oquabLearningTransferringMidlevel2014}, classification of weld defects \citep{liuWeldDefectImages2018}, and many other domain-specific image recognition tasks that are more specific than the ImageNet data on which it was trained.
+\autoref{fig:VGG16-structure} shows the architecture of VGG16; with transfer learning, the VGG16 model head is replaced with a trained model head specific to recognition of shoeprint class characteristics.
 
-Also mention ImageNet somewhere...
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{images/vgg16-shoe-nolabel}
+\caption[VGG16 structure]{VGG16 consists of five convolutional blocks that make up the model base.
+Each convolutional block contains an increasing number of increasingly complex features.
+After the convolutional blocks, the fully connected layers of the model head are used to make global connections between separate features.}\label{fig:VGG16-structure}
+\end{figure}
 
-\paragraph{LabelMe Shoe Annotations}
+\svp{In order to train a new head for the VGG16 model base, human-labeled images are necessary. The next section describes how these images were produced and used to fit our custom classifier. }
+
+\paragraph{Annotated Training Data}
+
+\svp{Start with classification scheme. Mention that it was developed in consultation with experts as well as based on past papers and classification schemes.}
+
+\begin{table}
+\centering
+
+\setlength\tabcolsep{1mm}
+\begin{tabular}{rccl}
+     Bowtie & \raisebox{-.5\height}{\includegraphics[width=.3\textwidth]{class_examples/bowtie_examples.png}} &
+     \raisebox{-.5\height}{\includegraphics[width=.3\textwidth]{class_examples/chevron_examples.png}} & Chevron \vspace{1mm}\\
+     Circle & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/circle_examples.png}} &
+     \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/line_examples.png}} & Line  \vspace{1mm}\\
+     Polygon & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/polygon_examples.png}} &
+     \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/quad_examples.png}} & Quad  \vspace{1mm}\\
+     Star & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/star_examples.png}} &
+     \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/text_examples.png}} & Text  \vspace{1mm}\\
+     Triangle & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/triangle_examples.png}} &
+      \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/other_examples.png}} & Other \\
+\end{tabular}
+\caption[Geometric elements used to classify tread patterns.]{A set of geometric elements used to classify tread patterns.
+Categories modified from \cite{grossVariabilitySignificanceClass2013}.}\label{tab:class-char-examples}
+\end{table}
+
+\svp{Discuss LabelMe and the annotation process. Multiple students working over the course of 6 months... etc.}
+
+\svp{It might be good to add in some (new) statistics relating brand information to common geometric features in tread - show, for instance, the top N brands with bowties vs.
+the top N brands with polygons.
+}
+
 Nine geometric class characteristics were chosen: Circle, ..., and Text.

### 2019-05-01 paper/paper.Rnw
At least begin to brainstorm--sorry

@@
 
-\svp{In order to train a new head for the VGG16 model base, human-labeled images are necessary. The next section describes how these images were produced and used to fit our custom classifier. }
+In order to train a new head for the VGG16 model base, human-labeled images are necessary. The next section describes how these images were produced and used to fit our custom classifier.
 
@@
 
-\begin{table}
-\centering
+\mt{\begin{itemize}
+\item Goal: Scheme that is broad enough to classify most shoes and specific enough to identify meaningful similarity between visually similar shoes
+\item Papers (Gross et all, 2013) suggested geometric features have discriminatory power
+\item Multi-label: label all shapes in a region
+\item Modified: e.g., polygon out of pentagon and hexagon
+\item Refer to table
+\end{itemize}}
 
-\setlength\tabcolsep{1mm}
-\begin{tabular}{rccl}
-     Bowtie & \raisebox{-.5\height}{\includegraphics[width=.3\textwidth]{class_examples/bowtie_examples.png}} &
-     \raisebox{-.5\height}{\includegraphics[width=.3\textwidth]{class_examples/chevron_examples.png}} & Chevron \vspace{1mm}\\
-     Circle & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/circle_examples.png}} &
-     \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/line_examples.png}} & Line  \vspace{1mm}\\
-     Polygon & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/polygon_examples.png}} &
-     \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/quad_examples.png}} & Quad  \vspace{1mm}\\
-     Star & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/star_examples.png}} &
-     \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/text_examples.png}} & Text  \vspace{1mm}\\
-     Triangle & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/triangle_examples.png}} &
-      \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/other_examples.png}} & Other \\
-\end{tabular}
-\caption[Geometric elements used to classify tread patterns.]{A set of geometric elements used to classify tread patterns.
+From CC:
+Class characteristics, as defined in Section 1.2, are characteristics which can be used to exclude shoes from a match at a crime scene, but cannot be used for individualized matching because they are shared by many shoes.
+A sufficiently well-defined set of features can separate shoes into make and model categories (Gross et al., 2013).
+Gross et al.
+(2013) define geometric features such as circle/oval, crepe, herringbone, hexagon, parallel lines, logo/lettering/numbering, perimeter lugs, star, and other.
+Working from these categories, a set of categories were assembled that is more suited to recognition by convolutional neural networks, as some of the definitions used in Gross et al.
+(2013) require spatial context which is not preserved during labeling (for example, lugs are required to be on the perimeter of the shoe).
+Table 2.1 shows three examples of each class.
+
+% \begin{table}
+% \centering
+%
+% \setlength\tabcolsep{1mm}
+% \begin{tabular}{rccl}
+%      Bowtie & \raisebox{-.5\height}{\includegraphics[width=.3\textwidth]{class_examples/bowtie_examples.png}} &
+%      \raisebox{-.5\height}{\includegraphics[width=.3\textwidth]{class_examples/chevron_examples.png}} & Chevron \vspace{1mm}\\
+%      Circle & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/circle_examples.png}} &
+%      \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/line_examples.png}} & Line  \vspace{1mm}\\
+%      Polygon & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/polygon_examples.png}} &
+%      \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/quad_examples.png}} & Quad  \vspace{1mm}\\
+%      Star & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/star_examples.png}} &
+%      \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/text_examples.png}} & Text  \vspace{1mm}\\
+%      Triangle & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/triangle_examples.png}} &
+%       \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/other_examples.png}} & Other \\
+% \end{tabular}
+% \caption[Geometric elements used to classify tread patterns.]{A set of geometric elements used to classify tread patterns.
 Categories modified from \cite{grossVariabilitySignificanceClass2013}.}\label{tab:class-char-examples}
-\end{table}
+% \end{table}
 
 \svp{Discuss LabelMe and the annotation process. Multiple students working over the course of 6 months... etc.}
+
+\mt{\begin{itemize}
+\item Mention that they're high school and/or undergrad lackeys? (Jk, students)
+\item Number of shoe images, number of labels
+\item Bar chart?
+\end{itemize}}
+
+\begin{figure}[hbt]
+\centering
+  \includegraphics[width=.9\textwidth]{images/LabelMe2.png}
+  \caption{An example of labeling images with LabelMe}
+\end{figure}
 

### 2019-05-13 paper/paper.Rnw
Add updates to methods

--- 2019-05-01 paper/paper.Rnw
@@
 \paragraph{VGG16}
-Developed by Oxford's Visual Graphics Group, VGG16 is a CNN with 16 ``functional" (i.e., convolutional and densely connected) layers and 5 ``structural" max pooling layers\citep{krizhevskyImageNetClassificationDeep2012}.
+Developed by Oxford's Visual Graphics Group, VGG16 is a CNN with 16 ``functional" (i.e., convolutional and densely connected) layers and 5 ``structural" max pooling layers \citep{krizhevskyImageNetClassificationDeep2012}.
 In contrast to other popular networks, like ResNet, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy.
@@
 
-\svp{In order to train a new head for the VGG16 model base, human-labeled images are necessary. The next section describes how these images were produced and used to fit our custom classifier. }
+<<vgg16layers, fig.width = 7, fig.height = 5.75>>=
+imgs <- list.files("./images/vgg16layers/", recursive = T, full.names = T)
+pngs <- lapply(imgs, function(x) png::readPNG(x)[46:1596, 32:1582, ])
+grobs <- lapply(pngs, grid::rasterGrob)
+gridExtra::grid.arrange(grobs = grobs, nrow = 5)
+@
+
+In order to train a new head for the VGG16 model base, human-labeled images are necessary. The next section describes how these images were produced and used to fit our custom classifier.
 
@@
 \svp{Start with classification scheme. Mention that it was developed in consultation with experts as well as based on past papers and classification schemes.}
+\mt{I don't know much about the consultation with experts...}
+
+\mt{Class characteristics are characteristics which can be used to exclude shoes from a match at a crime scene, but cannot be used for individualized matching because they are shared by many shoes.
+A sufficiently well-defined set of features can separate shoes into make and model categories \citep{grossVariabilitySignificanceClass2013}.
+\citet{grossVariabilitySignificanceClass2013} define geometric features such as circle/oval, crepe, herringbone, hexagon, parallel lines, logo/lettering/numbering, perimeter lugs, star, and other.
+Working from these categories, \fix{in addition to consultation with experts}, a set of categories were assembled that is more suited to recognition by convolutional neural networks, as some of the definitions used in \citet{grossVariabilitySignificanceClass2013} require spatial context which is not preserved during labeling (for example, lugs are required to be on the perimeter of the shoe).
+\autoref{tab:class-char-examples} shows three examples of each class.}
+
+\mt{
+\begin{description}
+\item [Bowtie] Bowtie shapes are roughly quadrilateral, with two opposite concave faces.
+The remaining two faces can be convex or straight, and the concave faces may have straight portions, so long as there is a concave region.
+Using this definition, shapes such as butterflies are included as bowties.
+\item [Chevron] Chevron shapes include repeating parallel lines as well as individual ``v" shapes. They may be angular but can also be curved.
+\item [Circle] Circles include ellipses and ovals; they must be round.
+\item [Line] Lines are repeated and parallel; a more general definition of a line would be difficult to differentiate from many other patterns. Lines can be mildly curved.
+\item [Polygon] Polygons are defined in this standard to have more than 4 sides. They include pentagons, hexagons, and octagons.
+\item [Quadrilateral] Quadrilaterals (quads) have four sides. They may have rounded or square corners.
+\item [Star] Stars are any shape with alternating concave and convex regions, or lines which emanate from a central point. ``X" and ``+" shapes are also classified as stars.
+\item [Text] Text is any shape which would be identified as text by a reasonable human.
+In most cases, the text on the outsole images used is made up of Latin alphabet characters; the model will likely not recognize text in other scripts (but could be trained if non-Latin text images could be obtained).
+\item [Triangle] Triangles are any three-sided figure.
+Like quadrilaterals, they can have rounded corners.
+In some cases, it is difficult to distinguish between a trapezoidal shape and a triangle when rounded corners are involved.
+\item [Other] Other features which were marked include logos, various textures (including crepe, stippling, etc.), and smooth regions with no discernible features.
+These regions are grouped and provide additional information---that none of the previous nine categories are present.
+\end{description}
+}
 
@@
 \begin{tabular}{rccl}
-     Bowtie & \raisebox{-.5\height}{\includegraphics[width=.3\textwidth]{class_examples/bowtie_examples.png}} &
-     \raisebox{-.5\height}{\includegraphics[width=.3\textwidth]{class_examples/chevron_examples.png}} & Chevron \vspace{1mm}\\
-     Circle & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/circle_examples.png}} &
-     \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/line_examples.png}} & Line  \vspace{1mm}\\
-     Polygon & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/polygon_examples.png}} &
-     \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/quad_examples.png}} & Quad  \vspace{1mm}\\
-     Star & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/star_examples.png}} &
-     \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/text_examples.png}} & Text  \vspace{1mm}\\
-     Triangle & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/triangle_examples.png}} &
-      \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{class_examples/other_examples.png}} & Other \\
+     Bowtie & \raisebox{-.5\height}{\includegraphics[width=.3\textwidth]{images/class_examples/bowtie_examples.png}} &
+     \raisebox{-.5\height}{\includegraphics[width=.3\textwidth]{images/class_examples/chevron_examples.png}} & Chevron \vspace{1mm}\\
+     Circle & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{images/class_examples/circle_examples.png}} &
+     \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{images/class_examples/line_examples.png}} & Line  \vspace{1mm}\\
+     Polygon & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{images/class_examples/polygon_examples.png}} &
+     \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{images/class_examples/quad_examples.png}} & Quad  \vspace{1mm}\\
+     Star & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{images/class_examples/star_examples.png}} &
+     \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{images/class_examples/text_examples.png}} & Text  \vspace{1mm}\\
+     Triangle & \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{images/class_examples/triangle_examples.png}} &
+      \raisebox{-.5\height}{\includegraphics[width=0.3\textwidth]{images/class_examples/other_examples.png}} & Other \\
 \end{tabular}
@@
 
-\svp{Discuss LabelMe and the annotation process. Multiple students working over the course of 6 months... etc.}
+\mt{Thousands of outsole images were scraped online shoe retail sites.
+These images were then uploaded for use in a tool called LabelMe (Russell et al., 2008), a labeling/annotating interface which allows users to easily select and label regions of an image.
+Over the course of six months, multiple students worked to label images of shoes according to the classification scheme using LabelMe.
+
+After annotation using the LabelMe software package, images are processed by an R script that identifies the minimum bounding rectangle of the region, crops the image to that region, and scales the cropped area to a 256 x 256 pixel image suitable for analysis by the convolutional neural network.
+During this process, aspect ratio is not preserved, though efforts are made to label regions
+which are relatively square to minimize the effect of this distortion. To date, 4371 shoes have been labeled, yielding 27135 multi-label images.}
+
+\begin{figure}[hbt]
+\centering
+  \includegraphics[width=.9\textwidth]{images/LabelMe2.png}
+  \caption{An example of labeling images with LabelMe}
+\end{figure}
+
+<<class-characteristic-barchart, eval = F, fig.width = 6, fig.height = 4, dpi = 300, fig.cap = "Distribution of classes in all labeled images.
+Quadrilaterals, lines, circles, text, and chevrons are relatively common; stars, polygons, and bowties are relatively rare.", fig.scap = "Distribution of classes in all labeled images.">>=
+annotated_imgs %>%
+  mutate(label_type = str_replace(label_type, "multi", "multiple")) %>%
+  mutate(labels = str_to_title(labels)) %>%
+  mutate(labels = factor(labels, levels = c("Quad", "Line", "Text", "Circle",
+                                            "Chevron", "Triangle", "Polygon",
+                                            "Star", "Bowtie", "Other"))) %>%
+  ggplot() +
+  geom_bar(aes(x = labels, fill = label_type), color = "black") +
+  scale_fill_manual("Labels", values = c("single" = "#6ba2b9",
+                                         "multiple" = "#2e5597")) +
+  coord_flip() +
+  ylab("# Labeled Images") +
+  xlab("")  +
+  ggtitle("Current Class Distribution (All Labeled Images)") +
+  mytheme +
+  theme(legend.position = c(1, 1),
+        legend.justification = c(1.03, 1.03),
+        legend.background = element_rect(fill = "white"))
+@
 
@@
 
-Nine geometric class characteristics were chosen: Circle, ..., and Text.
-Thousands of outsole images were scraped online shoe retail sites.
-These images were then uploaded for use in a tool called LabelMe (Russell et al., 2008), a labeling/annotating interface which allows users to easily select and label regions of an image.
-To date, 4371 shoes have been labeled, yielding 27135 multi-label images.
+\mt{Augmentation? Model training parameters?}
 
@@
 \paragraph{Overall Accuracy} % Model-wise and class-by-class, w/ confusion matrix
+
+\fix{Are we setting up this paper to generate images live, or pasting in static versions?}
+
+\mt{Figure 3.2 shows the ROC curve for the full model, and Figure 3.3 shows the curve for each class.
+The full model has an AUC of 0.88, and the AUC for individual classes ranges from 0.81 (for line) to 0.91 (for bowtie and text).
+While the class performances do vary slightly, each ROC curve is the same general shape and performs significantly better than random chance.}
 

### 2019-05-13 paper/paper.Rnw
slight changes

@@
 \paragraph{Transfer Learning}
-\mt{Neural networks are composed of sets of layers; the early layers, which are called the model base, contain feature detectors, and the final layers make up the classifier, or model head, which connects meaningful features to classification labels.} A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power.
+Neural networks are composed of sets of layers; the early layers, which are called the model base, contain feature detectors, and the final layers make up the classifier, or model head, which connects meaningful features to classification labels.
+A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power.
 Assembling data to train a network from scratch is often a nearly impossible task; the process of \emph{transfer learning} is a natural solution to this common problem.
@@
 
-\svp{Start with classification scheme. Mention that it was developed in consultation with experts as well as based on past papers and classification schemes.}
-\mt{I don't know much about the consultation with experts...}
+% \svp{Start with classification scheme. Mention that it was developed in consultation with experts as well as based on past papers and classification schemes.}
+% \mt{I don't know much about the consultation with experts...}
 
-\mt{Class characteristics are characteristics which can be used to exclude shoes from a match at a crime scene, but cannot be used for individualized matching because they are shared by many shoes.
-A sufficiently well-defined set of features can separate shoes into make and model categories \citep{grossVariabilitySignificanceClass2013}.
-\citet{grossVariabilitySignificanceClass2013} define geometric features such as circle/oval, crepe, herringbone, hexagon, parallel lines, logo/lettering/numbering, perimeter lugs, star, and other.
-Working from these categories, \fix{in addition to consultation with experts}, a set of categories were assembled that is more suited to recognition by convolutional neural networks, as some of the definitions used in \citet{grossVariabilitySignificanceClass2013} require spatial context which is not preserved during labeling (for example, lugs are required to be on the perimeter of the shoe).
+Class characteristics are characteristics which can be used to exclude shoes from a match at a crime scene, but cannot be used for individualized matching because they are shared by many shoes \svp{(e.g.
+all shoes of a specific brand and size)}.
+A sufficiently well-defined set of features can be used to separate shoes into make and model categories \citep{grossVariabilitySignificanceClass2013}; the set of features used in that study included circle/oval, crepe, herringbone, hexagon, parallel lines, logo/lettering/numbering, perimeter lugs, star, and other.
+After consulting with practitioners, we developed a set of categories suitable for automatic recognition by convolutional neural networks.
+These modifications were necessary because some of the definitions used in \citet{grossVariabilitySignificanceClass2013} require contextual spatial information which is not preserved during labeling (for example, lugs are required to be on the perimeter of the shoe).
 \autoref{tab:class-char-examples} shows three examples of each class.}
 
-\mt{
+The categories we use in this study are operationally defined as follows:
 \begin{description}
@@
 Using this definition, shapes such as butterflies are included as bowties.
+\svp{I'm still on the fence - we can just not mention this, or we could move butterflies to "star"...}
 \item [Chevron] Chevron shapes include repeating parallel lines as well as individual ``v" shapes. They may be angular but can also be curved.
@@
 \end{description}
-}
 
@@
 \centering
-
 \setlength\tabcolsep{1mm}

### 2019-05-13 paper/paper.Rnw
a few more changes

@@
 
-\mt{Thousands of outsole images were scraped online shoe retail sites.
-These images were then uploaded for use in a tool called LabelMe (Russell et al., 2008), a labeling/annotating interface which allows users to easily select and label regions of an image.
-Over the course of six months, multiple students worked to label images of shoes according to the classification scheme using LabelMe.
-
-After annotation using the LabelMe software package, images are processed by an R script that identifies the minimum bounding rectangle of the region, crops the image to that region, and scales the cropped area to a 256 x 256 pixel image suitable for analysis by the convolutional neural network.
-During this process, aspect ratio is not preserved, though efforts are made to label regions
-which are relatively square to minimize the effect of this distortion. To date, 4371 shoes have been labeled, yielding 27135 multi-label images.}
+\mt{Thousands of outsole images were scraped online shoe retail sites} \svp{and annotated using LabelMe, a tool for image annotation in computer vision problems\citep{labelme}}.
+After annotation, images are processed by an R script that identifies the minimum bounding rectangle of the region, crops the image to that region, and scales the cropped area to a 256 x 256 pixel image suitable for analysis by the convolutional neural network.
+During this process, aspect ratio is not preserved, though efforts are made to label regions which are relatively square to minimize the effect of this distortion.
+To date, 4371 shoes have been labeled, yielding 27135 multi-label images.}\fix{actual counts}
 
@@
 
-\mt{Augmentation? Model training parameters?}
+\mt{Augmentation? Model training parameters?}\svp{VERY briefly mention augmentation. Make sure to explain the difference between validation, training, and test sets}.
 

### 2019-05-20 paper/paper.Rnw
Get live results working

@@
 knitr::opts_chunk$set(echo = F, dev = "pdf", dpi = 1000)
+
+library(tidyverse)
+library(magrittr)
+library(tidyverse)
+library(ggplot2)
+library(keras)
+
+mytheme <- theme_bw()
+theme_set(mytheme)
+@
+
+<<results_setup, include = F>>=
+if("paper" %in% list.files()) {
+  wd <- file.path(getwd(), "paper")
+} else {
+  wd <- getwd()
+}
+
+codedir <- file.path(wd, "code")
+modeldir <- file.path(wd, "model")
+
+source(file.path(codedir, "Generate_Model_Images.R"))
+@
+
+<<labeled-data-setup, include = F>>=
+model_path <- file.path(modeldir, "TrainedModels")
+
+newest_model <- get_newest(dir = model_path, pattern = "weights.h5")
+newest_data_file <- file.path(modeldir, "RProcessedImages",
+                              newest_model$process_dir, "cropped_photos.Rdata")
+
+load(newest_data_file)
+source(file.path(codedir, "count_images.R"))
+@
+
+<<load-keras-model, include = F, eval = T, echo = F>>=
+model_dir <- newest_model$path
+load(list.files(model_dir, pattern = "-history.Rdata", full.names = T)[1])
+load(file.path(get_newest()$path, get_newest(pattern = "\\d.Rdata")$base_file))
 @
@@
 
-<<class-characteristic-barchart, eval = F, fig.width = 6, fig.height = 4, dpi = 300, fig.cap = "Distribution of classes in all labeled images.
+<<class-characteristic-barchart, fig.width = 6, fig.height = 4, dpi = 300, fig.cap = "Distribution of classes in all labeled images.
 Quadrilaterals, lines, circles, text, and chevrons are relatively common; stars, polygons, and bowties are relatively rare.", fig.scap = "Distribution of classes in all labeled images.">>=

### 2019-05-20 paper/code/Generate_Model_Images.R

@@ -1,0 +1,599 @@
+library(magrittr)
+library(purrr)
+library(dplyr)
+library(lubridate)
+library(stringr)
+library(jpeg)
+library(imager)
+library(tidyr)
+library(magick)
+library(viridis)
+library(ggcorrplot)
+library(deepviz)
+library(tidygraph)
+library(ggraph)
+library(keras)
+
+# --- Variables ----------------------------------------------------------------
+default_classes <- c(
+  "bowtie", "chevron", "circle", "line", "polygon",
+  "quad", "star", "text", "triangle"
+)
+
+# --- Functions ----------------------------------------------------------------
+
+# Get most recent model
+get_newest <- function(dir = file.path(modeldir, "TrainedModels"), pattern = "fullimage.rdata") {
+  dirlist <- list.dirs(path = dir) %>% rev()
+  newest_file <- NULL
+  startidx <- 1:pmin(length(dirlist) - 1, 10)
+  while (is.null(newest_file) & max(startidx) <= length(dirlist)) {
+    newest_file <- dirlist %>%
+      magrittr::extract(startidx) %>%
+      sapply(function(x) list.files(x, pattern = pattern, recursive = F, full.names = T)) %>%
+      unlist() %>%
+      as.character()
+
+    mtimes <- file.mtime(newest_file)
+
+    newest_file <- newest_file[which.max(mtimes)]
+
+    startidx <- pmin(max(startidx) + 1:10, length(dirlist))
+  }
+
+  path <- dirname(newest_file)
+  base_file <- basename(newest_file)
+  start_date <- str_extract(base_file, "\\d{4}-\\d{2}-\\d{2}[_ ]\\d{2}:\\d{2}:\\d{2}")
+  prefix <- str_remove(base_file, "-history.Rdata|-weights.h5|.h5|.Rdata|fullimage.rdata") %>%
+    str_remove(start_date) %>%
+    str_remove("^_")
+  process_dir <- basename(path)
+  return(list(path = path, base_file = base_file, prefix = prefix, start_date = start_date, process_dir = process_dir))
+}
+
+set_weights <- function(model_wts_file) {
+  input <- layer_input(shape = c(256, 256, 3))
+
+  conv_base <- application_vgg16(
+    weights = "imagenet",
+    include_top = FALSE,
+    input_tensor = input)
+
+  output <- conv_base$output %>%
+    layer_flatten(input_shape = input) %>%
+    layer_dense(units = 256,
+                activation = "relu",
[…]

### 2019-05-20 paper/code/count_images.R

@@ -1,0 +1,97 @@
+library(magrittr)
+library(tidyverse)
+library(ggplot2)
+library(keras)
+
+label_fixes <- c("triangles" = "triangle",
+                 "trianglee" = "triangle",
+                 "circlemtraignle" = "circle_triangle",
+                 "circle triangle" = "circle_triangle",
+                 "circleline" = "circle_line",
+                 "circle text" = "circle_text",
+                 "circle_elongated" = "circle",
+                 "cricle|cirle" = "circle",
+                 "chrevron" = "chevron",
+                 "lie" = "line",
+                 "texxt" = "text",
+                 "star quad" = "star_quad",
+                 "exc_idd" = "exclude",
+                 "quad?" = "quad",
+                 "qaud|qud" = "quad",
+                 "stars|start" = "star",
+                 "exlude" = "exclude",
+                 "rounded" = "",
+                 "smooth_texture" = "other",
+                 "octagon" = "polygon",
+                 "crepe" = "other",
+                 "hex" = "polygon",
+                 "smooth|hatching" = "other"
+)
+
+annotated_imgs <- select(dfunion, image, name) %>%
+  mutate(base_image = basename(image)) %>%
+  mutate(num_labels = str_count(name, "_") + 1,
+         annot_num = 1:n()) %>%
+  mutate(labels = str_remove_all(name, "\\(.{1,2}\\)") %>%
+           str_replace_all(label_fixes) %>%
+           str_split("_")) %>%
+  unnest(labels) %>%
+  mutate(labels = str_replace_all(labels, label_fixes)) %>%
+  mutate(label_type = ifelse(num_labels == 1, "single", "multi")) %>%
+  filter(labels != "") %>%
+  filter(labels != "exclude") %>%
+  filter(labels != "ribbon" & labels != "logo") %>%
+  filter(labels %in% c(default_classes, "other"))
+
+fix_annotation <- function(annot) {
+  useful_vars <- c("name", "deleted", "date", "id", "username")
+  new_vars <- set_names(useful_vars, c("name", "deleted", "date", "objID", "username"))
+
+  get_vars <- new_vars[new_vars %in% names(annot)]
+
+  if (!is.null(annot)) select(annot, !!get_vars) %>% as_tibble() else tibble()
+}
+
+# authors <- df %>%
[…]

### 2019-05-20 paper/code/vgg16_augment_onehot_automation.R

@@ -1,0 +1,229 @@
+library(magrittr)
+library(lubridate)
+library(stringr)
+library(jpeg)
+library(keras)
+
+use_backend("tensorflow")
+# install_keras()
+if (!exists("classes")) {
+  classes <- c(
+    "bowtie", "chevron", "circle", "line", "polygon",
+    "quad", "star", "text", "triangle"
+  )
+}
+
+if (!exists("modeldir")) {
+  modeldir <- here::here("model")
+}
+
+if (!exists("process_dir")) {
+  process_dir <- list.files(file.path(modeldir, "RProcessedImages")) %>%
+    as_datetime() %>%
+    max(na.rm = T) %>%
+    gsub("[^0-9\\ ]", "", .) %>%
+    gsub(" ", "-", .)
+} else {
+  process_dir <- file.path(modeldir, "RProcessedImages", process_dir)
+}
+
+if (!exists("aug_multiple")) {
+  aug_multiple <- 2
+}
+
+if (!exists("epochs")) {
+  epochs <- 30
+}
+
+process_dir <- gsub("[[:punct:]]models[[:punct:]]RProcessedImages[[:punct:]]{1,}", "\\1", process_dir)
+process_dir <- gsub("^[/\\\\]{1,}", "", process_dir)
+
+
+work_dir <- file.path(modeldir, "TrainedModels")
+start_date <- Sys.time() %>% gsub(" ", "_", .)
+model_dir <- file.path(work_dir, process_dir)
+dir.create(model_dir)
+
+name_file <- function(date, ext) {
+  pretrained_base <- "vgg16"
+  mod_type <- "onehotaug"
+  nclass <- paste0(length(classes), "class")
+  pixel_size <- "256"
+
+  filename <- paste(date, pretrained_base, mod_type, nclass,
+    pixel_size,
+    sep = "_"
+  )
+
+  file.path(model_dir, filename) %>%
+    paste0(., ext)
+}
+
+
+base_dir <- file.path(modeldir, "RProcessedImages", process_dir)
+train_dir <- file.path(base_dir, "train")
+train_aug_dir <- file.path(base_dir, "train")
+validation_dir <- file.path(base_dir, "validation")
+test_dir <- file.path(base_dir, "test")
+
+n_train <- length(list.files(train_dir))
+n_validation <- length(list.files(validation_dir))
+n_test <- length(list.files(test_dir))
+
[…]

### 2019-05-20 paper/paper.Rnw
updates to paper to make sure it compiles

@@
 <<results_setup, include = F>>=
-if("paper" %in% list.files()) {
+if ("paper" %in% list.files()) {
   wd <- file.path(getwd(), "paper")
@@
 These modifications were necessary because some of the definitions used in \citet{grossVariabilitySignificanceClass2013} require contextual spatial information which is not preserved during labeling (for example, lugs are required to be on the perimeter of the shoe).
-\autoref{tab:class-char-examples} shows three examples of each class.}
+\autoref{tab:class-char-examples} shows three examples of each class.
 
@@
 \mt{Thousands of outsole images were scraped online shoe retail sites} \svp{and annotated using LabelMe, a tool for image annotation in computer vision problems\citep{labelme}}.
-After annotation, images are processed by an R script that identifies the minimum bounding rectangle of the region, crops the image to that region, and scales the cropped area to a 256 x 256 pixel image suitable for analysis by the convolutional neural network.
+\mt{After annotation, images are processed by an R script that identifies the minimum bounding rectangle of the region, crops the image to that region, and scales the cropped area to a 256 x 256 pixel image suitable for analysis by the convolutional neural network.
 During this process, aspect ratio is not preserved, though efforts are made to label regions which are relatively square to minimize the effect of this distortion.

### 2019-05-20 paper/paper.Rnw
Push changes

--- 2019-05-20 paper/paper.Rnw
@@
 
-<<vgg16layers, fig.width = 7, fig.height = 5.75>>=
+<<vgg16layers, fig.width = 7, fig.height = 5.75, fig.cap = "A selection of filters from the convolutional layers of VGG16.">>=
 imgs <- list.files("./images/vgg16layers/", recursive = T, full.names = T)
@@
 
-\paragraph{Annotated Training Data}
+\paragraph{Annotated Training Data
 
-% \svp{Start with classification scheme. Mention that it was developed in consultation with experts as well as based on past papers and classification schemes.}
-% \mt{I don't know much about the consultation with experts...}
-
-Class characteristics are characteristics which can be used to exclude shoes from a match at a crime scene, but cannot be used for individualized matching because they are shared by many shoes \svp{(e.g.
-all shoes of a specific brand and size)}.
+Class characteristics are characteristics which can be used to exclude shoes from a match at a crime scene, but cannot be used for individualized matching because they are shared by many shoes (e.g., all shoes of a specific brand and size).
 A sufficiently well-defined set of features can be used to separate shoes into make and model categories \citep{grossVariabilitySignificanceClass2013}; the set of features used in that study included circle/oval, crepe, herringbone, hexagon, parallel lines, logo/lettering/numbering, perimeter lugs, star, and other.
@@
 
-\mt{Thousands of outsole images were scraped online shoe retail sites} \svp{and annotated using LabelMe, a tool for image annotation in computer vision problems\citep{labelme}}.
+Thousands of outsole images were scraped from online shoe retail sites and annotated using LabelMe, a tool for image annotation in computer vision problems \citep{labelme}.
 After annotation, images are processed by an R script that identifies the minimum bounding rectangle of the region, crops the image to that region, and scales the cropped area to a 256 x 256 pixel image suitable for analysis by the convolutional neural network.
 During this process, aspect ratio is not preserved, though efforts are made to label regions which are relatively square to minimize the effect of this distortion.
-To date, 4371 shoes have been labeled, yielding 27135 multi-label images.}\fix{actual counts}
+To date, \Sexpr{unique(ann_df$base_image) %>% length()} shoes have been labeled, yielding \Sexpr{nrow(dfunion)} multi-label images.
 
@@
 
-\fix{Are we setting up this paper to generate images live, or pasting in static versions?}
-
 \mt{Figure 3.2 shows the ROC curve for the full model, and Figure 3.3 shows the curve for each class.
@@
 
-\section*{References}
-
 \bibliography{mybibfile}

### 2019-05-20 paper/paper.Rnw
Fix parentheses

@@
 
-\paragraph{Annotated Training Data
+\paragraph{Annotated Training Data}
 

### 2019-05-24 paper/paper.Rnw
Add augmentation and training/val explanation

@@
 
-\mt{Augmentation? Model training parameters?}\svp{VERY briefly mention augmentation. Make sure to explain the difference between validation, training, and test sets}.
+\mt{Labeled images are scarce relative to the amount of data necessary to train a neural network.
+One solution to this scarcity is to artificially enlarge the data set using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}.
+Augmentation is the transformation of original input data using image operations such as cropping, zoom, skew, rotation, and color balance modification in order to distort or alter the image while maintaining the essential features corresponding to the label.
+This process reduces the potential for overfitting the model to the specific set of image data used during the training process, and also increases the amount of data available for training; examples of pre- and post-augmentation images are shown in \autoref{fig:augmented}.}
+
+\begin{figure}
+\centering
+\includegraphics[width=.1\textwidth]{augmentation/bowtie(R)-1-ugg-sienna-enamel-blue_product_8726365_color_120557.jpg}\hfill
+\includegraphics[width=.1\textwidth]{augmentation/aug_bowtie(R)-1-ugg-sienna-enamel-blue_product_8726365_color_120557_0_8344.jpg}\hfill\hfill\hfill
+\includegraphics[width=.1\textwidth]{augmentation/quad-2-nike-kids-downshifter-7-infant-toddler-gunsmoke-sunset-pulse-atmosphere-grey_product_8800875_color_720973.jpg}\hfill
+\includegraphics[width=.1\textwidth]{augmentation/aug_quad-2-nike-kids-downshifter-7-infant-toddler-gunsmoke-sunset-pulse-atmosphere-grey_product_8800875_color_720973_0_7533.jpg}\hfill\hfill\hfill
+\includegraphics[width=.1\textwidth]{augmentation/quad-2-puma-suede-classic-x-mac-three-port-royale-port-royale_product_9123838_color_251426.jpg}\hfill
+\includegraphics[width=.1\textwidth]{augmentation/aug_quad-2-puma-suede-classic-x-mac-three-port-royale-port-royale_product_9123838_color_251426_0_5689.jpg}\hfill\hfill\hfill
+\includegraphics[width=.1\textwidth]{augmentation/text-1-bernie-mev-kids-catwalk-little-kid-big-kid-multi-camo_product_9084647_color_87089.jpg}\hfill
+\includegraphics[width=.1\textwidth]{augmentation/aug_text-1-bernie-mev-kids-catwalk-little-kid-big-kid-multi-camo_product_9084647_color_87089_0_1891.jpg}
+\caption[Original and augmented images.]{Four sets of original (left) and augmented (right) labeled images.}\label{fig:augmented}
+\end{figure}
+
+Model training is analogous to the human learning process.
+A child learns to identify dogs by being presented with many labeled examples, where the parent uses the label "dog" for an animal walking by.
+That child's understanding of dogs is then measured by how many dogs the child is able to correctly identify, and also by the number of cats it mistakenly calls "dog".
+Similarly, CNN training is a series of stages where, in each, the model learns from a set of labeled training data and that learning is  evaluated through intermediate predictions of validation data.
+
+The \Sexpr{nrow(dfunion)} images were split such that 60\% were used for training.
+Since the categories do not exist in equal proportion in the labeled data, the training data were weighted by proportion during the training process to prevent the loss function from being overwhelmed by more frequent categories.
+Of the remaining 40\% of data, half were used for validation, to monitor the training process, and the remaining data were for testing the performance of the fitted model.
 

### 2019-05-31 paper/paper.Rnw
Center figures

--- 2019-05-20 paper/paper.Rnw
@@
 <<setup, include = F>>=
-knitr::opts_chunk$set(echo = F, dev = "pdf", dpi = 1000)
+knitr::opts_chunk$set(echo = F, dev = "pdf", dpi = 1000, message = F, warning = F, fig.align = "center")
 
@@
 \begin{keyword}
-footwear,class characteristics,computer vision, neural networks
+footwear, class characteristics, computer vision, neural networks
 \end{keyword}
@@
 As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks.
-\fix{add a few VGG16 filters here to demonstrate the generic filters in layers 1-4} \mt{All of them?
-Four per block?}
+\autoref{fig:vgg16layers} \fix{shows a selection of filters from VGG16, where each filter detects the presence of specific shapes or colors.
+Filters in earlier convolutional blocks detect simple features, such as colors and blobs, while later filters detect more complex combinations of features.
+Thus, a single image is made up of some combination of the hundreds of features that are detected by the filters in a CNN.}
 
@@
 Using this definition, shapes such as butterflies are included as bowties.
-\svp{I'm still on the fence - we can just not mention this, or we could move butterflies to "star"...}
+\svp{I'm still on the fence - we can just not mention this, or we could move butterflies to "star"...} \mt{I think it makes a lot of sense to move them to star.
+That means redoing the poster/slides examples though...
+grrrrrr}
 \item [Chevron] Chevron shapes include repeating parallel lines as well as individual ``v" shapes. They may be angular but can also be curved.
@@
 
-\mt{Augmentation? Model training parameters?}\svp{VERY briefly mention augmentation. Make sure to explain the difference between validation, training, and test sets}.
+Labeled images are scarce relative to the amount of data necessary to train a neural network.
+One solution to this scarcity is to artificially enlarge the data set using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}.
+Augmentation is the transformation of original input data using image operations such as cropping, zoom, skew, rotation, and color balance modification in order to distort or alter the image while maintaining the essential features corresponding to the label.
+This process reduces the potential for overfitting the model to the specific set of image data used during the training process, and also increases the amount of data available for training; examples of pre- and post-augmentation images are shown in \autoref{fig:augmented}.
+
+\begin{figure}
+\centering
+\includegraphics[width=.1\textwidth]{images/augmentation/bowtie(R)-1-ugg-sienna-enamel-blue_product_8726365_color_120557.jpg}\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/aug_bowtie(R)-1-ugg-sienna-enamel-blue_product_8726365_color_120557_0_8344.jpg}\hfill\hfill\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/quad-2-nike-kids-downshifter-7-infant-toddler-gunsmoke-sunset-pulse-atmosphere-grey_product_8800875_color_720973.jpg}\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/aug_quad-2-nike-kids-downshifter-7-infant-toddler-gunsmoke-sunset-pulse-atmosphere-grey_product_8800875_color_720973_0_7533.jpg}\hfill\hfill\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/quad-2-puma-suede-classic-x-mac-three-port-royale-port-royale_product_9123838_color_251426.jpg}\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/aug_quad-2-puma-suede-classic-x-mac-three-port-royale-port-royale_product_9123838_color_251426_0_5689.jpg}\hfill\hfill\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/text-1-bernie-mev-kids-catwalk-little-kid-big-kid-multi-camo_product_9084647_color_87089.jpg}\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/aug_text-1-bernie-mev-kids-catwalk-little-kid-big-kid-multi-camo_product_9084647_color_87089_0_1891.jpg}
+\caption[Original and augmented images.]{Four sets of original (left) and augmented (right) labeled images.}\label{fig:augmented}
+\end{figure}
+
+\mt{Model training is analogous to the human learning process.
+For example, a child learns to identify dogs by being presented with many labeled examples, such as when their parent uses the label ``dog" for an animal walking by.
+That child's understanding of dogs is then measured by how many dogs the child is able to correctly identify, and also by the number of cats it mistakenly calls ``dog".
+Similarly, CNN training is a series of stages, or ``epochs", where the model learns from the set of labeled training data, and that learning is measured through intermediate predictions of validation data.}
+
+The \Sexpr{nrow(dfunion)} images were split such that 60\% were used for training.
+Since the categories do not exist in equal proportion in the labeled data, the training data were weighted by proportion during the training process to prevent the loss function from being overwhelmed by more frequent categories.
+Of the remaining 40\% of data, half were used for validation, to monitor the training process, and the remaining data were for testing the performance of the fitted model.
 
 \section{Results}
+
+\subsection{Model Training}
+<<training-accuracy, fig.width = 7.5, fig.height = 5, out.width = ".75\\textwidth", fig.cap = "Training and validation accuracy and loss for each epoch of the fitting process.
+Training and validation accuracy reach 89.5\\% around epoch 9.
+After that point, validation loss remains the same and training loss decreases slightly, while validation accuracy increases more slowly than training accuracy.", fig.scap = "Training and validation accuracy and loss during each epoch.", fig.pos = 'h'>>=
+data.frame(history$metrics) %>%
+  mutate(epoch = 1:n()) %>%
+  gather(key = "measure", value = "value", -epoch) %>%
+  mutate(Type = ifelse(str_detect(measure, "val"), "Validation", "Training"),
+         measure = ifelse(str_detect(measure, "acc"), "Accuracy", "Loss")) %>%
+  # bind_rows(tibble(epoch = NA, value =  .6, measure = "Accuracy", Type = "Validation")) %>%
+  # bind_rows(tibble(epoch = NA, value =  .33, measure = "Loss", Type = "Validation")) %>%
+  ggplot(aes(x = epoch, y = value, color = Type)) +
+  geom_point() +
+  geom_smooth(se = F) +
+  facet_grid(measure~., scales = "free_y", switch = "both") +
+  theme_bw() +
+  scale_y_continuous("") +
+  scale_x_continuous("Epoch") +
+  ggtitle("CoNNOR Training Performance") + mytheme +
+  theme(axis.title.y = element_blank(), legend.position = c(1, .5),
+        legend.justification = c(1.03, -0.05),
+        legend.background = element_rect(fill = "white"))
+@
+
+\autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process.
+Overfitting, or fitting a model which performs too well on the training data relative to the validation data, is seen when the validation loss starts to increase after reaching a global minimum.
+Alternately, underfitting occurs when the validation accuracy is still increasing when model optimization is terminated.
+Neither of these outcomes appears in \autoref{fig:training-accuracy}, indicating that the model optimization process was halted at an appropriate epoch.
+
+\FloatBarrier
+
+\subsection{Model Accuracy}\label{sec:model-accuracy}
+
 \paragraph{Overall Accuracy} % Model-wise and class-by-class, w/ confusion matrix
+\autoref{fig:roc-code} shows the ROC curve for the full model, and \autoref{fig:auc-code} shows the curve for each class.
+The full model has an AUC of 0.88, and the AUC for individual classes ranges from 0.81 (for line) to 0.91 (for bowtie and text).
+While the class performances do vary slightly, each ROC curve is the same general shape and performs significantly better than random chance.
 
-\mt{Figure 3.2 shows the ROC curve for the full model, and Figure 3.3 shows the curve for each class.
-The full model has an AUC of 0.88, and the AUC for individual classes ranges from 0.81 (for line) to 0.91 (for bowtie and text).
-While the class performances do vary slightly, each ROC curve is the same general shape and performs significantly better than random chance.}
+<<roc-code, fig.width = 5, fig.height = 5, out.width = ".5\\textwidth", fig.cap = "ROC curve showing overall model performance.", fig.scap = "Overall model performance ROC curve.">>=
+library(pROC)
+pred_df <- as_tibble(preds) %>% gather(key = feature, value = value)
+test_labs_df <- as_tibble(test_labs) %>% gather(key = feature, value = value)
+whole_model_roc <- roc(test_labs_df$value, pred_df$value)
+
+whole_model_roc_df <- tibble(tpr = whole_model_roc$sensitivities,
+                             fpr = 1 - whole_model_roc$specificities,
+                             thresholds = whole_model_roc$thresholds,
+                             auc = whole_model_roc$auc[1]) %>%
+  nest(tpr, fpr, thresholds, .key =  "roc_plot") %>%
+  mutate(eer = purrr::map(roc_plot, eer))
+ggplot() +
+  geom_line(aes(x = fpr, y = tpr), data = unnest(whole_model_roc_df, roc_plot), size = 1.25) +
+  geom_label(aes(x = 1, y = .07, label = sprintf("AUC: %0.2f", auc)), hjust = 1, vjust = -0.2, data = whole_model_roc_df) +
+  geom_point(aes(x = fpr, y = tpr, color = "Equal Error Rate"), data = unnest(whole_model_roc_df, eer), size = 2) +
+  scale_color_manual("", values = "black") +
+  scale_x_continuous("False Positive Rate", breaks = c(0, .25, .5, .75, 1), labels = c("0.0", "", "0.5", "", "1.0")) +
+  scale_y_continuous("True Positive Rate", breaks = c(0, .25, .5, .75, 1), labels = c("0.0", "", "0.5", "", "1.0")) +
+  ggtitle("CoNNOR Test Set Performance (All Classes)") +
+  coord_fixed() + mytheme +
+  theme(legend.position = c(1, 0), legend.justification = c(1.01, -0.01), legend.title = element_blank(), legend.background = element_rect(fill = "white"))
+
+@
+
+<<auc-code, fig.width = 8, fig.height = 6, out.width = "\\textwidth", fig.cap = "Class-by-class ROC curves.
+AUC is area under the curve, a measure of overall model performance.
+Equal error rates are marked, indicating the position at which there is equal probability of a false positive or false negative error.", fig.scap = "Class-by-class ROC curves.">>=
+aucs <- plot_onehot_roc(preds, test_labs, str_to_title(classes))
+thresholds <- purrr::map_dbl(aucs$data$eer, ~.$thresholds)
+aucs$data$thresholds <- thresholds
+
+ggplot() +
+  geom_line(aes(x = fpr, y = tpr), data = unnest(aucs$data, roc_plot), size = 1.25) +
+  geom_label(aes(x = 1, y = 0, label = sprintf("AUC: %0.2f\nEER: %0.2f", auc, thresholds)), hjust = 1, vjust = -0.02, data = aucs$data) +
+  geom_point(aes(x = fpr, y = tpr, color = "Equal Error\nRate (EER)"), data = unnest(aucs$data, eer), size = 2.5) +
+  scale_color_manual("", values = "black") +
+  facet_wrap(~class) +
+  scale_x_continuous("False Positive Rate", breaks = c(0, .25, .5, .75, 1), labels = c("0.0", "", "0.5", "", "1.0")) +
+  scale_y_continuous("True Positive Rate", breaks = c(0, .25, .5, .75, 1), labels = c("0.0", "", "0.5", "", "1.0")) +
+  ggtitle("CoNNOR Test Set Performance") +
+  facet_wrap(~class, nrow = 2) +
+  coord_fixed() +
+  theme(legend.position = c(1, 0), legend.justification = c(1, 0))
+@
+
+\FloatBarrier
+
+As discussed in Section~\ref{sec:model-eval}, a confusion matrix is a cross-tabulation of the labels assigned to an image and the labels predicted by the model.
+In a single-label problem, an image belongs to only one category and the model predicts only one label, so applying the wrong label is equivalent to failing to apply the correct label.
+Thus, every possible combination of true label and applied prediction is represented in the two-dimensional/traditional confusion matrix.
+Extending the confusion matrix to more complex problems requires additional considerations.
+
+Although some efforts to generalize confusion matrices to multi-class problems do exist in the literature \citep{landgrebeEfficientMulticlassROC2008}, it appears that confusion matrices have not previously been applied to multi-label classification problems.
+Since a single image may belong to a combination of the $n$ categories, it is no longer true that a false positive and a false negative are equivalent (see \autoref{tab:error-3class-problem}).
+The confusion matrix presented in \autoref{fig:confusion_matrix} is like a traditional confusion matrix in that all values along the diagonal represent the proportion of true positives captured within each category.
+The off-diagonal values, however, are adjusted from the traditional method to remove the effect of any true positives from the calculation of false positive proportions.
+For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any image that truly contains both triangles and quadrilaterals is removed before calculating the proportion of false quadrilateral labels.
+
+<<confusion_matrix, fig.width = 9, fig.height = 8, out.width = "\\textwidth", fig.cap = "Confusion matrix, showing on the diagonal the correct classification rate and on the off-diagonal, classification errors.
+Note that in multi-label images, correct off-diagonal labels have been excluded from the calculation of false positives.", fig.scap = "Confusion matrix, with correct and incorrect model classifications.", dpi = 600>>=
+get_confusion_matrix(predictions = preds, classes = classes,
+                     test_labels = test_labs, threshold = thresholds) %>%
+  set_names(str_to_title(classes)) %>%
+  ggcorrplot(., hc.order = F, outline.col = "white", lab = T) +
+  scale_fill_gradient("Classification\nRate", low = "white",
+                      high = "cornflowerblue", limits = c(0, 1)) +
+  scale_x_discrete("Image Label") + scale_y_discrete("Prediction") +
+  theme(axis.title.x = element_text(size = 14),
+        axis.title.y = element_text(size = 14, angle = 90, vjust = 1)) +
+  ggtitle("CoNNOR Multi-Class Confusion Matrix: Test Set Performance") +
+  theme_bw() +
+  theme(panel.grid.major = element_line(color = "grey50"),
+        panel.grid.minor = element_line(color = "grey60")) +
+  theme(plot.margin = grid::unit(c(0,0,0,0), "mm"),
+        plot.background = element_rect(fill = "transparent", color = NA),
+        plot.subtitle = element_blank(), plot.caption = element_blank(),
+        panel.spacing = unit(c(0, 0, 0, 0), "mm"))
+@
+
+This modified confusion matrix preserves the ability to identify patterns in over- and under-predictions; for instance, the horizontal band in \autoref{fig:confusion_matrix} indicates that quadrilaterals are overrepresented, as they are predicted more often than they should be for every true label.
+Similarly, polygons, stars, and triangles are often underpredicted relative to the images which actually contain these categories since they produce a large number of false positive predictions into other categories, as evidenced by the vertical bands.
+
+Another interesting artifact in \autoref{fig:confusion_matrix} is that circles are often misclassified as text, and text is often mislabeled as containing circles.
+Not all of the images used in model training are accurately labeled: in some cases, the labels have not been updated after labeling guidelines have changed.
+The confusion between text and circles is a natural consequence of the overlap between geometric shapes and components of written text; it is not indicative of a lack of fit as much as a weakness in the classification system used to define the model.
+
+\autoref{fig:confusion_matrix} shows the general weaknesses of the model, but does not necessarily suggest where to look as to \emph{why} the model is not performing appropriately.
+In Sections~\ref{sec:model-consistency} and \ref{sec:heatmap}, alternative visualizations are used to better understand the weaknesses of the model when applied to actual images used to train, test, and validate the model.
 

### 2019-06-01 paper/paper.Rnw
light editing

@@
 
-\mt{Model training is analogous to the human learning process.
+Model training is analogous to the human learning process.
 For example, a child learns to identify dogs by being presented with many labeled examples, such as when their parent uses the label ``dog" for an animal walking by.
-That child's understanding of dogs is then measured by how many dogs the child is able to correctly identify, and also by the number of cats it mistakenly calls ``dog".
-Similarly, CNN training is a series of stages, or ``epochs", where the model learns from the set of labeled training data, and that learning is measured through intermediate predictions of validation data.}
+That child's understanding of dogs is then measured by how many dogs the child is able to correctly identify, and also by the number of other animals it mistakenly calls ``dog".
+Similarly, CNN training is a series of stages, or ``epochs", where the model learns from the set of labeled training data, and that learning is measured through intermediate predictions of validation data.
 
@@
 
-\FloatBarrier
-
-\subsection{Model Accuracy}\label{sec:model-accuracy}
+\subsection{Model Accuracy}
 
 \paragraph{Overall Accuracy} % Model-wise and class-by-class, w/ confusion matrix
+\svp{You need to explain these terms.
+Your CC was written for statisticians, this paper is for practitioners.
+You need a MUCH gentler introduction to the idea of accuracy, errors, and ROC curves, as well as a lot more definitions for terms.}
 \autoref{fig:roc-code} shows the ROC curve for the full model, and \autoref{fig:auc-code} shows the curve for each class.
@@
 
-\FloatBarrier
-
-As discussed in Section~\ref{sec:model-eval}, a confusion matrix is a cross-tabulation of the labels assigned to an image and the labels predicted by the model.
+A confusion matrix is a cross-tabulation of the labels assigned to an image and the labels predicted by the model.
 In a single-label problem, an image belongs to only one category and the model predicts only one label, so applying the wrong label is equivalent to failing to apply the correct label.
@@
 \autoref{fig:confusion_matrix} shows the general weaknesses of the model, but does not necessarily suggest where to look as to \emph{why} the model is not performing appropriately.
-In Sections~\ref{sec:model-consistency} and \ref{sec:heatmap}, alternative visualizations are used to better understand the weaknesses of the model when applied to actual images used to train, test, and validate the model.
+%In Sections~\ref{sec:model-consistency} and \ref{sec:heatmap}, alternative visualizations are used to better understand the weaknesses of the model when applied to actual images used to train, test, and validate the model.
 

### 2019-06-01 paper/paper.Rnw
Additional comments

@@
 
+\svp{Pros and cons of confusion matrix here? Maybe we just don't show the actual proportions on the graph? I'd like to discuss this more in person.}
+\svp{If we do include the confusion matrix, this section needs to be MUCH more clearly explained for practitioners. Cross-tabulation isn't a common vocab word...}
+
 A confusion matrix is a cross-tabulation of the labels assigned to an image and the labels predicted by the model.

### 2019-06-01 paper/paper.Rnw
more comments and edits

@@
 As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks.
-\autoref{fig:vgg16layers} \fix{shows a selection of filters from VGG16, where each filter detects the presence of specific shapes or colors.
+\autoref{fig:vgg16layers} shows a selection of filters from VGG16, where each filter detects the presence of specific shapes or colors.
 Filters in earlier convolutional blocks detect simple features, such as colors and blobs, while later filters detect more complex combinations of features.
-Thus, a single image is made up of some combination of the hundreds of features that are detected by the filters in a CNN.}
+Thus, a single image is made up of some combination of the hundreds of features that are detected by the filters in a CNN.
 
@@
 
-\paragraph{VGG16}
-Developed by Oxford's Visual Graphics Group, VGG16 is a CNN with 16 ``functional" (i.e., convolutional and densely connected) layers and 5 ``structural" max pooling layers \citep{krizhevskyImageNetClassificationDeep2012}.
-In contrast to other popular networks, like ResNet, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy.
-The simplicity of this structure provides the ability to peer into the inner workings of the network for diagnostic purposes, providing a distinct advantage over more complicated network structures with slightly higher accuracy ratings.
-VGG16 is a common choice for transfer learning because of this structural simplicity; it has been used for detection of text in natural images \citep{zhongDeepTextUnifiedFramework2016}, medical imaging classification \citep{oquabLearningTransferringMidlevel2014}, classification of weld defects \citep{liuWeldDefectImages2018}, and many other domain-specific image recognition tasks that are more specific than the ImageNet data on which it was trained.
-\autoref{fig:VGG16-structure} shows the architecture of VGG16; with transfer learning, the VGG16 model head is replaced with a trained model head specific to recognition of shoeprint class characteristics.
-
 \begin{figure}
@@
 \end{figure}
+
+\paragraph{VGG16}
+Developed by Oxford's Visual Graphics Group, VGG16 is a CNN with 16 ``functional" (i.e., convolutional and densely connected) layers and 5 ``structural" max pooling layers \citep{krizhevskyImageNetClassificationDeep2012}.
+In contrast to other popular networks, like ResNet, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy.
+The simplicity of this structure provides the ability to peer into the inner workings of the network for diagnostic purposes, providing a distinct advantage over more complicated network structures with slightly higher accuracy ratings.
+VGG16 is a common choice for transfer learning because of this structural simplicity; it has been used for detection of text in natural images \citep{zhongDeepTextUnifiedFramework2016}, medical imaging classification \citep{oquabLearningTransferringMidlevel2014}, classification of weld defects \citep{liuWeldDefectImages2018}, and many other domain-specific image recognition tasks that are more specific than the ImageNet data on which it was trained.
+\autoref{fig:VGG16-structure} shows the architecture of VGG16; with transfer learning, the VGG16 model head is replaced with a trained model head specific to recognition of shoeprint class characteristics.
 

### 2019-06-02 paper/paper.Rnw
Add brand information

--- 2019-06-01 paper/paper.Rnw
@@
 
+<<>>=
+img_names <- dfunion$filename %>% basename() %>% as.vector
+
+labs <- sapply(classes, grepl, x = str_extract(img_names, "^.*?-\\d?-"))
+str <- gsub("^.*?-\\d{0,2}-", "", img_names)
+
+# Extract first, first two, and first three words of string
+brand_1 <- str_extract(str, "^.*?-")
+brand_2 <- str_extract(str, "^.*?-.*?-")
+brand_3 <- str_extract(str, "^.*?-.*?-.*?-")
+
+# These are the beginning of brands that need an extra word extracted
+multi_brands <- c("the", "the-north", "polo", "polo-ralph", "5", "5-11",
+                  "1", "to", "dr", "la", "new", "old", "under",
+                  "steve", "bernie", "cole", "tory",
+                  "harley", "kristin", "eric",
+                  "spring", "chinese", "dirty")
+
+brand <- ifelse(brand_1 %in% paste0(multi_brands, "-"),
+                ifelse(brand_2 %in% paste0(multi_brands, "-"),
+                       brand_3, brand_2),
+                brand_1) %>%
+  gsub(. , pattern = "-", replacement = " ") %>%
+  trimws(., "right"); rm(brand_1, brand_2, brand_3, multi_brands)
+
+df_images <- cbind(tibble(str, brand), labs)
+
+df_shoes <- df_images %>%
+  group_by(str, brand) %>%
+  summarize_at(vars(bowtie:triangle), any)
+
+top_brands <- df_shoes %>%
+  group_by(brand) %>%
+  summarize(n = n()) %>%
+  arrange(desc(n))
+
+df_summ <- filter(df_shoes, brand %in% head(top_brands$brand, 8)) %>%
+  group_by(brand) %>%
+  summarize(n = n(),
+            bowtie = sum(bowtie)/n, chevron = sum(chevron)/n,
+            circle = sum(circle)/n, line = sum(line)/n,
+            polygon = sum(polygon)/n, quad = sum(quad)/n,
+            star = sum(star)/n, text = sum(text)/n, triangle = sum(triangle)/n)
+
+df_summ_long <- df_summ %>%
+  gather(key = shape, value = prop, -brand, -n) %>%
+  arrange(brand)
+
+df_summ_long$brand <- as.factor(df_summ_long$brand)
+df_summ_long$shape <- factor(df_summ_long$shape, levels = classes)
+
+# Line plot
+ggplot(df_summ_long) + geom_point(aes(x = prop, y = shape, color = brand)) +
+  geom_path(aes(x = prop, y = shape, group = brand, color = brand)) +
+  ggtitle("Shape proportions by brand")
+
+# unicode_symbols <- c(bowtie = "\u26FB", chevron = "\uFE3D",
+#                      circle =  "\u23FA", line = "\u2225",
+#                      polygon = "\u2B23", quad = "\u25A0",
+#                      star = "\u2605", text = "a", triangle = "\u25BC",
+#                      other = "\u003F")
+
+ggplot(df_summ_long, aes(x = brand, y = prop, shape = factor(shape),
+                         color = factor(shape))) +
+  geom_point(size = 4) +
+  scale_shape_manual(values = c("\u26FB","\uFE3D","\u23FA","\u2225",
+                                "\u2B23","\u25A0","\u2605","a","\u25BC")) +
+  theme_bw() +
+  coord_flip() +
+  ggtitle("Proportion of labeled shoes containing each shape by brand")
+@
+
 Labeled images are scarce relative to the amount of data necessary to train a neural network.
@@
 You need a MUCH gentler introduction to the idea of accuracy, errors, and ROC curves, as well as a lot more definitions for terms.}
-\autoref{fig:roc-code} shows the ROC curve for the full model, and \autoref{fig:auc-code} shows the curve for each class.
+\autoref{fig:overall-roc} shows the ROC curve for the full model, and \autoref{fig:class-roc} shows the curve for each class.
 The full model has an AUC of 0.88, and the AUC for individual classes ranges from 0.81 (for line) to 0.91 (for bowtie and text).
@@
 
-<<roc-code, fig.width = 5, fig.height = 5, out.width = ".5\\textwidth", fig.cap = "ROC curve showing overall model performance.", fig.scap = "Overall model performance ROC curve.">>=
+<<overall-roc, fig.width = 5, fig.height = 5, out.width = ".5\\textwidth", fig.cap = "ROC curve showing overall model performance.", fig.scap = "Overall model performance ROC curve.">>=
 library(pROC)
@@
 
-<<auc-code, fig.width = 8, fig.height = 6, out.width = "\\textwidth", fig.cap = "Class-by-class ROC curves.
+<<class-roc, fig.width = 8, fig.height = 6, out.width = "\\textwidth", fig.cap = "Class-by-class ROC curves.
 AUC is area under the curve, a measure of overall model performance.
@@
 Although some efforts to generalize confusion matrices to multi-class problems do exist in the literature \citep{landgrebeEfficientMulticlassROC2008}, it appears that confusion matrices have not previously been applied to multi-label classification problems.
-Since a single image may belong to a combination of the $n$ categories, it is no longer true that a false positive and a false negative are equivalent (see \autoref{tab:error-3class-problem}).
-The confusion matrix presented in \autoref{fig:confusion_matrix} is like a traditional confusion matrix in that all values along the diagonal represent the proportion of true positives captured within each category.
-The off-diagonal values, however, are adjusted from the traditional method to remove the effect of any true positives from the calculation of false positive proportions.
+Since a single image may belong to a combination of the $n$ categories, it is no longer true that a false positive and a false negative are equivalent.
+In the confusion matrix presented in \autoref{fig:confusion_matrix}, the values along the diagonal represent the proportion of true positives captured within each category.
+The off-diagonal values represent the proportion of false positives for each shape pairing; however, the values are adjusted to remove the effect of any true positives from the calculation of false positive proportions.
 For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any image that truly contains both triangles and quadrilaterals is removed before calculating the proportion of false quadrilateral labels.
@@
 \autoref{fig:confusion_matrix} shows the general weaknesses of the model, but does not necessarily suggest where to look as to \emph{why} the model is not performing appropriately.
-%In Sections~\ref{sec:model-consistency} and \ref{sec:heatmap}, alternative visualizations are used to better understand the weaknesses of the model when applied to actual images used to train, test, and validate the model.
 

### 2019-06-03 paper/paper.Rnw
Add relevant parts of CC conclusion

@@
 
-<<>>=
+<<brand-plots>>=
 img_names <- dfunion$filename %>% basename() %>% as.vector
@@
 
+The goal of this research was to develop a method to automatically identify geometric class characteristics of shoe outsoles.
+Such a method takes the ``impossible" problem of quantifying outsole features in a given population and lays the groundwork for a tractable solution.
+A set of geometric class characteristics was defined to both broadly classify a large variety of shoes and to narrow down similarity into a manageable number and type of features for further use in a shoeprint analysis.
+The final set of features used in this research achieves both goals quite well, and has potential for much finer-grained analysis when recognized features are combined with spatial information and relationships.
+
+CoNNOR was developed using the convolutional base of VGG16 and a newly trained classifier/head to identify the defined set of geometric features.
+After training, CoNNOR performs well on the data provided.
+In general, the model is able to identify many well-defined geometric shapes in the images both consistently and accurately.
+The prediction accuracy of this model provides the ability to compute statistics for the frequency of given features in a well-defined population.
+
 \section{Future Work}
+
+Although CoNNOR performs well in its current state, there are still a number of ways to potentially improve prediction accuracy.
+For one, it is clear from Section~\ref{sec:model-consistency} that image contrast still plays a large role in how well the model classifies the geometric shapes present.
+Thus, exploring methods of color correction, such as histogram normalization, may prove useful for eliminating the effect of contrast on predictions.
+Geometric features are also relatively simple with respect to the features that are being detected by the final convolutional block of VGG16, as in \autoref{fig:vgg16layers}; it is quite possible that prediction accuracy could improve by directly classifying the features that are output by the fourth block, rather than using the full convolutional base.
 

### 2019-06-03 paper/paper.Rnw
comments

@@
 
+\svp{Need to show image contrast stuff to use this paragraph.}
 Although CoNNOR performs well in its current state, there are still a number of ways to potentially improve prediction accuracy.
-For one, it is clear from Section~\ref{sec:model-consistency} that image contrast still plays a large role in how well the model classifies the geometric shapes present.
+For one, it is clear that image contrast still plays a large role in how well the model classifies the geometric shapes present.
 Thus, exploring methods of color correction, such as histogram normalization, may prove useful for eliminating the effect of contrast on predictions.

### 2019-06-03 paper/paper.Rnw
Try to get unicode to show up

--- 2019-06-03 paper/paper.Rnw
@@
 
-<<brand-plots>>=
+<<brand-plots, device = "cairo">>=
 img_names <- dfunion$filename %>% basename() %>% as.vector
@@
 Although CoNNOR performs well in its current state, there are still a number of ways to potentially improve prediction accuracy.
-For one, it is clear from Section~\ref{sec:model-consistency} that image contrast still plays a large role in how well the model classifies the geometric shapes present.
+For one, image contrast still plays a large role in how well the model classifies the geometric shapes present.
 Thus, exploring methods of color correction, such as histogram normalization, may prove useful for eliminating the effect of contrast on predictions.

### 2019-06-03 paper/paper.Rnw

--- 2019-06-03 paper/paper.Rnw
@@
 Although CoNNOR performs well in its current state, there are still a number of ways to potentially improve prediction accuracy.
-For one, it is clear that image contrast still plays a large role in how well the model classifies the geometric shapes present.
+For one, image contrast  plays a large role in how well the model classifies the geometric shapes present.
 Thus, exploring methods of color correction, such as histogram normalization, may prove useful for eliminating the effect of contrast on predictions.

### 2019-06-03 paper/paper.Rnw
structure/outline comments

@@
 In this paper, we discuss an alternate approach that uses an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images.
-Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain.
+Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain."
+\svp{Add 1-2 sentences connecting feature sets to the goal of reducing workload - possibly cite solemate as an example of manual feature identification}
 
 \section{Materials and Methods}
+\svp{Short bit (1-2 sentences) about low-level feature detectors and why they fail with this problem.}
+
+\svp{Motivate the use of CNNs directly - complex model, higher level of feature aggregation. Mention pre-trained networks to transition.}
+
+\svp{VGG16/ImageNet specifically. Note that the structure of the model is distinct from the fitted weights.}
+
+\svp{Transfer Learning - using CNNs with less data.
+Make the point that this is what makes our method different from some of the others we've already cited - training a new classifier allows us to work within already-existing semantic contexts that examiners use.}
+
 \paragraph{Transfer Learning}
@@
 
+\svp{Moving towards ... how did we train the model}
 \paragraph{Annotated Training Data}
@@
 
+\svp{This section header might need to move? Haven't decided yet...}
 \section{Results}
@@
 
+\svp{Starting into How CoNNOR performs}
 \subsection{Model Accuracy}
@@
 \section{Conclusions}
-
+\svp{Rework this to address 3 main conclusions - 1: NNs work for automatic feature ID, 2: Different levels of feature variability within classes, and 3: correcting for contrast is important.}
 The goal of this research was to develop a method to automatically identify geometric class characteristics of shoe outsoles.

### 2019-06-03 paper/paper.Rnw
Add plot for brands within shapes

--- 2019-06-03 paper/paper.Rnw
@@
 
+<<brands-within-shapes>>=
+shapes <- lapply(X = classes, function(x){
+  out <- df_shoes[df_shoes[,x, drop = T],]
+  out$shape <- x
+  out
+})
+names(shapes) <- classes
+
+shapes %>% map_dfr(., function(x) {
+  x %>% group_by(brand) %>%
+    summarize(n = n()) %>%
+    arrange(desc(n)) %>%
+    mutate(prop = n / sum(n),
+           shape = x$shape[1:length(n)]) %>%
+    head(10)
+  }) %>%
+  ggplot() +
+  geom_text(aes(x = as.factor(shape),
+                               y = prop, label = brand,
+                               color = as.factor(shape)))
+@
+
 Labeled images are scarce relative to the amount of data necessary to train a neural network.

### 2019-06-03 paper/paper.Rnw
Update intro and talk about feature detectors

@@
 In this paper, we discuss an alternate approach that uses an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images.
-Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain."
-\svp{Add 1-2 sentences connecting feature sets to the goal of reducing workload - possibly cite solemate as an example of manual feature identification}
+Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain.
+%\svp{Add 1-2 sentences connecting feature sets to the goal of reducing workload - possibly cite solemate as an example of manual feature identification}
+\mt{In addition, much of the existing software to assess geometric similarity, such as SoleMate \fix{how to cite software?}, requires examiners to identify and label features manually; automating classification within the currently used feature space could reduce or eliminate the need for such manual classification.}
 
 \section{Materials and Methods}
-\svp{Short bit (1-2 sentences) about low-level feature detectors and why they fail with this problem.}
+%\svp{Short bit (1-2 sentences) about low-level feature detectors and why they fail with this problem.}
+\mt{Shoeprint evidence from crime scenes is most commonly collected in the form of a photograph, so any useful method to automatically identify outsole characteristics must take the form of an image analysis task.
+There are a number of methods that may be employed to identify shapes and features in an image, such as generalized Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}.
+While these methods are useful in identifying these specific features at a low level, they are very computationally intensive and only identify features on a very small scale; as a result, they cannot reliably identify large geometric shapes like those that may be found in an outsole image.}
+\fix{Include?
+(my gut says its not necessary):} \mt{These classical methods are also extremely sensitive to lighting or color changes.
+As one or more models would be required for each geometric shape, classical methods would require use of additional machine learning models such as random forests to aggregate low-level features into functional geometries found on outsoles.}
 
-\svp{Motivate the use of CNNs directly - complex model, higher level of feature aggregation. Mention pre-trained networks to transition.}
+%\svp{Motivate the use of CNNs directly - complex model, higher level of feature aggregation. Mention pre-trained networks to transition.}
+Convolutional neural networks (CNNs) are widely recognized as superior for novel image classification.
+CNNs are a form of artificial neural network which make use of the image convolution operator used by many low-level feature extraction methods.%\footnote{The Google Trends interest graph for CNNs and computer vision shows a massive increase in convolutional neural networks between mid-2014 and 2018 \url{https://trends.google.com/trends/explore?date=2010-01-01\%202019-02-18&q=convolutional\%20neural\%20network,computer\%20vision}}.
+As CNNs have evolved, their architecture has become more complex, but the fundamental reliance on the image convolution operator sets CNNs apart from other artificial neural networks \citep{guRecentAdvancesConvolutional2018}.
+CNNs have deep architectures that can be trained to identify complex patterns, but they are structurally similar to the human visual architecture and output binary or probabilistic predictions for given labels that are readily interpretable.
+As CNNs make use of labeled training data, the predictions generated are for features which are similar to those identified by humans, resulting in models with greater face validity.
+Once a CNN is trained, it is relatively fast and easy to apply the model to new images and obtain classifications.
 

### 2019-06-03 paper/paper.Rnw
Address comments

@@
 CNNs are a form of artificial neural network which make use of the image convolution operator used by many low-level feature extraction methods.%\footnote{The Google Trends interest graph for CNNs and computer vision shows a massive increase in convolutional neural networks between mid-2014 and 2018 \url{https://trends.google.com/trends/explore?date=2010-01-01\%202019-02-18&q=convolutional\%20neural\%20network,computer\%20vision}}.
-As CNNs have evolved, their architecture has become more complex, but the fundamental reliance on the image convolution operator sets CNNs apart from other artificial neural networks \citep{guRecentAdvancesConvolutional2018}.
+%As CNNs have evolved, their architecture has become more complex, but the fundamental reliance on the image convolution operator sets CNNs apart from other artificial neural networks \citep{guRecentAdvancesConvolutional2018}.
 CNNs have deep architectures that can be trained to identify complex patterns, but they are structurally similar to the human visual architecture and output binary or probabilistic predictions for given labels that are readily interpretable.
@@
 
-\svp{VGG16/ImageNet specifically. Note that the structure of the model is distinct from the fitted weights.}
-
-\svp{Transfer Learning - using CNNs with less data.
+%\svp{VGG16/ImageNet specifically. Note that the structure of the model is distinct from the fitted weights.}
+%\svp{Transfer Learning - using CNNs with less data.
 Make the point that this is what makes our method different from some of the others we've already cited - training a new classifier allows us to work within already-existing semantic contexts that examiners use.}
+%\mt{Now that I can see the whole dang methods section, I think I prefer the order of the methods section as it is now, rather than the order suggested by the above comments, unless you disagree}
 
@@
 As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks.
-\autoref{fig:vgg16layers} shows a selection of filters from VGG16, where each filter detects the presence of specific shapes or colors.
+\autoref{fig:vgg16layers} shows a selection of filters from VGG16, where each filter detects the presence of specific shapes and colors.
 Filters in earlier convolutional blocks detect simple features, such as colors and blobs, while later filters detect more complex combinations of features.
-Thus, a single image is made up of some combination of the hundreds of features that are detected by the filters in a CNN.
+Each of the five convolutional blocks has between 128 and 1,536 filters, and a single image is made up of some combination of the thousands of features that are detected by those filters.
 
-Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated \citep{oquabLearningTransferringMidlevel2014}.
+\fix{Transition needed?: The structure of the model is distinct from the fitted weights.} Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated \citep{oquabLearningTransferringMidlevel2014}.
 In many cases, the entire model base is used, and fitted with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with diffierent sets of image data.

### 2019-06-06 paper/paper.Rnw
Changing methods tone

@@
 Make the point that this is what makes our method different from some of the others we've already cited - training a new classifier allows us to work within already-existing semantic contexts that examiners use.}
-%\mt{Now that I can see the whole dang methods section, I think I prefer the order of the methods section as it is now, rather than the order suggested by the above comments, unless you disagree}
+
+%\mt{Now that I can see the whole dang methods section, I think I prefer the order of the methods section as it is now, rather than the order suggested by the above comments, unless you feel strongly otherwise.}
 
@@
 \autoref{fig:vgg16layers} shows a selection of filters from VGG16, where each filter detects the presence of specific shapes and colors.
-Filters in earlier convolutional blocks detect simple features, such as colors and blobs, while later filters detect more complex combinations of features.
+\mt{Each row of filters in \autoref{fig:vgg16layers} corresponds to one of the five convolutional blocks of VGG16.} Filters in earlier convolutional blocks detect simple features, such as colors and blobs, while later filters detect more complex combinations of features.
 Each of the five convolutional blocks has between 128 and 1,536 filters, and a single image is made up of some combination of the thousands of features that are detected by those filters.
 
-\fix{Transition needed?: The structure of the model is distinct from the fitted weights.} Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated \citep{oquabLearningTransferringMidlevel2014}.
+%\fix{Transition needed?: The structure of the model is distinct from the fitted weights.}
+\mt{Each layer in the model structure is comprised of weights and connections between weights that have been optimized during training.} Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated \citep{oquabLearningTransferringMidlevel2014}.
 In many cases, the entire model base is used, and fitted with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with diffierent sets of image data.
@@
 
-\svp{Moving towards ... how did we train the model}
+%\svp{Moving towards ... how did we train the model}
 \paragraph{Annotated Training Data}
@@
 
-\svp{It might be good to add in some (new) statistics relating brand information to common geometric features in tread - show, for instance, the top N brands with bowties vs.
+%\svp{It might be good to add in some (new) statistics relating brand information to common geometric features in tread - show, for instance, the top N brands with bowties vs.
 the top N brands with polygons.
 }
+
+\mt{The brand plots are still displaying weirdly for me, and all the figures are really squished and cut off?}
 
@@
 
-Labeled images are scarce relative to the amount of data necessary to train a neural network.
-One solution to this scarcity is to artificially enlarge the data set using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}.
+%Labeled images are scarce relative to the amount of data necessary to train a neural network.
+\mt{Although transfer learning substantially reduces the amount of data required to train a CNN, l}abeled images are scarce \fix{[difficult to generate]} relative to the amount of data necessary to train a neural network \mt{with adequate performance}.
+\fix{} One solution to this scarcity is to artificially enlarge the data set using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}.
 Augmentation is the transformation of original input data using image operations such as cropping, zoom, skew, rotation, and color balance modification in order to distort or alter the image while maintaining the essential features corresponding to the label.
@@
 
-The \Sexpr{nrow(dfunion)} images were split such that 60\% were used for training.
-Since the categories do not exist in equal proportion in the labeled data, the training data were weighted by proportion during the training process to prevent the loss function from being overwhelmed by more frequent categories.
+The \Sexpr{nrow(dfunion)} images were split such that 60\% were used for training. Since \mt{some categories are much more common than others,}
+%the categories do not exist in equal proportion in the labeled data,
+the training data were weighted by proportion during the training process to prevent the \mt{model learning} from being overwhelmed by more frequent categories.
 Of the remaining 40\% of data, half were used for validation, to monitor the training process, and the remaining data were for testing the performance of the fitted model.
@@
 \autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process.
-Overfitting, or fitting a model which performs too well on the training data relative to the validation data, is seen when the validation loss starts to increase after reaching a global minimum.
-Alternately, underfitting occurs when the validation accuracy is still increasing when model optimization is terminated.
+Overfitting \mt{occurs when a model learns the training data so well that its understanding of the categories becomes specific to only the training cases, which, in turn, leads to poor prediction of new images that were not in the training set.}
+%Overfitting, or fitting a model which performs too well on the training data relative to the validation data,
+Overfitting is seen when the validation loss starts to increase after reaching a global minimum.
+Alternately, underfitting occurs when the validation accuracy is still increasing when model optimization is terminated\mt{, because model performance is still improving with continued training}.
 Neither of these outcomes appears in \autoref{fig:training-accuracy}, indicating that the model optimization process was halted at an appropriate epoch.
 
-\svp{Starting into How CoNNOR performs}
+%\svp{Starting into How CoNNOR performs}
 \subsection{Model Accuracy}

### 2019-06-06 paper/paper.Rnw
Add heatmaps

@@
 
+\fix{How much should be said about loss and loss functions?}
+\Mt{A loss function is a function that measures how far model predictions are from the labels given to an image for each image in the validation and test sets; the model "learns" by updating its predictions in such a way that will minimize this loss.}
+
 \autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process.
@@
 %Overfitting, or fitting a model which performs too well on the training data relative to the validation data,
-Overfitting is seen when the validation loss starts to increase after reaching a global minimum.
-Alternately, underfitting occurs when the validation accuracy is still increasing when model optimization is terminated\mt{, because model performance is still improving with continued training}.
+Overfitting is suspected if the validation loss starts to increase after reaching a global minimum.
+Alternately, underfitting occurs if the validation accuracy is still increasing when model optimization is terminated\mt{, because model performance is still improving with continued training}.
 Neither of these outcomes appears in \autoref{fig:training-accuracy}, indicating that the model optimization process was halted at an appropriate epoch.
@@
 \paragraph{Overall Accuracy} % Model-wise and class-by-class, w/ confusion matrix
-\svp{You need to explain these terms.
+%\svp{You need to explain these terms.
 Your CC was written for statisticians, this paper is for practitioners.
 You need a MUCH gentler introduction to the idea of accuracy, errors, and ROC curves, as well as a lot more definitions for terms.}
-\autoref{fig:overall-roc} shows the ROC curve for the full model, and \autoref{fig:class-roc} shows the curve for each class.
+\mt{Receiver Operating Characteristic (ROC) curves are a diagnostic plot that compares the false positive rate against the true positive rate for a classification method, and the Area Under the Curve (AUC) quantifies the shape of the ROC curve.
+Perfect prediction would be indicated by a right angle along the upper left corner of the plot, with a corresponding AUC of 1, and a diagonal line with an AUC of 0.5 indicates that the classification method performs no better than random chance.} \autoref{fig:overall-roc} shows the ROC curve for the model across all classes, and \autoref{fig:class-roc} shows the curve for each class.
 The full model has an AUC of 0.88, and the AUC for individual classes ranges from 0.81 (for line) to 0.91 (for bowtie and text).
-While the class performances do vary slightly, each ROC curve is the same general shape and performs significantly better than random chance.
+While the class performances do vary slightly, each ROC curve is the same general shape and performs \fix{significantly} better than random chance.
 
@@
 
-A confusion matrix is a cross-tabulation of the labels assigned to an image and the labels predicted by the model.
-In a single-label problem, an image belongs to only one category and the model predicts only one label, so applying the wrong label is equivalent to failing to apply the correct label.
+\mt{ROC curves are useful for summarizing the accuracy of model predictions, but they do not indicate which classes are being assigned the false positive or false negative labels when accuracy is not perfect.
+A confusion matrix is a helpful diagnostic tool to understand which classes are commonly confused for other classes.} A confusion matrix is a \mt{summary of the relationships between} the labels assigned to an image and the labels predicted by the model.
+%In a single-label problem, an image belongs to only one category and the model predicts only one label, so applying the wrong label is equivalent to failing to apply the correct label.
 Thus, every possible combination of true label and applied prediction is represented in the two-dimensional/traditional confusion matrix.
 Extending the confusion matrix to more complex problems requires additional considerations.
-
-Although some efforts to generalize confusion matrices to multi-class problems do exist in the literature \citep{landgrebeEfficientMulticlassROC2008}, it appears that confusion matrices have not previously been applied to multi-label classification problems.
+%Although some efforts to generalize confusion matrices to multi-class problems do exist in the literature \citep{landgrebeEfficientMulticlassROC2008}, it appears that confusion matrices have not previously been applied to multi-label classification problems.
 Since a single image may belong to a combination of the $n$ categories, it is no longer true that a false positive and a false negative are equivalent.
@@
 
-This modified confusion matrix preserves the ability to identify patterns in over- and under-predictions; for instance, the horizontal band in \autoref{fig:confusion_matrix} indicates that quadrilaterals are overrepresented, as they are predicted more often than they should be for every true label.
-Similarly, polygons, stars, and triangles are often underpredicted relative to the images which actually contain these categories since they produce a large number of false positive predictions into other categories, as evidenced by the vertical bands.
+%This modified confusion matrix preserves the ability to identify patterns in over- and under-predictions; for instance,
+The horizontal band in \autoref{fig:confusion_matrix} indicates that quadrilaterals are predicted more often than they should be for every true label. Similarly, polygons, stars, and triangles
+%are often underpredicted relative to the images which actually contain these categories since they
+roduce a large number of false positive predictions into other categories, as evidenced by the vertical bands.
 
@@
 
-\autoref{fig:confusion_matrix} shows the general weaknesses of the model, but does not necessarily suggest where to look as to \emph{why} the model is not performing appropriately.
+\autoref{fig:confusion_matrix} shows \mt{which classes are being predicted incorrectly, but does not necessarily suggest \emph{why} the confusion is taking place.
+To investigate further, it is useful to} \fix{see what the model is seeing.}
+%the general weaknesses of the model, but does not necessarily suggest where to look as to \emph{why} the model is not performing appropriately.
 
 \paragraph{Model Diagnostics}
+
+\mt{A heatmap is a visual diagnostic tool that displays the portions of an image that are most significant to classification for a given class.
+Heatmaps provide diagnostic information: using these graphics, we can identify which locations in an image contribute most to CoNNOR's predictions.} In \fix{the figures in this subsection,} heatmaps are shown for a selection of three classes, which are either contained in the image or have high output probabilities, alongside the original image.
+
+\fix{I picked the heatmaps I thought illustrated the point best, but didn't modify them heavily yet.
+How many do we want to include, and do we want to highlight model challenges or just focus on the point that CoNNOR sees what we see?}
+
+<<chevron-circle-heatmap, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing chevrons and circles.", fig.cap = "Chevrons are identified by sharp corners and parallel lines, but the thick line between the two shape sections is falsely identified as a quadrilateral.", fig.pos = "!t">>=
+heatmaps <- list.files(file.path(wd, "images", "heatmaps"),
+                       pattern = "^heatmap")
+knitr::include_graphics(heatmaps[grepl("chevron-circle", heatmaps)])
+@
+
+<<text-seychelles-heatmap, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing text.", fig.cap = "This heatmap illustrates that the high probability of a circle is a result of the highly curled tails of the 'S' in the text.">>=
+knitr::include_graphics(heatmaps[grepl("text", heatmaps)])
+@
+
+<<corks-heatmap, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing none of the target classes.", fig.cap = "This outsole shows repeated outlines of corks, which do not fall into any of the nine classes; the concave corners where the corks overlap are predicted to be stars, and the round edges lead to predictions of circles.">>=
+knitr::include_graphics(heatmaps[grepl("corks", heatmaps)])
+@
 

### 2019-06-07 paper/paper.Rnw
Add consistency plots, explain ROC even more gently, and add explanation of EER

@@
 load(file.path(get_newest()$path, get_newest(pattern = "\\d.Rdata")$base_file))
+
+model_wts_file <- file.path(newest_model$path, newest_model$base_file)
+loaded_model <- set_weights(model_wts_file)
 @
@@
 %\svp{Add 1-2 sentences connecting feature sets to the goal of reducing workload - possibly cite solemate as an example of manual feature identification}
-\mt{In addition, much of the existing software to assess geometric similarity, such as SoleMate \fix{how to cite software?}, requires examiners to identify and label features manually; automating classification within the currently used feature space could reduce or eliminate the need for such manual classification.}
+\mt{In addition, much of the existing software to assess geometric similarity, such as SoleMate \fix{(how to cite software?)}, requires examiners to identify and label features manually; automating classification within the currently used feature space could reduce or eliminate the need for such manual classification.}
 
@@
 There are a number of methods that may be employed to identify shapes and features in an image, such as generalized Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}.
-While these methods are useful in identifying these specific features at a low level, they are very computationally intensive and only identify features on a very small scale; as a result, they cannot reliably identify large geometric shapes like those that may be found in an outsole image.}
-\fix{Include?
-(my gut says its not necessary):} \mt{These classical methods are also extremely sensitive to lighting or color changes.
-As one or more models would be required for each geometric shape, classical methods would require use of additional machine learning models such as random forests to aggregate low-level features into functional geometries found on outsoles.}
+While these methods are useful in identifying these specific features at a low level, they are very computationally intensive and only identify features on a very small scale; as a result, they cannot reliably identify large geometric shapes like those that may be found in an outsole image.} \fix{Necessary?: In addition, as one or more models would be required for each geometric shape, classical methods would require use of additional machine learning models such as random forests to aggregate low-level features into functional geometries found on outsoles.}
 
@@
 Convolutional neural networks (CNNs) are widely recognized as superior for novel image classification.
-CNNs are a form of artificial neural network which make use of the image convolution operator used by many low-level feature extraction methods.%\footnote{The Google Trends interest graph for CNNs and computer vision shows a massive increase in convolutional neural networks between mid-2014 and 2018 \url{https://trends.google.com/trends/explore?date=2010-01-01\%202019-02-18&q=convolutional\%20neural\%20network,computer\%20vision}}.
-%As CNNs have evolved, their architecture has become more complex, but the fundamental reliance on the image convolution operator sets CNNs apart from other artificial neural networks \citep{guRecentAdvancesConvolutional2018}.
+CNNs are a form of artificial neural network which make use of the image convolution operator used by many low-level feature extraction methods, \mt{with the additional ability to aggregate such features and meaningfully connect them to a pre-determined set of labels}.
 CNNs have deep architectures that can be trained to identify complex patterns, but they are structurally similar to the human visual architecture and output binary or probabilistic predictions for given labels that are readily interpretable.
@@
 Once a CNN is trained, it is relatively fast and easy to apply the model to new images and obtain classifications.
-
-%\svp{VGG16/ImageNet specifically. Note that the structure of the model is distinct from the fitted weights.}
-%\svp{Transfer Learning - using CNNs with less data.
-Make the point that this is what makes our method different from some of the others we've already cited - training a new classifier allows us to work within already-existing semantic contexts that examiners use.}
-
-%\mt{Now that I can see the whole dang methods section, I think I prefer the order of the methods section as it is now, rather than the order suggested by the above comments, unless you feel strongly otherwise.}
 
@@
 
-%\fix{Transition needed?: The structure of the model is distinct from the fitted weights.}
-\mt{Each layer in the model structure is comprised of weights and connections between weights that have been optimized during training.} Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated \citep{oquabLearningTransferringMidlevel2014}.
+\mt{Each layer in the model structure is comprised of sets of weights and connections that have been optimized during training.} Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated \citep{oquabLearningTransferringMidlevel2014}.
 In many cases, the entire model base is used, and fitted with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with diffierent sets of image data.
@@
 
-%\svp{It might be good to add in some (new) statistics relating brand information to common geometric features in tread - show, for instance, the top N brands with bowties vs.
-the top N brands with polygons.
-}
-
-\mt{The brand plots are still displaying weirdly for me, and all the figures are really squished and cut off?}
-
-<<brand-plots, device = "cairo">>=
+<<brand-plots, device = "cairo", include = F>>=
 img_names <- dfunion$filename %>% basename() %>% as.vector
@@
 
-# Line plot
-ggplot(df_summ_long) + geom_point(aes(x = prop, y = shape, color = brand)) +
-  geom_path(aes(x = prop, y = shape, group = brand, color = brand)) +
-  ggtitle("Shape proportions by brand")
+# # Line plot
+# ggplot(df_summ_long) + geom_point(aes(x = prop, y = shape, color = brand)) +
+#   geom_path(aes(x = prop, y = shape, group = brand, color = brand)) +
+#   ggtitle("Shape proportions by brand")
 
@@
   theme_bw() +
+  theme(legend.position = "none") +
   coord_flip() +
@@
                                y = prop, label = brand,
-                               color = as.factor(shape)))
+                               color = as.factor(shape))) +
+  theme(legend.position = "none")
 @
@@
 %Labeled images are scarce relative to the amount of data necessary to train a neural network.
-\mt{Although transfer learning substantially reduces the amount of data required to train a CNN, l}abeled images are scarce \fix{[difficult to generate]} relative to the amount of data necessary to train a neural network \mt{with adequate performance}.
-\fix{} One solution to this scarcity is to artificially enlarge the data set using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}.
+\mt{Although transfer learning substantially reduces the amount of data required to train a CNN, l}abeled images are \mt{still difficult or time-consuming to generate} %scarce \fix{[difficult to generate]}
+relative to the amount of data necessary to train \mt{even a small} neural network \mt{adequately.
+Another common} solution to data scarcity is to artificially enlarge the data set using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}.
 Augmentation is the transformation of original input data using image operations such as cropping, zoom, skew, rotation, and color balance modification in order to distort or alter the image while maintaining the essential features corresponding to the label.
@@
 \fix{How much should be said about loss and loss functions?}
-\Mt{A loss function is a function that measures how far model predictions are from the labels given to an image for each image in the validation and test sets; the model "learns" by updating its predictions in such a way that will minimize this loss.}
+\mt{A loss function is a function that measures how far model predictions are from the given labels for each image in the validation and test sets; the model ``learns" by updating its predictions in such a way that will minimize this loss.}
 
@@
 
-%\svp{Starting into How CoNNOR performs}
 \subsection{Model Accuracy}
@@
 You need a MUCH gentler introduction to the idea of accuracy, errors, and ROC curves, as well as a lot more definitions for terms.}
-\mt{Receiver Operating Characteristic (ROC) curves are a diagnostic plot that compares the false positive rate against the true positive rate for a classification method, and the Area Under the Curve (AUC) quantifies the shape of the ROC curve.
-Perfect prediction would be indicated by a right angle along the upper left corner of the plot, with a corresponding AUC of 1, and a diagonal line with an AUC of 0.5 indicates that the classification method performs no better than random chance.} \autoref{fig:overall-roc} shows the ROC curve for the model across all classes, and \autoref{fig:class-roc} shows the curve for each class.
+
+\mt{For each image in the test set, the ``true labels" are nine human-assigned labels of 0 or 1 corresponding to each of the nine shape categories.
+When the CNN predicts which shapes are in the image, however, it assigns a probability between 0 and 1 for each shape category, with a total of nine probabilities per test image.
+Thus, to determine the accuracy of the model predictions, one must choose a threshold to the probabilities, such that a predicted probability above the threshold indicates a ``yes" prediction from the model and a probability below the threshold indicates a ``no".
+If the threshold is high, \fix{moderate} probabilities will be treated as a ``no", which increases the number of true negative and false negative predictions.
+Conversely, if the threshold is too low, \fix{moderate} probabilites will be considered as ``yes", which increases the true and false positive predictions in the data.
+An appropriate threshold must be chosen to produce a high true positive rate while also ideally keeping the false positive rate low.}
+
+\mt{Receiver Operating Characteristic (ROC) curves are a type diagnostic plot that compares the false positive rate against the true positive rate for a classification method, computed by comparing the two rates at a number of thresholds, and the Area Under the Curve (AUC) quantifies the shape of the ROC curve.
+Perfect prediction would be indicated by a right angle along the upper left corner of the plot, with a corresponding AUC of 1, and a diagonal line with an AUC of 0.5 would indicate that the classification method performs no better than random chance.} \autoref{fig:overall-roc} shows the ROC curve for \fix{the model} across all classes, and \autoref{fig:class-roc} shows the curve for each class.
 The full model has an AUC of 0.88, and the AUC for individual classes ranges from 0.81 (for line) to 0.91 (for bowtie and text).
 While the class performances do vary slightly, each ROC curve is the same general shape and performs \fix{significantly} better than random chance.
+\mt{The points in \autoref{fig:class-roc} represent the equal error rate (EER), which is the threshold where there is equal probability of a false positive or false negative error.
+Thus, for each category, any model prediction with probability above the EER for that shape is considered a positive prediction.}
 
@@
 \svp{Pros and cons of confusion matrix here? Maybe we just don't show the actual proportions on the graph? I'd like to discuss this more in person.}
-\svp{If we do include the confusion matrix, this section needs to be MUCH more clearly explained for practitioners. Cross-tabulation isn't a common vocab word...}
 
@@
 A confusion matrix is a helpful diagnostic tool to understand which classes are commonly confused for other classes.} A confusion matrix is a \mt{summary of the relationships between} the labels assigned to an image and the labels predicted by the model.
-%In a single-label problem, an image belongs to only one category and the model predicts only one label, so applying the wrong label is equivalent to failing to apply the correct label.
-Thus, every possible combination of true label and applied prediction is represented in the two-dimensional/traditional confusion matrix.
-Extending the confusion matrix to more complex problems requires additional considerations.
-%Although some efforts to generalize confusion matrices to multi-class problems do exist in the literature \citep{landgrebeEfficientMulticlassROC2008}, it appears that confusion matrices have not previously been applied to multi-label classification problems.
-Since a single image may belong to a combination of the $n$ categories, it is no longer true that a false positive and a false negative are equivalent.
+
 In the confusion matrix presented in \autoref{fig:confusion_matrix}, the values along the diagonal represent the proportion of true positives captured within each category.
-The off-diagonal values represent the proportion of false positives for each shape pairing; however, the values are adjusted to remove the effect of any true positives from the calculation of false positive proportions.
+The off-diagonal values represent the proportion of false positives for each shape pairing; \mt{however, since a single image may truly contain multiple shapes}, the values are adjusted to remove the effect of any true positives from the calculation of false positive proportions.
 For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any image that truly contains both triangles and quadrilaterals is removed before calculating the proportion of false quadrilateral labels.
@@
 
-%This modified confusion matrix preserves the ability to identify patterns in over- and under-predictions; for instance,
-The horizontal band in \autoref{fig:confusion_matrix} indicates that quadrilaterals are predicted more often than they should be for every true label. Similarly, polygons, stars, and triangles
-%are often underpredicted relative to the images which actually contain these categories since they
-roduce a large number of false positive predictions into other categories, as evidenced by the vertical bands.
-
+The horizontal band in \autoref{fig:confusion_matrix} indicates that quadrilaterals are predicted more often than they should be for every true label.
+Similarly, polygons, stars, and triangles produce a large number of false positive predictions into other categories, as evidenced by the vertical bands.
 Another interesting artifact in \autoref{fig:confusion_matrix} is that circles are often misclassified as text, and text is often mislabeled as containing circles.
-Not all of the images used in model training are accurately labeled: in some cases, the labels have not been updated after labeling guidelines have changed.
+Not all of the images used in model training are accurately labeled: in some cases, the labels have not \mt{yet} been updated after labeling guidelines have changed.
 The confusion between text and circles is a natural consequence of the overlap between geometric shapes and components of written text; it is not indicative of a lack of fit as much as a weakness in the classification system used to define the model.
@@
 
+\paragraph{Model Predictions}
+
+ \mt{What else to say here?}
+
+<<exemplars-consistency, fig.cap = "Examples", fig.pos = "p!", out.width = "100%", cache = T, fig.pos = "!h">>=
+exemplar <- list.files(file.path(wd, "images", "consistency", "exemplar"), full.names = T)
+pred_prob_plot(exemplar, loaded_model)
+@
+
+\fix{Can also look at many examples within same shape categories.}
+
+<<chevrons-consistency, fig.cap = "Chevrons", fig.pos = "p!", out.width = "100%", cache = T>>=
+chevron <- list.files(file.path(wd, "images", "consistency", "chevron"), full.names = T)
+pred_prob_plot(chevron, loaded_model)
+@
+
+<<circle-text-consistency, fig.cap = "Text with circles or circle-like shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
+circle_text <- list.files(file.path(wd, "images", "consistency", "circle-text"), full.names = T)
+pred_prob_plot(circle_text, loaded_model)
+@
+
 \paragraph{Model Diagnostics}
 
-\mt{A heatmap is a visual diagnostic tool that displays the portions of an image that are most significant to classification for a given class.
-Heatmaps provide diagnostic information: using these graphics, we can identify which locations in an image contribute most to CoNNOR's predictions.} In \fix{the figures in this subsection,} heatmaps are shown for a selection of three classes, which are either contained in the image or have high output probabilities, alongside the original image.
+\mt{A heatmap is a visual diagnostic tool that highlights the portions of an image that are most significant to classification for a given class.
+Using these graphics, we can identify which locations in an image contribute most to the model's predictions.} In \fix{the figures in this subsection,} heatmaps are shown for a selection of three classes, which are either contained in the image or have high output probabilities, alongside the original image.
 
@@
 
-<<chevron-circle-heatmap, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing chevrons and circles.", fig.cap = "Chevrons are identified by sharp corners and parallel lines, but the thick line between the two shape sections is falsely identified as a quadrilateral.", fig.pos = "!t">>=
+\FloatBarrier
+
+<<chevron-circle-heatmap, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing chevrons and circles.", fig.cap = "Chevrons are identified by sharp corners and parallel lines, but the thick line between the two shape sections is falsely identified as a quadrilateral.", fig.pos = "!h">>=
 heatmaps <- list.files(file.path(wd, "images", "heatmaps"),
-                       pattern = "^heatmap")
+                       pattern = "^heatmap", full.names = T)
 knitr::include_graphics(heatmaps[grepl("chevron-circle", heatmaps)])
@@
 @
+
+\FloatBarrier
 

### 2019-06-07 paper/paper.Rnw
Update

@@
 Thus, to determine the accuracy of the model predictions, one must choose a threshold to the probabilities, such that a predicted probability above the threshold indicates a ``yes" prediction from the model and a probability below the threshold indicates a ``no".
-If the threshold is high, \fix{moderate} probabilities will be treated as a ``no", which increases the number of true negative and false negative predictions.
+If the threshold is too high, \fix{moderate} probabilities will be treated as a ``no", which increases the number of true negative and false negative predictions.
 Conversely, if the threshold is too low, \fix{moderate} probabilites will be considered as ``yes", which increases the true and false positive predictions in the data.

### 2019-06-07 paper/paper.Rnw
Get rid of floatbarrier - it won't compile with that in there.
comments

@@
 %\svp{Add 1-2 sentences connecting feature sets to the goal of reducing workload - possibly cite solemate as an example of manual feature identification}
-\mt{In addition, much of the existing software to assess geometric similarity, such as SoleMate \fix{(how to cite software?)}, requires examiners to identify and label features manually; automating classification within the currently used feature space could reduce or eliminate the need for such manual classification.}
+In addition, much of the existing software to assess geometric similarity, such as SoleMate\footnote{\url{http://www.fosterfreeman.com/trace-evidence/357-sicar-6-solemate-2.html}}, requires examiners to identify and label features manually; automating classification within the currently used feature space could reduce or eliminate the need for such manual classification.
 
@@
 %\svp{Short bit (1-2 sentences) about low-level feature detectors and why they fail with this problem.}
-\mt{Shoeprint evidence from crime scenes is most commonly collected in the form of a photograph, so any useful method to automatically identify outsole characteristics must take the form of an image analysis task.
-There are a number of methods that may be employed to identify shapes and features in an image, such as generalized Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}.
-While these methods are useful in identifying these specific features at a low level, they are very computationally intensive and only identify features on a very small scale; as a result, they cannot reliably identify large geometric shapes like those that may be found in an outsole image.} \fix{Necessary?: In addition, as one or more models would be required for each geometric shape, classical methods would require use of additional machine learning models such as random forests to aggregate low-level features into functional geometries found on outsoles.}
-
+\mt{Shoeprint evidence from crime scenes is most commonly collected in the form of a photograph, so any useful method to automatically identify outsole characteristics must take the form of an image analysis task.}\svp{We're not working with prints, though - please try to avoid the temptation to delve too far into prints.
+"Ultimately, most pattern evidence is stored in photographic form; as a result, we must use image analysis methods to create numerical features from the original visual record." or something to that effect.}
+There are a number of methods that may be employed to identify shapes and features in an image, such as \svp{Fourier transforms, }Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}.
+While these methods are useful in identifying these specific features at a low level, they can be computationally intensive and only identify features on a very small scale; as a result, they cannot reliably identify large geometric shapes like those that may be found in an outsole image.
 %\svp{Motivate the use of CNNs directly - complex model, higher level of feature aggregation. Mention pre-trained networks to transition.}
 Convolutional neural networks (CNNs) are widely recognized as superior for novel image classification.
-CNNs are a form of artificial neural network which make use of the image convolution operator used by many low-level feature extraction methods, \mt{with the additional ability to aggregate such features and meaningfully connect them to a pre-determined set of labels}.
-CNNs have deep architectures that can be trained to identify complex patterns, but they are structurally similar to the human visual architecture and output binary or probabilistic predictions for given labels that are readily interpretable.
+CNNs are a form of artificial neural network which make use of the image convolution operator used by many low-level feature extraction methods, with the additional ability to aggregate such features and meaningfully connect them to a pre-determined set of labels.
+CNNs have deep architectures that can be trained to identify complex patterns, but they are structurally similar to the \svp{architecture of the }human visual \svp{system} and output binary or probabilistic predictions for given labels that are readily interpretable.
 As CNNs make use of labeled training data, the predictions generated are for features which are similar to those identified by humans, resulting in models with greater face validity.
@@
 As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks.
-\autoref{fig:vgg16layers} shows a selection of filters from VGG16, where each filter detects the presence of specific shapes and colors.
-\mt{Each row of filters in \autoref{fig:vgg16layers} corresponds to one of the five convolutional blocks of VGG16.} Filters in earlier convolutional blocks detect simple features, such as colors and blobs, while later filters detect more complex combinations of features.
+\svp{You need to talk about the model structure of vgg16 before you start talking about layers.
+Talk of convolutional blocks doesn't really make sense until you show the layer structure.}\autoref{fig:vgg16layers} shows a selection of filters from VGG16, where each filter detects the presence of specific shapes and colors.
+Each row of filters in \autoref{fig:vgg16layers} corresponds to one of the five convolutional blocks of VGG16.
+Filters in earlier convolutional blocks detect simple features, such as colors and blobs, while later filters detect more complex combinations of features.
 Each of the five convolutional blocks has between 128 and 1,536 filters, and a single image is made up of some combination of the thousands of features that are detected by those filters.
 
-\mt{Each layer in the model structure is comprised of sets of weights and connections that have been optimized during training.} Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated \citep{oquabLearningTransferringMidlevel2014}.
+Each layer in the model structure is comprised of sets of weights and connections that have been optimized during training.
+Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated \citep{oquabLearningTransferringMidlevel2014}.
 In many cases, the entire model base is used, and fitted with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with diffierent sets of image data.
@@
 
-In this paper, we will demonstrate the use of transfer learning to automatically identify features in shoe treads which are used by forensic examiners, using a general-purpose CNN with relatively simple structure, VGG16, and a database of labeled shoe tread images we have assembled for this project, which are both described below.
-This approach differs from other approaches \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017} in automatic footwear identification which use the output from the model base directly and do not attempt to add human-friendly contextual information with an additional classifier.
+\fix{In this paper, we will use transfer learning to automatically identify features in shoe treads which are used by forensic examiners, leveraging a general-purpose CNN with relatively simple structure, VGG16, and a database of labeled shoe tread images we have assembled for this project, which are both described below.} \svp{This sentence is a bit awkward, because you're introducing too many things.
+The paragraph is good, but I don't think that it makes sense to have it here.
+Introduce VGG16, then use this paragraph as a transition to talking about the actual model we fit.} This approach differs from other approaches \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017} in automatic footwear identification which use the output from the model base directly and do not attempt to add human-friendly contextual information with an additional classifier.
 
@@
 
-<<vgg16layers, fig.width = 7, fig.height = 5.75, fig.cap = "A selection of filters from the convolutional layers of VGG16.">>=
+<<vgg16layers, fig.width = 7, fig.height = 5.75, fig.cap = "A selection of filters from the convolutional layers of VGG16.\\svp{Label the rows with block numbers please}">>=
 imgs <- list.files("./images/vgg16layers/", recursive = T, full.names = T)
@@
 The remaining two faces can be convex or straight, and the concave faces may have straight portions, so long as there is a concave region.
-Using this definition, shapes such as butterflies are included as bowties.
+%Using this definition, shapes such as butterflies are included as bowties.
 \svp{I'm still on the fence - we can just not mention this, or we could move butterflies to "star"...} \mt{I think it makes a lot of sense to move them to star.
 That means redoing the poster/slides examples though...
-grrrrrr}
+grrrrrr} I'm just commenting that bit out for now.
 \item [Chevron] Chevron shapes include repeating parallel lines as well as individual ``v" shapes. They may be angular but can also be curved.
@@
 \item [Other] Other features which were marked include logos, various textures (including crepe, stippling, etc.), and smooth regions with no discernible features.
-These regions are grouped and provide additional information---that none of the previous nine categories are present.
+These regions are grouped and provide additional information - that none of the previous nine categories are present.
 \end{description}
@@
 
-\FloatBarrier
-
 <<chevron-circle-heatmap, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing chevrons and circles.", fig.cap = "Chevrons are identified by sharp corners and parallel lines, but the thick line between the two shape sections is falsely identified as a quadrilateral.", fig.pos = "!h">>=
@@
 @
-
-\FloatBarrier
 

### 2019-06-07 paper/paper.Rnw
move brand plots to the part of the paper where we're talking about different results - it's more useful there to explain why performance differs for different classes.

@@
 
+%Labeled images are scarce relative to the amount of data necessary to train a neural network.
+Although transfer learning substantially reduces the amount of data required to train a CNN \svp{by several orders of magnitude}, \fix{labeled images are still difficult or time-consuming to generate relative to the amount of data necessary to train even a small neural network adequately.
+Another common solution to data scarcity is to artificially enlarge the data set using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}.} \svp{I would lose the last clause of the first sentence and combine it with the next, e.g.
+to make most efficient use of the labeled data, we enlarge the training data using a process called image augmentation.}
+ Augmentation is the transformation of original input data using image operations such as cropping, zoom, skew, rotation, and color balance modification in order to distort or alter the image while maintaining the essential features corresponding to the label.
+This process reduces the potential for overfitting the model to the specific set of image data used during the training process, and also increases the amount of data available for training.
+Examples of pre- and post-augmentation images are shown in \autoref{fig:augmented}.
+
+\begin{figure}
+\centering
+\includegraphics[width=.1\textwidth]{images/augmentation/bowtie(R)-1-ugg-sienna-enamel-blue_product_8726365_color_120557.jpg}\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/aug_bowtie(R)-1-ugg-sienna-enamel-blue_product_8726365_color_120557_0_8344.jpg}\hfill\hfill\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/quad-2-nike-kids-downshifter-7-infant-toddler-gunsmoke-sunset-pulse-atmosphere-grey_product_8800875_color_720973.jpg}\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/aug_quad-2-nike-kids-downshifter-7-infant-toddler-gunsmoke-sunset-pulse-atmosphere-grey_product_8800875_color_720973_0_7533.jpg}\hfill\hfill\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/quad-2-puma-suede-classic-x-mac-three-port-royale-port-royale_product_9123838_color_251426.jpg}\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/aug_quad-2-puma-suede-classic-x-mac-three-port-royale-port-royale_product_9123838_color_251426_0_5689.jpg}\hfill\hfill\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/text-1-bernie-mev-kids-catwalk-little-kid-big-kid-multi-camo_product_9084647_color_87089.jpg}\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/aug_text-1-bernie-mev-kids-catwalk-little-kid-big-kid-multi-camo_product_9084647_color_87089_0_1891.jpg}
+\caption[Original and augmented images.]{Four sets of original (left) and augmented (right) labeled images.}\label{fig:augmented}
+\end{figure}
+
+Model training is analogous to the human learning process.
+For example, a child learns to identify dogs by being presented with many labeled examples, such as when their parent uses the label ``dog" for an animal walking by.
+That child's understanding of dogs is then measured by how many dogs the child is able to correctly identify, and also by the number of other animals it mistakenly calls ``dog".
+Similarly, CNN training is a series of stages, or ``epochs", where the model learns from the set of labeled training data, and that learning is measured through intermediate predictions on validation data.
+
+The \Sexpr{nrow(dfunion)} images were split such that 60\% were used for training. Since \mt{some categories are much more common than others,}\svp{I like the equal proportion phrasing better myself}
+%the categories do not exist in equal proportion in the labeled data,
+the training data were weighted by proportion during the training process to \svp{ensure that the model can identify both rare and common geometric shapes}.%from being overwhelmed by more frequent categories.
+Of the remaining 40\% of data, half were used for validation, to monitor the training process, and the remaining data were for testing the performance of the fitted model.
+
+\svp{This section header might need to move? Haven't decided yet...}
+\section{Results}
+
+\subsection{Model Training}
+<<training-accuracy, fig.width = 7.5, fig.height = 5, out.width = ".75\\textwidth", fig.cap = "Training and validation accuracy and loss for each epoch of the fitting process.
+Training and validation accuracy reach 89.5\\% around epoch 9.
+After that point, validation loss remains the same and training loss decreases slightly, while validation accuracy increases more slowly than training accuracy.", fig.scap = "Training and validation accuracy and loss during each epoch.", fig.pos = 'h'>>=
+data.frame(history$metrics) %>%
+  mutate(epoch = 1:n()) %>%
+  gather(key = "measure", value = "value", -epoch) %>%
+  mutate(Type = ifelse(str_detect(measure, "val"), "Validation", "Training"),
+         measure = ifelse(str_detect(measure, "acc"), "Accuracy", "Loss")) %>%
+  # bind_rows(tibble(epoch = NA, value =  .6, measure = "Accuracy", Type = "Validation")) %>%
+  # bind_rows(tibble(epoch = NA, value =  .33, measure = "Loss", Type = "Validation")) %>%
+  ggplot(aes(x = epoch, y = value, color = Type)) +
+  geom_point() +
+  geom_smooth(se = F) +
+  facet_grid(measure~., scales = "free_y", switch = "both") +
+  theme_bw() +
+  scale_y_continuous("") +
+  scale_x_continuous("Epoch") +
+  ggtitle("CoNNOR Training Performance") + mytheme +
+  theme(axis.title.y = element_blank(), legend.position = c(1, .5),
+        legend.justification = c(1.03, -0.05),
+        legend.background = element_rect(fill = "white"))
+@
+
+\fix{How much should be said about loss and loss functions?}\svp{}
+A loss function is a function that measures how far model predictions are from the given labels for each image in the validation and test sets; the model ``learns" by updating its predictions to minimize the loss.
+
+\autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process.
+Overfitting occurs when a model learns the training data so well that its understanding of the categories becomes specific to the training cases, which, in turn, leads to poor prediction of new images that were not in the training set.
+%Overfitting, or fitting a model which performs too well on the training data relative to the validation data,
+\svp{get rid of passive voice here.
+Overfitting occurs when the validation loss begins to increase after reaching a global minimum} Overfitting is suspected if the validation loss starts to increase after reaching a global minimum.
+Alternately, underfitting occurs if the validation accuracy is still increasing when model optimization is terminated\svp{(for instance, between epochs 1 and 10 in \autoref{fig:training-accuracy})}, because model performance is still improving with continued training.
+\svp{In \autoref{fig:training-accuracy}, validation accuracy levels off after epoch 20, and validation loss has not yet begun to increase at epoch 30,} indicating that the model optimization process was halted at an appropriate epoch.
+
+\subsection{Model Accuracy}
+
+\paragraph{Overall Accuracy} % Model-wise and class-by-class, w/ confusion matrix
+%\svp{You need to explain these terms.
+Your CC was written for statisticians, this paper is for practitioners.
+You need a MUCH gentler introduction to the idea of accuracy, errors, and ROC curves, as well as a lot more definitions for terms.}
+
+For each image in the test set, the ``true labels" are nine human-assigned labels of 0 or 1 corresponding to \svp{the presence or absence of} each of the nine shape categories \svp{in the image}.
+When the CNN predicts which shapes are in the image, however, it assigns a probability between 0 and 1 for each shape category, with a total of nine probabilities per test image.
+Thus, to determine the accuracy of the model predictions, \svp{fix weird phrasing - we must select a threshold for each label to discretize the model predictions} \fix{one must choose a threshold to the probabilities, such that a predicted probability above the threshold indicates a ``yes" prediction from the model and a probability below the threshold indicates a ``no".} If the threshold is too high, \fix{moderate} probabilities will be treated as a ``no", which increases the number of true negative and false negative predictions.
+Conversely, if the threshold is too low, \fix{moderate} probabilites will be considered as ``yes", which increases the true and false positive predictions in the data.
+An appropriate threshold must be chosen to produce a high true positive rate while also ideally keeping the false positive rate low.
+
+Receiver Operating Characteristic (ROC) curves are a type diagnostic plot \svp{which compare} the false positive rate \svp{to} the true positive rate for a classification method.\svp{Make this two sentences so that it's less convoluted}\fix{, computed by comparing the two rates at a number of thresholds, and the Area Under the Curve (AUC) quantifies the shape of the ROC curve.} Perfect prediction would be indicated by a right angle along the upper left corner of the plot, with a corresponding AUC of 1, and a diagonal line with an AUC of 0.5 would indicate that the classification method performs no better than random chance.
+\autoref{fig:overall-roc} shows the ROC curve for \fix{the model} across all classes, and \autoref{fig:class-roc} shows the curve for each class.
+The full model has an AUC of 0.88, and the AUC for individual classes ranges from 0.81 (for lines) to 0.91 (for bowtie and text).
+While the class performances do vary slightly, each ROC curve is the same general shape and performs significantly better than random chance.
+The points in \autoref{fig:class-roc} represent the equal error rate (EER), which is the threshold where there is equal probability of a false positive or false negative error.
+\svp{We will use the class-wise EER to set a threshold corresponding to each class, so that }any model prediction with probability above the EER for that shape is considered a positive prediction.
+
+<<overall-roc, fig.width = 5, fig.height = 5, out.width = ".5\\textwidth", fig.cap = "ROC curve showing overall model performance.", fig.scap = "Overall model performance ROC curve.">>=
+library(pROC)
+pred_df <- as_tibble(preds) %>% gather(key = feature, value = value)
+test_labs_df <- as_tibble(test_labs) %>% gather(key = feature, value = value)
+whole_model_roc <- roc(test_labs_df$value, pred_df$value)
+
+whole_model_roc_df <- tibble(tpr = whole_model_roc$sensitivities,
+                             fpr = 1 - whole_model_roc$specificities,
+                             thresholds = whole_model_roc$thresholds,
+                             auc = whole_model_roc$auc[1]) %>%
+  nest(tpr, fpr, thresholds, .key =  "roc_plot") %>%
+  mutate(eer = purrr::map(roc_plot, eer))
+ggplot() +
+  geom_line(aes(x = fpr, y = tpr), data = unnest(whole_model_roc_df, roc_plot), size = 1.25) +
+  geom_label(aes(x = 1, y = .07, label = sprintf("AUC: %0.2f", auc)), hjust = 1, vjust = -0.2, data = whole_model_roc_df) +
+  geom_point(aes(x = fpr, y = tpr, color = "Equal Error Rate"), data = unnest(whole_model_roc_df, eer), size = 2) +
+  scale_color_manual("", values = "black") +
+  scale_x_continuous("False Positive Rate", breaks = c(0, .25, .5, .75, 1), labels = c("0.0", "", "0.5", "", "1.0")) +
+  scale_y_continuous("True Positive Rate", breaks = c(0, .25, .5, .75, 1), labels = c("0.0", "", "0.5", "", "1.0")) +
+  ggtitle("CoNNOR Test Set Performance (All Classes)") +
+  coord_fixed() + mytheme +
+  theme(legend.position = c(1, 0), legend.justification = c(1.01, -0.01), legend.title = element_blank(), legend.background = element_rect(fill = "white"))
+
+@
+
+<<class-roc, fig.width = 8, fig.height = 6, out.width = "\\textwidth", fig.cap = "Class-by-class ROC curves.
+AUC is area under the curve, a measure of overall model performance.
+Equal error rates are marked, indicating the position at which there is equal probability of a false positive or false negative error.", fig.scap = "Class-by-class ROC curves.">>=
+aucs <- plot_onehot_roc(preds, test_labs, str_to_title(classes))
+thresholds <- purrr::map_dbl(aucs$data$eer, ~.$thresholds)
+aucs$data$thresholds <- thresholds
+
+ggplot() +
+  geom_line(aes(x = fpr, y = tpr), data = unnest(aucs$data, roc_plot), size = 1.25) +
+  geom_label(aes(x = 1, y = 0, label = sprintf("AUC: %0.2f\nEER: %0.2f", auc, thresholds)), hjust = 1, vjust = -0.02, data = aucs$data) +
+  geom_point(aes(x = fpr, y = tpr, color = "Equal Error\nRate (EER)"), data = unnest(aucs$data, eer), size = 2.5) +
+  scale_color_manual("", values = "black") +
+  facet_wrap(~class) +
+  scale_x_continuous("False Positive Rate", breaks = c(0, .25, .5, .75, 1), labels = c("0.0", "", "0.5", "", "1.0")) +
+  scale_y_continuous("True Positive Rate", breaks = c(0, .25, .5, .75, 1), labels = c("0.0", "", "0.5", "", "1.0")) +
+  ggtitle("CoNNOR Test Set Performance") +
+  facet_wrap(~class, nrow = 2) +
+  coord_fixed() +
+  theme(legend.position = c(1, 0), legend.justification = c(1, 0))
+@
+
+\svp{Pros and cons of confusion matrix here? Maybe we just don't show the actual proportions on the graph? I'd like to discuss this more in person.}
+
+\svp{I'd focus more on the fact that ROC curves are a general summary of the model's performance, but we must also examine the specific cases where the model misclassifies objects to determine if there is a pattern.}
+\mt{ROC curves are useful for summarizing the accuracy of model predictions, but they do not indicate which classes are being assigned the false positive or false negative labels when accuracy is not perfect.
+A confusion matrix is a helpful diagnostic tool to understand which classes are commonly confused for other classes.} A confusion matrix is a \mt{summary of the relationships between} the labels assigned to an image and the labels predicted by the model.
+
+In the confusion matrix presented in \autoref{fig:confusion_matrix}, the values along the diagonal represent the proportion of true positives captured within each category.
+The off-diagonal values represent the proportion of false positives for each shape pairing; however, since a single image may truly contain multiple shapes, the values \svp{have been} adjusted to remove the effect of any true positives from the calculation of false positive proportions.
+\svp{I can't remember exactly how we're doing that, but we may want to look more closely at that - if we set everything to NA instead of 0, the rate calculations would be higher than expected}.
+For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any image that truly contains both triangles and quadrilaterals is removed before calculating the proportion of false quadrilateral labels.
+
+<<confusion_matrix, fig.width = 9, fig.height = 8, out.width = "\\textwidth", fig.cap = "Confusion matrix, showing on the diagonal the correct classification rate and on the off-diagonal, classification errors.
+Note that in multi-label images, correct off-diagonal labels have been excluded from the calculation of false positives.", fig.scap = "Confusion matrix, with correct and incorrect model classifications.", dpi = 600>>=
+get_confusion_matrix(predictions = preds, classes = classes,
+                     test_labels = test_labs, threshold = thresholds) %>%
+  set_names(str_to_title(classes)) %>%
+  ggcorrplot(., hc.order = F, outline.col = "white", lab = T) +
+  scale_fill_gradient("Classification\nRate", low = "white",
+                      high = "cornflowerblue", limits = c(0, 1)) +
+  scale_x_discrete("Image Label") + scale_y_discrete("Prediction") +
+  theme(axis.title.x = element_text(size = 14),
+        axis.title.y = element_text(size = 14, angle = 90, vjust = 1)) +
+  ggtitle("CoNNOR Multi-Class Confusion Matrix: Test Set Performance") +
+  theme_bw() +
+  theme(panel.grid.major = element_line(color = "grey50"),
+        panel.grid.minor = element_line(color = "grey60")) +
+  theme(plot.margin = grid::unit(c(0,0,0,0), "mm"),
+        plot.background = element_rect(fill = "transparent", color = NA),
+        plot.subtitle = element_blank(), plot.caption = element_blank(),
+        panel.spacing = unit(c(0, 0, 0, 0), "mm"))
+@
+
+The horizontal band in \autoref{fig:confusion_matrix} indicates that quadrilaterals are predicted more often than they should be for every true label.
+Similarly, polygons, stars, and triangles produce a large number of false positive predictions into other categories, as evidenced by the vertical bands.
+Another interesting artifact in \autoref{fig:confusion_matrix} is that circles are often misclassified as text, and text is often mislabeled as containing circles.
+Not all of the images used in model training are accurately labeled: in some cases, the labels have not \mt{yet} been updated after labeling guidelines have changed.
+The confusion between text and circles is a natural consequence of the overlap between geometric shapes and components of written text; it is not indicative of a lack of fit as much as a weakness in the classification system used to define the model.
+
+\autoref{fig:confusion_matrix} shows \svp{systematic errors in class predictions}, but does not necessarily suggest \emph{why} the confusion is taking place.
+To assess the reasons for model failure, \svp{we look at the images to} \fix{see what the model is seeing.}
+%the general weaknesses of the model, but does not necessarily suggest where to look as to \emph{why} the model is not performing appropriately.
+
+\svp{I would move heatmaps up and out of the diagnostic section and transition into that, possibly}
+
+\svp{Three main model exploration chunks: consistency, heatmaps, brand (motivated by the fact that some classes perform so badly because the exemplars are so variable) - need to figure out the best sequence to transition}
+
+\paragraph{Model Predictions}
+
+ \mt{What else to say here?}
+
+<<exemplars-consistency, fig.cap = "Examples", fig.pos = "p!", out.width = "100%", cache = T, fig.pos = "!h">>=
+exemplar <- list.files(file.path(wd, "images", "consistency", "exemplar"), full.names = T)
+pred_prob_plot(exemplar, loaded_model)
+@
+
+\fix{Can also look at many examples within same shape categories.}
+
+<<chevrons-consistency, fig.cap = "Chevrons", fig.pos = "p!", out.width = "100%", cache = T>>=
+chevron <- list.files(file.path(wd, "images", "consistency", "chevron"), full.names = T)
+pred_prob_plot(chevron, loaded_model)
+@
+
+<<circle-text-consistency, fig.cap = "Text with circles or circle-like shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
+circle_text <- list.files(file.path(wd, "images", "consistency", "circle-text"), full.names = T)
+pred_prob_plot(circle_text, loaded_model)
+@
+
+\svp{I've moved this down here because I think it fits better in the model exploration section - using this to explain why performance differs - intraclass variability.}
+
+\paragraph{Brand Investigations}
+
 <<brand-plots, device = "cairo", include = F>>=
@@
 
-%Labeled images are scarce relative to the amount of data necessary to train a neural network.
-\mt{Although transfer learning substantially reduces the amount of data required to train a CNN, l}abeled images are \mt{still difficult or time-consuming to generate} %scarce \fix{[difficult to generate]}
-relative to the amount of data necessary to train \mt{even a small} neural network \mt{adequately.
-Another common} solution to data scarcity is to artificially enlarge the data set using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}.
-Augmentation is the transformation of original input data using image operations such as cropping, zoom, skew, rotation, and color balance modification in order to distort or alter the image while maintaining the essential features corresponding to the label.
-This process reduces the potential for overfitting the model to the specific set of image data used during the training process, and also increases the amount of data available for training; examples of pre- and post-augmentation images are shown in \autoref{fig:augmented}.
-
-\begin{figure}
-\centering
-\includegraphics[width=.1\textwidth]{images/augmentation/bowtie(R)-1-ugg-sienna-enamel-blue_product_8726365_color_120557.jpg}\hfill
-\includegraphics[width=.1\textwidth]{images/augmentation/aug_bowtie(R)-1-ugg-sienna-enamel-blue_product_8726365_color_120557_0_8344.jpg}\hfill\hfill\hfill
-\includegraphics[width=.1\textwidth]{images/augmentation/quad-2-nike-kids-downshifter-7-infant-toddler-gunsmoke-sunset-pulse-atmosphere-grey_product_8800875_color_720973.jpg}\hfill
-\includegraphics[width=.1\textwidth]{images/augmentation/aug_quad-2-nike-kids-downshifter-7-infant-toddler-gunsmoke-sunset-pulse-atmosphere-grey_product_8800875_color_720973_0_7533.jpg}\hfill\hfill\hfill
-\includegraphics[width=.1\textwidth]{images/augmentation/quad-2-puma-suede-classic-x-mac-three-port-royale-port-royale_product_9123838_color_251426.jpg}\hfill
-\includegraphics[width=.1\textwidth]{images/augmentation/aug_quad-2-puma-suede-classic-x-mac-three-port-royale-port-royale_product_9123838_color_251426_0_5689.jpg}\hfill\hfill\hfill
-\includegraphics[width=.1\textwidth]{images/augmentation/text-1-bernie-mev-kids-catwalk-little-kid-big-kid-multi-camo_product_9084647_color_87089.jpg}\hfill
-\includegraphics[width=.1\textwidth]{images/augmentation/aug_text-1-bernie-mev-kids-catwalk-little-kid-big-kid-multi-camo_product_9084647_color_87089_0_1891.jpg}
-\caption[Original and augmented images.]{Four sets of original (left) and augmented (right) labeled images.}\label{fig:augmented}
-\end{figure}
-
-Model training is analogous to the human learning process.
-For example, a child learns to identify dogs by being presented with many labeled examples, such as when their parent uses the label ``dog" for an animal walking by.
-That child's understanding of dogs is then measured by how many dogs the child is able to correctly identify, and also by the number of other animals it mistakenly calls ``dog".
-Similarly, CNN training is a series of stages, or ``epochs", where the model learns from the set of labeled training data, and that learning is measured through intermediate predictions of validation data.
-
-The \Sexpr{nrow(dfunion)} images were split such that 60\% were used for training. Since \mt{some categories are much more common than others,}
-%the categories do not exist in equal proportion in the labeled data,
-the training data were weighted by proportion during the training process to prevent the \mt{model learning} from being overwhelmed by more frequent categories.
-Of the remaining 40\% of data, half were used for validation, to monitor the training process, and the remaining data were for testing the performance of the fitted model.
-
-\svp{This section header might need to move? Haven't decided yet...}
-\section{Results}
-
-\subsection{Model Training}
-<<training-accuracy, fig.width = 7.5, fig.height = 5, out.width = ".75\\textwidth", fig.cap = "Training and validation accuracy and loss for each epoch of the fitting process.
-Training and validation accuracy reach 89.5\\% around epoch 9.
-After that point, validation loss remains the same and training loss decreases slightly, while validation accuracy increases more slowly than training accuracy.", fig.scap = "Training and validation accuracy and loss during each epoch.", fig.pos = 'h'>>=
-data.frame(history$metrics) %>%
-  mutate(epoch = 1:n()) %>%
-  gather(key = "measure", value = "value", -epoch) %>%
-  mutate(Type = ifelse(str_detect(measure, "val"), "Validation", "Training"),
-         measure = ifelse(str_detect(measure, "acc"), "Accuracy", "Loss")) %>%
-  # bind_rows(tibble(epoch = NA, value =  .6, measure = "Accuracy", Type = "Validation")) %>%
-  # bind_rows(tibble(epoch = NA, value =  .33, measure = "Loss", Type = "Validation")) %>%
-  ggplot(aes(x = epoch, y = value, color = Type)) +
-  geom_point() +
-  geom_smooth(se = F) +
-  facet_grid(measure~., scales = "free_y", switch = "both") +
-  theme_bw() +
-  scale_y_continuous("") +
-  scale_x_continuous("Epoch") +
-  ggtitle("CoNNOR Training Performance") + mytheme +
-  theme(axis.title.y = element_blank(), legend.position = c(1, .5),
-        legend.justification = c(1.03, -0.05),
-        legend.background = element_rect(fill = "white"))
-@
-
-\fix{How much should be said about loss and loss functions?}
-\mt{A loss function is a function that measures how far model predictions are from the given labels for each image in the validation and test sets; the model ``learns" by updating its predictions in such a way that will minimize this loss.}
-
-\autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process.
-Overfitting \mt{occurs when a model learns the training data so well that its understanding of the categories becomes specific to only the training cases, which, in turn, leads to poor prediction of new images that were not in the training set.}
-%Overfitting, or fitting a model which performs too well on the training data relative to the validation data,
-Overfitting is suspected if the validation loss starts to increase after reaching a global minimum.
-Alternately, underfitting occurs if the validation accuracy is still increasing when model optimization is terminated\mt{, because model performance is still improving with continued training}.
-Neither of these outcomes appears in \autoref{fig:training-accuracy}, indicating that the model optimization process was halted at an appropriate epoch.
-
-\subsection{Model Accuracy}
-
-\paragraph{Overall Accuracy} % Model-wise and class-by-class, w/ confusion matrix
-%\svp{You need to explain these terms.
-Your CC was written for statisticians, this paper is for practitioners.
-You need a MUCH gentler introduction to the idea of accuracy, errors, and ROC curves, as well as a lot more definitions for terms.}
-
-\mt{For each image in the test set, the ``true labels" are nine human-assigned labels of 0 or 1 corresponding to each of the nine shape categories.
-When the CNN predicts which shapes are in the image, however, it assigns a probability between 0 and 1 for each shape category, with a total of nine probabilities per test image.
-Thus, to determine the accuracy of the model predictions, one must choose a threshold to the probabilities, such that a predicted probability above the threshold indicates a ``yes" prediction from the model and a probability below the threshold indicates a ``no".
-If the threshold is too high, \fix{moderate} probabilities will be treated as a ``no", which increases the number of true negative and false negative predictions.
-Conversely, if the threshold is too low, \fix{moderate} probabilites will be considered as ``yes", which increases the true and false positive predictions in the data.
-An appropriate threshold must be chosen to produce a high true positive rate while also ideally keeping the false positive rate low.}
-
-\mt{Receiver Operating Characteristic (ROC) curves are a type diagnostic plot that compares the false positive rate against the true positive rate for a classification method, computed by comparing the two rates at a number of thresholds, and the Area Under the Curve (AUC) quantifies the shape of the ROC curve.
-Perfect prediction would be indicated by a right angle along the upper left corner of the plot, with a corresponding AUC of 1, and a diagonal line with an AUC of 0.5 would indicate that the classification method performs no better than random chance.} \autoref{fig:overall-roc} shows the ROC curve for \fix{the model} across all classes, and \autoref{fig:class-roc} shows the curve for each class.
-The full model has an AUC of 0.88, and the AUC for individual classes ranges from 0.81 (for line) to 0.91 (for bowtie and text).
-While the class performances do vary slightly, each ROC curve is the same general shape and performs \fix{significantly} better than random chance.
-\mt{The points in \autoref{fig:class-roc} represent the equal error rate (EER), which is the threshold where there is equal probability of a false positive or false negative error.
-Thus, for each category, any model prediction with probability above the EER for that shape is considered a positive prediction.}
-
-<<overall-roc, fig.width = 5, fig.height = 5, out.width = ".5\\textwidth", fig.cap = "ROC curve showing overall model performance.", fig.scap = "Overall model performance ROC curve.">>=
-library(pROC)
-pred_df <- as_tibble(preds) %>% gather(key = feature, value = value)
-test_labs_df <- as_tibble(test_labs) %>% gather(key = feature, value = value)
-whole_model_roc <- roc(test_labs_df$value, pred_df$value)
-
-whole_model_roc_df <- tibble(tpr = whole_model_roc$sensitivities,
-                             fpr = 1 - whole_model_roc$specificities,
-                             thresholds = whole_model_roc$thresholds,
-                             auc = whole_model_roc$auc[1]) %>%
-  nest(tpr, fpr, thresholds, .key =  "roc_plot") %>%
-  mutate(eer = purrr::map(roc_plot, eer))
-ggplot() +
-  geom_line(aes(x = fpr, y = tpr), data = unnest(whole_model_roc_df, roc_plot), size = 1.25) +
-  geom_label(aes(x = 1, y = .07, label = sprintf("AUC: %0.2f", auc)), hjust = 1, vjust = -0.2, data = whole_model_roc_df) +
-  geom_point(aes(x = fpr, y = tpr, color = "Equal Error Rate"), data = unnest(whole_model_roc_df, eer), size = 2) +
-  scale_color_manual("", values = "black") +
-  scale_x_continuous("False Positive Rate", breaks = c(0, .25, .5, .75, 1), labels = c("0.0", "", "0.5", "", "1.0")) +
-  scale_y_continuous("True Positive Rate", breaks = c(0, .25, .5, .75, 1), labels = c("0.0", "", "0.5", "", "1.0")) +
-  ggtitle("CoNNOR Test Set Performance (All Classes)") +
-  coord_fixed() + mytheme +
-  theme(legend.position = c(1, 0), legend.justification = c(1.01, -0.01), legend.title = element_blank(), legend.background = element_rect(fill = "white"))
-
-@
-
-<<class-roc, fig.width = 8, fig.height = 6, out.width = "\\textwidth", fig.cap = "Class-by-class ROC curves.
-AUC is area under the curve, a measure of overall model performance.
-Equal error rates are marked, indicating the position at which there is equal probability of a false positive or false negative error.", fig.scap = "Class-by-class ROC curves.">>=
-aucs <- plot_onehot_roc(preds, test_labs, str_to_title(classes))
-thresholds <- purrr::map_dbl(aucs$data$eer, ~.$thresholds)
-aucs$data$thresholds <- thresholds
-
-ggplot() +
-  geom_line(aes(x = fpr, y = tpr), data = unnest(aucs$data, roc_plot), size = 1.25) +
-  geom_label(aes(x = 1, y = 0, label = sprintf("AUC: %0.2f\nEER: %0.2f", auc, thresholds)), hjust = 1, vjust = -0.02, data = aucs$data) +
-  geom_point(aes(x = fpr, y = tpr, color = "Equal Error\nRate (EER)"), data = unnest(aucs$data, eer), size = 2.5) +
-  scale_color_manual("", values = "black") +
-  facet_wrap(~class) +
-  scale_x_continuous("False Positive Rate", breaks = c(0, .25, .5, .75, 1), labels = c("0.0", "", "0.5", "", "1.0")) +
-  scale_y_continuous("True Positive Rate", breaks = c(0, .25, .5, .75, 1), labels = c("0.0", "", "0.5", "", "1.0")) +
-  ggtitle("CoNNOR Test Set Performance") +
-  facet_wrap(~class, nrow = 2) +
-  coord_fixed() +
-  theme(legend.position = c(1, 0), legend.justification = c(1, 0))
-@
-
-\svp{Pros and cons of confusion matrix here? Maybe we just don't show the actual proportions on the graph? I'd like to discuss this more in person.}
-
-\mt{ROC curves are useful for summarizing the accuracy of model predictions, but they do not indicate which classes are being assigned the false positive or false negative labels when accuracy is not perfect.
-A confusion matrix is a helpful diagnostic tool to understand which classes are commonly confused for other classes.} A confusion matrix is a \mt{summary of the relationships between} the labels assigned to an image and the labels predicted by the model.
-
-In the confusion matrix presented in \autoref{fig:confusion_matrix}, the values along the diagonal represent the proportion of true positives captured within each category.
-The off-diagonal values represent the proportion of false positives for each shape pairing; \mt{however, since a single image may truly contain multiple shapes}, the values are adjusted to remove the effect of any true positives from the calculation of false positive proportions.
-For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any image that truly contains both triangles and quadrilaterals is removed before calculating the proportion of false quadrilateral labels.
-
-<<confusion_matrix, fig.width = 9, fig.height = 8, out.width = "\\textwidth", fig.cap = "Confusion matrix, showing on the diagonal the correct classification rate and on the off-diagonal, classification errors.
-Note that in multi-label images, correct off-diagonal labels have been excluded from the calculation of false positives.", fig.scap = "Confusion matrix, with correct and incorrect model classifications.", dpi = 600>>=
-get_confusion_matrix(predictions = preds, classes = classes,
-                     test_labels = test_labs, threshold = thresholds) %>%
-  set_names(str_to_title(classes)) %>%
-  ggcorrplot(., hc.order = F, outline.col = "white", lab = T) +
-  scale_fill_gradient("Classification\nRate", low = "white",
-                      high = "cornflowerblue", limits = c(0, 1)) +
-  scale_x_discrete("Image Label") + scale_y_discrete("Prediction") +
-  theme(axis.title.x = element_text(size = 14),
-        axis.title.y = element_text(size = 14, angle = 90, vjust = 1)) +
-  ggtitle("CoNNOR Multi-Class Confusion Matrix: Test Set Performance") +
-  theme_bw() +
-  theme(panel.grid.major = element_line(color = "grey50"),
-        panel.grid.minor = element_line(color = "grey60")) +
-  theme(plot.margin = grid::unit(c(0,0,0,0), "mm"),
-        plot.background = element_rect(fill = "transparent", color = NA),
-        plot.subtitle = element_blank(), plot.caption = element_blank(),
-        panel.spacing = unit(c(0, 0, 0, 0), "mm"))
-@
-
-The horizontal band in \autoref{fig:confusion_matrix} indicates that quadrilaterals are predicted more often than they should be for every true label.
-Similarly, polygons, stars, and triangles produce a large number of false positive predictions into other categories, as evidenced by the vertical bands.
-Another interesting artifact in \autoref{fig:confusion_matrix} is that circles are often misclassified as text, and text is often mislabeled as containing circles.
-Not all of the images used in model training are accurately labeled: in some cases, the labels have not \mt{yet} been updated after labeling guidelines have changed.
-The confusion between text and circles is a natural consequence of the overlap between geometric shapes and components of written text; it is not indicative of a lack of fit as much as a weakness in the classification system used to define the model.
-
-\autoref{fig:confusion_matrix} shows \mt{which classes are being predicted incorrectly, but does not necessarily suggest \emph{why} the confusion is taking place.
-To investigate further, it is useful to} \fix{see what the model is seeing.}
-%the general weaknesses of the model, but does not necessarily suggest where to look as to \emph{why} the model is not performing appropriately.
-
-\paragraph{Model Predictions}
-
- \mt{What else to say here?}
-
-<<exemplars-consistency, fig.cap = "Examples", fig.pos = "p!", out.width = "100%", cache = T, fig.pos = "!h">>=
-exemplar <- list.files(file.path(wd, "images", "consistency", "exemplar"), full.names = T)
-pred_prob_plot(exemplar, loaded_model)
-@
-
-\fix{Can also look at many examples within same shape categories.}
-
-<<chevrons-consistency, fig.cap = "Chevrons", fig.pos = "p!", out.width = "100%", cache = T>>=
-chevron <- list.files(file.path(wd, "images", "consistency", "chevron"), full.names = T)
-pred_prob_plot(chevron, loaded_model)
-@
-
-<<circle-text-consistency, fig.cap = "Text with circles or circle-like shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
-circle_text <- list.files(file.path(wd, "images", "consistency", "circle-text"), full.names = T)
-pred_prob_plot(circle_text, loaded_model)
-@
-
 \paragraph{Model Diagnostics}
 
-\mt{A heatmap is a visual diagnostic tool that highlights the portions of an image that are most significant to classification for a given class.
-Using these graphics, we can identify which locations in an image contribute most to the model's predictions.} In \fix{the figures in this subsection,} heatmaps are shown for a selection of three classes, which are either contained in the image or have high output probabilities, alongside the original image.
+A heatmap \fix{please add the technical name for heatmaps - look in the distill.io papers?} is a visual diagnostic tool that highlights the portions of an image that are most significant to classification for a given class.
+Using these graphics, we can identify which locations in an image contribute most to the model's predictions.
+In \fix{the figures in this subsection,} heatmaps are shown for a selection of three classes, which are either contained in the image or have high output probabilities, alongside the original image.
 
 \fix{I picked the heatmaps I thought illustrated the point best, but didn't modify them heavily yet.
-How many do we want to include, and do we want to highlight model challenges or just focus on the point that CoNNOR sees what we see?}
+How many do we want to include, and do we want to highlight model challenges or just focus on the point that CoNNOR sees what we see?} \svp{I think it's worth highlighting the fact that we'll never get a model that's perfect because that curly s is pretty much a circle, and there will always be weird shapes that partially activate different categories.}
 
+\svp{We may want to explore using the border of the square to show whether the model actually assigns that label - like the bullet plots}
 <<chevron-circle-heatmap, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing chevrons and circles.", fig.cap = "Chevrons are identified by sharp corners and parallel lines, but the thick line between the two shape sections is falsely identified as a quadrilateral.", fig.pos = "!h">>=

### 2019-06-10 paper/paper.Rnw
Incorporate edits and add brief results outline

@@
 Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain.
-%\svp{Add 1-2 sentences connecting feature sets to the goal of reducing workload - possibly cite solemate as an example of manual feature identification}
 In addition, much of the existing software to assess geometric similarity, such as SoleMate\footnote{\url{http://www.fosterfreeman.com/trace-evidence/357-sicar-6-solemate-2.html}}, requires examiners to identify and label features manually; automating classification within the currently used feature space could reduce or eliminate the need for such manual classification.
@@
 \section{Materials and Methods}
-%\svp{Short bit (1-2 sentences) about low-level feature detectors and why they fail with this problem.}
-\mt{Shoeprint evidence from crime scenes is most commonly collected in the form of a photograph, so any useful method to automatically identify outsole characteristics must take the form of an image analysis task.}\svp{We're not working with prints, though - please try to avoid the temptation to delve too far into prints.
-"Ultimately, most pattern evidence is stored in photographic form; as a result, we must use image analysis methods to create numerical features from the original visual record." or something to that effect.}
-There are a number of methods that may be employed to identify shapes and features in an image, such as \svp{Fourier transforms, }Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}.
+
+\mt{Ultimately, most pattern evidence is stored in photographic form; as a result, we must use image analysis methods to create numerical features from the original visual record.} There are a number of methods that may be employed to identify shapes and features in an image, such as Fourier transforms, Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}.
 While these methods are useful in identifying these specific features at a low level, they can be computationally intensive and only identify features on a very small scale; as a result, they cannot reliably identify large geometric shapes like those that may be found in an outsole image.
+
 %\svp{Motivate the use of CNNs directly - complex model, higher level of feature aggregation. Mention pre-trained networks to transition.}
@@
 CNNs are a form of artificial neural network which make use of the image convolution operator used by many low-level feature extraction methods, with the additional ability to aggregate such features and meaningfully connect them to a pre-determined set of labels.
-CNNs have deep architectures that can be trained to identify complex patterns, but they are structurally similar to the \svp{architecture of the }human visual \svp{system} and output binary or probabilistic predictions for given labels that are readily interpretable.
+CNNs have deep architectures that can be trained to identify complex patterns, but they are structurally similar to the architecture of the human visual system and output binary or probabilistic predictions for given labels that are readily interpretable.
 As CNNs make use of labeled training data, the predictions generated are for features which are similar to those identified by humans, resulting in models with greater face validity.
@@
 As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks.
-\svp{You need to talk about the model structure of vgg16 before you start talking about layers.
-Talk of convolutional blocks doesn't really make sense until you show the layer structure.}\autoref{fig:vgg16layers} shows a selection of filters from VGG16, where each filter detects the presence of specific shapes and colors.
+\svp{You need to talk about the model structure of vgg16 before you start talking about layers. Talk of convolutional blocks doesn't really make sense until you show the layer structure.}
+\mt{I agree, but I thought that this was where we were arguing that the ``eyes" are transferrable, so I wanted to talk about the filters to back that up.}
+\autoref{fig:vgg16layers} shows a selection of filters from VGG16, where each filter detects the presence of specific shapes and colors.
 Each row of filters in \autoref{fig:vgg16layers} corresponds to one of the five convolutional blocks of VGG16.
@@
 
-<<vgg16layers, fig.width = 7, fig.height = 5.75, fig.cap = "A selection of filters from the convolutional layers of VGG16.\\svp{Label the rows with block numbers please}">>=
-imgs <- list.files("./images/vgg16layers/", recursive = T, full.names = T)
+<<vgg16layers, fig.width = 7, fig.height = 5.75, fig.cap = "A selection of filters from the convolutional layers of VGG16.">>=
+imgs <- list.files(file.path(wd, "images", "vgg16layers"),
+                   recursive = T, full.names = T)
 pngs <- lapply(imgs, function(x) png::readPNG(x)[46:1596, 32:1582, ])
 grobs <- lapply(pngs, grid::rasterGrob)
-gridExtra::grid.arrange(grobs = grobs, nrow = 5)
+filters <- gridExtra::arrangeGrob(grobs = grobs, nrow = 5)
+ggsave(filename = file.path(wd, "images", "unlabeled_filters.png"),
+       plot = filters, device = "png", width = 6, height = 5.3)
+
+library(magick)
+b <- image_blank(width = 300, height = 1590)
+f <- image_read(file.path(wd, "images", "unlabeled_filters.png"))
+f_l <- image_append(c(b, f)) %>%
+  image_annotate("Block 1:", location = "+60+159",
+                 color = "black", size = 60) %>%
+  image_annotate("Block 2:", location = "+60+477",
+                 color = "black", size = 60) %>%
+  image_annotate("Block 3:", location = "+60+795",
+                 color = "black", size = 60) %>%
+  image_annotate("Block 4:", location = "+60+1113",
+                 color = "black", size = 60) %>%
+  image_annotate("Block 5:", location = "+60+1431",
+                 color = "black", size = 60) %>%
+  image_scale(geometry = "600x")
+image_write(image = f_l, path = file.path(wd, "images", "labeled_filters.png"),
+            format = "png")
+knitr::include_graphics(file.path(wd, "images", "labeled_filters.png"))
 @
@@
 %Using this definition, shapes such as butterflies are included as bowties.
-\svp{I'm still on the fence - we can just not mention this, or we could move butterflies to "star"...} \mt{I think it makes a lot of sense to move them to star.
-That means redoing the poster/slides examples though...
-grrrrrr} I'm just commenting that bit out for now.
 \item [Chevron] Chevron shapes include repeating parallel lines as well as individual ``v" shapes. They may be angular but can also be curved.
@@
 These regions are grouped and provide additional information - that none of the previous nine categories are present.
+\mt{Why no EM dash?
+``---"}
 \end{description}
@@
 %Labeled images are scarce relative to the amount of data necessary to train a neural network.
-Although transfer learning substantially reduces the amount of data required to train a CNN \svp{by several orders of magnitude}, \fix{labeled images are still difficult or time-consuming to generate relative to the amount of data necessary to train even a small neural network adequately.
-Another common solution to data scarcity is to artificially enlarge the data set using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}.} \svp{I would lose the last clause of the first sentence and combine it with the next, e.g.
+Although transfer learning substantially reduces the amount of data required to train a CNN by several orders of magnitude, \fix{labeled images are still difficult or time-consuming to generate relative to the amount of data necessary to train even a small neural network adequately.
+Another common solution to data scarcity is to artificially enlarge the data set using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}.}
+\svp{I would lose the last clause of the first sentence and combine it with the next, e.g.
 to make most efficient use of the labeled data, we enlarge the training data using a process called image augmentation.}
+\mt{I don't think the first half of the first sentence is relevant here without the second.
+Should I use a different transition?
+I was trying to put augmentation in the context of transfer learning; i.e., transfer learning helps, but getting data is hard, so we use augmentation to get more, with the added benefit of improving generalizability.}
 Augmentation is the transformation of original input data using image operations such as cropping, zoom, skew, rotation, and color balance modification in order to distort or alter the image while maintaining the essential features corresponding to the label.
@@
 
-The \Sexpr{nrow(dfunion)} images were split such that 60\% were used for training. Since \mt{some categories are much more common than others,}\svp{I like the equal proportion phrasing better myself}
-%the categories do not exist in equal proportion in the labeled data,
-the training data were weighted by proportion during the training process to \svp{ensure that the model can identify both rare and common geometric shapes}.%from being overwhelmed by more frequent categories.
+The \Sexpr{nrow(dfunion)} images were split such that 60\% were used for training. Since the categories do not exist in equal proportion in the labeled data,
+the training data were weighted by proportion during the training process to ensure that the model can identify both rare and common geometric shapes.
 Of the remaining 40\% of data, half were used for validation, to monitor the training process, and the remaining data were for testing the performance of the fitted model.
@@
 %Overfitting, or fitting a model which performs too well on the training data relative to the validation data,
-\svp{get rid of passive voice here.
-Overfitting occurs when the validation loss begins to increase after reaching a global minimum} Overfitting is suspected if the validation loss starts to increase after reaching a global minimum.
-Alternately, underfitting occurs if the validation accuracy is still increasing when model optimization is terminated\svp{(for instance, between epochs 1 and 10 in \autoref{fig:training-accuracy})}, because model performance is still improving with continued training.
-\svp{In \autoref{fig:training-accuracy}, validation accuracy levels off after epoch 20, and validation loss has not yet begun to increase at epoch 30,} indicating that the model optimization process was halted at an appropriate epoch.
+Overfitting occurs when the validation loss begins to increase after reaching a global minimum.
+Alternately, underfitting occurs if the validation accuracy is still increasing when model optimization is terminated (for instance, between epochs 1 and 10 in \autoref{fig:training-accuracy}), because model performance is still improving with continued training.
+In \autoref{fig:training-accuracy}, validation accuracy levels off after epoch 20, and validation loss has not yet begun to increase at epoch 30, indicating that the model optimization process was halted at an appropriate epoch.
 
 \subsection{Model Accuracy}
+
+\mt{\textbf{OUTLINE OF RESULTS}
+\begin{itemize}
+\item ROC shows that, in general, model performance is better than chance and broadly compares class performance.
+\item Show model predictions primarily to provide a few examples to back up the claim that predictions generally look good, but also include one or two case studies that motivate class-by-class diagnostics
+\item Confusion matrix shows which classes are most accurate or most problematic.
+However, when we see that one shape is commonly misclasified as another, the conf mat doesn't let us do anything more than speculate the cause.
+\item Heatmaps help us look at specific instances of classification to 1) verify that the model is using appropriate features when it classifies correctly and 2) diagnose the cause of confusion when the model classifies incorrectly
+\item (Potentially switch the order of prediction examples and the confusion matrix, so the conf mat indicates systemic class confusion and then back that up with the examples?)
+\end{itemize}
+}
 
@@
 When the CNN predicts which shapes are in the image, however, it assigns a probability between 0 and 1 for each shape category, with a total of nine probabilities per test image.
-Thus, to determine the accuracy of the model predictions, \svp{fix weird phrasing - we must select a threshold for each label to discretize the model predictions} \fix{one must choose a threshold to the probabilities, such that a predicted probability above the threshold indicates a ``yes" prediction from the model and a probability below the threshold indicates a ``no".} If the threshold is too high, \fix{moderate} probabilities will be treated as a ``no", which increases the number of true negative and false negative predictions.
+Thus, to determine the accuracy of the model predictions, we must select a threshold for each label to discretize the model predictions.
+%such that a predicted probability above the threshold indicates a ``yes" prediction from the model and a probability below the threshold indicates a ``no".}
+If the threshold is too high, \fix{moderate} probabilities will be treated as a ``no", which increases the number of true negative and false negative predictions.
 Conversely, if the threshold is too low, \fix{moderate} probabilites will be considered as ``yes", which increases the true and false positive predictions in the data.
@@
 
-Receiver Operating Characteristic (ROC) curves are a type diagnostic plot \svp{which compare} the false positive rate \svp{to} the true positive rate for a classification method.\svp{Make this two sentences so that it's less convoluted}\fix{, computed by comparing the two rates at a number of thresholds, and the Area Under the Curve (AUC) quantifies the shape of the ROC curve.} Perfect prediction would be indicated by a right angle along the upper left corner of the plot, with a corresponding AUC of 1, and a diagonal line with an AUC of 0.5 would indicate that the classification method performs no better than random chance.
+%\mt{I don't immediately see how to separate the below sentence into two without making it really clunky...
+I figured that in the context of thresholds (previous paragraph), it would be easy to just stick in that those thresholds are used to compute the curve.}
+
+%Better?: Receiver Operating Characteristic (ROC) curves are a type diagnostic plot which compare the false positive rate to the true positive rate for a classification method at a number of thresholds, and the Area Under the Curve (AUC) quantifies the shape of the ROC curve.
+
+Receiver Operating Characteristic (ROC) curves are a type diagnostic plot which compare the false positive rate to the true positive rate for a classification method.\svp{Make this two sentences so that it's less convoluted}\fix{, computed by comparing the two rates at a number of thresholds, and the Area Under the Curve (AUC) quantifies the shape of the ROC curve.} Perfect prediction would be indicated by a right angle along the upper left corner of the plot, with a corresponding AUC of 1, and a diagonal line with an AUC of 0.5 would indicate that the classification method performs no better than random chance.
 \autoref{fig:overall-roc} shows the ROC curve for \fix{the model} across all classes, and \autoref{fig:class-roc} shows the curve for each class.
@@
 The points in \autoref{fig:class-roc} represent the equal error rate (EER), which is the threshold where there is equal probability of a false positive or false negative error.
-\svp{We will use the class-wise EER to set a threshold corresponding to each class, so that }any model prediction with probability above the EER for that shape is considered a positive prediction.
+We will use the class-wise EER to set a threshold corresponding to each class, so that any model prediction with probability above the EER for that shape is considered a positive prediction.
 
@@
 \svp{I'd focus more on the fact that ROC curves are a general summary of the model's performance, but we must also examine the specific cases where the model misclassifies objects to determine if there is a pattern.}
-\mt{ROC curves are useful for summarizing the accuracy of model predictions, but they do not indicate which classes are being assigned the false positive or false negative labels when accuracy is not perfect.
+\mt{ROC curves provide a general summary of model accuracy, but they do not indicate which classes are being assigned the false positive or false negative labels when accuracy is not perfect.
 A confusion matrix is a helpful diagnostic tool to understand which classes are commonly confused for other classes.} A confusion matrix is a \mt{summary of the relationships between} the labels assigned to an image and the labels predicted by the model.
@@
 In the confusion matrix presented in \autoref{fig:confusion_matrix}, the values along the diagonal represent the proportion of true positives captured within each category.
-The off-diagonal values represent the proportion of false positives for each shape pairing; however, since a single image may truly contain multiple shapes, the values \svp{have been} adjusted to remove the effect of any true positives from the calculation of false positive proportions.
-\svp{I can't remember exactly how we're doing that, but we may want to look more closely at that - if we set everything to NA instead of 0, the rate calculations would be higher than expected}.
+The off-diagonal values represent the proportion of false positives for each shape pairing; however, since a single image may truly contain multiple shapes, the values have been adjusted to remove the effect of any true positives from the calculation of false positive proportions.
+\svp{I can't remember exactly how we're doing that, but we may want to look more closely at that - if we set everything to NA instead of 0, the rate calculations would be higher than expected} \mt{They're being set to NA, and then we're calculating proportions with na.rm = T.
+I don't understand what you mean about higher rates than expected...}.
 For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any image that truly contains both triangles and quadrilaterals is removed before calculating the proportion of false quadrilateral labels.
@@
 
-\autoref{fig:confusion_matrix} shows \svp{systematic errors in class predictions}, but does not necessarily suggest \emph{why} the confusion is taking place.
-To assess the reasons for model failure, \svp{we look at the images to} \fix{see what the model is seeing.}
-%the general weaknesses of the model, but does not necessarily suggest where to look as to \emph{why} the model is not performing appropriately.
+\autoref{fig:confusion_matrix} shows systematic errors in class predictions, but does not necessarily suggest \emph{why} the confusion is taking place.
+To assess the reasons for model failure, we look at the images to \fix{see what the model is seeing.}
 

### 2019-06-10 paper/paper.Rnw
need to fix explanation of conf matrix

@@
 The off-diagonal values represent the proportion of false positives for each shape pairing; however, since a single image may truly contain multiple shapes, the values have been adjusted to remove the effect of any true positives from the calculation of false positive proportions.
-\svp{I can't remember exactly how we're doing that, but we may want to look more closely at that - if we set everything to NA instead of 0, the rate calculations would be higher than expected} \mt{They're being set to NA, and then we're calculating proportions with na.rm = T.
+\svp{As a result of this adjustment, the probabilities off-diagonal are the proportion ...
+need to explain the conditional probability we're showing here} %\svp{I can't remember exactly how we're doing that, but we may want to look more closely at that - if we set everything to NA instead of 0, the rate calculations would be higher than expected} \mt{They're being set to NA, and then we're calculating proportions with na.rm = T.
 I don't understand what you mean about higher rates than expected...}.

### 2019-06-10 paper/paper.Rnw
Add and restructure consistency images, update outline of results, and other miscellaneous updates

@@
 
-\mt{Ultimately, most pattern evidence is stored in photographic form; as a result, we must use image analysis methods to create numerical features from the original visual record.} There are a number of methods that may be employed to identify shapes and features in an image, such as Fourier transforms, Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}.
+Ultimately, most pattern evidence is stored in photographic form; as a result, we must use image analysis methods to create numerical features from the original visual record.
+There are a number of methods that may be employed to identify shapes and features in an image, such as Fourier transforms, Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}.
 While these methods are useful in identifying these specific features at a low level, they can be computationally intensive and only identify features on a very small scale; as a result, they cannot reliably identify large geometric shapes like those that may be found in an outsole image.
 
-%\svp{Motivate the use of CNNs directly - complex model, higher level of feature aggregation. Mention pre-trained networks to transition.}
 Convolutional neural networks (CNNs) are widely recognized as superior for novel image classification.
@@
 \paragraph{Transfer Learning}
+%\mt{Training a CNN from scratch is hard.
+There are CNNs trained on complex images, and transfer learning lets us use those weights.
+In particular, we're using VGG16.
+We can look at the filters of VGG16 to support our claim that it will be good for the task.}
 Neural networks are composed of sets of layers; the early layers, which are called the model base, contain feature detectors, and the final layers make up the classifier, or model head, which connects meaningful features to classification labels.
@@
 
-Many convolutional neural networks have been trained on a set of images known as ImageNet; these images are hierarchically labeled, so the same image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity \citep{dengImageNetLargeScaleHierarchical}.
+Many convolutional neural networks have been trained on a \mt{sub}set of images \mt{from the database} ImageNet; these images are hierarchically labeled, so the same image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity \citep{dengImageNetLargeScaleHierarchical}.
 CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously.
 As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks.
-\svp{You need to talk about the model structure of vgg16 before you start talking about layers. Talk of convolutional blocks doesn't really make sense until you show the layer structure.}
-\mt{I agree, but I thought that this was where we were arguing that the ``eyes" are transferrable, so I wanted to talk about the filters to back that up.}
-\autoref{fig:vgg16layers} shows a selection of filters from VGG16, where each filter detects the presence of specific shapes and colors.
-Each row of filters in \autoref{fig:vgg16layers} corresponds to one of the five convolutional blocks of VGG16.
-Filters in earlier convolutional blocks detect simple features, such as colors and blobs, while later filters detect more complex combinations of features.
-Each of the five convolutional blocks has between 128 and 1,536 filters, and a single image is made up of some combination of the thousands of features that are detected by those filters.
+\fix{More to be said later.}
 
@@
 It has been successfully applied in automatic classification of medical images \citep{shinDeepConvolutionalNeural2016} as well as in various applications of shoe forensics \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017}.
-
-\fix{In this paper, we will use transfer learning to automatically identify features in shoe treads which are used by forensic examiners, leveraging a general-purpose CNN with relatively simple structure, VGG16, and a database of labeled shoe tread images we have assembled for this project, which are both described below.} \svp{This sentence is a bit awkward, because you're introducing too many things.
-The paragraph is good, but I don't think that it makes sense to have it here.
-Introduce VGG16, then use this paragraph as a transition to talking about the actual model we fit.} This approach differs from other approaches \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017} in automatic footwear identification which use the output from the model base directly and do not attempt to add human-friendly contextual information with an additional classifier.
 
@@
 \paragraph{VGG16}
-Developed by Oxford's Visual Graphics Group, VGG16 is a CNN with 16 ``functional" (i.e., convolutional and densely connected) layers and 5 ``structural" max pooling layers \citep{krizhevskyImageNetClassificationDeep2012}.
+Developed by Oxford's Visual Graphics Group, VGG16 is a \mt{pre-trained} CNN with 16 ``functional" (i.e., convolutional and densely connected) layers and 5 ``structural" max pooling layers \citep{krizhevskyImageNetClassificationDeep2012}.
 In contrast to other popular networks, like ResNet, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy.
@@
 \autoref{fig:VGG16-structure} shows the architecture of VGG16; with transfer learning, the VGG16 model head is replaced with a trained model head specific to recognition of shoeprint class characteristics.
+
+\mt{The model base of VGG16 is made up of five convolutional blocks alternated with max pooling layers.
+Each convolutional block is made up of two or three convolutional layers, and each layer is made up of hundreds of filters, such that each of the five convolutional blocks has between 128 and 1,536 filters.
+Strictly speaking, a filter is a set of numerical weights that, when applied to an image, quantifies the presence of the specific shapes and/or colors that the filter is trained to detect.
+\autoref{fig:vgg16layers} shows a selection of filters from VGG16, with each row corresponding to one of the five convolutional blocks.
+Filters in earlier convolutional blocks detect simple features, such as colors and blobs, while later filters detect more complex combinations of features.
+When considered together, the series of filters perform the process of feature detection, such that a single image is made up of some combination of the thousands of features that are detected by filters.
+Although the filters of VGG16 have been trained and optimized on the ImageNet Data, the large number of filters span a wide variety of shapes and features} \fix{which are still useful/relevant to classifying outsoles.
+In a sense, this is much like teaching humans to classify objects they have never seen before: an object may be unfamiliar, but our eyes can still detect its color, shape, and texture because we have learned these ideas from other sources.}
 
@@
 f_l <- image_append(c(b, f)) %>%
-  image_annotate("Block 1:", location = "+60+159",
+  image_annotate("Block 1:", location = "+60+139",
                  color = "black", size = 60) %>%
-  image_annotate("Block 2:", location = "+60+477",
+  image_annotate("Block 2:", location = "+60+457",
                  color = "black", size = 60) %>%
-  image_annotate("Block 3:", location = "+60+795",
+  image_annotate("Block 3:", location = "+60+775",
                  color = "black", size = 60) %>%
-  image_annotate("Block 4:", location = "+60+1113",
+  image_annotate("Block 4:", location = "+60+1093",
                  color = "black", size = 60) %>%
-  image_annotate("Block 5:", location = "+60+1431",
+  image_annotate("Block 5:", location = "+60+1411",
                  color = "black", size = 60) %>%
@@
 
-In order to train a new head for the VGG16 model base, human-labeled images are necessary. The next section describes how these images were produced and used to fit our custom classifier.
+In this paper, we demonstrate the use of transfer learning to automatically identify features in shoe treads that are used by forensic examiners.
+To do this, we leverage the relatively simple structure and generalizability of VGG16 and train a new classifier using a database of labeled shoe tread images we have assembled for this project.
+This approach differs from other approaches \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017} in automatic footwear identification that use the output from the model base directly and do not attempt to add human-friendly contextual information with an additional classifier.
+The next section describes how the database of images was produced and used to fit our custom classifier.
 
@@
 \item [Other] Other features which were marked include logos, various textures (including crepe, stippling, etc.), and smooth regions with no discernible features.
-These regions are grouped and provide additional information - that none of the previous nine categories are present.
-\mt{Why no EM dash?
-``---"}
+These regions are grouped and provide the additional information that none of the previous nine categories are present.
 \end{description}
@@
 
-<<class-characteristic-barchart, fig.width = 6, fig.height = 4, dpi = 300, fig.cap = "Distribution of classes in all labeled images.
+<<class-characteristic-barchart, fig.width = 6, fig.height = 4, dpi = 300, out.width = "80%", fig.cap = "Distribution of classes in all labeled images.
 Quadrilaterals, lines, circles, text, and chevrons are relatively common; stars, polygons, and bowties are relatively rare.", fig.scap = "Distribution of classes in all labeled images.">>=
@@
 %Labeled images are scarce relative to the amount of data necessary to train a neural network.
-Although transfer learning substantially reduces the amount of data required to train a CNN by several orders of magnitude, \fix{labeled images are still difficult or time-consuming to generate relative to the amount of data necessary to train even a small neural network adequately.
-Another common solution to data scarcity is to artificially enlarge the data set using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}.}
-\svp{I would lose the last clause of the first sentence and combine it with the next, e.g.
-to make most efficient use of the labeled data, we enlarge the training data using a process called image augmentation.}
-\mt{I don't think the first half of the first sentence is relevant here without the second.
-Should I use a different transition?
-I was trying to put augmentation in the context of transfer learning; i.e., transfer learning helps, but getting data is hard, so we use augmentation to get more, with the added benefit of improving generalizability.}
+Although transfer learning substantially reduces the amount of data required to train a CNN by several orders of magnitude, labeled images are still \fix{difficult or time-consuming} to generate relative to the amount of data necessary to train even a small neural network adequately.
+Thus, to make the most efficient use of the existing labeled data, we enlarge the training data using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}.
 Augmentation is the transformation of original input data using image operations such as cropping, zoom, skew, rotation, and color balance modification in order to distort or alter the image while maintaining the essential features corresponding to the label.
@@
 
-\fix{How much should be said about loss and loss functions?}\svp{}
-A loss function is a function that measures how far model predictions are from the given labels for each image in the validation and test sets; the model ``learns" by updating its predictions to minimize the loss.
+In machine learning, a loss function is a function that measures how far model predictions are from the given labels for each image in the validation and test sets; the model ``learns" by updating its predictions to minimize the loss.
+\autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process.
 
-\autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process.
-Overfitting occurs when a model learns the training data so well that its understanding of the categories becomes specific to the training cases, which, in turn, leads to poor prediction of new images that were not in the training set.
-%Overfitting, or fitting a model which performs too well on the training data relative to the validation data,
-Overfitting occurs when the validation loss begins to increase after reaching a global minimum.
+Overfitting is when a model learns the training data so well that its understanding of the categories becomes specific to the training cases, which, in turn, leads to poor prediction of new images that were not in the training set.
+In training, overfitting is indicated when the validation loss begins to increase after reaching a global minimum.
 Alternately, underfitting occurs if the validation accuracy is still increasing when model optimization is terminated (for instance, between epochs 1 and 10 in \autoref{fig:training-accuracy}), because model performance is still improving with continued training.
@@
 \begin{itemize}
-\item ROC shows that, in general, model performance is better than chance and broadly compares class performance.
-\item Show model predictions primarily to provide a few examples to back up the claim that predictions generally look good, but also include one or two case studies that motivate class-by-class diagnostics
-\item Confusion matrix shows which classes are most accurate or most problematic.
-However, when we see that one shape is commonly misclasified as another, the conf mat doesn't let us do anything more than speculate the cause.
-\item Heatmaps help us look at specific instances of classification to 1) verify that the model is using appropriate features when it classifies correctly and 2) diagnose the cause of confusion when the model classifies incorrectly
-\item (Potentially switch the order of prediction examples and the confusion matrix, so the conf mat indicates systemic class confusion and then back that up with the examples?)
+\item ROC
+  \begin{itemize}
+  \item Overall: model performance is better than chance
+  \item Classes: broad comparison of class performance
+  \end{itemize}
+\item Examples
+  \begin{itemize}
+  \item Show what predictions look like
+  \item Introduce the idea that some classifications aren't perfect
+  \item Motivate systematic class-by-class confusion
+  \end{itemize}
+\item Confusion Matrix
+  \begin{itemize}
+  \item Show patterns in misclassification
+  \item Conf mat doesn't tell us why certain predictions are wrong
+  \end{itemize}
+\item In this paper, we identify two main reasons for error: some classes are more variable, and some classes share physical features (I.e., the model has to learn a lot of features to encapsulate a class with lots of variety, but then other classes share those features)
+\item Brands
+  \begin{itemize}
+  \item Bowties/polygons/stars are mostly a few brands, but the top quad/line brands still only take up a small proportion of all quads (.04 or less)
+  \end{itemize}
+\item Heatmaps
+  \begin{itemize}
+  \item Verify that the model is using appropriate features when predictions are correct
+  \item Show examples for strong misclassification, but make the case that the mistakes are understandable
+  \end{itemize}
 \end{itemize}
@@
 
-\paragraph{Overall Accuracy} % Model-wise and class-by-class, w/ confusion matrix
-%\svp{You need to explain these terms.
-Your CC was written for statisticians, this paper is for practitioners.
-You need a MUCH gentler introduction to the idea of accuracy, errors, and ROC curves, as well as a lot more definitions for terms.}
+\paragraph{Overall Accuracy}
 
-For each image in the test set, the ``true labels" are nine human-assigned labels of 0 or 1 corresponding to \svp{the presence or absence of} each of the nine shape categories \svp{in the image}.
+For each image in the test set, the ``true labels" are nine human-assigned labels of 0 or 1 corresponding to the presence or absence of each of the nine shape categories in the image.
 When the CNN predicts which shapes are in the image, however, it assigns a probability between 0 and 1 for each shape category, with a total of nine probabilities per test image.
 Thus, to determine the accuracy of the model predictions, we must select a threshold for each label to discretize the model predictions.
-%such that a predicted probability above the threshold indicates a ``yes" prediction from the model and a probability below the threshold indicates a ``no".}
 If the threshold is too high, \fix{moderate} probabilities will be treated as a ``no", which increases the number of true negative and false negative predictions.
@@
 
-%\mt{I don't immediately see how to separate the below sentence into two without making it really clunky...
-I figured that in the context of thresholds (previous paragraph), it would be easy to just stick in that those thresholds are used to compute the curve.}
-
-%Better?: Receiver Operating Characteristic (ROC) curves are a type diagnostic plot which compare the false positive rate to the true positive rate for a classification method at a number of thresholds, and the Area Under the Curve (AUC) quantifies the shape of the ROC curve.
-
-Receiver Operating Characteristic (ROC) curves are a type diagnostic plot which compare the false positive rate to the true positive rate for a classification method.\svp{Make this two sentences so that it's less convoluted}\fix{, computed by comparing the two rates at a number of thresholds, and the Area Under the Curve (AUC) quantifies the shape of the ROC curve.} Perfect prediction would be indicated by a right angle along the upper left corner of the plot, with a corresponding AUC of 1, and a diagonal line with an AUC of 0.5 would indicate that the classification method performs no better than random chance.
+Receiver Operating Characteristic (ROC) curves are a type diagnostic plot which compare the false positive rate to the true positive rate for a classification method, and the Area Under the Curve (AUC) quantifies the shape of the ROC curve.
+Perfect prediction would be indicated by a right angle along the upper left corner of the plot, with a corresponding AUC of 1, and a diagonal line with an AUC of 0.5 would indicate that the classification method performs no better than random chance.
 \autoref{fig:overall-roc} shows the ROC curve for \fix{the model} across all classes, and \autoref{fig:class-roc} shows the curve for each class.
@@
 
-\svp{Pros and cons of confusion matrix here? Maybe we just don't show the actual proportions on the graph? I'd like to discuss this more in person.}
+\paragraph{Model Predictions}
+\mt{ROC curves provide a general summary of model accuracy.
+To learn more about how specific images are classified, we can view images from the test set along with their respective nine model predictions.
+\autoref{fig:single-exemplars} shows model predictions for images that contain only one type of shape, and \autoref{fig:multi-exemplars} shows similar predictions for images containing multiple shapes.
+Most predictions for the single-label images are straightforward, but there is much less consensus of model opinion in the multi-label images.
+}
 
-\svp{I'd focus more on the fact that ROC curves are a general summary of the model's performance, but we must also examine the specific cases where the model misclassifies objects to determine if there is a pattern.}
-\mt{ROC curves provide a general summary of model accuracy, but they do not indicate which classes are being assigned the false positive or false negative labels when accuracy is not perfect.
-A confusion matrix is a helpful diagnostic tool to understand which classes are commonly confused for other classes.} A confusion matrix is a \mt{summary of the relationships between} the labels assigned to an image and the labels predicted by the model.
+<<single-exemplars, fig.cap = "Examples containing a single shape", fig.pos = "p!", out.width = "100%", cache = T>>=
+archetype <- list.files(file.path(wd, "images", "consistency", "single-exemplar"), full.names = T)
+pred_prob_plot(archetype, loaded_model, sort = F)
+@
+
+<<multi-exemplars, fig.cap = "Examples containing multiple shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
+exemplar <- list.files(file.path(wd, "images", "consistency", "multi-exemplar"), full.names = T)
+pred_prob_plot(exemplar, loaded_model)
+@
+
+\mt{These figures provide representative examples of classifications across multiple categories; it is also useful to examine many cases within a given shape. \autoref{fig:chevrons-consistency}}
+
+<<chevrons-consistency, fig.cap = "Chevrons", fig.pos = "p!", out.width = "100%", cache = T>>=
+chevron <- list.files(file.path(wd, "images", "consistency", "chevron"), full.names = T)
+pred_prob_plot(rev(chevron), loaded_model, sort = F)
+@
+
+<<circle-text-consistency, fig.cap = "Text with circles or circle-like shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
+circle_text <- list.files(file.path(wd, "images", "consistency", "circle-text"), full.names = T)
+pred_prob_plot(rev(circle_text), loaded_model, sort = F)
+@
+
+%\svp{Pros and cons of confusion matrix here? Maybe we just don't show the actual proportions on the graph? I'd like to discuss this more in person.}
+
+\fix{Now that we've seen specific examples, it is useful to summarize the systematic errors by class.} A confusion matrix is a helpful diagnostic tool to understand which classes are commonly confused for other classes.} A confusion matrix is a \mt{summary of the relationships between} the labels assigned to an image and the labels predicted by the model.
 
@@
 \svp{As a result of this adjustment, the probabilities off-diagonal are the proportion ...
-need to explain the conditional probability we're showing here} %\svp{I can't remember exactly how we're doing that, but we may want to look more closely at that - if we set everything to NA instead of 0, the rate calculations would be higher than expected} \mt{They're being set to NA, and then we're calculating proportions with na.rm = T.
-I don't understand what you mean about higher rates than expected...}.
+need to explain the conditional probability we're showing here} %\svp{I can't remember exactly how we're doing that, but we may want to look more closely at that - if we set everything to NA instead of 0, the rate calculations would be higher than expected}
 For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any image that truly contains both triangles and quadrilaterals is removed before calculating the proportion of false quadrilateral labels.
@@
 \svp{Three main model exploration chunks: consistency, heatmaps, brand (motivated by the fact that some classes perform so badly because the exemplars are so variable) - need to figure out the best sequence to transition}
-
-\paragraph{Model Predictions}
-
- \mt{What else to say here?}
-
-<<exemplars-consistency, fig.cap = "Examples", fig.pos = "p!", out.width = "100%", cache = T, fig.pos = "!h">>=
-exemplar <- list.files(file.path(wd, "images", "consistency", "exemplar"), full.names = T)
-pred_prob_plot(exemplar, loaded_model)
-@
-
-\fix{Can also look at many examples within same shape categories.}
-
-<<chevrons-consistency, fig.cap = "Chevrons", fig.pos = "p!", out.width = "100%", cache = T>>=
-chevron <- list.files(file.path(wd, "images", "consistency", "chevron"), full.names = T)
-pred_prob_plot(chevron, loaded_model)
-@
-
-<<circle-text-consistency, fig.cap = "Text with circles or circle-like shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
-circle_text <- list.files(file.path(wd, "images", "consistency", "circle-text"), full.names = T)
-pred_prob_plot(circle_text, loaded_model)
-@
-
-\svp{I've moved this down here because I think it fits better in the model exploration section - using this to explain why performance differs - intraclass variability.}
 

### 2019-06-10 paper/paper.Rnw
Update outline and keep reorganizing and filling in results

@@
 Strictly speaking, a filter is a set of numerical weights that, when applied to an image, quantifies the presence of the specific shapes and/or colors that the filter is trained to detect.
-\autoref{fig:vgg16layers} shows a selection of filters from VGG16, with each row corresponding to one of the five convolutional blocks.
+\fix{Necessary?
+(Or just skip...): Visualizing a filter then refers to visualizing the features to which the filter provides its maximum response.} \autoref{fig:vgg16layers} shows a selection of filters from VGG16, with each row corresponding to one of the five convolutional blocks.
 Filters in earlier convolutional blocks detect simple features, such as colors and blobs, while later filters detect more complex combinations of features.
@@
 
-%\svp{Moving towards ... how did we train the model}
 \paragraph{Annotated Training Data}
@@
 
-%Labeled images are scarce relative to the amount of data necessary to train a neural network.
 Although transfer learning substantially reduces the amount of data required to train a CNN by several orders of magnitude, labeled images are still \fix{difficult or time-consuming} to generate relative to the amount of data necessary to train even a small neural network adequately.
@@
   \item Show patterns in misclassification
-  \item Conf mat doesn't tell us why certain predictions are wrong
+  \item Motivates further understanding of common mistakes
   \end{itemize}
-\item In this paper, we identify two main reasons for error: some classes are more variable, and some classes share physical features (I.e., the model has to learn a lot of features to encapsulate a class with lots of variety, but then other classes share those features)
+\item Two main reasons for error: some classes are more variable, and some classes share physical features (I.e., the model has to learn a lot of features to encapsulate a class with lots of variety, but then other classes share those features)
 \item Brands
   \begin{itemize}
-  \item Bowties/polygons/stars are mostly a few brands, but the top quad/line brands still only take up a small proportion of all quads (.04 or less)
+  \item Bowties/polygons/stars are mostly found in only a few brands, but the top N quad/line brands only take up a small proportion of all quads/lines (.04 or less)
+  \item In general, the most difficult classes to classify are those that are highly variable
   \end{itemize}
@@
   \item Verify that the model is using appropriate features when predictions are correct
-  \item Show examples for strong misclassification, but make the case that the mistakes are understandable
+  \item Show examples for confident misclassifications (like Seychelles), but make the case that the mistakes are understandable
   \end{itemize}
@@
 
-For each image in the test set, the ``true labels" are nine human-assigned labels of 0 or 1 corresponding to the presence or absence of each of the nine shape categories in the image.
+For each image in the test set, the ``true" labels are nine human-assigned labels of 0 or 1 corresponding to the presence or absence of each of the nine shape categories in the image.
 When the CNN predicts which shapes are in the image, however, it assigns a probability between 0 and 1 for each shape category, with a total of nine probabilities per test image.
@@
 \paragraph{Model Predictions}
-\mt{ROC curves provide a general summary of model accuracy.
-To learn more about how specific images are classified, we can view images from the test set along with their respective nine model predictions.
-\autoref{fig:single-exemplars} shows model predictions for images that contain only one type of shape, and \autoref{fig:multi-exemplars} shows similar predictions for images containing multiple shapes.
-Most predictions for the single-label images are straightforward, but there is much less consensus of model opinion in the multi-label images.
-}
+\mt{The ROC curves provide only a general summary of model accuracy; to learn more about how specific images are classified, we can view images from the test set along with their respective nine model predictions.} \autoref{fig:single-exemplars} shows model predictions for images that contain only one type of shape, and \autoref{fig:multi-exemplars} shows similar predictions for images containing multiple shapes.
+Most predictions for the single-label images are straightforward; all of the archetypical shapes are corectly identified by the model, with only small confusions like those between quad/polygon and circle/text.
+\fix{A later section will discuss possible causes for this confusion.}
 
@@
 
+The multi-label images also perform quite well.
+\fix{Find a few more examples here and discuss them.
+Maybe a few similar ones to show consistency matters more than explicit label.
+Make the point that this task is messy and shape definitions are hard, but the model is doing pretty well given the circumstances.}
+
 <<multi-exemplars, fig.cap = "Examples containing multiple shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
@@
 
-\mt{These figures provide representative examples of classifications across multiple categories; it is also useful to examine many cases within a given shape. \autoref{fig:chevrons-consistency}}
+\fix{These figures provide representative examples of classifications across multiple categories; it is also useful to examine many cases within a given shape.} \mt{\autoref{fig:chevrons-consistency} shows a number of images containing chevrons.
+The top four rows show regularly repeating chevron patterns and seems to indicate that the model optimally identifies chevrons of a certain width; wide chevrons cross into the realm of bowties, whereas narrow chevrons become indistinguishable from lines.
+The next six rows show individual and closed chevrons.
+There is a significant amount of confusion between these wider chevrons and quadrilaterals.
+Finally, the bottom four rows show less-than-standard chevrons.
+The rounded chevron has a much lower prediction than those with sharper angles, but still exceeds the EER for chevron; the white chevrons are confounded with overlaid circles, which introduces a complexity that the model attributes to text.}
 
@@
 @
+
+\fix{Describe \autoref{fig:circle-text-consistency}...... The final two don't contain a true circle. IDK why the TOMS logo totally misses circle.}
 
@@
 
-\fix{Now that we've seen specific examples, it is useful to summarize the systematic errors by class.} A confusion matrix is a helpful diagnostic tool to understand which classes are commonly confused for other classes.} A confusion matrix is a \mt{summary of the relationships between} the labels assigned to an image and the labels predicted by the model.
-
-In the confusion matrix presented in \autoref{fig:confusion_matrix}, the values along the diagonal represent the proportion of true positives captured within each category.
+mt{\paragraph{Accuracy by class}}
+\fix{Many of the trends in the examples in figures \ref{fig:single-exemplars}-\ref{fig:circle-text-consistency} are not unique.
+A confusion matrix is a type of visualization that summarizes the relationships between the labels assigned to an image and the labels predicted by the model, and is a helpful diagnostic tool to understand which classes are systematically misclassified.} \mt{The confusion matrix for \fix{our model} is presented in \autoref{fig:confusion_matrix}}.
+The values along the diagonal represent the proportion of true positives captured within each category.
 The off-diagonal values represent the proportion of false positives for each shape pairing; however, since a single image may truly contain multiple shapes, the values have been adjusted to remove the effect of any true positives from the calculation of false positive proportions.
-\svp{As a result of this adjustment, the probabilities off-diagonal are the proportion ...
-need to explain the conditional probability we're showing here} %\svp{I can't remember exactly how we're doing that, but we may want to look more closely at that - if we set everything to NA instead of 0, the rate calculations would be higher than expected}
+%\svp{As a result of this adjustment, the probabilities off-diagonal are the proportion ... need to explain the conditional probability we're showing here}
+\fix{As a result of this adjustment, the off-diagonal values are the proportion of images within a true class that are falsely labeled as containing another class that has not been identified in the image.}%\svp{I can't remember exactly how we're doing that, but we may want to look more closely at that - if we set everything to NA instead of 0, the rate calculations would be higher than expected}
 For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any image that truly contains both triangles and quadrilaterals is removed before calculating the proportion of false quadrilateral labels.
@@
 
-\autoref{fig:confusion_matrix} shows systematic errors in class predictions, but does not necessarily suggest \emph{why} the confusion is taking place.
+%\autoref{fig:confusion_matrix} shows systematic errors in class predictions, but does not necessarily suggest \emph{why} the confusion is taking place.
 To assess the reasons for model failure, we look at the images to \fix{see what the model is seeing.}
 
-\svp{I would move heatmaps up and out of the diagnostic section and transition into that, possibly}
+%\svp{I would move heatmaps up and out of the diagnostic section and transition into that, possibly}
 
-\svp{Three main model exploration chunks: consistency, heatmaps, brand (motivated by the fact that some classes perform so badly because the exemplars are so variable) - need to figure out the best sequence to transition}
+%\svp{Three main model exploration chunks: consistency, heatmaps, brand (motivated by the fact that some classes perform so badly because the exemplars are so variable) - need to figure out the best sequence to transition}
 
@@
 
-A heatmap \fix{please add the technical name for heatmaps - look in the distill.io papers?} is a visual diagnostic tool that highlights the portions of an image that are most significant to classification for a given class.
+Belonging to a category of techniques called class activation map (CAM) visualizations, a heatmap is a visual diagnostic tool that highlights the portions of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}.
 Using these graphics, we can identify which locations in an image contribute most to the model's predictions.

### 2019-06-12 paper/paper.Rnw
Add brand ideas

@@
 Although the filters of VGG16 have been trained and optimized on the ImageNet Data, the large number of filters span a wide variety of shapes and features} \fix{which are still useful/relevant to classifying outsoles.
-In a sense, this is much like teaching humans to classify objects they have never seen before: an object may be unfamiliar, but our eyes can still detect its color, shape, and texture because we have learned these ideas from other sources.}
+In a sense, this is much like teaching humans to classify objects they have never seen before: an object may be unfamiliar, but our eyes can still detect its color, shape, and texture because we have learned these ideas from other sources.} \fix{Say somewhere: Max pooling layers aggregate the features in one block before passing them to the next block, which helps reduce the computational load of training the model and helps to generalize the relationship between features.}
 
@@
 
-Although transfer learning substantially reduces the amount of data required to train a CNN by several orders of magnitude, labeled images are still \fix{difficult or time-consuming} to generate relative to the amount of data necessary to train even a small neural network adequately.
+Transfer learning substantially reduces the amount of data required to train a CNN by several orders of magnitude, but labeled images are still difficult to generate on a large scale.
 Thus, to make the most efficient use of the existing labeled data, we enlarge the training data using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}.
@@
 Overfitting is when a model learns the training data so well that its understanding of the categories becomes specific to the training cases, which, in turn, leads to poor prediction of new images that were not in the training set.
-In training, overfitting is indicated when the validation loss begins to increase after reaching a global minimum.
+During the training process, overfitting is indicated when the validation loss begins to increase after reaching a global minimum.
 Alternately, underfitting occurs if the validation accuracy is still increasing when model optimization is terminated (for instance, between epochs 1 and 10 in \autoref{fig:training-accuracy}), because model performance is still improving with continued training.
@@
 
-%\autoref{fig:confusion_matrix} shows systematic errors in class predictions, but does not necessarily suggest \emph{why} the confusion is taking place.
-To assess the reasons for model failure, we look at the images to \fix{see what the model is seeing.}
-
 %\svp{I would move heatmaps up and out of the diagnostic section and transition into that, possibly}
@@
 \paragraph{Brand Investigations}
+
+One possible explanation for the poor performance of some categories is the amount of variability within the shapes.
+\fix{For example, as seen in fig, a large proportion of bowties come from Birks; thus, the model learns bowties from Birks well, which leads to more accurate predictions of new bowtie images that come from Birks.
+There is no similar dominance in categories like quads.
+Instead, quads are highly variable  and come from a wide range of brands, so it is harder for the model to learn which set of features designate a quadrilateral.
+Since a large number of features are then associated with quads, the model gets more easily confused.
+When it sees some of those same features in other classes, it tries to guess quad.
+This is further explored in the next section.}
+
+\mt{That was a really quick and dirty version of what I'm trying to say... Am I missing any big points, before I try to expand it and clean up the technical part of the writing?}
 
@@
 
+Heatmaps
+  \begin{itemize}
+  \item Verify that the model is using appropriate features when predictions are correct
+  \item Show examples for confident misclassifications (like Seychelles), but make the case that the mistakes are understandable
+  \end{itemize}
+
 Belonging to a category of techniques called class activation map (CAM) visualizations, a heatmap is a visual diagnostic tool that highlights the portions of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}.

### 2019-06-12 paper/paper.Rnw
Add contrast to results outline

@@
   \end{itemize}
-\item Two main reasons for error: some classes are more variable, and some classes share physical features (I.e., the model has to learn a lot of features to encapsulate a class with lots of variety, but then other classes share those features)
+\item Three main reasons for error: image contrast, some classes are more variable, and some classes share physical features (I.e., the model has to learn a lot of features to encapsulate a class with lots of variety, but then other classes share those features)
+\item Contrast
 \item Brands
@@
 %\svp{As a result of this adjustment, the probabilities off-diagonal are the proportion ... need to explain the conditional probability we're showing here}
-\fix{As a result of this adjustment, the off-diagonal values are the proportion of images within a true class that are falsely labeled as containing another class that has not been identified in the image.}%\svp{I can't remember exactly how we're doing that, but we may want to look more closely at that - if we set everything to NA instead of 0, the rate calculations would be higher than expected}
+\fix{As a result of this adjustment, the off-diagonal values are the proportion of images within a true class that are falsely labeled as containing another class that has not been identified in the image.
+}%\svp{I can't remember exactly how we're doing that, but we may want to look more closely at that - if we set everything to NA instead of 0, the rate calculations would be higher than expected}
 For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any image that truly contains both triangles and quadrilaterals is removed before calculating the proportion of false quadrilateral labels.
@@
 One possible explanation for the poor performance of some categories is the amount of variability within the shapes.
-\fix{For example, as seen in fig, a large proportion of bowties come from Birks; thus, the model learns bowties from Birks well, which leads to more accurate predictions of new bowtie images that come from Birks.
+\fix{For example, as seen in \autoref{fig:brands-within-shapes}, a large proportion of bowties come from Birks; thus, the model learns bowties from Birks well, which leads to more accurate predictions of new bowtie images that come from Birks.
 There is no similar dominance in categories like quads.
@@
 
-Heatmaps
-  \begin{itemize}
-  \item Verify that the model is using appropriate features when predictions are correct
-  \item Show examples for confident misclassifications (like Seychelles), but make the case that the mistakes are understandable
-  \end{itemize}
-
 Belonging to a category of techniques called class activation map (CAM) visualizations, a heatmap is a visual diagnostic tool that highlights the portions of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}.

### 2019-06-13 paper/code/Generate_Model_Images.R
Add EER option to pred_prob_plot()

@@ -92,7 +92,13 @@ predict_new <- function(img_path, model, classes = default_classes) {
   })
 }
 
-pred_prob_plot <- function(img_path, model, classes = default_classes, sort = T) {
+pred_prob_plot <- function(img_path, model, classes = default_classes,
+                           sort = T, eer = NULL) {
+  if(!is.null(eer)) {
+    assertthat::assert_that(is.numeric(eer))
+    assertthat::assert_that(length(eer) == length(classes))
+  }
+
   img_preds <- predict_new(img_path, model)
 
   img_preds_mat <- purrr::map_dfc(img_preds, "pred") %>%
@@ -114,25 +120,58 @@ pred_prob_plot <- function(img_path, model, classes = default_classes, sort = T)
       mutate(idx = row_number())
   }
 
+  if(is.null(eer)) {
+    eer <- rep(1, length(classes))
+  }
+
+  above_eer <- apply(as.matrix(img_preds_df)[,classes], 1, function(x){
+      as.numeric(as.numeric(x) > eer)
+    }) %>% t() %>% set_colnames(classes) %>%
+      as.data.frame() %>%
+      cbind(., idx = img_preds_df$idx) %>%
+      gather(key = "class", value = "yes", -idx)
+
   img_preds_df_long <- img_preds_df %>%
     select(idx, 1:length(classes)) %>%
     gather(key = "class", value = value, -idx)
 
+  img_preds_df_long <- merge(img_preds_df_long, above_eer,
+                             by = c("class", "idx"))
+  img_preds_df_long$fill <- img_preds_df_long$value + img_preds_df_long$yes
+
   features_xraster <- purrr::map2(img_preds_df$path, img_preds_df$idx,
                                   ~annotation_custom(grid::rasterGrob(readJPEG(.x), interpolate = T),
-                                                     ymin = .y - .5, ymax = .y + .5,
-                                                     xmin = -0.5, xmax = 0.5))
+                                                     ymin = .y - .5,
+                                                     ymax = .y + .5,
+                                                     xmin = -0.5,
+                                                     xmax = 0.5))
+
+  if(sum(eer) == length(classes)) {
+    cols <- c("0" = "black", "1" = "black")
+    w <- h <- 1
+  } else {
[…]

### 2019-06-13 paper/paper.Rnw
Add updated heatmap process and code

@@
 modeldir <- file.path(wd, "model")
+imgdir <- file.path(wd, "images")
 
@@
 
+<<eers, include = F>>=
+eers <- purrr::map_dbl(aucs$data$eer, function(x){x$thresholds}) %>%
+  setNames(classes)
+@
+
 <<single-exemplars, fig.cap = "Examples containing a single shape", fig.pos = "p!", out.width = "100%", cache = T>>=
-archetype <- list.files(file.path(wd, "images", "consistency", "single-exemplar"), full.names = T)
-pred_prob_plot(archetype, loaded_model, sort = F)
+single_exemplars <- list.files(file.path(wd, "images", "consistency",
+                                         "single-exemplar"), full.names = T)
+pred_prob_plot(single_exemplars, loaded_model, sort = F, eer = eers)
 @
@@
 <<multi-exemplars, fig.cap = "Examples containing multiple shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
-exemplar <- list.files(file.path(wd, "images", "consistency", "multi-exemplar"), full.names = T)
-pred_prob_plot(exemplar, loaded_model)
+multi_exemplars <- list.files(file.path(wd, "images", "consistency", "multi-exemplar"), full.names = T)
+pred_prob_plot(multi_exemplars, loaded_model, eer = eers)
 @
@@
 chevron <- list.files(file.path(wd, "images", "consistency", "chevron"), full.names = T)
-pred_prob_plot(rev(chevron), loaded_model, sort = F)
+pred_prob_plot(rev(chevron), loaded_model, sort = F, eer = eers)
 @
@@
 circle_text <- list.files(file.path(wd, "images", "consistency", "circle-text"), full.names = T)
-pred_prob_plot(rev(circle_text), loaded_model, sort = F)
+pred_prob_plot(rev(circle_text), loaded_model, sort = F, eer = eers)
 @
@@
 %\svp{Three main model exploration chunks: consistency, heatmaps, brand (motivated by the fact that some classes perform so badly because the exemplars are so variable) - need to figure out the best sequence to transition}
+
+\paragraph{Contrast}
+
+<<low-contrast-consistency, fig.cap = "Images with low contrast, before and after correction", fig.pos = "p!", out.width = "100%", cache = T>>=
+lowcontrast <- list.files(file.path(wd, "images", "consistency", "low-contrast"), full.names = T)
+pred_prob_plot(lowcontrast, loaded_model, sort = F)
+@
 
@@
 In \fix{the figures in this subsection,} heatmaps are shown for a selection of three classes, which are either contained in the image or have high output probabilities, alongside the original image.
+\fix{Many of these heatmaps illustrate that multiple classes share certain features, which understandably confuses the model.}
 
-\fix{I picked the heatmaps I thought illustrated the point best, but didn't modify them heavily yet.
+%\fix{I picked the heatmaps I thought illustrated the point best, but didn't modify them heavily yet.
 How many do we want to include, and do we want to highlight model challenges or just focus on the point that CoNNOR sees what we see?} \svp{I think it's worth highlighting the fact that we'll never get a model that's perfect because that curly s is pretty much a circle, and there will always be weird shapes that partially activate different categories.}
 
-\svp{We may want to explore using the border of the square to show whether the model actually assigns that label - like the bullet plots}
+\fix{Susan, please generate these heatmaps and move them to the "static" folder.
+My computer still has issues with the transparent layer of the heatmap, so it's probably easiest to do what we did for my CC and have you generate the images so they turn out correctly.
+Once I can see them all, I'll decide which three heatmaps best illustrate the point that shape categories have similar features that (understandably) confuse the model and I'll write about those specific heatmaps.}
+
+<<setup-heatmap>>=
+imgs <- list.files(file.path(wd, "images", "heatmaps", "raw"), full.names = T)
+@
+
+<<generate-heatmaps, echo = F>>=
+# My computer has troubles with the transparency of the overlay...
+# This code is functional, but highest quality will come from generating these
+# images on Bigfoot and loading them statically.
+
+h_imgdir <- file.path(imgdir, "heatmaps", "processed")
+if(!dir.exists(h_imgdir)) {dir.create(h_imgdir)}
+
+calc_heatmap(imgs[which(grepl("chevron-circle", imgs))], loaded_model) %>%
+  prune_heatmap(c("chevron", "quad", "circle")) %>%
+  create_composite(save_file = T, outdir = h_imgdir)
+
+calc_heatmap(imgs[which(grepl("corks", imgs))], loaded_model) %>%
+  prune_heatmap(c("star", "text", "circle")) %>%
+  create_composite(save_file = T, outdir = h_imgdir)
+
+calc_heatmap(imgs[which(grepl("text", imgs))], loaded_model) %>%
+  prune_heatmap(c("text", "circle", "line")) %>%
+  create_composite(save_file = T, outdir = h_imgdir)
+
+calc_heatmap(imgs[which(grepl("triangle", imgs))], loaded_model) %>%
+  prune_heatmap(c("triangle", "quad", "chevron")) %>%
+  create_composite(save_file = T, outdir = h_imgdir)
+
+calc_heatmap(imgs[which(grepl("star", imgs))], loaded_model) %>%
+  prune_heatmap(c("star", "bowtie", "polygon")) %>%
+  create_composite(save_file = T, outdir = h_imgdir)
+
+calc_heatmap(imgs[which(grepl("polygon", imgs))], loaded_model) %>%
+  prune_heatmap(c("star", "chevron", "polygon")) %>%
+  create_composite(save_file = T, outdir = h_imgdir)
+
+heatmaps <- list.files(imgdir, pattern = "^heatmap")
+@
+
 <<chevron-circle-heatmap, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing chevrons and circles.", fig.cap = "Chevrons are identified by sharp corners and parallel lines, but the thick line between the two shape sections is falsely identified as a quadrilateral.", fig.pos = "!h">>=
-heatmaps <- list.files(file.path(wd, "images", "heatmaps"),
+heatmaps <- list.files(file.path(wd, "images", "heatmaps", "static"),
                        pattern = "^heatmap", full.names = T)
@@
 \svp{Rework this to address 3 main conclusions - 1: NNs work for automatic feature ID, 2: Different levels of feature variability within classes, and 3: correcting for contrast is important.}
+\mt{I'll do this once Results is stable, so I can make it match the rest of the story.}
 The goal of this research was to develop a method to automatically identify geometric class characteristics of shoe outsoles.

### 2019-06-14 paper/paper.Rnw
A few changes/comments, updated images

@@
 
-Many convolutional neural networks have been trained on a \mt{sub}set of images \mt{from the database} ImageNet; these images are hierarchically labeled, so the same image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity \citep{dengImageNetLargeScaleHierarchical}.
+Many convolutional neural networks have been trained on a subset of images \svp{from ImageNet}; these images are hierarchically labeled, so the same image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity \citep{dengImageNetLargeScaleHierarchical}.
 CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously.
@@
 \paragraph{VGG16}
-Developed by Oxford's Visual Graphics Group, VGG16 is a \mt{pre-trained} CNN with 16 ``functional" (i.e., convolutional and densely connected) layers and 5 ``structural" max pooling layers \citep{krizhevskyImageNetClassificationDeep2012}.
+Developed by Oxford's Visual Graphics Group, VGG16 is a pre-trained CNN with 16 ``functional" (i.e., convolutional and densely connected) layers and 5 ``structural" max pooling layers \citep{krizhevskyImageNetClassificationDeep2012}.
 In contrast to other popular networks, like ResNet, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy.
@@
 
-\mt{The model base of VGG16 is made up of five convolutional blocks alternated with max pooling layers.
-Each convolutional block is made up of two or three convolutional layers, and each layer is made up of hundreds of filters, such that each of the five convolutional blocks has between 128 and 1,536 filters.
-Strictly speaking, a filter is a set of numerical weights that, when applied to an image, quantifies the presence of the specific shapes and/or colors that the filter is trained to detect.
-\fix{Necessary?
-(Or just skip...): Visualizing a filter then refers to visualizing the features to which the filter provides its maximum response.} \autoref{fig:vgg16layers} shows a selection of filters from VGG16, with each row corresponding to one of the five convolutional blocks.
-Filters in earlier convolutional blocks detect simple features, such as colors and blobs, while later filters detect more complex combinations of features.
-When considered together, the series of filters perform the process of feature detection, such that a single image is made up of some combination of the thousands of features that are detected by filters.
-Although the filters of VGG16 have been trained and optimized on the ImageNet Data, the large number of filters span a wide variety of shapes and features} \fix{which are still useful/relevant to classifying outsoles.
-In a sense, this is much like teaching humans to classify objects they have never seen before: an object may be unfamiliar, but our eyes can still detect its color, shape, and texture because we have learned these ideas from other sources.} \fix{Say somewhere: Max pooling layers aggregate the features in one block before passing them to the next block, which helps reduce the computational load of training the model and helps to generalize the relationship between features.}
+\svp{The base of VGG16 is made up of five blocks, which consist of two or three convolutional layers and one max pooling layer which aggregates the signal from the convolutional layers.
+Each convolutional layer} is made up of hundreds of filters, such that each of the five convolutional blocks has between 128 and 1,536 filters.
+A filter is a set of numerical weights that, when applied to an image, quantifies the presence of the specific shapes and/or colors that the filter is trained to detect.
+\autoref{fig:vgg16layers} shows a selection of filters from VGG16, with each row corresponding to one of the five convolutional blocks.
+Filters in earlier convolutional blocks detect simple features, such as colors, \svp{lines, corners,} and blobs, while later filters detect more complex combinations of features.
+\svp{This aggregation of features through successive convolutional layers mirrors the process of complex feature detection in human vision.} \svp{Any complex feature can be disassembled into a set of simple features; similarly, the successive sets of layers detect increasingly detailed features.} %When considered together, the series of filters perform the process of feature detection, such that a single image is made up of some combination of the thousands of features that are detected by filters.
+Although the filters of VGG16 have been trained and optimized on the ImageNet Data, the large number of filters span a wide variety of shapes and features, \svp{many of which are useful when examining tread patterns.
+Just as humans can detect the features of an unfamiliar object (consituent parts, edges, textures, and corners), a CNN trained on a wide set of objects can be used to classify unrelated images, such as shoe soles.} %\fix{which are still useful/relevant to classifying outsoles.
+In a sense, this is much like teaching humans to classify objects they have never seen before: an object may be unfamiliar, but our eyes can still detect its color, shape, and texture because we have learned these ideas from other sources.} \fix{Say somewhere: Max pooling layers aggregate the features in one block before passing them to the next block, which helps reduce the computational load of training the model and helps to generalize the relationship between features.} % I've tried to do this concisely - if you don't like it, please feel free to revert the changes.
 
@@
 
-Receiver Operating Characteristic (ROC) curves are a type diagnostic plot which compare the false positive rate to the true positive rate for a classification method, and the Area Under the Curve (AUC) quantifies the shape of the ROC curve.
+Receiver Operating Characteristic (ROC) curves are a type diagnostic plot which compare the false positive rate to the true positive rate for a classification method, and the Area Under the Curve (AUC) \svp{is a summary measure which} quantifies the \svp{overall accuracy of the model}.
+%It doesn't really summarize the shape - you can have very different shapes of curves with the same AUC value.
 Perfect prediction would be indicated by a right angle along the upper left corner of the plot, with a corresponding AUC of 1, and a diagonal line with an AUC of 0.5 would indicate that the classification method performs no better than random chance.
@@
 The full model has an AUC of 0.88, and the AUC for individual classes ranges from 0.81 (for lines) to 0.91 (for bowtie and text).
-While the class performances do vary slightly, each ROC curve is the same general shape and performs significantly better than random chance.
+\fix{pull these numbers out of the model using \\Sexpr} While the class performances do vary slightly, each ROC curve is the same general shape and performs significantly better than random chance.
 The points in \autoref{fig:class-roc} represent the equal error rate (EER), which is the threshold where there is equal probability of a false positive or false negative error.
@@
 \paragraph{Model Predictions}
-\mt{The ROC curves provide only a general summary of model accuracy; to learn more about how specific images are classified, we can view images from the test set along with their respective nine model predictions.} \autoref{fig:single-exemplars} shows model predictions for images that contain only one type of shape, and \autoref{fig:multi-exemplars} shows similar predictions for images containing multiple shapes.
+The ROC curves provide only a general summary of model accuracy; to learn more about how specific images are classified, we can examine images from the test set along with their respective nine model predictions.
+\autoref{fig:single-exemplars} shows model predictions for images that contain only one type of shape, and \autoref{fig:multi-exemplars} shows similar predictions for images containing multiple shapes.
 Most predictions for the single-label images are straightforward; all of the archetypical shapes are corectly identified by the model, with only small confusions like those between quad/polygon and circle/text.
@@
 
-\fix{These figures provide representative examples of classifications across multiple categories; it is also useful to examine many cases within a given shape.} \mt{\autoref{fig:chevrons-consistency} shows a number of images containing chevrons.
+\fix{These figures provide representative examples of classifications across multiple categories; it is also useful to examine many cases within a given shape.} \autoref{fig:chevrons-consistency} shows a number of images containing chevrons.
 The top four rows show regularly repeating chevron patterns and seems to indicate that the model optimally identifies chevrons of a certain width; wide chevrons cross into the realm of bowties, whereas narrow chevrons become indistinguishable from lines.
@@
 Finally, the bottom four rows show less-than-standard chevrons.
-The rounded chevron has a much lower prediction than those with sharper angles, but still exceeds the EER for chevron; the white chevrons are confounded with overlaid circles, which introduces a complexity that the model attributes to text.}
+The rounded chevron has a much lower prediction than those with sharper angles, but still exceeds the EER for chevron; the white chevrons are confounded with overlaid circles, which introduces a complexity that the model attributes to text.
 

### 2019-06-14 paper/paper.Rnw
Light editing, try to structure results

--- 2019-06-13 paper/paper.Rnw
@@
   \end{itemize}
+  \item (IDK where to put this, but I think it's a really important point that consistency matters more than accuracy.
+As long as a shape is predicted consistently, that provides information about outsole similarity, even if it's not 100\% correct.
+But, if phrased wrong, this undermines the claim that our model can replicate the feature detection examiners are doing...)
 \end{itemize}
@@
 Thus, to determine the accuracy of the model predictions, we must select a threshold for each label to discretize the model predictions.
-If the threshold is too high, \fix{moderate} probabilities will be treated as a ``no", which increases the number of true negative and false negative predictions.
-Conversely, if the threshold is too low, \fix{moderate} probabilites will be considered as ``yes", which increases the true and false positive predictions in the data.
+If the threshold is too low, moderate probabilites will \mt{always meet the threshold}, which increases the true positive and false positive predictions in the test data.
+Conversely, if the threshold is too high, moderate probabilities will \mt{fail to meet the threshold,} which increases the number of true and false negative predictions.
 An appropriate threshold must be chosen to produce a high true positive rate while also ideally keeping the false positive rate low.
+\mt{Is it worth introducing the acronyms TPR and FPR?}
 
@@
 ggplot() +
-  geom_line(aes(x = fpr, y = tpr), data = unnest(whole_model_roc_df, roc_plot), size = 1.25) +
-  geom_label(aes(x = 1, y = .07, label = sprintf("AUC: %0.2f", auc)), hjust = 1, vjust = -0.2, data = whole_model_roc_df) +
-  geom_point(aes(x = fpr, y = tpr, color = "Equal Error Rate"), data = unnest(whole_model_roc_df, eer), size = 2) +
+  geom_line(aes(x = fpr, y = tpr), data = unnest(whole_model_roc_df, roc_plot),
+            size = 1.25) +
+  geom_label(aes(x = 1, y = .07, label = sprintf("AUC: %0.2f", auc)), hjust = 1,
+             vjust = -0.2, data = whole_model_roc_df) +
+  geom_point(aes(x = fpr, y = tpr, color = "Equal Error Rate"),
+             data = unnest(whole_model_roc_df, eer), size = 2) +
   scale_color_manual("", values = "black") +
-  scale_x_continuous("False Positive Rate", breaks = c(0, .25, .5, .75, 1), labels = c("0.0", "", "0.5", "", "1.0")) +
-  scale_y_continuous("True Positive Rate", breaks = c(0, .25, .5, .75, 1), labels = c("0.0", "", "0.5", "", "1.0")) +
+  scale_x_continuous("False Positive Rate", breaks = c(0, .25, .5, .75, 1),
+                     labels = c("0.0", "", "0.5", "", "1.0")) +
+  scale_y_continuous("True Positive Rate", breaks = c(0, .25, .5, .75, 1),
+                     labels = c("0.0", "", "0.5", "", "1.0")) +
   ggtitle("CoNNOR Test Set Performance (All Classes)") +
   coord_fixed() + mytheme +
-  theme(legend.position = c(1, 0), legend.justification = c(1.01, -0.01), legend.title = element_blank(), legend.background = element_rect(fill = "white"))
+  theme(legend.position = c(1, 0), legend.justification = c(1.01, -0.01),
+        legend.title = element_blank(),
+        legend.background = element_rect(fill = "white"))
 
@@
 \paragraph{Model Predictions}
-\mt{The ROC curves provide only a general summary of model accuracy; to learn more about how specific images are classified, we can view images from the test set along with their respective nine model predictions.} \autoref{fig:single-exemplars} shows model predictions for images that contain only one type of shape, and \autoref{fig:multi-exemplars} shows similar predictions for images containing multiple shapes.
-Most predictions for the single-label images are straightforward; all of the archetypical shapes are corectly identified by the model, with only small confusions like those between quad/polygon and circle/text.
-\fix{A later section will discuss possible causes for this confusion.}
+\mt{The ROC curves provide only a general summary of model accuracy; to learn more about how specific images are classified, we can view images from the test set along with their respective nine model predictions.} \autoref{fig:single-exemplars} shows model predictions for images that contain only one type of shape, and \autoref{fig:multi-exemplars} shows predictions for images containing multiple shapes.
+Most predictions for the single-class images are \fix{promising}; there are some confusions between circle/text and polygon/quad, but all of the true shapes are corectly identified by the model.
+\fix{The multi-label images also perform quite well; again, true labels are almost all present, with a small number of false positive labels.
+[Expand on specific cases more, once determining how many examples to keep and how much to highlight errors.] A later section will discuss possible causes for these class confusions.}
 
@@
 
-The multi-label images also perform quite well.
-\fix{Find a few more examples here and discuss them.
-Maybe a few similar ones to show consistency matters more than explicit label.
-Make the point that this task is messy and shape definitions are hard, but the model is doing pretty well given the circumstances.}
-
 <<multi-exemplars, fig.cap = "Examples containing multiple shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
-multi_exemplars <- list.files(file.path(wd, "images", "consistency", "multi-exemplar"), full.names = T)
-pred_prob_plot(multi_exemplars, loaded_model, eer = eers)
+multi_exemplars <- list.files(file.path(wd, "images", "consistency",
+                                        "multi-exemplar"), full.names = T)
+pred_prob_plot(rev(multi_exemplars), loaded_model, sort = F, eer = eers)
 @
 
-\fix{These figures provide representative examples of classifications across multiple categories; it is also useful to examine many cases within a given shape.} \mt{\autoref{fig:chevrons-consistency} shows a number of images containing chevrons.
+\fix{Figures \ref{fig:single-exemplars} and \ref{multi-exemplars} provide representative examples of classifications across multiple categories; it is also useful to examine many cases within a given shape.} \mt{\autoref{fig:chevrons-consistency} shows a number of images containing chevrons.
 The top four rows show regularly repeating chevron patterns and seems to indicate that the model optimally identifies chevrons of a certain width; wide chevrons cross into the realm of bowties, whereas narrow chevrons become indistinguishable from lines.
@@
 
+\mt{If We end up making the claim that consistency matters more than accuracy, we can use \autoref{fig:ugg-consistency} to show that consistently predicting line isn't strictly right but does establish similarity between the images.
+If we don't emphasize the consistency point as much, the ugg plot is overkill.}
+
 <<circle-text-consistency, fig.cap = "Text with circles or circle-like shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
@@
 
+<<ugg-consistency, fig.cap = "Text with circles or circle-like shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
+ugg <- list.files(file.path(wd, "images", "consistency", "ugg-star-logo"), full.names = T)
+pred_prob_plot(rev(ugg), loaded_model, sort = F, eer = eers)
+@
+
 %\svp{Pros and cons of confusion matrix here? Maybe we just don't show the actual proportions on the graph? I'd like to discuss this more in person.}
 
-mt{\paragraph{Accuracy by class}}
-\fix{Many of the trends in the examples in figures \ref{fig:single-exemplars}-\ref{fig:circle-text-consistency} are not unique.
-A confusion matrix is a type of visualization that summarizes the relationships between the labels assigned to an image and the labels predicted by the model, and is a helpful diagnostic tool to understand which classes are systematically misclassified.} \mt{The confusion matrix for \fix{our model} is presented in \autoref{fig:confusion_matrix}}.
+\mt{\paragraph{Accuracy by class}}
+\fix{Figures \ref{fig:single-exemplars}-\ref{fig:circle-text-consistency} show predictions for a number of individual images.
+Many of the misclassifications in these examples are not just seen in individual images, but rather exist systematically in a class.} A confusion matrix is a type of visualization that summarizes the relationships between the labels assigned to an image and the labels predicted by the model across all test data, and is a helpful diagnostic tool to understand which classes are systematically misclassified.} \mt{The confusion matrix for \fix{our model} is presented in \autoref{fig:confusion_matrix}}.
 The values along the diagonal represent the proportion of true positives captured within each category.
@@
 For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any image that truly contains both triangles and quadrilaterals is removed before calculating the proportion of false quadrilateral labels.
+\mt{Note that this matrix is not symmetric---the proportion of images containing text that are falsely predicted to contain circles is not necessarily equal to the proportion of images containing circles that are falsely predicted to contain text.}
 
@@
 Another interesting artifact in \autoref{fig:confusion_matrix} is that circles are often misclassified as text, and text is often mislabeled as containing circles.
-Not all of the images used in model training are accurately labeled: in some cases, the labels have not \mt{yet} been updated after labeling guidelines have changed.
+
+\fix{(Transition) We have identified three possible causes for systematic misclassification: incorrect labels, low contrast, and variability within classes/overlap between classes.}
+
+%One reason for these misclassifications is that not all of the images used in model training are accurately labeled: in some cases, the labels have not \mt{yet} been updated after labeling guidelines have changed.
 The confusion between text and circles is a natural consequence of the overlap between geometric shapes and components of written text; it is not indicative of a lack of fit as much as a weakness in the classification system used to define the model.
-
-%\svp{I would move heatmaps up and out of the diagnostic section and transition into that, possibly}
-
-%\svp{Three main model exploration chunks: consistency, heatmaps, brand (motivated by the fact that some classes perform so badly because the exemplars are so variable) - need to figure out the best sequence to transition}
 
@@
 
+\fix{Low contrast images are difficult to predict because the features are not prominent.
+The model has 'learned' that predicting diffuse probabilities across the classes minimizes loss, so it tends to assign diffuse probabilities to low contrast images.
+Contrast correction seems to improve predictions, but the diffuse probability strategy persists, so we will incorporate contrast correction into model training.}
+
 <<low-contrast-consistency, fig.cap = "Images with low contrast, before and after correction", fig.pos = "p!", out.width = "100%", cache = T>>=
 lowcontrast <- list.files(file.path(wd, "images", "consistency", "low-contrast"), full.names = T)
-pred_prob_plot(lowcontrast, loaded_model, sort = F)
+pred_prob_plot(lowcontrast, loaded_model, sort = F, eer = eers)
 @
@@
 My computer still has issues with the transparent layer of the heatmap, so it's probably easiest to do what we did for my CC and have you generate the images so they turn out correctly.
-Once I can see them all, I'll decide which three heatmaps best illustrate the point that shape categories have similar features that (understandably) confuse the model and I'll write about those specific heatmaps.}
+Once I can see them all, I'll decide which three (or so) heatmaps best illustrate the point that shape categories have similar features that (understandably) confuse the model and I'll write about those specific heatmaps.}
 
@@
 
-<<generate-heatmaps, echo = F>>=
+<<generate-heatmaps, echo = F, eval = F>>=
 # My computer has troubles with the transparency of the overlay...

### 2019-06-17 paper/paper.Rnw

@@
 \paragraph{Model Predictions}
-\mt{The ROC curves provide only a general summary of model accuracy; to learn more about how specific images are classified, we can view images from the test set along with their respective nine model predictions.} \autoref{fig:single-exemplars} shows model predictions for images that contain only one type of shape, and \autoref{fig:multi-exemplars} shows predictions for images containing multiple shapes.
+The ROC curves provide only a general summary of model accuracy; to learn more about how specific images are classified, we can view images from the test set along with their respective nine model predictions.
+\autoref{fig:single-exemplars} shows model predictions for images that contain only one type of shape, and \autoref{fig:multi-exemplars} shows predictions for images containing multiple shapes.
 Most predictions for the single-class images are \fix{promising}; there are some confusions between circle/text and polygon/quad, but all of the true shapes are corectly identified by the model.
@@
 
-\fix{Figures \ref{fig:single-exemplars} and \ref{multi-exemplars} provide representative examples of classifications across multiple categories; it is also useful to examine many cases within a given shape.} \mt{\autoref{fig:chevrons-consistency} shows a number of images containing chevrons.
+\fix{Figures \ref{fig:single-exemplars} and \ref{multi-exemplars} provide representative examples of classifications across multiple categories; it is also useful to examine many cases within a given shape.} \autoref{fig:chevrons-consistency} shows a number of images containing chevrons.
 The top four rows show regularly repeating chevron patterns and seems to indicate that the model optimally identifies chevrons of a certain width; wide chevrons cross into the realm of bowties, whereas narrow chevrons become indistinguishable from lines.

### 2019-06-17 paper/paper.Rnw
Fixing references

@@
 
-\fix{Figures \ref{fig:single-exemplars} and \ref{multi-exemplars} provide representative examples of classifications across multiple categories; it is also useful to examine many cases within a given shape.} \autoref{fig:chevrons-consistency} shows a number of images containing chevrons.
+\fix{Figures \ref{fig:single-exemplars} and \ref{fig:multi-exemplars} provide representative examples of classifications across multiple categories; it is also useful to examine many cases within a given shape.} \autoref{fig:chevrons-consistency} shows a number of images containing chevrons.
 The top four rows show regularly repeating chevron patterns and seems to indicate that the model optimally identifies chevrons of a certain width; wide chevrons cross into the realm of bowties, whereas narrow chevrons become indistinguishable from lines.
@@
 \fix{Figures \ref{fig:single-exemplars}-\ref{fig:circle-text-consistency} show predictions for a number of individual images.
-Many of the misclassifications in these examples are not just seen in individual images, but rather exist systematically in a class.} A confusion matrix is a type of visualization that summarizes the relationships between the labels assigned to an image and the labels predicted by the model across all test data, and is a helpful diagnostic tool to understand which classes are systematically misclassified.} \mt{The confusion matrix for \fix{our model} is presented in \autoref{fig:confusion_matrix}}.
+Many of the misclassifications in these examples are not just seen in individual images, but rather exist systematically in a class.} A confusion matrix is a type of visualization that summarizes the relationships between the labels assigned to an image and the labels predicted by the model across all test data, and is a helpful diagnostic tool to understand which classes are systematically misclassified.
+\mt{The confusion matrix for \fix{our model} is presented in \autoref{fig:confusion_matrix}}.
 The values along the diagonal represent the proportion of true positives captured within each category.
@@
 
-<<brands-within-shapes>>=
+<<brands-within-shapes, fig.cap = "Caption here">>=
 shapes <- lapply(X = classes, function(x){

### 2019-06-17 paper/paper.Rnw
fix sexpr issues

@@
 The full model has an AUC of 0.88, and the AUC for individual classes ranges from 0.81 (for lines) to 0.91 (for bowtie and text).
-\fix{pull these numbers out of the model using \\Sexpr} While the class performances do vary slightly, each ROC curve is the same general shape and performs significantly better than random chance.
+\fix{pull these numbers out of the model using Sexpr} While the class performances do vary slightly, each ROC curve is the same general shape and performs significantly better than random chance.
 The points in \autoref{fig:class-roc} represent the equal error rate (EER), which is the threshold where there is equal probability of a false positive or false negative error.

### 2019-06-17 paper/paper.Rnw
switch to citeauthor - see if that does anything?

@@
 After consulting with practitioners, we developed a set of categories suitable for automatic recognition by convolutional neural networks.
-These modifications were necessary because some of the definitions used in \citet{grossVariabilitySignificanceClass2013} require contextual spatial information which is not preserved during labeling (for example, lugs are required to be on the perimeter of the shoe).
+These modifications were necessary because some of the definitions used in \citeauthor{grossVariabilitySignificanceClass2013} require contextual spatial information which is not preserved during labeling (for example, lugs are required to be on the perimeter of the shoe).
 \autoref{tab:class-char-examples} shows three examples of each class.

### 2019-06-17 paper/paper.Rnw
some rephrasing

@@
 Thousands of outsole images were scraped from online shoe retail sites and annotated using LabelMe, a tool for image annotation in computer vision problems \citep{labelme}.
-After annotation, images are processed by an R script that identifies the minimum bounding rectangle of the region, crops the image to that region, and scales the cropped area to a 256 x 256 pixel image suitable for analysis by the convolutional neural network.
-During this process, aspect ratio is not preserved, though efforts are made to label regions which are relatively square to minimize the effect of this distortion.
+After annotation, the minimum bounding rectangle of the region \svp{is identified} and the image is cropped to that area; subsequently, the cropped image is scaled to 256 x 256 pixels.} During this process, aspect ratio is not preserved, though efforts are made to label regions which are relatively square to minimize the effect of this distortion.
 To date, \Sexpr{unique(ann_df$base_image) %>% length()} shoes have been labeled, yielding \Sexpr{nrow(dfunion)} multi-label images.
@@
 the training data were weighted by proportion during the training process to ensure that the model can identify both rare and common geometric shapes.
-Of the remaining 40\% of data, half were used for validation, to monitor the training process, and the remaining data were for testing the performance of the fitted model.
+Of the remaining 40\% of data, half were used for validation, to monitor the training process, and the remaining data were for testing the performance of the final model.
 
@@
 \subsection{Model Training}
+\svp{During model fitting, the model ``learns" from the training set and is evaluated on the validation set, which consists of previously unseen images (e.g.
+images not used during the training process).
+Learning occurs by minimization of the ``Loss", which is a function of the distance between the model predictions and the image labels.
+As the loss decreases, the model's accuracy (the proportion of correct labels) increases.
+During model training, we expect that the accuracy for the training set will increase beyond the accuracy for the validation set; the validation accuracy will eventually level off.
+Similarly, we expect the loss for the training set and the validation set to decrease initially.} \autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process.
+% In machine learning, a loss function is a function that measures how far model predictions are from the given labels for each image in the validation and test sets; the model ``learns" by updating its predictions to minimize the loss.
+\autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process.
 <<training-accuracy, fig.width = 7.5, fig.height = 5, out.width = ".75\\textwidth", fig.cap = "Training and validation accuracy and loss for each epoch of the fitting process.
@@
 @
-
-In machine learning, a loss function is a function that measures how far model predictions are from the given labels for each image in the validation and test sets; the model ``learns" by updating its predictions to minimize the loss.
-\autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process.
 

### 2019-06-17 paper/paper.Rnw
force a specific model version

@@
 
-newest_model <- get_newest(dir = model_path, pattern = "weights.h5")
+# newest_model <- get_newest(dir = model_path, pattern = "weights.h5")
+newest_model <- list(path = file.path(model_path, "20190514-155453"),
+    base_file = "2019-05-14_19:45:22_vgg16_onehotaug_9class_256-weights.h5",
+    prefix = "vgg16_onehotaug_9class_256", start_date = "2019-05-14_19:45:22",
+    process_dir = "20190514-155453")
 newest_data_file <- file.path(modeldir, "RProcessedImages",

### 2019-06-17 paper/paper.Rnw
phrasing

@@
 Thousands of outsole images were scraped from online shoe retail sites and annotated using LabelMe, a tool for image annotation in computer vision problems \citep{labelme}.
-After annotation, the minimum bounding rectangle of the region \svp{is identified} and the image is cropped to that area; subsequently, the cropped image is scaled to 256 x 256 pixels.} During this process, aspect ratio is not preserved, though efforts are made to label regions which are relatively square to minimize the effect of this distortion.
+After annotation, the minimum bounding rectangle of the region \svp{is identified and the image is cropped to that area; subsequently, the cropped image is scaled to 256 x 256 pixels.} During this process, aspect ratio is not preserved, though efforts are made to label regions which are relatively square to minimize the effect of this distortion.
 To date, \Sexpr{unique(ann_df$base_image) %>% length()} shoes have been labeled, yielding \Sexpr{nrow(dfunion)} multi-label images.

### 2019-06-17 paper/paper.Rnw
Fix brand plot

@@
 
-top_brands <- df_shoes %>%
-  group_by(brand) %>%
-  summarize(n = n()) %>%
-  arrange(desc(n))
+df_shoes_long <- df_shoes %>%
+  tidyr::gather(key = feature, value = value, -str, -brand) %>%
+  ungroup() %>%
+  tidyr::nest(c(str, value)) %>%
+  mutate(n_true = purrr::map_int(data, ~sum(.$value)),
+         n_shoes_brand = purrr::map_int(data, ~length(.$value)))
 
-df_summ <- filter(df_shoes, brand %in% head(top_brands$brand, 8)) %>%
-  group_by(brand) %>%
-  summarize(n = n(),
-            bowtie = sum(bowtie)/n, chevron = sum(chevron)/n,
-            circle = sum(circle)/n, line = sum(line)/n,
-            polygon = sum(polygon)/n, quad = sum(quad)/n,
-            star = sum(star)/n, text = sum(text)/n, triangle = sum(triangle)/n)
+brand_prop_by_feature <- df_shoes_long %>%
+  group_by(feature) %>%
+  mutate(prop = n_true/sum(n_true)) %>%
+  arrange(feature, desc(prop)) %>%
+  mutate(brand_cat = ifelse(row_number() <= 10, brand, "other")) %>%
+  group_by(feature) %>%
+  mutate(rank = pmin(row_number(), 11))
+brand_prop_by_feature_10 <- brand_prop_by_feature %>%
+  group_by(feature, brand_cat, rank) %>%
+  summarize(prop = sum(prop)) %>%
+  ungroup() %>%
+  group_by(feature) %>%
+  arrange(feature, rank) %>%
+  mutate(cprop = cumsum(prop),
+         cprop2 = lag(cprop, 1, default = 0),
+         mid = (cprop + cprop2)/2,
+         rank = ifelse(rank == 11, NA, rank))
 
-df_summ_long <- df_summ %>%
-  gather(key = shape, value = prop, -brand, -n) %>%
-  arrange(brand)
-
-df_summ_long$brand <- as.factor(df_summ_long$brand)
-df_summ_long$shape <- factor(df_summ_long$shape, levels = classes)
-
-# # Line plot
-# ggplot(df_summ_long) + geom_point(aes(x = prop, y = shape, color = brand)) +
-#   geom_path(aes(x = prop, y = shape, group = brand, color = brand)) +
-#   ggtitle("Shape proportions by brand")
-
-# unicode_symbols <- c(bowtie = "\u26FB", chevron = "\uFE3D",
-#                      circle =  "\u23FA", line = "\u2225",
-#                      polygon = "\u2B23", quad = "\u25A0",
-#                      star = "\u2605", text = "a", triangle = "\u25BC",
-#                      other = "\u003F")
-
-ggplot(df_summ_long, aes(x = brand, y = prop, shape = factor(shape),
-                         color = factor(shape))) +
-  geom_point(size = 4) +
-  scale_shape_manual(values = c("\u26FB","\uFE3D","\u23FA","\u2225",
-                                "\u2B23","\u25A0","\u2605","a","\u25BC")) +
-  theme_bw() +
-  theme(legend.position = "none") +
-  coord_flip() +
-  ggtitle("Proportion of labeled shoes containing each shape by brand")
-@
-
-<<brands-within-shapes, fig.cap = "Caption here">>=
-shapes <- lapply(X = classes, function(x){
-  out <- df_shoes[df_shoes[,x, drop = T],]
-  out$shape <- x
-  out
-})
-names(shapes) <- classes
-
-shapes %>% map_dfr(., function(x) {
-  x %>% group_by(brand) %>%
-    summarize(n = n()) %>%
-    arrange(desc(n)) %>%
-    mutate(prop = n / sum(n),
-           shape = x$shape[1:length(n)]) %>%
-    head(10)
-  }) %>%
-  ggplot() +
-  geom_text(aes(x = as.factor(shape),
-                               y = prop, label = brand,
-                               color = as.factor(shape))) +
-  theme(legend.position = "none")
+ggplot(data = brand_prop_by_feature_10) +
+  geom_bar(aes(x = feature, y = prop, fill = factor(rank)), position = "stack", stat = "identity", color = "black") +
+  geom_text(aes(x = feature, y = 1-mid, label = brand_cat), data = filter(brand_prop_by_feature_10, prop > .05), size = 3) +
+  scale_fill_brewer(palette = "Paired", guide = F, na.value = "grey") +
+  scale_y_continuous("Proportion") + scale_x_discrete("") +
+  ggtitle("Proportion of Labeled Features by Brand")
 @

### 2019-06-17 paper/paper.Rnw
Prune examples

--- 2019-06-17 paper/paper.Rnw
@@
 
-# newest_model <- get_newest(dir = model_path, pattern = "weights.h5")
-newest_model <- list(path = file.path(model_path, "20190514-155453"),
-    base_file = "2019-05-14_19:45:22_vgg16_onehotaug_9class_256-weights.h5",
-    prefix = "vgg16_onehotaug_9class_256", start_date = "2019-05-14_19:45:22",
-    process_dir = "20190514-155453")
+newest_model <- get_newest(dir = model_path, pattern = "weights.h5")
 newest_data_file <- file.path(modeldir, "RProcessedImages",
@@
 After consulting with practitioners, we developed a set of categories suitable for automatic recognition by convolutional neural networks.
-These modifications were necessary because some of the definitions used in \citeauthor{grossVariabilitySignificanceClass2013} require contextual spatial information which is not preserved during labeling (for example, lugs are required to be on the perimeter of the shoe).
+These modifications were necessary because some of the definitions used in \citet{grossVariabilitySignificanceClass2013} require contextual spatial information which is not preserved during labeling (for example, lugs are required to be on the perimeter of the shoe).
 \autoref{tab:class-char-examples} shows three examples of each class.
@@
 Thousands of outsole images were scraped from online shoe retail sites and annotated using LabelMe, a tool for image annotation in computer vision problems \citep{labelme}.
-After annotation, the minimum bounding rectangle of the region \svp{is identified and the image is cropped to that area; subsequently, the cropped image is scaled to 256 x 256 pixels.} During this process, aspect ratio is not preserved, though efforts are made to label regions which are relatively square to minimize the effect of this distortion.
+After annotation, images are processed by an R script that identifies the minimum bounding rectangle of the region, crops the image to that region, and scales the cropped area to a 256 x 256 pixel image suitable for analysis by the convolutional neural network.
+During this process, aspect ratio is not preserved, though efforts are made to label regions which are relatively square to minimize the effect of this distortion.
 To date, \Sexpr{unique(ann_df$base_image) %>% length()} shoes have been labeled, yielding \Sexpr{nrow(dfunion)} multi-label images.
@@
 the training data were weighted by proportion during the training process to ensure that the model can identify both rare and common geometric shapes.
-Of the remaining 40\% of data, half were used for validation, to monitor the training process, and the remaining data were for testing the performance of the final model.
+Of the remaining 40\% of data, half were used for validation, to monitor the training process, and the remaining data were for testing the performance of the fitted model.
 
@@
 \subsection{Model Training}
-\svp{During model fitting, the model ``learns" from the training set and is evaluated on the validation set, which consists of previously unseen images (e.g.
-images not used during the training process).
-Learning occurs by minimization of the ``Loss", which is a function of the distance between the model predictions and the image labels.
-As the loss decreases, the model's accuracy (the proportion of correct labels) increases.
-During model training, we expect that the accuracy for the training set will increase beyond the accuracy for the validation set; the validation accuracy will eventually level off.
-Similarly, we expect the loss for the training set and the validation set to decrease initially.} \autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process.
-% In machine learning, a loss function is a function that measures how far model predictions are from the given labels for each image in the validation and test sets; the model ``learns" by updating its predictions to minimize the loss.
-\autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process.
 <<training-accuracy, fig.width = 7.5, fig.height = 5, out.width = ".75\\textwidth", fig.cap = "Training and validation accuracy and loss for each epoch of the fitting process.
@@
 @
+
+In machine learning, a loss function is a function that measures how far model predictions are from the given labels for each image in the validation and test sets; the model ``learns" by updating its predictions to minimize the loss.
+\autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process.
 
@@
                                          "single-exemplar"), full.names = T)
-pred_prob_plot(single_exemplars, loaded_model, sort = F, eer = eers)
-@
-
-<<multi-exemplars, fig.cap = "Examples containing multiple shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
 multi_exemplars <- list.files(file.path(wd, "images", "consistency",
                                         "multi-exemplar"), full.names = T)
-pred_prob_plot(rev(multi_exemplars), loaded_model, sort = F, eer = eers)
+pred_prob_plot(rev(c(single_exemplars, multi_exemplars)),
+               loaded_model, sort = F, eer = eers)
 @

### 2019-06-18 paper/paper.Rnw
Update examples to come from test/validation of newer model instance

@@
 
-# newest_model <- get_newest(dir = model_path, pattern = "weights.h5")
+if(grepl("!School", model_path)) {
+  newest_model <- get_newest(dir = model_path, pattern = "weights.h5")
+} else {
   newest_model <- list(path = file.path(model_path, "20190514-155453"),
@@
       process_dir = "20190514-155453")
+}
+
 newest_data_file <- file.path(modeldir, "RProcessedImages",
@@
 \paragraph{Transfer Learning}
-%\mt{Training a CNN from scratch is hard.
-There are CNNs trained on complex images, and transfer learning lets us use those weights.
+%\mt{Goal: Training a CNN from scratch is hard.
+There are CNNs trained on complex images to detect a wide range of features, and transfer learning lets us use those feature detectors.
 In particular, we're using VGG16.
@@
 
-Many convolutional neural networks have been trained on a subset of images \svp{from ImageNet}; these images are hierarchically labeled, so the same image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity \citep{dengImageNetLargeScaleHierarchical}.
+Many convolutional neural networks have been trained on a subset of images from \mt{an online image database called ImageNet \citep{dengImageNetLargeScaleHierarchical}.
+In particular, there is a standard set of images from ImageNet, consisting of 1.2 million images and spanning 1,000 catetories, that is commonly used in computer vision tasks; these images provide both a wide variety of images for training complex image-classification models and a standardized way to compare performance across different models.
+Images from ImageNet} are hierarchically labeled, so \mt{an} image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity.
 CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously.
 As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks.
-\fix{More to be said later.}
+\mt{Later in this paper, we will examine a selection of filters from a CNN trained on ImageNet that demonstrates the generalizability of pre-trained models.}
 
@@
 \paragraph{VGG16}
-Developed by Oxford's Visual Graphics Group, VGG16 is a pre-trained CNN with 16 ``functional" (i.e., convolutional and densely connected) layers and 5 ``structural" max pooling layers \citep{krizhevskyImageNetClassificationDeep2012}.
+Developed by Oxford's Visual Graphics Group, VGG16 is a \mt{CNN that has been trained on ImageNet.
+VGG16 has} 16 ``functional" (i.e., convolutional and densely connected) layers and 5 ``structural" max pooling layers.
+\citep{krizhevskyImageNetClassificationDeep2012}.
 In contrast to other popular networks, like ResNet, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy.
@@
 
-\svp{The base of VGG16 is made up of five blocks, which consist of two or three convolutional layers and one max pooling layer which aggregates the signal from the convolutional layers.
-Each convolutional layer} is made up of hundreds of filters, such that each of the five convolutional blocks has between 128 and 1,536 filters.
+The base of VGG16 is made up of five blocks, which consist of two or three convolutional layers and one max pooling layer which aggregates the signal from the convolutional layers.
+Each convolutional layer is made up of hundreds of filters, such that each of the five convolutional blocks has between 128 and 1,536 filters.
 A filter is a set of numerical weights that, when applied to an image, quantifies the presence of the specific shapes and/or colors that the filter is trained to detect.
 \autoref{fig:vgg16layers} shows a selection of filters from VGG16, with each row corresponding to one of the five convolutional blocks.
-Filters in earlier convolutional blocks detect simple features, such as colors, \svp{lines, corners,} and blobs, while later filters detect more complex combinations of features.
-\svp{This aggregation of features through successive convolutional layers mirrors the process of complex feature detection in human vision.} \svp{Any complex feature can be disassembled into a set of simple features; similarly, the successive sets of layers detect increasingly detailed features.} %When considered together, the series of filters perform the process of feature detection, such that a single image is made up of some combination of the thousands of features that are detected by filters.
-Although the filters of VGG16 have been trained and optimized on the ImageNet Data, the large number of filters span a wide variety of shapes and features, \svp{many of which are useful when examining tread patterns.
-Just as humans can detect the features of an unfamiliar object (consituent parts, edges, textures, and corners), a CNN trained on a wide set of objects can be used to classify unrelated images, such as shoe soles.} %\fix{which are still useful/relevant to classifying outsoles.
-In a sense, this is much like teaching humans to classify objects they have never seen before: an object may be unfamiliar, but our eyes can still detect its color, shape, and texture because we have learned these ideas from other sources.} \fix{Say somewhere: Max pooling layers aggregate the features in one block before passing them to the next block, which helps reduce the computational load of training the model and helps to generalize the relationship between features.} % I've tried to do this concisely - if you don't like it, please feel free to revert the changes.
+Filters in earlier convolutional blocks detect simple features, such as colors, lines, corners, and blobs, while later filters detect more complex combinations of features.
+This aggregation of features through successive convolutional layers mirrors the process of complex feature detection in human vision.
+Any complex feature can be disassembled into a set of simple features; similarly, the successive sets of layers detect increasingly detailed features.
+
+Although the filters of VGG16 have been trained and optimized on the ImageNet data, the large number of filters span a wide variety of shapes and features, many of which are useful when examining tread patterns.
+Just as humans can detect the features of an unfamiliar object (consituent parts, edges, textures, and corners), a CNN trained on a wide set of objects can be used to classify unrelated images, such as shoe soles.
 
 <<vgg16layers, fig.width = 7, fig.height = 5.75, fig.cap = "A selection of filters from the convolutional layers of VGG16.">>=
+
+if(!file.exists(file.path(wd, "images", "unlabeled_filters.png"))) {
   imgs <- list.files(file.path(wd, "images", "vgg16layers"),
@@
          plot = filters, device = "png", width = 6, height = 5.3)
+}
 
+if(!file.exists(file.path(wd, "images", "labeled_filters.png"))) {
   library(magick)
@@
               format = "png")
+}
+
 knitr::include_graphics(file.path(wd, "images", "labeled_filters.png"))
@@
 Thousands of outsole images were scraped from online shoe retail sites and annotated using LabelMe, a tool for image annotation in computer vision problems \citep{labelme}.
-After annotation, the minimum bounding rectangle of the region \svp{is identified and the image is cropped to that area; subsequently, the cropped image is scaled to 256 x 256 pixels.} During this process, aspect ratio is not preserved, though efforts are made to label regions which are relatively square to minimize the effect of this distortion.
+After annotation, the minimum bounding rectangle of the region is identified and the image is cropped to that area; subsequently, the cropped image is scaled to 256 x 256 pixels.
+During this process, aspect ratio is not preserved, though efforts are made to label regions which are relatively square to minimize the effect of this distortion.
 To date, \Sexpr{unique(ann_df$base_image) %>% length()} shoes have been labeled, yielding \Sexpr{nrow(dfunion)} multi-label images.
@@
 \subsection{Model Training}
-\svp{During model fitting, the model ``learns" from the training set and is evaluated on the validation set, which consists of previously unseen images (e.g.
+During model fitting, the model ``learns" from the training set and is evaluated on the validation set, which consists of previously unseen images (i.e.
 images not used during the training process).
-Learning occurs by minimization of the ``Loss", which is a function of the distance between the model predictions and the image labels.
+Learning occurs by minimization of the ``loss", which is a function of the distance between the model predictions and the image labels.
 As the loss decreases, the model's accuracy (the proportion of correct labels) increases.
 During model training, we expect that the accuracy for the training set will increase beyond the accuracy for the validation set; the validation accuracy will eventually level off.
-Similarly, we expect the loss for the training set and the validation set to decrease initially.} \autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process.
-% In machine learning, a loss function is a function that measures how far model predictions are from the given labels for each image in the validation and test sets; the model ``learns" by updating its predictions to minimize the loss.
+Similarly, we expect the loss for the training set and the validation set to decrease initially.
 \autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process.
+
 <<training-accuracy, fig.width = 7.5, fig.height = 5, out.width = ".75\\textwidth", fig.cap = "Training and validation accuracy and loss for each epoch of the fitting process.
@@
 \end{itemize}
-  \item (IDK where to put this, but I think it's a really important point that consistency matters more than accuracy.
-As long as a shape is predicted consistently, that provides information about outsole similarity, even if it's not 100\% correct.
-But, if phrased wrong, this undermines the claim that our model can replicate the feature detection examiners are doing...)
 \end{itemize}
@@
 Thus, to determine the accuracy of the model predictions, we must select a threshold for each label to discretize the model predictions.
-If the threshold is too low, moderate probabilites will \mt{always meet the threshold}, which increases the true positive and false positive predictions in the test data.
-Conversely, if the threshold is too high, moderate probabilities will \mt{fail to meet the threshold,} which increases the number of true and false negative predictions.
+If the threshold is too low, moderate probabilites will be more likely to exceed the threshold, which increases the true positive and false positive predictions in the test data.
+Conversely, if the threshold is too high, moderate probabilities will more often fail to meet the threshold, which increases the number of true and false negative predictions.
 An appropriate threshold must be chosen to produce a high true positive rate while also ideally keeping the false positive rate low.
-\mt{Is it worth introducing the acronyms TPR and FPR?}
 
-Receiver Operating Characteristic (ROC) curves are a type diagnostic plot which compare the false positive rate to the true positive rate for a classification method, and the Area Under the Curve (AUC) \svp{is a summary measure which} quantifies the \svp{overall accuracy of the model}.
-%It doesn't really summarize the shape - you can have very different shapes of curves with the same AUC value.
+Receiver Operating Characteristic (ROC) curves are a type diagnostic plot which compare the false positive rate to the true positive rate for a classification method, and the Area Under the Curve (AUC) is a summary measure which quantifies the overall accuracy of the model.
 Perfect prediction would be indicated by a right angle along the upper left corner of the plot, with a corresponding AUC of 1, and a diagonal line with an AUC of 0.5 would indicate that the classification method performs no better than random chance.
-\autoref{fig:overall-roc} shows the ROC curve for \fix{the model} across all classes, and \autoref{fig:class-roc} shows the curve for each class.
-The full model has an AUC of 0.88, and the AUC for individual classes ranges from 0.81 (for lines) to 0.91 (for bowtie and text).
-\fix{pull these numbers out of the model using Sexpr} While the class performances do vary slightly, each ROC curve is the same general shape and performs significantly better than random chance.
-The points in \autoref{fig:class-roc} represent the equal error rate (EER), which is the threshold where there is equal probability of a false positive or false negative error.
-We will use the class-wise EER to set a threshold corresponding to each class, so that any model prediction with probability above the EER for that shape is considered a positive prediction.
+\autoref{fig:overall-roc} shows the ROC curve for our model across all classes, and \autoref{fig:class-roc} shows the curve for each class.
 
@@
         legend.background = element_rect(fill = "white"))
-
 @
@@
 
-\paragraph{Model Predictions}
-The ROC curves provide only a general summary of model accuracy; to learn more about how specific images are classified, we can view images from the test set along with their respective nine model predictions.
-\autoref{fig:single-exemplars} shows model predictions for images that contain only one type of shape, and \autoref{fig:multi-exemplars} shows predictions for images containing multiple shapes.
-Most predictions for the single-class images are \fix{promising}; there are some confusions between circle/text and polygon/quad, but all of the true shapes are corectly identified by the model.
-\fix{The multi-label images also perform quite well; again, true labels are almost all present, with a small number of false positive labels.
-[Expand on specific cases more, once determining how many examples to keep and how much to highlight errors.] A later section will discuss possible causes for these class confusions.}
+The full model has an AUC of \Sexpr{round(whole_model_roc_df$auc,2)}, and the AUC for individual classes ranges from \Sexpr{round(min(aucs$data$auc),2)} (for \Sexpr{sprintf("%s",classes[which.min(aucs$data$auc)])}) to \Sexpr{round(max(aucs$data$auc),2)}  (for \Sexpr{sprintf("%s",classes[which.max(aucs$data$auc)])}).
+While the class performances do vary slightly, each ROC curve is the same general shape and performs significantly better than random chance.
+The points in \autoref{fig:class-roc} represent the equal error rate (EER), which is the threshold where there is equal probability of a false positive or false negative error.
+We will use the class-wise \mt{classwise?} EER to set a threshold corresponding to each class, so that any model prediction with probability above the EER for that shape is considered a positive prediction.
+
+\paragraph{Model Predictions---Examples}
+The ROC curves provide only a general summary of model accuracy; to learn more about how specific images are classified, we can view images from the test set along with their respective \mt{model-assigned} nine class predictions.
+\autoref{fig:exemplars} and \autoref{fig:circle-text-consistency} show model predictions for a selection of images.
+\mt{Each image shown in the following examples belongs to either the validation or test data, so none of the images were used during model training.
+In these figures, the values in each row are the predicted probabilities for the nine classes in the corresponding image image, such that each column represents predictions for a given class.
+The saturation of blue color corresponds to the strength of the predicted probability, and the navy blue border around an image indicates that the predicted probability is above the EER threshold for the given class.
+Note that the EER is not the same for all classes; in other words, the same probablity value assigned across different categories may yield different conclusions about whether the model predicts the shape is present or absent in the image.}
+
+The first rows of \autoref{fig:exemplars} are images which contain only one of the nine classes.
+Most of the predictions for these images are \mt{promising; there are some false-positive labels that indicate confusions between circle/text and polygon/quad, but all of the true shapes are corectly identified by the model above EER.
+The next image belongs to the ``other" category, as it does not contain any of the nine classes; the model correctly assigns near-zero probabilities to each class.
+The final rows are of images that belong to more than one class category.
+Again, predictions appear quite good, with most of the prominent shapes correctly identified above EER.
+\fix{These are mostly new examples from our last meeting because many of the previous images were in the training set...
+Grrrrr.....
+I have two or three too many, not sure about color contrast within images, etc., feel free to delete a few or we can discuss them.} [Expand on specific misclassifications more, after determining which examples to delete.] A later section will discuss possible causes for these class confusions.}
 
@@
 
-<<single-exemplars, fig.cap = "Examples containing a single shape", fig.pos = "p!", out.width = "100%", cache = T>>=
+<<exemplars, fig.cap = "Examples containing a single shape", fig.pos = "p!", out.width = "100%", cache = F>>=
 single_exemplars <- list.files(file.path(wd, "images", "consistency",
@@
 
-\fix{Figures \ref{fig:single-exemplars} and \ref{fig:multi-exemplars} provide representative examples of classifications across multiple categories; it is also useful to examine many cases within a given shape.} \autoref{fig:chevrons-consistency} shows a number of images containing chevrons.
-The top four rows show regularly repeating chevron patterns and seems to indicate that the model optimally identifies chevrons of a certain width; wide chevrons cross into the realm of bowties, whereas narrow chevrons become indistinguishable from lines.
-The next six rows show individual and closed chevrons.
-There is a significant amount of confusion between these wider chevrons and quadrilaterals.
-Finally, the bottom four rows show less-than-standard chevrons.
-The rounded chevron has a much lower prediction than those with sharper angles, but still exceeds the EER for chevron; the white chevrons are confounded with overlaid circles, which introduces a complexity that the model attributes to text.
+\mt{While figure \ref{fig:exemplars} shows representative examples of classifications across multiple shape categories, it is also useful to examine many cases within a given shape to get a sense of model consistency.} \mt{\autoref{fig:circle-text-consistency} shows a number of images containing circles and/or text.
+The top four rows show text contained within a separate, distinct circle shape, and the next five rows show circles represented as the letter ``o" in text.
+The remaining images contain only one of either text or circles, as well as other possible shape categories.
+In particular, the first two of these images contain text that use the capital letter ``G", the next two images contain no circular shapes, and the final \fix{N} images contain pure circles that do not appear to belong to text.
+The model is able to identify the circle in all cases where a circle is truly present, regardless of whether the circle is part of the text or it is distinct; however, the circle prediction is also strong in three of the four images that do not contain a circle, which indicates that the model is having dificulty distinguishing the two categories.
+This confusion is likely due to the frequency of images which contain both circles and text: text is present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("circle", dfunion$name)), 2)}\% of our labeled images containing circles, and circles are present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("text", dfunion$name)), 2)}\% of our images containing text.} In addition, the distinction between a circle and an ``o" comes from the context around the shape, which is much more difficult for the model to parse than simply detecting a shape.
+More importantly, this behavior does not suggest poor model performance as much as a weakness of the classification scheme, which will be discussed in a later section.
 
-<<chevrons-consistency, fig.cap = "Chevrons", fig.pos = "p!", out.width = "100%", cache = T>>=
-chevron <- list.files(file.path(wd, "images", "consistency", "chevron"), full.names = T)
-pred_prob_plot(rev(chevron), loaded_model, sort = F, eer = eers)
-@
-
-\fix{Describe \autoref{fig:circle-text-consistency}...... The final two don't contain a true circle. IDK why the TOMS logo totally misses circle.}
-
-\mt{If We end up making the claim that consistency matters more than accuracy, we can use \autoref{fig:ugg-consistency} to show that consistently predicting line isn't strictly right but does establish similarity between the images.
-If we don't emphasize the consistency point as much, the ugg plot is overkill.}
-
-<<circle-text-consistency, fig.cap = "Text with circles or circle-like shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
+<<circle-text-consistency, fig.cap = "Text with circles or circle-like shapes", fig.pos = "p!", out.width = "100%", cache = F>>=
 circle_text <- list.files(file.path(wd, "images", "consistency", "circle-text"), full.names = T)
@@
 
-<<ugg-consistency, fig.cap = "Text with circles or circle-like shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
-ugg <- list.files(file.path(wd, "images", "consistency", "ugg-star-logo"), full.names = T)
-pred_prob_plot(rev(ugg), loaded_model, sort = F, eer = eers)
-@
-
 %\svp{Pros and cons of confusion matrix here? Maybe we just don't show the actual proportions on the graph? I'd like to discuss this more in person.}
 
-\mt{\paragraph{Accuracy by class}}
-\fix{Figures \ref{fig:single-exemplars}-\ref{fig:circle-text-consistency} show predictions for a number of individual images.
-Many of the misclassifications in these examples are not just seen in individual images, but rather exist systematically in a class.} A confusion matrix is a type of visualization that summarizes the relationships between the labels assigned to an image and the labels predicted by the model across all test data, and is a helpful diagnostic tool to understand which classes are systematically misclassified.
+\mt{\paragraph{Confusion Matrix}}
+\mt{Figures \ref{fig:exemplars} and \ref{fig:circle-text-consistency} show predictions for a number of individual images.
+Many of the misclassifications in these examples, however, are not just seen in individual images, but rather exist systematically between certain classes.} A confusion matrix is a type of visualization that summarizes the relationships between the labels assigned to an image and the labels predicted by the model across all test data, and is a helpful diagnostic tool to understand which classes are systematically misclassified.
 \mt{The confusion matrix for \fix{our model} is presented in \autoref{fig:confusion_matrix}}.
+%mt{Each value in the matrix represents the probability of assigning the label associated with that value's corresponding row, conditional upon the image truly belonging to the category associated with the value's column.
+Thus, }
 The values along the diagonal represent the proportion of true positives captured within each category.
 The off-diagonal values represent the proportion of false positives for each shape pairing; however, since a single image may truly contain multiple shapes, the values have been adjusted to remove the effect of any true positives from the calculation of false positive proportions.
-%\svp{As a result of this adjustment, the probabilities off-diagonal are the proportion ... need to explain the conditional probability we're showing here}
-\fix{As a result of this adjustment, the off-diagonal values are the proportion of images within a true class that are falsely labeled as containing another class that has not been identified in the image.
-}%\svp{I can't remember exactly how we're doing that, but we may want to look more closely at that - if we set everything to NA instead of 0, the rate calculations would be higher than expected}
-For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any image that truly contains both triangles and quadrilaterals is removed before calculating the proportion of false quadrilateral labels.
+For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any images that truly contains both triangles and quadrilaterals are removed before calculating the proportion of false quadrilateral labels.
+\mt{As a result of this adjustment, the off-diagonal values of the confusion matrix are the proportion of images within a true class (column-wise) that are falsely labeled as containing another class that has not been labeled in the image (row-wise).}%\svp{I can't remember exactly how we're doing that, but we may want to look more closely at that - if we set everything to NA instead of 0, the rate calculations would be higher than expected}
  \mt{Note that this matrix is not symmetric---the proportion of images containing text that are falsely predicted to contain circles is not necessarily equal to the proportion of images containing circles that are falsely predicted to contain text.}
@@
 The horizontal band in \autoref{fig:confusion_matrix} indicates that quadrilaterals are predicted more often than they should be for every true label.
-Similarly, polygons, stars, and triangles produce a large number of false positive predictions into other categories, as evidenced by the vertical bands.
-Another interesting artifact in \autoref{fig:confusion_matrix} is that circles are often misclassified as text, and text is often mislabeled as containing circles.
+Similarly, polygons, stars, and triangles produce a large number of false positive predictions into other categories, as evidenced by the\mt{ir} vertical bands.
+It is also apparent that circles and text are commonly confused, which supports Another interesting artifact in \autoref{fig:confusion_matrix} is that circles are often misclassified as text, and text is often mislabeled as containing circles.
 
@@
 
-<<brand-plots, device = "cairo", include = F>>=
+<<brand-plots, out.width = "100%", fig.cap = "Proportions of top 10 shoe brands in our labeled dataset within each shape category, sorted by prevalence.
+Brand text is present when the brand makes up more than five percent of labeled shoes for the shape category.
+Some shapes, like bowtie and polygon, are relatively dominated by a small number of brands while other shapes, such as text and quad, are comprised of a large number of brands.
+Typically, model predictions are best for shape categories without much variability, which, in general, coincides with shapes dominated by a small number of brands.">>=
 img_names <- dfunion$filename %>% basename() %>% as.vector

### 2019-06-19 paper/code/Generate_Model_Images.R
Update pred_prob_plot to have legend for EER

@@ -137,7 +137,7 @@ pred_prob_plot <- function(img_path, model, classes = default_classes,
 
   img_preds_df_long <- merge(img_preds_df_long, above_eer,
                              by = c("class", "idx"))
-  img_preds_df_long$fill <- img_preds_df_long$value + img_preds_df_long$yes
+  img_preds_df_long$yes <- factor(img_preds_df_long$yes, labels = c("No", "Yes"))
 
   features_xraster <- purrr::map2(img_preds_df$path, img_preds_df$idx,
                                   ~annotation_custom(grid::rasterGrob(readJPEG(.x), interpolate = T),
@@ -147,16 +147,16 @@ pred_prob_plot <- function(img_path, model, classes = default_classes,
                                                      xmax = 0.5))
 
   if(sum(eer) == length(classes)) {
-    cols <- c("0" = "black", "1" = "black")
+    cols <- c("No" = "black", "Yes" = "black")
     w <- h <- 1
   } else {
-    cols <- c("0" = "white", "1" = "navyblue")
+    cols <- c("No" = "white", "Yes" = "navyblue")
     w <- h <- .975
   }
 
   ggplot() +
     geom_tile(aes(y = idx, x = as.numeric(factor(class)),
-                  fill = value, color = factor(yes),
+                  fill = value, color = yes,
                   width = w, height = h),
               data = img_preds_df_long) +
     geom_text(aes(x = as.numeric(factor(class)), y = idx,
@@ -164,14 +164,17 @@ pred_prob_plot <- function(img_path, model, classes = default_classes,
               hjust = 0.5, vjust = 0.5) +
     scale_fill_gradient("Predicted\nProbability", low = "white",
                         high = "cornflowerblue",limits = c(0, 1)) +
-    scale_color_manual(values = cols, guide = "none") +
+    scale_color_manual(values = cols) +
     features_xraster +
     coord_fixed() +
     scale_x_continuous(limits = c(-0.5, 9.5), breaks = 1:9,
                        labels = default_classes, expand = c(0,0)) +
     scale_y_continuous(limits = c(.5, length(img_path) + .5), expand = c(0,0)) +
     theme(axis.text.y = element_blank(), axis.title = element_blank(),
-          axis.ticks.y = element_blank())
+          axis.ticks.y = element_blank()) +
+    guides(color = guide_legend(title = "Above EER", order = 2,
[…]

### 2019-06-19 paper/code/Generate_Model_Images.R
Don't add legend if no EER provided

@@ -154,7 +154,7 @@ pred_prob_plot <- function(img_path, model, classes = default_classes,
     w <- h <- .975
   }
 
-  ggplot() +
+  g <- ggplot() +
     geom_tile(aes(y = idx, x = as.numeric(factor(class)),
                   fill = value, color = yes,
                   width = w, height = h),
@@ -164,17 +164,23 @@ pred_prob_plot <- function(img_path, model, classes = default_classes,
               hjust = 0.5, vjust = 0.5) +
     scale_fill_gradient("Predicted\nProbability", low = "white",
                         high = "cornflowerblue",limits = c(0, 1)) +
-    scale_color_manual(values = cols) +
+    scale_color_manual(values = cols, guide = "none") +
     features_xraster +
     coord_fixed() +
     scale_x_continuous(limits = c(-0.5, 9.5), breaks = 1:9,
                        labels = default_classes, expand = c(0,0)) +
     scale_y_continuous(limits = c(.5, length(img_path) + .5), expand = c(0,0)) +
     theme(axis.text.y = element_blank(), axis.title = element_blank(),
-          axis.ticks.y = element_blank()) +
-    guides(color = guide_legend(title = "Above EER", order = 2,
-                                override.aes = list(fill = NA)),
-           fill = guide_colorbar(order = 1))
+          axis.ticks.y = element_blank())
+
+  if(sum(eer) == length(classes)) {
+    l <- geom_blank()
+  } else {
+    l <- guides(color = guide_legend(title = "Above EER", order = 2,
+                                     override.aes = list(fill = NA)),
+                fill = guide_colorbar(order = 1))
+  }
+     g + l
 }
 
 # --- # Heatmaps # -------------------------------------------------------------

### 2019-06-19 paper/paper.Rnw
Wrote first full pass at results

@@
 
-<<exemplars, fig.cap = "Examples containing a single shape", fig.pos = "p!", out.width = "100%", cache = F>>=
+<<exemplars, fig.cap = "Examples containing a single shape", fig.pos = "p!", out.width = "100%", cache = T>>=
 single_exemplars <- list.files(file.path(wd, "images", "consistency",
@@
 The model is able to identify the circle in all cases where a circle is truly present, regardless of whether the circle is part of the text or it is distinct; however, the circle prediction is also strong in three of the four images that do not contain a circle, which indicates that the model is having dificulty distinguishing the two categories.
-This confusion is likely due to the frequency of images which contain both circles and text: text is present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("circle", dfunion$name)), 2)}\% of our labeled images containing circles, and circles are present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("text", dfunion$name)), 2)}\% of our images containing text.} In addition, the distinction between a circle and an ``o" comes from the context around the shape, which is much more difficult for the model to parse than simply detecting a shape.
-More importantly, this behavior does not suggest poor model performance as much as a weakness of the classification scheme, which will be discussed in a later section.
+This confusion is likely due to the frequency of images which contain both circles and text: text is present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("circle", dfunion$name)), 2)}\% of our labeled images containing circles, and circles are present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("text", dfunion$name)), 2)}\% of our images containing text.
+In addition, the distinction between a circle and an ``o" comes from the context around the shape, which is much more difficult for the model to parse than simply detecting a shape.
+More importantly, this behavior does not suggest poor model performance as much as a weakness of the classification scheme, which will be discussed in a later section.}
 
-<<circle-text-consistency, fig.cap = "Text with circles or circle-like shapes", fig.pos = "p!", out.width = "100%", cache = F>>=
+\mt{Side note: I keep futzing with this section. I am going to forcibly move on and welcome Susan's feedback, and not allow myself to change any more of the consistency plots.}
+
+<<circle-text-consistency, fig.cap = "Text with circles or circle-like shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
 circle_text <- list.files(file.path(wd, "images", "consistency", "circle-text"), full.names = T)
@@
 Thus, }
-The values along the diagonal represent the proportion of true positives captured within each category.
+The values along the diagonal \mt{of the matrix} represent the proportion of true positives captured within each category.
 The off-diagonal values represent the proportion of false positives for each shape pairing; however, since a single image may truly contain multiple shapes, the values have been adjusted to remove the effect of any true positives from the calculation of false positive proportions.
 For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any images that truly contains both triangles and quadrilaterals are removed before calculating the proportion of false quadrilateral labels.
-\mt{As a result of this adjustment, the off-diagonal values of the confusion matrix are the proportion of images within a true class (column-wise) that are falsely labeled as containing another class that has not been labeled in the image (row-wise).}%\svp{I can't remember exactly how we're doing that, but we may want to look more closely at that - if we set everything to NA instead of 0, the rate calculations would be higher than expected}
- \mt{Note that this matrix is not symmetric---the proportion of images containing text that are falsely predicted to contain circles is not necessarily equal to the proportion of images containing circles that are falsely predicted to contain text.}
+\mt{As a result of this adjustment, the off-diagonal values of the confusion matrix are the proportion of images within a true class (column-wise) that are falsely labeled as containing another class that has not been labeled in the image (row-wise).
+Note that this matrix is not symmetric---the proportion of images containing text that are falsely predicted to contain circles is not necessarily equal to the proportion of images containing circles that are falsely predicted to contain text.}
 
@@
 The horizontal band in \autoref{fig:confusion_matrix} indicates that quadrilaterals are predicted more often than they should be for every true label.
-Similarly, polygons, stars, and triangles produce a large number of false positive predictions into other categories, as evidenced by the\mt{ir} vertical bands.
-It is also apparent that circles and text are commonly confused, which supports Another interesting artifact in \autoref{fig:confusion_matrix} is that circles are often misclassified as text, and text is often mislabeled as containing circles.
+Similarly, polygons, stars, and triangles produce a large number of false positive predictions into other categories, as evidenced by the vertical bands \mt{for these categories}.
+It is also apparent that circles and text are commonly confused, which supports our earlier obsesrvartions from \autoref{fig:circle-text-consistency} that these two classes are difficult for the model to distinguish.
 
-\fix{(Transition) We have identified three possible causes for systematic misclassification: incorrect labels, low contrast, and variability within classes/overlap between classes.}
+\subsection{Model Diagnostics}
+After examining model performance, both in specific cases and systematically across classes, there seem to be two primary reasons for misclassifications: poor image contrast, and \fix{overlapping features between categories.}
 
-%One reason for these misclassifications is that not all of the images used in model training are accurately labeled: in some cases, the labels have not \mt{yet} been updated after labeling guidelines have changed.
-The confusion between text and circles is a natural consequence of the overlap between geometric shapes and components of written text; it is not indicative of a lack of fit as much as a weakness in the classification system used to define the model.
+\paragraph{Image Contrast}
+\mt{As described earlier, the images used for this research were web-scraped from online retailers before being labeled and cropped.
+One consequence of using outsole images instead of impressions is the impact of colors in the image.
+Features from outsoles that are primarily black or white in color are typically not as prominent as features from images with more varied color schemes; as a result, the features detected by the model base} \fix{are not well defined, which is propogated through both model training and prediction of new images.} \mt{In other words, the model 'learns' that, when there are few meaningful features with which to determine which shapes are present, the best strategy for minimizing loss is to predict diffuse probabilities across many classes, as seen in \autoref{fig:low-contrast-consistency}.}
 
-\paragraph{Contrast}
-
-\fix{Low contrast images are difficult to predict because the features are not prominent.
-The model has 'learned' that predicting diffuse probabilities across the classes minimizes loss, so it tends to assign diffuse probabilities to low contrast images.
-Contrast correction seems to improve predictions, but the diffuse probability strategy persists, so we will incorporate contrast correction into model training.}
-
-<<low-contrast-consistency, fig.cap = "Images with low contrast, before and after correction", fig.pos = "p!", out.width = "100%", cache = T>>=
-lowcontrast <- list.files(file.path(wd, "images", "consistency", "low-contrast"), full.names = T)
+<<low-contrast-consistency, fig.cap = "Images with low contrast", fig.pos = "p!", out.width = "100%", cache = F>>=
+lowcontrast <- list.files(file.path(wd, "images", "consistency", "low-contrast2"), full.names = T)
 pred_prob_plot(lowcontrast, loaded_model, sort = F, eer = eers)
@@
 
-\paragraph{Brand Investigations}
+\mt{Note to self: If the final heatmap choices have strong indication of 'searching' the edges of images, change the problem ``low contrast" (above) to ``poor image quality" so this section can cover both contrast issues \emph{and} the consequences of cropping (which changes the image and, thus, the prediction).
+Otherwise, if there's no compelling reason to address cropping, I couldn't find many examples during a quick glance through test images, so it doesn't seem worth trying too hard to add it.}
 
-One possible explanation for the poor performance of some categories is the amount of variability within the shapes.
-\fix{For example, as seen in \autoref{fig:brands-within-shapes}, a large proportion of bowties come from Birks; thus, the model learns bowties from Birks well, which leads to more accurate predictions of new bowtie images that come from Birks.
-There is no similar dominance in categories like quads.
-Instead, quads are highly variable  and come from a wide range of brands, so it is harder for the model to learn which set of features designate a quadrilateral.
-Since a large number of features are then associated with quads, the model gets more easily confused.
-When it sees some of those same features in other classes, it tries to guess quad.
-This is further explored in the next section.}
+%The confusion between text and circles is a natural consequence of the overlap between geometric shapes and components of written text; it is not indicative of a lack of fit as much as a weakness in the classification system used to define the model.
 
-\mt{That was a really quick and dirty version of what I'm trying to say... Am I missing any big points, before I try to expand it and clean up the technical part of the writing?}
+\paragraph{Feature Overlap}
+%\paragraph{Brand Investigations}
+\mt{Given the way we have defined our categories, there are a number of features (such as lines, angles, and curves) that are not unique to only one category. For example:
+\begin{itemize}
+\item a triangle may be made up of three acute angles, but a quad can also contain three acute angles.
+\item the only distinction between an ``o" and a circle is whether there is more text around it.
+\item the only difference between lines and thin repeating quads are the ends of the shapes.
+\end{itemize}
+}
+
+\mt{To detect these shapes, the model is doing a huge amount of feature integration beyond just detecting the presence of raw lines, angles, and curves.
+A correct classification requires the detection and integration of (often) relatively large features (with respect to filters only a few pixels wide, and also that one feature may take up the whole 256x256 pixel image).
+So the model already has a tough time learning to predict an image holistically.}
+
+\mt{This issue is exacerbated by the variability of images within a class.
+A quadrilateral isn't just a square or rectangle with 4 right angles; it can also be a parallelogram, diamond, trapezoid, chevron, or anything else that has four sides.
+In general, some of our identified shapes are very common on outsoles and appear in our data in many different forms, which further widens the sets of features the model must learn to associate with a single class.
+Unfortunately, by associating even more possible features for a single category, the number of features that then overlap between categories also incresases.}
+
+\mt{\autoref{fig:class-characteristic-barchart} shows the number of images labeled within each category; the largest categories also tend to have the most variability in its labeled images.
+This is supported by \autoref{fig:brand-plots}, which breaks down, by proportion, the top ten most frequent brands in the labeled data for each class.
+The figure shows that, for example,} a relatively large proportion of bowties in our labeled data are contributed by images of Birkenstocks.
+Since the bowties on Birkenstocks tend to be similar in shape and size, the model is able to create a relatively stable definition of bowties and thus predicts new images of bowties relatively well.
+There is no similar consistency in categories like quadrilaterals.
+\autoref{fig:brand-plots} indicates that no one brand makes up more than 3\% of our labeled quadrilateral images.
+This suggests that the quadrilaterals in our labeled data are more likely to come in a variety of shapes and sizes, so it is more difficult for the model to learn which set\mt{s} of features designate a quadrilateral.
 
@@
 
-\paragraph{Model Diagnostics}
+\mt{Note to self: I feel like I want the heatmaps concentrated in one spot, and it would feel like a strange interjection to explain them earlier.
+However, I mostly see including them as a way to support the claims we've been making throughout the paper, so I could definitely imagine an argument for including them earlier for specific examples so that saying ``the model sees..." is less speculatory...}
 
-Belonging to a category of techniques called class activation map (CAM) visualizations, a heatmap is a visual diagnostic tool that highlights the portions of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}.
+\mt{Until this point, we have mostly been speculating as to what the model ``sees" when making its predictions.
+In fact, it is relatively easy to verify these claims using heatmaps.} Belonging to a category of techniques called class activation map (CAM) visualizations, a heatmap is a visual diagnostic tool that highlights the portions of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}.
 Using these graphics, we can identify which locations in an image contribute most to the model's predictions.
 In \fix{the figures in this subsection,} heatmaps are shown for a selection of three classes, which are either contained in the image or have high output probabilities, alongside the original image.
-\fix{Many of these heatmaps illustrate that multiple classes share certain features, which understandably confuses the model.}
+\fix{Many of these heatmaps illustrate that multiple classes share certain features, which understandably confuses the model.} \fix{ADD EER BORDER TO HEATMAP CODE....
+grrrr}
 
-%\fix{I picked the heatmaps I thought illustrated the point best, but didn't modify them heavily yet.
-How many do we want to include, and do we want to highlight model challenges or just focus on the point that CoNNOR sees what we see?} \svp{I think it's worth highlighting the fact that we'll never get a model that's perfect because that curly s is pretty much a circle, and there will always be weird shapes that partially activate different categories.}
-
-\fix{Susan, please generate these heatmaps and move them to the "static" folder.
-My computer still has issues with the transparent layer of the heatmap, so it's probably easiest to do what we did for my CC and have you generate the images so they turn out correctly.
-Once I can see them all, I'll decide which three (or so) heatmaps best illustrate the point that shape categories have similar features that (understandably) confuse the model and I'll write about those specific heatmaps.}
+%\fix{Chosen heatmaps are up for debate!}
 
@@
   prune_heatmap(c("chevron", "quad", "circle")) %>%
-  create_composite(save_file = T, outdir = h_imgdir)
+  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
 
@@
   prune_heatmap(c("star", "text", "circle")) %>%
-  create_composite(save_file = T, outdir = h_imgdir)
+  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
 
@@
   prune_heatmap(c("text", "circle", "line")) %>%
-  create_composite(save_file = T, outdir = h_imgdir)
+  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
 
@@
   prune_heatmap(c("triangle", "quad", "chevron")) %>%
-  create_composite(save_file = T, outdir = h_imgdir)
+  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
 
@@
   prune_heatmap(c("star", "bowtie", "polygon")) %>%
-  create_composite(save_file = T, outdir = h_imgdir)
+  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
 
@@
   prune_heatmap(c("star", "chevron", "polygon")) %>%
-  create_composite(save_file = T, outdir = h_imgdir)
+  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
+
+calc_heatmap(imgs[which(grepl("other", imgs))], loaded_model) %>%
+  #prune_heatmap(c("star", "chevron", "polygon")) %>%
+  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
 
@@
 knitr::include_graphics(heatmaps[grepl("corks", heatmaps)])
+@
+
+<<puzzle-heatmap, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing none of the target classes.", fig.cap = "This outsole shows repeated outlines of puzzle pieces, which do not fall into any of the nine classes.">>=
+knitr::include_graphics(heatmaps[grepl("other", heatmaps)])
 @

### 2019-06-19 paper/paper.Rnw
Update outlines and add points to conclusion

@@
 \begin{itemize}
-\item ROC
+\item ROC: shows model performance is generally good, gives broad comparison of class performances
+\item Examples: show what predictions look like, the model isn't perfect, and motivate that some confusions are systematic between specific classes
+\item Confusion Matrix: show patterns in misclassification, motivate diagnostics to explain the cause of common mistakes
+\item Two main reasons for error: image contrast, and overlapping features between classes
 \begin{itemize}
-\item Overall: model performance is better than chance
-\item Classes: broad comparison of class performance
+\item Low contrast: the model can't see features, which affects training and prediction
+\item Overlapping features:
+\begin{itemize}
+\item Many classes share individual features (e.g., lines, angles, curves), so the model already has to integrate features and learn context (e.g., circle vs o)
+\item Some classes are also much more variable (see brand plots), which means the model has to widen its definitions of certain shapes, which makes confusion between classes worse
 \end{itemize}
-\item Examples
-\begin{itemize}
-\item Show what predictions look like
-\item Introduce the idea that some classifications aren't perfect
-\item Motivate systematic class-by-class confusion
 \end{itemize}
-\item Confusion Matrix
-\begin{itemize}
-\item Show patterns in misclassification
-\item Motivates further understanding of common mistakes
-\end{itemize}
-\item Three main reasons for error: image contrast, some classes are more variable, and some classes share physical features (I.e., the model has to learn a lot of features to encapsulate a class with lots of variety, but then other classes share those features)
-\item Contrast
-\item Brands
-\begin{itemize}
-\item Bowties/polygons/stars are mostly found in only a few brands, but the top N quad/line brands only take up a small proportion of all quads/lines (.04 or less)
-\item In general, the most difficult classes to classify are those that are highly variable
-\end{itemize}
-\item Heatmaps
-\begin{itemize}
-\item Verify that the model is using appropriate features when predictions are correct
-\item Show examples for confident misclassifications (like Seychelles), but make the case that the mistakes are understandable
-\end{itemize}
+\item Heatmaps: 1) verify that the model is using category-appropriate features to make decisions and 2) show examples of understandable mistakes
 \end{itemize}
@@
 \subsection{Model Diagnostics}
-After examining model performance, both in specific cases and systematically across classes, there seem to be two primary reasons for misclassifications: poor image contrast, and \fix{overlapping features between categories.}
+After examining model performance, both in specific cases and systematically across classes, there seem to be \mt{two primary reasons for misclassifications: poor image contrast, and overlapping features between categories.}
 
@@
 
-\mt{To detect these shapes, the model is doing a huge amount of feature integration beyond just detecting the presence of raw lines, angles, and curves.
+\mt{To detect these shapes, the model is doing a significant amount of feature integration beyond just detecting the presence of raw lines, angles, and curves.
 A correct classification requires the detection and integration of (often) relatively large features (with respect to filters only a few pixels wide, and also that one feature may take up the whole 256x256 pixel image).
@@
 The figure shows that, for example,} a relatively large proportion of bowties in our labeled data are contributed by images of Birkenstocks.
-Since the bowties on Birkenstocks tend to be similar in shape and size, the model is able to create a relatively stable definition of bowties and thus predicts new images of bowties relatively well.
+Since the bowties on Birkenstocks tend to be similar in shape and size, the model is able to create a relatively stable definition of bowties and thus predicts new images of bowties well.
 There is no similar consistency in categories like quadrilaterals.
@@
 \svp{Rework this to address 3 main conclusions - 1: NNs work for automatic feature ID, 2: Different levels of feature variability within classes, and 3: correcting for contrast is important.}
-\mt{I'll do this once Results is stable, so I can make it match the rest of the story.}
 The goal of this research was to develop a method to automatically identify geometric class characteristics of shoe outsoles.
-Such a method takes the ``impossible" problem of quantifying outsole features in a given population and lays the groundwork for a tractable solution.
-A set of geometric class characteristics was defined to both broadly classify a large variety of shoes and to narrow down similarity into a manageable number and type of features for further use in a shoeprint analysis.
+Such a method takes the difficult problem of quantifying outsole features in a given population and lays the groundwork for a tractable solution.
+A set of geometric class characteristics was defined, based on feature sets currently used by examiners, to both broadly classify a large variety of shoes and to narrow down similarity into a manageable number and type of features for further use in a shoeprint analysis.
 The final set of features used in this research achieves both goals quite well, and has potential for much finer-grained analysis when recognized features are combined with spatial information and relationships.
 
-CoNNOR was developed using the convolutional base of VGG16 and a newly trained classifier/head to identify the defined set of geometric features.
-After training, CoNNOR performs well on the data provided.
-In general, the model is able to identify many well-defined geometric shapes in the images both consistently and accurately.
-The prediction accuracy of this model provides the ability to compute statistics for the frequency of given features in a well-defined population.
+Our model was developed using the convolutional base of VGG16 and a newly trained classifier/head to identify the defined set of geometric features.
+After training, the model performs well on the data provided.
+In general, the model is able to identify many well-defined geometric shapes in the images.
 
-\section{Future Work}
+\mt{
+\begin{itemize}
+\item Many model errors are a result of overlapping geometric features between classes.
+The variability within certain classes exacerbates the difficulty in defining categories, leading to frequent confusion between circle/text and triangle/quad/polygon
+\item Model predictions also indicate that contrast correction is important, and model should be retrained on cc'd images
+\item Mention cutting off conv block?
+\end{itemize}
+}
 
-\svp{Need to show image contrast stuff to use this paragraph.}
-Although CoNNOR performs well in its current state, there are still a number of ways to potentially improve prediction accuracy.
+%\section{Future Work}
+%\svp{Need to show image contrast stuff to use this paragraph.}
+%Although CoNNOR performs well in its current state, there are still a number of ways to potentially improve prediction accuracy.
 For one, image contrast  plays a large role in how well the model classifies the geometric shapes present.

### 2019-06-19 paper/paper.Rnw
Make heatmap goals explicit

@@
 In \fix{the figures in this subsection,} heatmaps are shown for a selection of three classes, which are either contained in the image or have high output probabilities, alongside the original image.
-\fix{Many of these heatmaps illustrate that multiple classes share certain features, which understandably confuses the model.} \fix{ADD EER BORDER TO HEATMAP CODE....
-grrrr}
+\fix{Many of these heatmaps illustrate that multiple classes share certain features, which understandably confuses the model.} \fix{Add EER border to heatmaps}
+
+\fix{Choose three heatmaps:
+\begin{itemize}
+\item An image with multiple true shapes that the model clearly predicts well, to prove that it has learned the features we think it should
+\item An image with high quad/tri/poly confusion, but the model uses appropriate features to make educated guesses
+\item An image with no true classes but would yield false positives (e.g., corks, puzzle pieces) to show that there will always be weird shapes that our classification can't be expected to cover, but our model will still behave reasonably enough when encountering these cases.
+\end{itemize}}
 

### 2019-06-20 paper/paper.Rnw
Some not terribly significant phrasing tweaks

@@
 Some of these features may include brand logos and size markings, while others are more generic geometric shapes, such as circles, triangles, stars, and chevrons.
-Class characteristic matches can be used for elimination, but are not sufficient for identification, as the characteristics are shared by all shoes with the same size, make, and model \citep{grossVariabilitySignificanceClass2013,bodziakfootwear2000}.
+Class characteristic\svp{s} can be used for elimination, but are not sufficient for identification, as the characteristics are shared by all shoes with the same size, make, and model \citep{grossVariabilitySignificanceClass2013,bodziakfootwear2000}.
 Several attempts have been made to automatically identify features in both prints and shoe tread images, with low-level image analysis methods including fractal decomposition \citep{alexanderAutomaticClassificationRecognition1999}, scale-invariant feature recognition \citep{pavlouAutomaticExtractionClassification2006}, Fourier-Mellin transformation \citep{guehamAutomaticRecognitionShoeprints2008}, and other classical image analysis methods.
-Some of these low-level methods perform relatively well in good conditions \citep{luostarinenMeasuringAccuracyAutomatic2014a} but have degraded performance under conditions which are suboptimal, including those commonly found at crimescenes.
+Some of these low-level methods perform relatively well in good conditions \citep{luostarinenMeasuringAccuracyAutomatic2014a} but have degraded performance under conditions which are suboptimal, including those commonly found at crime scenes.
 
 In the wider field of computer vision, attention has turned to more robust methods for image recognition, such as convolutional neural networks (CNNs), which are capable of achieving near-human accuracy under even degraded image conditions \citep{krizhevskyImageNetClassificationDeep2012}.
-These networks are designed to mimic the process of human vision, and typically involve application of sets of filters, some of which mimic the filters used in the low-level techniques used by early automatic classification attempts.
+These networks are designed to mimic the process of human vision, and typically involve application of sets of filters, some of which mimic the filters used in the low-level techniques used \svp{in} early automatic classification attempts.
 CNNs have been applied in footwear forensics, making use of more general neural networks optimized to detect objects found in natural scenes, such as trees, animals, buildings, and cars; the pre-trained network's features are then used to compare shoes or prints to determine how well they match \citep{kongCrossDomainImageMatching2019a,kongCrossDomainForensicShoeprint2017,zhangAdaptingConvolutionalNeural2017}.
@@
 \section{Materials and Methods}
-
+\fix{We may want to mention some of the methods in the introduction for continuity sake...}
 Ultimately, most pattern evidence is stored in photographic form; as a result, we must use image analysis methods to create numerical features from the original visual record.
 There are a number of methods that may be employed to identify shapes and features in an image, such as Fourier transforms, Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}.
-While these methods are useful in identifying these specific features at a low level, they can be computationally intensive and only identify features on a very small scale; as a result, they cannot reliably identify large geometric shapes like those that may be found in an outsole image.
+While these methods are useful in identifying these specific features at a low level, they can be computationally intensive and only identify features on a very small scale; as a result, they cannot reliably identify large geometric shapes like those that may be found in a \svp{high resolution} outsole image.
 
-Convolutional neural networks (CNNs) are widely recognized as superior for novel image classification.
+Convolutional neural networks (CNNs) are widely recognized as superior for novel image classification \svp{and feature detection}.
 CNNs are a form of artificial neural network which make use of the image convolution operator used by many low-level feature extraction methods, with the additional ability to aggregate such features and meaningfully connect them to a pre-determined set of labels.
@@
 We can look at the filters of VGG16 to support our claim that it will be good for the task.}
-Neural networks are composed of sets of layers; the early layers, which are called the model base, contain feature detectors, and the final layers make up the classifier, or model head, which connects meaningful features to classification labels.
+Neural networks are composed of sets of layers; the early layers, which are called the model base, contain feature detectors, and the final layers make up the classifier, or model head, which connects meaningful features to classification labels \svp{(see \autoref{fig:VGG16-structure} for an example of the structure of a CNN)}.
 A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power.
-Assembling data to train a network from scratch is often a nearly impossible task; the process of \emph{transfer learning} is a natural solution to this common problem.
+Assembling \svp{sufficient} data to train a network from scratch is \svp{a gargantuan} task; the process of \emph{transfer learning} is a natural solution to this common problem.
 Transfer learning leverages the modularity of neural networks, that is, that the model base can be separated from the classifier which produces predictions.
+\svp{Transfer learning approaches use the base of a model trained on more general image data and train a new model head or classifier using a smaller set of domain-specific labeled images.}
 
-Many convolutional neural networks have been trained on a subset of images from \mt{an online image database called ImageNet \citep{dengImageNetLargeScaleHierarchical}.
+\fix{This paragraph may not be necessary, depending on the level of detail we need to get into.}
+Many convolutional neural networks have been trained on a subset of images from an online image database called ImageNet \citep{dengImageNetLargeScaleHierarchical}.
 In particular, there is a standard set of images from ImageNet, consisting of 1.2 million images and spanning 1,000 catetories, that is commonly used in computer vision tasks; these images provide both a wide variety of images for training complex image-classification models and a standardized way to compare performance across different models.
-Images from ImageNet} are hierarchically labeled, so \mt{an} image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity.
+Images from ImageNet are hierarchically labeled, so \mt{an} image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity.
 CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously.
 As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks.
-\mt{Later in this paper, we will examine a selection of filters from a CNN trained on ImageNet that demonstrates the generalizability of pre-trained models.}
+\mt{Later in this paper, we will examine a selection of filters from a CNN trained on ImageNet that demonstrates the generalizability of pre-trained models.}\fix{Consider citing the figure number; e.g.
+Figure xxx shows a selection of filters from VGG16 which demonstrate the levels of detail in the initial layers of a model trained on generic data.
+The only issue is that this means figures are cited out of order...
+not sure whether that's kosher.}
 
+\fix{This section repeats bits from the last two paragraphs. Consider reordering things so that there is less duplication? What's the best way to do that coherently?}
 Each layer in the model structure is comprised of sets of weights and connections that have been optimized during training.
 Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated \citep{oquabLearningTransferringMidlevel2014}.
-In many cases, the entire model base is used, and fitted with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with diffierent sets of image data.
+In many cases, the entire model base is used with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with \svp{smaller, domain-specific} sets of image data.
 Transfer learning allows CNNs to be applied to smaller datasets of several thousand images, reduces the amount of computational time required to fit the model, and provides boosted performance compared to training a new model from scratch \citep{yosinskiHowTransferableAre2014}.
@@
 \paragraph{VGG16}
-Developed by Oxford's Visual Graphics Group, VGG16 is a \mt{CNN that has been trained on ImageNet.
-VGG16 has} 16 ``functional" (i.e., convolutional and densely connected) layers and 5 ``structural" max pooling layers.
+\fix{You haven't introduced what convolutional and densely connected layers are, so lets leave that part out for now}
+Developed by Oxford's Visual Graphics Group, VGG16 is a CNN trained on ImageNet.
+VGG16 has 16 ``functional" (i.e., convolutional and densely connected) layers and 5 ``structural" max pooling layers.
 \citep{krizhevskyImageNetClassificationDeep2012}.
@@
 
+\fix{I don't really want to get into convolutional layers, but if we're going to mention them we probably need to say something about them. Thoughts?}
 The base of VGG16 is made up of five blocks, which consist of two or three convolutional layers and one max pooling layer which aggregates the signal from the convolutional layers.

### 2019-06-20 paper/paper.Rnw
Repeat sentence?

@@
 \paragraph{Annotated Training Data}
-
+\fix{The first sentence is a repeat from earlier in the paper, so we may want to move directly into the feature set we're using. Thoughts?}
 Class characteristics are characteristics which can be used to exclude shoes from a match at a crime scene, but cannot be used for individualized matching because they are shared by many shoes (e.g., all shoes of a specific brand and size).
@@
 After consulting with practitioners, we developed a set of categories suitable for automatic recognition by convolutional neural networks.
-These modifications were necessary because some of the definitions used in \citeauthor{grossVariabilitySignificanceClass2013} require contextual spatial information which is not preserved during labeling (for example, lugs are required to be on the perimeter of the shoe).
+These modifications were necessary because some of the definitions used in Gross et al.
+(2013) require contextual spatial information which is not preserved during labeling (for example, lugs are required to be on the perimeter of the shoe).
 \autoref{tab:class-char-examples} shows three examples of each class.

### 2019-06-20 paper/paper.Rnw
More slight tweaks in phrasing

@@
 
-Overfitting is when a model learns the training data so well that its understanding of the categories becomes specific to the training cases, which, in turn, leads to poor prediction of new images that were not in the training set.
+Overfitting \svp{occurs} when a model learns the training data so well that its understanding of the categories becomes specific to the training cases, which, in turn, leads to poor prediction of new images that were not in the training set.
 During the training process, overfitting is indicated when the validation loss begins to increase after reaching a global minimum.

### 2019-06-20 paper/paper.Rnw
Slight restructuring of ROC/AUC paragraph... nothing of substance changed, but I think this might be a bit easier to read if you're not already familiar with ROC/AUC.

@@
 An appropriate threshold must be chosen to produce a high true positive rate while also ideally keeping the false positive rate low.
-
-Receiver Operating Characteristic (ROC) curves are a type diagnostic plot which compare the false positive rate to the true positive rate for a classification method, and the Area Under the Curve (AUC) is a summary measure which quantifies the overall accuracy of the model.
+\svp{One way to set an optimal threshold is to utilize Receiver Operating Characteristic (ROC) curves, } a type diagnostic plot which compare the false positive rate to the true positive rate for a classification method.
+\autoref{fig:overall-roc} shows the ROC curve for our model across all classes, and \autoref{fig:class-roc} shows the curve for each class.
+\svp{ROC curves can be summerized by the } Area Under the Curve (AUC), which quantifies the overall accuracy of the model.
 Perfect prediction would be indicated by a right angle along the upper left corner of the plot, with a corresponding AUC of 1, and a diagonal line with an AUC of 0.5 would indicate that the classification method performs no better than random chance.
-\autoref{fig:overall-roc} shows the ROC curve for our model across all classes, and \autoref{fig:class-roc} shows the curve for each class.
 

### 2019-06-20 paper/paper.Rnw
ROC + examples comments and changes

@@
 The full model has an AUC of \Sexpr{round(whole_model_roc_df$auc,2)}, and the AUC for individual classes ranges from \Sexpr{round(min(aucs$data$auc),2)} (for \Sexpr{sprintf("%s",classes[which.min(aucs$data$auc)])}) to \Sexpr{round(max(aucs$data$auc),2)}  (for \Sexpr{sprintf("%s",classes[which.max(aucs$data$auc)])}).
-While the class performances do vary slightly, each ROC curve is the same general shape and performs significantly better than random chance.
+While the class performances do vary slightly, each ROC curve \svp{indicates that the model} performs significantly better than random chance.
 The points in \autoref{fig:class-roc} represent the equal error rate (EER), which is the threshold where there is equal probability of a false positive or false negative error.
-We will use the class-wise \mt{classwise?} EER to set a threshold corresponding to each class, so that any model prediction with probability above the EER for that shape is considered a positive prediction.
+We will use \svp{a class specific} EER to set \svp{an individual} threshold \svp{for} each class, so that any model prediction with probability above the EER for that shape is considered a positive prediction.
 
 \paragraph{Model Predictions---Examples}
-The ROC curves provide only a general summary of model accuracy; to learn more about how specific images are classified, we can view images from the test set along with their respective \mt{model-assigned} nine class predictions.
+The ROC curves provide only a general summary of model accuracy; to learn more about how specific images are classified, we can view images from the test set along with \svp{the nine class predictions assigned by the model}.
 \autoref{fig:exemplars} and \autoref{fig:circle-text-consistency} show model predictions for a selection of images.
-\mt{Each image shown in the following examples belongs to either the validation or test data, so none of the images were used during model training.
+Each image shown in the following examples belongs to either the validation or test data, so none of the images were used during model training.
 In these figures, the values in each row are the predicted probabilities for the nine classes in the corresponding image image, such that each column represents predictions for a given class.
 The saturation of blue color corresponds to the strength of the predicted probability, and the navy blue border around an image indicates that the predicted probability is above the EER threshold for the given class.
-Note that the EER is not the same for all classes; in other words, the same probablity value assigned across different categories may yield different conclusions about whether the model predicts the shape is present or absent in the image.}
+Note that the EER is not the same for all classes; in other words, the same probablity value assigned across different categories may yield different conclusions about whether the model predicts the shape is present or absent in the image.
 
 The first rows of \autoref{fig:exemplars} are images which contain only one of the nine classes.
-Most of the predictions for these images are \mt{promising; there are some false-positive labels that indicate confusions between circle/text and polygon/quad, but all of the true shapes are corectly identified by the model above EER.
+Most of the predictions for these images are promising; there are some false-positive labels that indicate confusions between circle/text and polygon/quad, but all of the true shapes are corectly identified by the model above EER.
 The next image belongs to the ``other" category, as it does not contain any of the nine classes; the model correctly assigns near-zero probabilities to each class.
 The final rows are of images that belong to more than one class category.
-Again, predictions appear quite good, with most of the prominent shapes correctly identified above EER.
+Again, predictions \svp{are largely accurate}, with most of the prominent shapes correctly identified above EER.
 \fix{These are mostly new examples from our last meeting because many of the previous images were in the training set...
 Grrrrr.....
-I have two or three too many, not sure about color contrast within images, etc., feel free to delete a few or we can discuss them.} [Expand on specific misclassifications more, after determining which examples to delete.] A later section will discuss possible causes for these class confusions.}
-
+I have two or three too many, not sure about color contrast within images, etc., feel free to delete a few or we can discuss them.
+[Expand on specific misclassifications more, after determining which examples to delete.]}
+\fix{We need to make sure that the x axis labels aren't crammed together...
+Overall the examples are good, but I think I might get rid of the all-white DC image, and possibly the one right below it as well.
+We can discuss on Monday.}
 <<eers, include = F>>=
@@
 
-<<exemplars, fig.cap = "Examples containing a single shape", fig.pos = "p!", out.width = "100%", cache = T>>=
+<<exemplars, fig.cap = "Examples containing a single shape", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 8, fig.height = 12>>=
 single_exemplars <- list.files(file.path(wd, "images", "consistency",
@@
 
-\mt{While figure \ref{fig:exemplars} shows representative examples of classifications across multiple shape categories, it is also useful to examine many cases within a given shape to get a sense of model consistency.} \mt{\autoref{fig:circle-text-consistency} shows a number of images containing circles and/or text.
-The top four rows show text contained within a separate, distinct circle shape, and the next five rows show circles represented as the letter ``o" in text.
-The remaining images contain only one of either text or circles, as well as other possible shape categories.
-In particular, the first two of these images contain text that use the capital letter ``G", the next two images contain no circular shapes, and the final \fix{N} images contain pure circles that do not appear to belong to text.
+While figure \ref{fig:exemplars} shows representative examples of classifications across multiple shape categories, it is also useful to examine many cases within a given shape to get a sense of model consistency.
+\autoref{fig:circle-text-consistency} shows a number of images containing circles and/or text.
+\fix{this isn't accurate to what I'm seeing...} \fix{The top four rows show text contained within a separate, distinct circle shape, and the next five rows show circles represented as the letter ``o" in text.
+The remaining images contain only one of either text or circles, as well as other possible shape categories.} In particular, the first two of these images contain text that use the capital letter ``G", the next two images contain no circular shapes, and the final \fix{N} images contain pure circles that do not appear to belong to text.
 The model is able to identify the circle in all cases where a circle is truly present, regardless of whether the circle is part of the text or it is distinct; however, the circle prediction is also strong in three of the four images that do not contain a circle, which indicates that the model is having dificulty distinguishing the two categories.
-This confusion is likely due to the frequency of images which contain both circles and text: text is present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("circle", dfunion$name)), 2)}\% of our labeled images containing circles, and circles are present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("text", dfunion$name)), 2)}\% of our images containing text.
+\svp{While the circle predictions are above the EER for letters such as C and S, these predictions are generally much lower than those where distinct circles or Os are seen, indicating that there are some indications that features are not as strong in the quasi-circle shapes, but that the model is not able to completely separate Os from C, S, and G shapes which are also very circular.} This confusion is \svp{complicated} due to the frequency of images which contain both circles and text: text is present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("circle", dfunion$name)), 2)}\% of our labeled images containing circles, and circles are present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("text", dfunion$name)), 2)}\% of our images containing text.
 In addition, the distinction between a circle and an ``o" comes from the context around the shape, which is much more difficult for the model to parse than simply detecting a shape.
-More importantly, this behavior does not suggest poor model performance as much as a weakness of the classification scheme, which will be discussed in a later section.}
+More importantly, this behavior does not suggest poor model performance as much as a weakness of the classification scheme, which will be discussed in a later section.
 
-\mt{Side note: I keep futzing with this section. I am going to forcibly move on and welcome Susan's feedback, and not allow myself to change any more of the consistency plots.}
+% \mt{Side note: I keep futzing with this section.
+I am going to forcibly move on and welcome Susan's feedback, and not allow myself to change any more of the consistency plots.} \svp{Good strategy.
+Biggest feedback is to change the text to actually match the image as it is currently, but I get why you stopped and I'm proud of you for noticing that it was a problem and finding a solution.
+:) Nice job.}
 
@@
 
-\mt{\paragraph{Confusion Matrix}}
-\mt{Figures \ref{fig:exemplars} and \ref{fig:circle-text-consistency} show predictions for a number of individual images.
-Many of the misclassifications in these examples, however, are not just seen in individual images, but rather exist systematically between certain classes.} A confusion matrix is a type of visualization that summarizes the relationships between the labels assigned to an image and the labels predicted by the model across all test data, and is a helpful diagnostic tool to understand which classes are systematically misclassified.
-\mt{The confusion matrix for \fix{our model} is presented in \autoref{fig:confusion_matrix}}.
-%mt{Each value in the matrix represents the probability of assigning the label associated with that value's corresponding row, conditional upon the image truly belonging to the category associated with the value's column.
+\paragraph{Confusion Matrix}
+Figures \ref{fig:exemplars} and \ref{fig:circle-text-consistency} show predictions for a number of individual images.
+Many of the misclassifications in these examples, however, are not just seen in individual images, but rather exist systematically between certain classes.
+% Nicely phrased!!!
+I couldn't have done it better myself.
+A confusion matrix is a type of visualization that summarizes the relationships between the labels assigned to an image and the labels predicted by the model across all test data, and is a helpful diagnostic tool to understand which classes are systematically misclassified.
+The confusion matrix for the fitted model is presented in \autoref{fig:confusion_matrix.} %mt{Each value in the matrix represents the probability of assigning the label associated with that value's corresponding row, conditional upon the image truly belonging to the category associated with the value's column.
 Thus, }
-The values along the diagonal \mt{of the matrix} represent the proportion of true positives captured within each category.
+The values along the diagonal of the matrix represent the proportion of true positives captured within each category.
 The off-diagonal values represent the proportion of false positives for each shape pairing; however, since a single image may truly contain multiple shapes, the values have been adjusted to remove the effect of any true positives from the calculation of false positive proportions.
-For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any images that truly contains both triangles and quadrilaterals are removed before calculating the proportion of false quadrilateral labels.
-\mt{As a result of this adjustment, the off-diagonal values of the confusion matrix are the proportion of images within a true class (column-wise) that are falsely labeled as containing another class that has not been labeled in the image (row-wise).
-Note that this matrix is not symmetric---the proportion of images containing text that are falsely predicted to contain circles is not necessarily equal to the proportion of images containing circles that are falsely predicted to contain text.}
+\svp{Thus, this might be considered a conditional confusion matrix, because we condition the off diagonal probabilities on the label not being correct.} For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any images that truly contains both triangles and quadrilaterals are removed before calculating the proportion of false quadrilateral labels.
+As a result of this adjustment, the off-diagonal values of the confusion matrix are the proportion of images within a true class (column-wise) that are falsely labeled as containing another class that has not been labeled in the image (row-wise).
+Note that this matrix is not symmetric: the proportion of images containing text that are falsely predicted to contain circles is not necessarily equal to the proportion of images containing circles that are falsely predicted to contain text.
 

### 2019-06-20 paper/paper.Rnw
Confusion matrix comments and changes; model diagnostics phrasing and accepting changes.

@@
 The horizontal band in \autoref{fig:confusion_matrix} indicates that quadrilaterals are predicted more often than they should be for every true label.
-Similarly, polygons, stars, and triangles produce a large number of false positive predictions into other categories, as evidenced by the vertical bands \mt{for these categories}.
-It is also apparent that circles and text are commonly confused, which supports our earlier obsesrvartions from \autoref{fig:circle-text-consistency} that these two classes are difficult for the model to distinguish.
+Similarly, polygons, stars, and triangles produce a \svp{moderate} number of false positive predictions \svp{in} other categories, as evidenced by the vertical bands for these categories.
+\fix{I don't really see it for stars...} It is also apparent that circles and text are commonly confused, which supports our earlier obsesrvartions from \autoref{fig:circle-text-consistency} that these two classes are difficult for the model to distinguish, \svp{in part because circles are relatively common components of latin script characters.}
 
 \subsection{Model Diagnostics}
-After examining model performance, both in specific cases and systematically across classes, there seem to be \mt{two primary reasons for misclassifications: poor image contrast, and overlapping features between categories.}
+After examining model performance, both in specific cases and systematically across classes, there seem to be two primary reasons for misclassifications: poor image contrast, and overlapping features between categories.
 
 \paragraph{Image Contrast}
-\mt{As described earlier, the images used for this research were web-scraped from online retailers before being labeled and cropped.
+As described earlier, the images used for this research were \svp{obtained} from online retailers before being labeled and cropped.
 One consequence of using outsole images instead of impressions is the impact of colors in the image.
-Features from outsoles that are primarily black or white in color are typically not as prominent as features from images with more varied color schemes; as a result, the features detected by the model base} \fix{are not well defined, which is propogated through both model training and prediction of new images.} \mt{In other words, the model 'learns' that, when there are few meaningful features with which to determine which shapes are present, the best strategy for minimizing loss is to predict diffuse probabilities across many classes, as seen in \autoref{fig:low-contrast-consistency}.}
+Features from outsoles that are primarily black or white in color are typically not as prominent as features from images with more varied color schemes; as a result, \svp{the features in low contrast images are not easily resolved by the filters; this} is propogated through both model training and prediction of new images.
+In other words, the model 'learns' that when there are few meaningful features with which to determine which shapes are present, the best strategy for minimizing loss is to predict diffuse probabilities across many classes, as seen in \autoref{fig:low-contrast-consistency}.
 
@@
 \mt{Note to self: If the final heatmap choices have strong indication of 'searching' the edges of images, change the problem ``low contrast" (above) to ``poor image quality" so this section can cover both contrast issues \emph{and} the consequences of cropping (which changes the image and, thus, the prediction).
-Otherwise, if there's no compelling reason to address cropping, I couldn't find many examples during a quick glance through test images, so it doesn't seem worth trying too hard to add it.}
+Otherwise, if there's no compelling reason to address cropping, I couldn't find many examples during a quick glance through test images, so it doesn't seem worth trying too hard to add it.} \svp{I think the edge convergence is more a numerical artifact than an edge effect, personally, so I wouldn't mention it.}
 
@@
 %\paragraph{Brand Investigations}
-\mt{Given the way we have defined our categories, there are a number of features (such as lines, angles, and curves) that are not unique to only one category. For example:
+We have defined our categories \svp{through consultation with practitioners and modifications of previous research, but as a result}, there are a number of features (such as lines, angles, and curves) which are not unique to only one category.
+For example:
 \begin{itemize}
@@
 \end{itemize}
-}
 
-\mt{To detect these shapes, the model is doing a significant amount of feature integration beyond just detecting the presence of raw lines, angles, and curves.
-A correct classification requires the detection and integration of (often) relatively large features (with respect to filters only a few pixels wide, and also that one feature may take up the whole 256x256 pixel image).
-So the model already has a tough time learning to predict an image holistically.}
+To detect these shapes, the model is doing a significant amount of feature integration beyond just detecting the presence of raw lines, angles, and curves.
+A correct classification requires the detection and integration of (often) relatively large features (with respect to filters only a few pixels wide); \svp{in addition, features can be relatively small relative to the 256x256 image, or may take up the entire image.} This wide variation in relative size of the features, combined with the variability of different features across classes, leads to a number of common (but understandable) misclassifications.
 
-\mt{This issue is exacerbated by the variability of images within a class.
-A quadrilateral isn't just a square or rectangle with 4 right angles; it can also be a parallelogram, diamond, trapezoid, chevron, or anything else that has four sides.
+\svp{This is one area where the CNN's predictions do not match human labels; in part, this is because even with some wider context, the CNN does not use both bottom-up feature integration and top-down object identification based on wider experience - as a result, it cannot easily differentiate between text and constituent shapes within the text.}
+
+This issue is exacerbated by the variability of images within a class.
+A quadrilateral isn't just a square or rectangle with 4 right angles; it can also be a parallelogram, diamond, trapezoid, part of a chevron, or anything else that has four sides.
 In general, some of our identified shapes are very common on outsoles and appear in our data in many different forms, which further widens the sets of features the model must learn to associate with a single class.
-Unfortunately, by associating even more possible features for a single category, the number of features that then overlap between categories also incresases.}
+Unfortunately, by associating even more possible features for a single category, the number of features that then overlap between categories also increases.
 
-\mt{\autoref{fig:class-characteristic-barchart} shows the number of images labeled within each category; the largest categories also tend to have the most variability in its labeled images.
+\autoref{fig:class-characteristic-barchart} shows the number of images labeled within each category; the largest categories also tend to have the most variability in labeled images.
 This is supported by \autoref{fig:brand-plots}, which breaks down, by proportion, the top ten most frequent brands in the labeled data for each class.
-The figure shows that, for example,} a relatively large proportion of bowties in our labeled data are contributed by images of Birkenstocks.
+The figure shows that, for example, a relatively large proportion of bowties in our labeled data are contributed by images of Birkenstocks.
 Since the bowties on Birkenstocks tend to be similar in shape and size, the model is able to create a relatively stable definition of bowties and thus predicts new images of bowties well.
@@
 \autoref{fig:brand-plots} indicates that no one brand makes up more than 3\% of our labeled quadrilateral images.
-This suggests that the quadrilaterals in our labeled data are more likely to come in a variety of shapes and sizes, so it is more difficult for the model to learn which set\mt{s} of features designate a quadrilateral.
+This suggests that the quadrilaterals in our labeled data are more likely to come in a variety of shapes and sizes, so it is more difficult for the model to learn which set or sets of features \svp{correspond to} a quadrilateral.
 
-<<brand-plots, out.width = "100%", fig.cap = "Proportions of top 10 shoe brands in our labeled dataset within each shape category, sorted by prevalence.
+<<brand-plots, out.width = "100%", fig.width = 7, fig.height = 5, fig.cap = "Proportions of top 10 shoe brands in our labeled dataset within each shape category, sorted by prevalence.
 Brand text is present when the brand makes up more than five percent of labeled shoes for the shape category.

### 2019-06-20 paper/paper.Rnw
Comment out outline - this is pretty coherent, so I'm not sure we need it in the pdf anymore

@@
 \subsection{Model Accuracy}
-
-\mt{\textbf{OUTLINE OF RESULTS}
-\begin{itemize}
-\item ROC: shows model performance is generally good, gives broad comparison of class performances
-\item Examples: show what predictions look like, the model isn't perfect, and motivate that some confusions are systematic between specific classes
-\item Confusion Matrix: show patterns in misclassification, motivate diagnostics to explain the cause of common mistakes
-\item Two main reasons for error: image contrast, and overlapping features between classes
-\begin{itemize}
-\item Low contrast: the model can't see features, which affects training and prediction
-\item Overlapping features:
-\begin{itemize}
-\item Many classes share individual features (e.g., lines, angles, curves), so the model already has to integrate features and learn context (e.g., circle vs o)
-\item Some classes are also much more variable (see brand plots), which means the model has to widen its definitions of certain shapes, which makes confusion between classes worse
-\end{itemize}
-\end{itemize}
-\item Heatmaps: 1) verify that the model is using category-appropriate features to make decisions and 2) show examples of understandable mistakes
-\end{itemize}
-}
+%
+% \mt{\textbf{OUTLINE OF RESULTS}
+% \begin{itemize}
+% \item ROC: shows model performance is generally good, gives broad comparison of class performances
+% \item Examples: show what predictions look like, the model isn't perfect, and motivate that some confusions are systematic between specific classes
+% \item Confusion Matrix: show patterns in misclassification, motivate diagnostics to explain the cause of common mistakes
+% \item Two main reasons for error: image contrast, and overlapping features between classes
+% \begin{itemize}
+% \item Low contrast: the model can't see features, which affects training and prediction
+% \item Overlapping features:
+% \begin{itemize}
+% \item Many classes share individual features (e.g., lines, angles, curves), so the model already has to integrate features and learn context (e.g., circle vs o)
+% \item Some classes are also much more variable (see brand plots), which means the model has to widen its definitions of certain shapes, which makes confusion between classes worse
+% \end{itemize}
+% \end{itemize}
+% \item Heatmaps: 1) verify that the model is using category-appropriate features to make decisions and 2) show examples of understandable mistakes
+% \end{itemize}
+% }
 

### 2019-06-20 paper/paper.Rnw
Rephrasing and responding to comments with more comments :)

@@
 
-\mt{Note to self: I feel like I want the heatmaps concentrated in one spot, and it would feel like a strange interjection to explain them earlier.
+% \mt{Note to self: I feel like I want the heatmaps concentrated in one spot, and it would feel like a strange interjection to explain them earlier.
 However, I mostly see including them as a way to support the claims we've been making throughout the paper, so I could definitely imagine an argument for including them earlier for specific examples so that saying ``the model sees..." is less speculatory...}
 
-\mt{Until this point, we have mostly been speculating as to what the model ``sees" when making its predictions.
-In fact, it is relatively easy to verify these claims using heatmaps.} Belonging to a category of techniques called class activation map (CAM) visualizations, a heatmap is a visual diagnostic tool that highlights the portions of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}.
+\svp{We have proposed several hypotheses to explain why the model's predictions are not entirely consistent with manual labels on the images.
+We can test these hypotheses by examining what the model ``sees" when making its predictions, using heatmaps, or class activation maps (CAMs).}
+\svp{Heatmaps or Class Activation Maps} are a visual diagnostic tool that highlights the \svp{areas} of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}.
 Using these graphics, we can identify which locations in an image contribute most to the model's predictions.
 In \fix{the figures in this subsection,} heatmaps are shown for a selection of three classes, which are either contained in the image or have high output probabilities, alongside the original image.
-\fix{Many of these heatmaps illustrate that multiple classes share certain features, which understandably confuses the model.} \fix{Add EER border to heatmaps}
+\svp{In these heatmaps, we can confirm that multiple classes have common features, resulting in misclassifications}.
+\fix{Add EER border to heatmaps}
 
@@
 
-<<corks-heatmap, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing none of the target classes.", fig.cap = "This outsole shows repeated outlines of corks, which do not fall into any of the nine classes; the concave corners where the corks overlap are predicted to be stars, and the round edges lead to predictions of circles.">>=
+<<corks-heatmap, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing none of the target classes.", fig.cap = "This outsole shows repeated outlines of corks, which do not fall into any of the nine classes; the concave corners where the corks overlap are predicted to be stars, and the round edges lead to predictions of circles.\\svp{Not really...
+only 0.03 probability}">>=
 knitr::include_graphics(heatmaps[grepl("corks", heatmaps)])

### 2019-06-20 paper/paper.Rnw
Latex doesn't like underscores... change chunk name to use a dash instead. Also move punctuation outside the reference :)

@@
 A confusion matrix is a type of visualization that summarizes the relationships between the labels assigned to an image and the labels predicted by the model across all test data, and is a helpful diagnostic tool to understand which classes are systematically misclassified.
-The confusion matrix for the fitted model is presented in \autoref{fig:confusion_matrix.} %mt{Each value in the matrix represents the probability of assigning the label associated with that value's corresponding row, conditional upon the image truly belonging to the category associated with the value's column.
+The confusion matrix for the fitted model is presented in \autoref{fig:confusion-matrix}.
+%mt{Each value in the matrix represents the probability of assigning the label associated with that value's corresponding row, conditional upon the image truly belonging to the category associated with the value's column.
 Thus, }
@@
 
-<<confusion_matrix, fig.width = 9, fig.height = 8, out.width = "\\textwidth", fig.cap = "Confusion matrix, showing on the diagonal the correct classification rate and on the off-diagonal, classification errors.
+<<confusion-matrix, fig.width = 9, fig.height = 8, out.width = "\\textwidth", fig.cap = "Confusion matrix, showing on the diagonal the correct classification rate and on the off-diagonal, classification errors.
 Note that in multi-label images, correct off-diagonal labels have been excluded from the calculation of false positives.", fig.scap = "Confusion matrix, with correct and incorrect model classifications.", dpi = 600>>=

### 2019-06-20 paper/paper.Rnw
missed a reference!

@@
 
-The horizontal band in \autoref{fig:confusion_matrix} indicates that quadrilaterals are predicted more often than they should be for every true label.
+The horizontal band in \autoref{fig:confusion-matrix} indicates that quadrilaterals are predicted more often than they should be for every true label.
 Similarly, polygons, stars, and triangles produce a \svp{moderate} number of false positive predictions \svp{in} other categories, as evidenced by the vertical bands for these categories.

### 2019-06-20 paper/paper.Rnw
conclusions comments

@@
 \item Many model errors are a result of overlapping geometric features between classes.
-The variability within certain classes exacerbates the difficulty in defining categories, leading to frequent confusion between circle/text and triangle/quad/polygon
+The variability within certain classes exacerbates the difficulty in defining categories, leading to frequent confusion between circle/text and triangle/quad/polygon \svp{This is at least in part due to the features used by practitioners, and partially due to our generalization/operational definition.}
 \item Model predictions also indicate that contrast correction is important, and model should be retrained on cc'd images
-\item Mention cutting off conv block?
+\item Mention cutting off conv block? \svp{Sure, just do it in friendly terms}
 \end{itemize}

### 2019-06-20 paper/paper.Rnw
Accept edits

@@
 Some of these features may include brand logos and size markings, while others are more generic geometric shapes, such as circles, triangles, stars, and chevrons.
-Class characteristic\svp{s} can be used for elimination, but are not sufficient for identification, as the characteristics are shared by all shoes with the same size, make, and model \citep{grossVariabilitySignificanceClass2013,bodziakfootwear2000}.
+Class characteristics can be used for elimination, but are not sufficient for identification, as the characteristics are shared by all shoes with the same size, make, and model \citep{grossVariabilitySignificanceClass2013,bodziakfootwear2000}.
 Several attempts have been made to automatically identify features in both prints and shoe tread images, with low-level image analysis methods including fractal decomposition \citep{alexanderAutomaticClassificationRecognition1999}, scale-invariant feature recognition \citep{pavlouAutomaticExtractionClassification2006}, Fourier-Mellin transformation \citep{guehamAutomaticRecognitionShoeprints2008}, and other classical image analysis methods.
@@
 In the wider field of computer vision, attention has turned to more robust methods for image recognition, such as convolutional neural networks (CNNs), which are capable of achieving near-human accuracy under even degraded image conditions \citep{krizhevskyImageNetClassificationDeep2012}.
-These networks are designed to mimic the process of human vision, and typically involve application of sets of filters, some of which mimic the filters used in the low-level techniques used \svp{in} early automatic classification attempts.
+These networks are designed to mimic the process of human vision, and typically involve application of sets of filters, some of which mimic the filters used in the low-level techniques used in early automatic classification attempts.
 CNNs have been applied in footwear forensics, making use of more general neural networks optimized to detect objects found in natural scenes, such as trees, animals, buildings, and cars; the pre-trained network's features are then used to compare shoes or prints to determine how well they match \citep{kongCrossDomainImageMatching2019a,kongCrossDomainForensicShoeprint2017,zhangAdaptingConvolutionalNeural2017}.
@@
 There are a number of methods that may be employed to identify shapes and features in an image, such as Fourier transforms, Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}.
-While these methods are useful in identifying these specific features at a low level, they can be computationally intensive and only identify features on a very small scale; as a result, they cannot reliably identify large geometric shapes like those that may be found in a \svp{high resolution} outsole image.
+While these methods are useful in identifying these specific features at a low level, they can be computationally intensive and only identify features on a very small scale; as a result, they cannot reliably identify large geometric shapes like those that may be found in a high resolution outsole image.
 
-Convolutional neural networks (CNNs) are widely recognized as superior for novel image classification \svp{and feature detection}.
+Convolutional neural networks (CNNs) are widely recognized as superior for novel image classification and feature detection.
 CNNs are a form of artificial neural network which make use of the image convolution operator used by many low-level feature extraction methods, with the additional ability to aggregate such features and meaningfully connect them to a pre-determined set of labels.
@@
 We can look at the filters of VGG16 to support our claim that it will be good for the task.}
-Neural networks are composed of sets of layers; the early layers, which are called the model base, contain feature detectors, and the final layers make up the classifier, or model head, which connects meaningful features to classification labels \svp{(see \autoref{fig:VGG16-structure} for an example of the structure of a CNN)}.
+Neural networks are composed of sets of layers; the early layers, which are called the model base, contain feature detectors, and the final layers make up the classifier, or model head, which connects meaningful features to classification labels (see \autoref{fig:VGG16-structure} for an example of the structure of a CNN).
 A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power.
-Assembling \svp{sufficient} data to train a network from scratch is \svp{a gargantuan} task; the process of \emph{transfer learning} is a natural solution to this common problem.
+Assembling sufficient data to train a network from scratch is a gargantuan task; the process of \emph{transfer learning} is a natural solution to this common problem.
 Transfer learning leverages the modularity of neural networks, that is, that the model base can be separated from the classifier which produces predictions.
-\svp{Transfer learning approaches use the base of a model trained on more general image data and train a new model head or classifier using a smaller set of domain-specific labeled images.}
+Transfer learning approaches use the base of a model trained on more general image data and train a new model head or classifier using a smaller set of domain-specific labeled images.
 
@@
 In particular, there is a standard set of images from ImageNet, consisting of 1.2 million images and spanning 1,000 catetories, that is commonly used in computer vision tasks; these images provide both a wide variety of images for training complex image-classification models and a standardized way to compare performance across different models.
-Images from ImageNet are hierarchically labeled, so \mt{an} image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity.
+Images from ImageNet are hierarchically labeled, so an image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity.
 CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously.
@@
 The only issue is that this means figures are cited out of order...
-not sure whether that's kosher.}
+not sure whether that's kosher.}\mt{I think I used to have that, but then you said I hadn't introduced VGG16 yet so it didn't make sense here (which also totally makes sense).
+It's definitely a chicken and egg problem...}
 
@@
 Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated \citep{oquabLearningTransferringMidlevel2014}.
-In many cases, the entire model base is used with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with \svp{smaller, domain-specific} sets of image data.
+In many cases, the entire model base is used with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with smaller, domain-specific sets of image data.
 Transfer learning allows CNNs to be applied to smaller datasets of several thousand images, reduces the amount of computational time required to fit the model, and provides boosted performance compared to training a new model from scratch \citep{yosinskiHowTransferableAre2014}.
@@
 \paragraph{VGG16}
-\fix{You haven't introduced what convolutional and densely connected layers are, so lets leave that part out for now}
-Developed by Oxford's Visual Graphics Group, VGG16 is a CNN trained on ImageNet.
-VGG16 has 16 ``functional" (i.e., convolutional and densely connected) layers and 5 ``structural" max pooling layers.
+Developed by Oxford's Visual Graphics Group, VGG16 is a CNN trained on ImageNet
 \citep{krizhevskyImageNetClassificationDeep2012}.
-In contrast to other popular networks, like ResNet, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy.
+In contrast to other popular \mt{pre-trained networks}, like ResNet, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy.
 The simplicity of this structure provides the ability to peer into the inner workings of the network for diagnostic purposes, providing a distinct advantage over more complicated network structures with slightly higher accuracy ratings.
@@
 \fix{I don't really want to get into convolutional layers, but if we're going to mention them we probably need to say something about them. Thoughts?}
+\mt{I only want to give enough detail to briefly explain 1) how filters work as feature detectors and 2) why filters trained on not-shoes will work well on shoes.
+I think that point is made strongly by showing hierarchy in the blocks.
+I thought that the phrasing below gave just enough of a background to mention that filters are structured in layers, but that might raise more questions than it answers---maybe we could gloss over the layer-structure of the blocks and just mention they contain filters?}
 The base of VGG16 is made up of five blocks, which consist of two or three convolutional layers and one max pooling layer which aggregates the signal from the convolutional layers.
@@
 \fix{The first sentence is a repeat from earlier in the paper, so we may want to move directly into the feature set we're using. Thoughts?}
+\mt{Yeah, I think this kind of introduction belongs at the beginning of the paper. But I think we still want to mention the Gross features here and how we modified them.}
 Class characteristics are characteristics which can be used to exclude shoes from a match at a crime scene, but cannot be used for individualized matching because they are shared by many shoes (e.g., all shoes of a specific brand and size).
@@
 (2013) require contextual spatial information which is not preserved during labeling (for example, lugs are required to be on the perimeter of the shoe).
-\autoref{tab:class-char-examples} shows three examples of each class.
+\autoref{tab:class-char-examples} shows three examples of each class \mt{in our feature set}.
 
@@
 
-Overfitting \svp{occurs} when a model learns the training data so well that its understanding of the categories becomes specific to the training cases, which, in turn, leads to poor prediction of new images that were not in the training set.
+Overfitting occurs when a model learns the training data so well that its understanding of the categories becomes specific to the training cases, which, in turn, leads to poor prediction of new images that were not in the training set.
 During the training process, overfitting is indicated when the validation loss begins to increase after reaching a global minimum.
@@
 An appropriate threshold must be chosen to produce a high true positive rate while also ideally keeping the false positive rate low.
-\svp{One way to set an optimal threshold is to utilize Receiver Operating Characteristic (ROC) curves, } a type diagnostic plot which compare the false positive rate to the true positive rate for a classification method.
+One way to set an optimal threshold is to utilize Receiver Operating Characteristic (ROC) curves, a type of diagnostic plot which compare the false positive rate to the true positive rate for a classification method.
 \autoref{fig:overall-roc} shows the ROC curve for our model across all classes, and \autoref{fig:class-roc} shows the curve for each class.
-\svp{ROC curves can be summerized by the } Area Under the Curve (AUC), which quantifies the overall accuracy of the model.
+ROC curves can be summerized by the Area Under the Curve (AUC), which quantifies the overall accuracy of the model.
 Perfect prediction would be indicated by a right angle along the upper left corner of the plot, with a corresponding AUC of 1, and a diagonal line with an AUC of 0.5 would indicate that the classification method performs no better than random chance.
@@
 The full model has an AUC of \Sexpr{round(whole_model_roc_df$auc,2)}, and the AUC for individual classes ranges from \Sexpr{round(min(aucs$data$auc),2)} (for \Sexpr{sprintf("%s",classes[which.min(aucs$data$auc)])}) to \Sexpr{round(max(aucs$data$auc),2)}  (for \Sexpr{sprintf("%s",classes[which.max(aucs$data$auc)])}).
-While the class performances do vary slightly, each ROC curve \svp{indicates that the model} performs significantly better than random chance.
+While the class performances do vary slightly, each ROC curve indicates that the model performs significantly better than random chance.
 The points in \autoref{fig:class-roc} represent the equal error rate (EER), which is the threshold where there is equal probability of a false positive or false negative error.
-We will use \svp{a class specific} EER to set \svp{an individual} threshold \svp{for} each class, so that any model prediction with probability above the EER for that shape is considered a positive prediction.
+We will use a class specific EER to set an individual threshold for each class, so that any model prediction with probability above the EER for that shape is considered a positive prediction.
 
 \paragraph{Model Predictions---Examples}
-The ROC curves provide only a general summary of model accuracy; to learn more about how specific images are classified, we can view images from the test set along with \svp{the nine class predictions assigned by the model}.
+The ROC curves provide only a general summary of model accuracy; to learn more about how specific images are classified, we can view images from the test set along with the nine class predictions assigned by the model.
 \autoref{fig:exemplars} and \autoref{fig:circle-text-consistency} show model predictions for a selection of images.
 Each image shown in the following examples belongs to either the validation or test data, so none of the images were used during model training.
-In these figures, the values in each row are the predicted probabilities for the nine classes in the corresponding image image, such that each column represents predictions for a given class.
+In these figures, the values in each row are the predicted probabilities for the nine classes in the corresponding image, such that each column represents predictions for a given class.
 The saturation of blue color corresponds to the strength of the predicted probability, and the navy blue border around an image indicates that the predicted probability is above the EER threshold for the given class.
@@
 The final rows are of images that belong to more than one class category.
-Again, predictions \svp{are largely accurate}, with most of the prominent shapes correctly identified above EER.
+Again, predictions are largely accurate, with most of the prominent shapes correctly identified above EER.
 \fix{These are mostly new examples from our last meeting because many of the previous images were in the training set...
@@
 The model is able to identify the circle in all cases where a circle is truly present, regardless of whether the circle is part of the text or it is distinct; however, the circle prediction is also strong in three of the four images that do not contain a circle, which indicates that the model is having dificulty distinguishing the two categories.
-\svp{While the circle predictions are above the EER for letters such as C and S, these predictions are generally much lower than those where distinct circles or Os are seen, indicating that there are some indications that features are not as strong in the quasi-circle shapes, but that the model is not able to completely separate Os from C, S, and G shapes which are also very circular.} This confusion is \svp{complicated} due to the frequency of images which contain both circles and text: text is present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("circle", dfunion$name)), 2)}\% of our labeled images containing circles, and circles are present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("text", dfunion$name)), 2)}\% of our images containing text.
+\svp{While the circle predictions are above the EER for letters such as C and S, these predictions are generally much lower than those where distinct circles or Os are seen, indicating that there are some indications that features are not as strong in the quasi-circle shapes, but that the model is not able to completely separate Os from C, S, and G shapes which are also very circular.} This confusion is complicated \mt{by} the frequency of images which contain both circles and text: text is present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("circle", dfunion$name)), 2)}\% of our labeled images containing circles, and circles are present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("text", dfunion$name)), 2)}\% of our images containing text.
 In addition, the distinction between a circle and an ``o" comes from the context around the shape, which is much more difficult for the model to parse than simply detecting a shape.
@@
 :) Nice job.}
+\mt{As the above section is right now, the text should match the images...
+I had some issues with image sorting, because Windows will put 11 before 2 but not before 02, so maybe Linux has a different behavior that's resorting the images?
+I also cache them to speed up recompiling, so maybe something's up there?}
 
@@
 The off-diagonal values represent the proportion of false positives for each shape pairing; however, since a single image may truly contain multiple shapes, the values have been adjusted to remove the effect of any true positives from the calculation of false positive proportions.
-\svp{Thus, this might be considered a conditional confusion matrix, because we condition the off diagonal probabilities on the label not being correct.} For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any images that truly contains both triangles and quadrilaterals are removed before calculating the proportion of false quadrilateral labels.
+Thus, this might be considered a conditional confusion matrix, because we condition the off diagonal probabilities on the label not being correct.
+For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any images that truly contains both triangles and quadrilaterals are removed before calculating the proportion of false quadrilateral labels.
 As a result of this adjustment, the off-diagonal values of the confusion matrix are the proportion of images within a true class (column-wise) that are falsely labeled as containing another class that has not been labeled in the image (row-wise).
@@
 The horizontal band in \autoref{fig:confusion-matrix} indicates that quadrilaterals are predicted more often than they should be for every true label.
-Similarly, polygons, stars, and triangles produce a \svp{moderate} number of false positive predictions \svp{in} other categories, as evidenced by the vertical bands for these categories.
-\fix{I don't really see it for stars...} It is also apparent that circles and text are commonly confused, which supports our earlier obsesrvartions from \autoref{fig:circle-text-consistency} that these two classes are difficult for the model to distinguish, \svp{in part because circles are relatively common components of latin script characters.}
+Similarly, polygons and triangles produce a moderate number of false positive predictions in other categories, as evidenced by the vertical bands for these categories.
+It is also apparent that circles and text are commonly confused, which supports our earlier obsesrvartions from \autoref{fig:circle-text-consistency} that these two classes are difficult for the model to distinguish, in part because circles are relatively common components of latin script characters.
 
@@
 \paragraph{Image Contrast}
-As described earlier, the images used for this research were \svp{obtained} from online retailers before being labeled and cropped.
+As described earlier, the images used for this research were obtained \mt{:D} from online retailers before being labeled and cropped.
 One consequence of using outsole images instead of impressions is the impact of colors in the image.
-Features from outsoles that are primarily black or white in color are typically not as prominent as features from images with more varied color schemes; as a result, \svp{the features in low contrast images are not easily resolved by the filters; this} is propogated through both model training and prediction of new images.
+Features from outsoles that are primarily black or white in color are typically not as prominent as features from images with more varied color schemes; as a result, the features in low contrast images are not easily resolved by the filters; this is propogated through both model training and prediction of new images.
 In other words, the model 'learns' that when there are few meaningful features with which to determine which shapes are present, the best strategy for minimizing loss is to predict diffuse probabilities across many classes, as seen in \autoref{fig:low-contrast-consistency}.
@@
 
-\mt{Note to self: If the final heatmap choices have strong indication of 'searching' the edges of images, change the problem ``low contrast" (above) to ``poor image quality" so this section can cover both contrast issues \emph{and} the consequences of cropping (which changes the image and, thus, the prediction).
-Otherwise, if there's no compelling reason to address cropping, I couldn't find many examples during a quick glance through test images, so it doesn't seem worth trying too hard to add it.} \svp{I think the edge convergence is more a numerical artifact than an edge effect, personally, so I wouldn't mention it.}
-
 %The confusion between text and circles is a natural consequence of the overlap between geometric shapes and components of written text; it is not indicative of a lack of fit as much as a weakness in the classification system used to define the model.
@@
 %\paragraph{Brand Investigations}
-We have defined our categories \svp{through consultation with practitioners and modifications of previous research, but as a result}, there are a number of features (such as lines, angles, and curves) which are not unique to only one category.
+We have defined our categories through consultation with practitioners and modifications of previous research, but as a result, there are a number of features (such as lines, angles, and curves) which are not unique to only one category.
 For example:
@@
 To detect these shapes, the model is doing a significant amount of feature integration beyond just detecting the presence of raw lines, angles, and curves.
-A correct classification requires the detection and integration of (often) relatively large features (with respect to filters only a few pixels wide); \svp{in addition, features can be relatively small relative to the 256x256 image, or may take up the entire image.} This wide variation in relative size of the features, combined with the variability of different features across classes, leads to a number of common (but understandable) misclassifications.
+A correct classification requires the detection and integration of (often) relatively large features (with respect to filters only a few pixels wide); in addition, features can be relatively small relative to the 256x256 image, or may take up the entire image.
+This wide variation in relative size of the features, combined with the variability of different features across classes, leads to a number of common (but understandable) misclassifications.
 
-\svp{This is one area where the CNN's predictions do not match human labels; in part, this is because even with some wider context, the CNN does not use both bottom-up feature integration and top-down object identification based on wider experience - as a result, it cannot easily differentiate between text and constituent shapes within the text.}
+This is one area where the CNN's predictions do not match human labels; in part, this is because even with some wider context, the CNN does not use both bottom-up feature integration and top-down object identification based on wider experience - as a result, it cannot easily differentiate between text and constituent shapes within the text.
 
@@
 In general, some of our identified shapes are very common on outsoles and appear in our data in many different forms, which further widens the sets of features the model must learn to associate with a single class.
-Unfortunately, by associating even more possible features for a single category, the number of features that then overlap between categories also increases.
+Unfortunately, by associating even more possible features with a single category, the number of features that then overlap between categories also increases.
 
@@
 \autoref{fig:brand-plots} indicates that no one brand makes up more than 3\% of our labeled quadrilateral images.
-This suggests that the quadrilaterals in our labeled data are more likely to come in a variety of shapes and sizes, so it is more difficult for the model to learn which set or sets of features \svp{correspond to} a quadrilateral.
+This suggests that the quadrilaterals in our labeled data are more likely to come in a variety of shapes and sizes, so it is more difficult for the model to learn which set or sets of features correspond to a quadrilateral.
 
@@
 
-\svp{We have proposed several hypotheses to explain why the model's predictions are not entirely consistent with manual labels on the images.
-We can test these hypotheses by examining what the model ``sees" when making its predictions, using heatmaps, or class activation maps (CAMs).}
-\svp{Heatmaps or Class Activation Maps} are a visual diagnostic tool that highlights the \svp{areas} of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}.
+We have proposed several hypotheses to explain why the model's predictions are not entirely consistent with manual labels on the images.
+We can test these hypotheses by examining what the model ``sees" when making its predictions, using heatmaps, or class activation maps (CAMs).
+\fix{Heatmaps or Class Activation Maps} are a visual diagnostic tool that highlights the areas of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}.
 Using these graphics, we can identify which locations in an image contribute most to the model's predictions.
 In \fix{the figures in this subsection,} heatmaps are shown for a selection of three classes, which are either contained in the image or have high output probabilities, alongside the original image.
-\svp{In these heatmaps, we can confirm that multiple classes have common features, resulting in misclassifications}.
+In these heatmaps, we can confirm that multiple classes have common features, resulting in misclassifications.
 \fix{Add EER border to heatmaps}

### 2019-06-20 paper/paper.Rnw
Fix methods to address comments and remove redundancy

@@
 \modulolinenumbers[5]
+
+% Allow strikethrough text
+\usepackage[normalem]{ulem}
 
@@
 Neural networks are composed of sets of layers; the early layers, which are called the model base, contain feature detectors, and the final layers make up the classifier, or model head, which connects meaningful features to classification labels (see \autoref{fig:VGG16-structure} for an example of the structure of a CNN).
-A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power.
+\mt{Each of these layers is comprised of sets of weights and connections, or parameters, that are optimized during training.} A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power.
 Assembling sufficient data to train a network from scratch is a gargantuan task; the process of \emph{transfer learning} is a natural solution to this common problem.
-Transfer learning leverages the modularity of neural networks, that is, that the model base can be separated from the classifier which produces predictions.
-Transfer learning approaches use the base of a model trained on more general image data and train a new model head or classifier using a smaller set of domain-specific labeled images.
+Transfer learning leverages the modularity of neural networks, that is, that the model base that detects features can be separated from the classifier which produces predictions.
+Transfer learning approaches use the base of a model trained on more general image data and train a new model head or classifier using a smaller set of domain-specific labeled images \mt{\citep{oquabLearningTransferringMidlevel2014}, allowing CNNs to be applied to smaller datasets of several thousand images, reduces the amount of computational time required to fit the model, and provides boosted performance compared to training a new model from scratch \citep{yosinskiHowTransferableAre2014}.
+It has been successfully applied in automatic classification of medical images \citep{shinDeepConvolutionalNeural2016} as well as in various applications of shoe forensics \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017}.}
 
-\fix{This paragraph may not be necessary, depending on the level of detail we need to get into.}
+%\fix{This paragraph may not be necessary, depending on the level of detail we need to get into.}
+%\mt{I think briefly explaining the pre-training data helps support the idea that the model can learn to classify like we do, but it can definitely be said more concisely.}
 Many convolutional neural networks have been trained on a subset of images from an online image database called ImageNet \citep{dengImageNetLargeScaleHierarchical}.
 In particular, there is a standard set of images from ImageNet, consisting of 1.2 million images and spanning 1,000 catetories, that is commonly used in computer vision tasks; these images provide both a wide variety of images for training complex image-classification models and a standardized way to compare performance across different models.
-Images from ImageNet are hierarchically labeled, so an image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity.
-CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously.
+\mt{\sout{Images from ImageNet are hierarchically labeled, so an image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity.
+}}CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously.
 As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks.
-\mt{Later in this paper, we will examine a selection of filters from a CNN trained on ImageNet that demonstrates the generalizability of pre-trained models.}\fix{Consider citing the figure number; e.g.
+\mt{Later in this paper, we will examine a selection of filters from a CNN trained on ImageNet that demonstrates the generalizability of pre-trained models.} \fix{Consider citing the figure number; e.g.
 Figure xxx shows a selection of filters from VGG16 which demonstrate the levels of detail in the initial layers of a model trained on generic data.
 The only issue is that this means figures are cited out of order...
-not sure whether that's kosher.}\mt{I think I used to have that, but then you said I hadn't introduced VGG16 yet so it didn't make sense here (which also totally makes sense).
+not sure whether that's kosher.} \mt{I think I used to have that, but then you said I hadn't introduced VGG16 yet so it didn't make sense here (which also totally makes sense).
 It's definitely a chicken and egg problem...}
 
-\fix{This section repeats bits from the last two paragraphs. Consider reordering things so that there is less duplication? What's the best way to do that coherently?}
-Each layer in the model structure is comprised of sets of weights and connections that have been optimized during training.
-Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated \citep{oquabLearningTransferringMidlevel2014}.
-In many cases, the entire model base is used with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with smaller, domain-specific sets of image data.
-Transfer learning allows CNNs to be applied to smaller datasets of several thousand images, reduces the amount of computational time required to fit the model, and provides boosted performance compared to training a new model from scratch \citep{yosinskiHowTransferableAre2014}.
+%\fix{This section repeats bits from the last two paragraphs. Consider reordering things so that there is less duplication? What's the best way to do that coherently?}
+%\mt{Note to Susan: Most of the following paragraph has been Frankensteined into the 'Transfer learning' paragraph. If it sounds okay there, we can delete this paragraph.}
+%Each layer in the model structure is comprised of sets of weights and connections that have been optimized during training. \mt{moved to para 1}
+%Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated \citep{oquabLearningTransferringMidlevel2014}.
+\mt{citation used in para 1}
+%In many cases, the entire model base is used with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with smaller, domain-specific sets of image data.
+\mt{redundant to above, not used}
+%Transfer learning allows CNNs to be applied to smaller datasets of several thousand images, reduces the amount of computational time required to fit the model, and provides boosted performance compared to training a new model from scratch \citep{yosinskiHowTransferableAre2014}.
 It has been successfully applied in automatic classification of medical images \citep{shinDeepConvolutionalNeural2016} as well as in various applications of shoe forensics \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017}.
+\mt{moved to end of para 1}
 
@@
 \citep{krizhevskyImageNetClassificationDeep2012}.
-In contrast to other popular \mt{pre-trained networks}, like ResNet, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy.
+In contrast to other popular \mt{pre-trained} networks, like ResNet, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy.
 The simplicity of this structure provides the ability to peer into the inner workings of the network for diagnostic purposes, providing a distinct advantage over more complicated network structures with slightly higher accuracy ratings.

### 2019-06-20 paper/paper.Rnw
Heatmaps: options for EER border, prune by vector of classes or top N classes, sort alphabetically or by pred prob. Also updated heatmap code and examples (not necessarily final)

@@
 \fix{The first sentence is a repeat from earlier in the paper, so we may want to move directly into the feature set we're using. Thoughts?}
-\mt{Yeah, I think this kind of introduction belongs at the beginning of the paper. But I think we still want to mention the Gross features here and how we modified them.}
-Class characteristics are characteristics which can be used to exclude shoes from a match at a crime scene, but cannot be used for individualized matching because they are shared by many shoes (e.g., all shoes of a specific brand and size).
-A sufficiently well-defined set of features can be used to separate shoes into make and model categories \citep{grossVariabilitySignificanceClass2013}; the set of features used in that study included circle/oval, crepe, herringbone, hexagon, parallel lines, logo/lettering/numbering, perimeter lugs, star, and other.
+\mt{\sout{Class characteristics are characteristics which can be used to exclude shoes from a match at a crime scene, but cannot be used for individualized matching because they are shared by many shoes (e.g., all shoes of a specific brand and size).}} \mt{Existing research indicates that a} sufficiently well-defined set of features can be used to separate shoes into make and model categories \citep{grossVariabilitySignificanceClass2013}; the set of features used in that study included circle/oval, crepe, herringbone, hexagon, parallel lines, logo/lettering/numbering, perimeter lugs, star, and other.
 After consulting with practitioners, we developed a set of categories suitable for automatic recognition by convolutional neural networks.
@@
 
-Thousands of outsole images were scraped from online shoe retail sites and annotated using LabelMe, a tool for image annotation in computer vision problems \citep{labelme}.
+Thousands of outsole images were \mt{obtained} from online shoe retail sites and annotated using LabelMe, a tool for image annotation in computer vision problems \citep{labelme}.
 After annotation, the minimum bounding rectangle of the region is identified and the image is cropped to that area; subsequently, the cropped image is scaled to 256 x 256 pixels.
@@
 
-Transfer learning substantially reduces the amount of data required to train a CNN by several orders of magnitude, but labeled images are still difficult to generate on a large scale.
+Transfer learning reduces the amount of data required to train a CNN by several orders of magnitude, but labeled images are still difficult to generate on a large scale.
 Thus, to make the most efficient use of the existing labeled data, we enlarge the training data using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}.
@@
 
-<<circle-text-consistency, fig.cap = "Text with circles or circle-like shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
+<<circle-text-consistency, fig.cap = "Text with circles or circle-like shapes", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 8, fig.height = 12>>=
 circle_text <- list.files(file.path(wd, "images", "consistency", "circle-text"), full.names = T)
@@
 \paragraph{Image Contrast}
-As described earlier, the images used for this research were obtained \mt{:D} from online retailers before being labeled and cropped.
+As described earlier, the images used for this research were obtained from online retailers before being labeled and cropped.
 One consequence of using outsole images instead of impressions is the impact of colors in the image.
-Features from outsoles that are primarily black or white in color are typically not as prominent as features from images with more varied color schemes; as a result, the features in low contrast images are not easily resolved by the filters; this is propogated through both model training and prediction of new images.
+Features from outsoles that are primarily black or white in color are typically not as prominent as features from images with more varied color schemes.
+As a result, the features in low contrast images are not easily resolved by the filters; this is propogated through both model training and prediction of new images.
 In other words, the model 'learns' that when there are few meaningful features with which to determine which shapes are present, the best strategy for minimizing loss is to predict diffuse probabilities across many classes, as seen in \autoref{fig:low-contrast-consistency}.
@@
 
-\autoref{fig:class-characteristic-barchart} shows the number of images labeled within each category; the largest categories also tend to have the most variability in labeled images.
+The shape categories with the largest number of images, shown in \autoref{fig:class-characteristic-barchart}, also tend to have the most variability in labeled images.
 This is supported by \autoref{fig:brand-plots}, which breaks down, by proportion, the top ten most frequent brands in the labeled data for each class.
@@
 There is no similar consistency in categories like quadrilaterals.
-\autoref{fig:brand-plots} indicates that no one brand makes up more than 3\% of our labeled quadrilateral images.
+\autoref{fig:brand-plots} indicates that no one brand makes up more than 4\% of our labeled quadrilateral images.
 This suggests that the quadrilaterals in our labeled data are more likely to come in a variety of shapes and sizes, so it is more difficult for the model to learn which set or sets of features correspond to a quadrilateral.
@@
   mutate(rank = pmin(row_number(), 11))
+
 brand_prop_by_feature_10 <- brand_prop_by_feature %>%
@@
 ggplot(data = brand_prop_by_feature_10) +
-  geom_bar(aes(x = feature, y = prop, fill = factor(rank)), position = "stack", stat = "identity", color = "black") +
-  geom_text(aes(x = feature, y = 1-mid, label = brand_cat), data = filter(brand_prop_by_feature_10, prop > .05), size = 3) +
+  geom_bar(aes(x = feature, y = prop, fill = factor(rank)),
+           position = "stack", stat = "identity", color = "black") +
+  geom_text(aes(x = feature, y = 1-mid, label = brand_cat),
+            data = filter(brand_prop_by_feature_10, prop > .05), size = 2.5) +
   scale_fill_brewer(palette = "Paired", guide = F, na.value = "grey") +
@@
 
-\fix{Choose three heatmaps:
+\fix{Choose three (or four) heatmaps:
 \begin{itemize}
 \item An image with multiple true shapes that the model clearly predicts well, to prove that it has learned the features we think it should
+\item (Seychelles circle/text confusion)
 \item An image with high quad/tri/poly confusion, but the model uses appropriate features to make educated guesses
@@
 
-%\fix{Chosen heatmaps are up for debate!}
+\mt{The captions on the old heatmaps are wrong since we switched models. Pls ignore :)}
+
+\mt{I added EER options to heatmaps, but there are still some not ideal bits (it makes the absence of border look strange), so we can decide how to improve it or whether to use it}
 
@@
 
-<<generate-heatmaps, echo = F, eval = F>>=
+<<generate-heatmaps, echo = F, eval = T>>=
 # My computer has troubles with the transparency of the overlay...
@@
 
-calc_heatmap(imgs[which(grepl("chevron-circle", imgs))], loaded_model) %>%
-  prune_heatmap(c("chevron", "quad", "circle")) %>%
-  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
+# # Turn all images in "heatmaps/raw" to top-3 heatmaps
+# sapply(imgs, function(x) {
+#   calc_heatmap(x, loaded_model) %>%
+#     prune_heatmap(top_n = 3, sort = "value") %>%
+#     create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
+# })
+
+calc_heatmap(imgs[which(grepl("Ugg-grey", imgs))], loaded_model) %>%
+  prune_heatmap(top_n = 3, sort = "value") %>%
+  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F,
+                   eer = eers)
+
+calc_heatmap(imgs[which(grepl("text", imgs))], loaded_model) %>%
+  prune_heatmap(top_n = 3, sort = "value") %>%
+  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F,
+                   eer = eers)
+
+calc_heatmap(imgs[which(grepl("polygon", imgs))], loaded_model) %>%
+  prune_heatmap(top_n = 3, sort = "value") %>%
+  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F,
+                   eer = eers)
 
 calc_heatmap(imgs[which(grepl("corks", imgs))], loaded_model) %>%
-  prune_heatmap(c("star", "text", "circle")) %>%
-  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
+  prune_heatmap(top_n = 3, sort = "value") %>%
+  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F,
+                   eer = eers)
+#
+# calc_heatmap(imgs[which(grepl("chevron-circle", imgs))], loaded_model) %>%
+#   #prune_heatmap(c("chevron", "quad", "circle")) %>%
+#   create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
+#
+# calc_heatmap(imgs[which(grepl("triangle", imgs))], loaded_model) %>%
+#   #prune_heatmap(c("triangle", "quad", "chevron")) %>%
+#   create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
+#
+# calc_heatmap(imgs[which(grepl("star", imgs))], loaded_model) %>%
+#   #prune_heatmap(c("star", "bowtie", "polygon")) %>%
+#   create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
+#
+# calc_heatmap(imgs[which(grepl("other", imgs))], loaded_model) %>%
+#   #prune_heatmap(c("star", "chevron", "polygon")) %>%
+#   create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
 
-calc_heatmap(imgs[which(grepl("text", imgs))], loaded_model) %>%
-  prune_heatmap(c("text", "circle", "line")) %>%
-  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
-
-calc_heatmap(imgs[which(grepl("triangle", imgs))], loaded_model) %>%
-  prune_heatmap(c("triangle", "quad", "chevron")) %>%
-  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
-
-calc_heatmap(imgs[which(grepl("star", imgs))], loaded_model) %>%
-  prune_heatmap(c("star", "bowtie", "polygon")) %>%
-  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
-
-calc_heatmap(imgs[which(grepl("polygon", imgs))], loaded_model) %>%
-  prune_heatmap(c("star", "chevron", "polygon")) %>%
-  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
-
-calc_heatmap(imgs[which(grepl("other", imgs))], loaded_model) %>%
-  #prune_heatmap(c("star", "chevron", "polygon")) %>%
-  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
-
-heatmaps <- list.files(imgdir, pattern = "^heatmap")
+heatmaps <- list.files(h_imgdir, pattern = "^heatmap", full.names = T)
 @
 
-<<chevron-circle-heatmap, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing chevrons and circles.", fig.cap = "Chevrons are identified by sharp corners and parallel lines, but the thick line between the two shape sections is falsely identified as a quadrilateral.", fig.pos = "!h">>=
+<<current-heatmaps, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.cap = "Correct or reasonably incorrect heatmaps", echo = F, message = F, warning = F>>=
+knitr::include_graphics(heatmaps[grepl("Ugg-grey", heatmaps)])
+knitr::include_graphics(heatmaps[grepl("text", heatmaps)])
+knitr::include_graphics(heatmaps[grepl("polygon", heatmaps)])
+knitr::include_graphics(heatmaps[grepl("corks", heatmaps)])
+
+sapply(heatmaps, knitr::include_graphics)
+@
+
+<<chevron-circle-heatmap, eval = F, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing chevrons and circles.", fig.cap = "Chevrons are identified by sharp corners and parallel lines, but the thick line between the two shape sections is falsely identified as a quadrilateral.", fig.pos = "!h">>=
 heatmaps <- list.files(file.path(wd, "images", "heatmaps", "static"),
@@
 
-<<text-seychelles-heatmap, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing text.", fig.cap = "This heatmap illustrates that the high probability of a circle is a result of the highly curled tails of the 'S' in the text.">>=
+<<text-seychelles-heatmap, eval = F, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing text.", fig.cap = "This heatmap illustrates that the high probability of a circle is a result of the highly curled tails of the 'S' in the text.">>=
 knitr::include_graphics(heatmaps[grepl("text", heatmaps)])
@@
 
-<<corks-heatmap, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing none of the target classes.", fig.cap = "This outsole shows repeated outlines of corks, which do not fall into any of the nine classes; the concave corners where the corks overlap are predicted to be stars, and the round edges lead to predictions of circles.\\svp{Not really...
+<<corks-heatmap, eval = F, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing none of the target classes.", fig.cap = "This outsole shows repeated outlines of corks, which do not fall into any of the nine classes; the concave corners where the corks overlap are predicted to be stars, and the round edges lead to predictions of circles.\\svp{Not really...
 only 0.03 probability}">>=
@@
 
-<<puzzle-heatmap, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing none of the target classes.", fig.cap = "This outsole shows repeated outlines of puzzle pieces, which do not fall into any of the nine classes.">>=
+<<puzzle-heatmap, eval = F, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing none of the target classes.", fig.cap = "This outsole shows repeated outlines of puzzle pieces, which do not fall into any of the nine classes.">>=
 knitr::include_graphics(heatmaps[grepl("other", heatmaps)])

### 2019-06-20 paper/code/Generate_Model_Images.R

@@ -262,9 +262,26 @@ heatmap_overlay <- function(heatmap, geometry = geometry, width = 256, height = 
     image_resize(geometry, filter = "quadratic")
 }
 
-prune_heatmap <- function(full_heatmap, keep_classes) {
+# Choose specific classes or the number of classes to show
+prune_heatmap <- function(full_heatmap, keep_classes, top_n,
+                          sort = c("name", "value")) {
+
+  if(missing(keep_classes) && missing(top_n)) {
+    warning("No pruning criteria provided.")
+  }
+
+  sort <- match.arg(sort)
   tmp <- full_heatmap
-  idx <- which(tmp$classes %in% keep_classes)
+
+  if(!missing(top_n)) {
+    keep_classes <- names(head(sort(tmp$predictions, decreasing= T), top_n))
+  }
+
+  if(sort == "value") {
+    idx <- sapply(keep_classes, function(x){which(x == tmp$classes)})
+  } else {
+    idx <- which(tmp$classes %in% keep_classes)
+  }
   tmp$heatmap <- tmp$heatmap[idx, , ]
   tmp$predictions <- tmp$predictions[idx]
   tmp$truth <- tmp$truth[idx]
@@ -275,7 +292,7 @@ prune_heatmap <- function(full_heatmap, keep_classes) {
 create_composite <- function(heatmap_data, save_file = F, outdir = ".",
                              td = tempdir(), fixed_labels = T,
                              fail_file = file.path("Images", "poop.jpg"),
-                             color_by_label = T) {
+                             color_by_label = T, eer = NULL) {
 
   # Fix image dimensions
   dim(heatmap_data$img) <- dim(heatmap_data$img)[-1]
@@ -303,7 +320,16 @@ create_composite <- function(heatmap_data, save_file = F, outdir = ".",
   filelist <- rep(blank_img, n_classes)
   unmod_image <- image_composite(blank_img, image, offset = "+22+6") %>%
     # Add label
-    image_annotate("Original Image", color = "black", size = 30, location = "+0+3", gravity = "South")
+    image_annotate("Original Image", color = "black", size = 30,
+                   location = "+0+3", gravity = "South")
+
+  if(!is.null(eer)) {
+    above_eer <- heatmap_data$predictions > eer[heatmap_data$classes]
+  } else{
+    above_eer <- rep(NA, n_classes)
+  }
+
+
 
   for (j in 1:n_classes) {
 
[…]

### 2019-06-21 paper/paper.Rnw
Accepting changes

@@
 \paragraph{Transfer Learning}
-%\mt{Goal: Training a CNN from scratch is hard.
-There are CNNs trained on complex images to detect a wide range of features, and transfer learning lets us use those feature detectors.
-In particular, we're using VGG16.
-We can look at the filters of VGG16 to support our claim that it will be good for the task.}
 Neural networks are composed of sets of layers; the early layers, which are called the model base, contain feature detectors, and the final layers make up the classifier, or model head, which connects meaningful features to classification labels (see \autoref{fig:VGG16-structure} for an example of the structure of a CNN).
@@
 
-%\fix{This paragraph may not be necessary, depending on the level of detail we need to get into.}
-%\mt{I think briefly explaining the pre-training data helps support the idea that the model can learn to classify like we do, but it can definitely be said more concisely.}
 Many convolutional neural networks have been trained on a subset of images from an online image database called ImageNet \citep{dengImageNetLargeScaleHierarchical}.
 In particular, there is a standard set of images from ImageNet, consisting of 1.2 million images and spanning 1,000 catetories, that is commonly used in computer vision tasks; these images provide both a wide variety of images for training complex image-classification models and a standardized way to compare performance across different models.
-\mt{\sout{Images from ImageNet are hierarchically labeled, so an image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity.
-}}CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously.
+CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously.
 As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks.
@@
 It's definitely a chicken and egg problem...}
-
-%\fix{This section repeats bits from the last two paragraphs. Consider reordering things so that there is less duplication? What's the best way to do that coherently?}
-%\mt{Note to Susan: Most of the following paragraph has been Frankensteined into the 'Transfer learning' paragraph. If it sounds okay there, we can delete this paragraph.}
-%Each layer in the model structure is comprised of sets of weights and connections that have been optimized during training. \mt{moved to para 1}
-%Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated \citep{oquabLearningTransferringMidlevel2014}.
-\mt{citation used in para 1}
-%In many cases, the entire model base is used with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with smaller, domain-specific sets of image data.
-\mt{redundant to above, not used}
-%Transfer learning allows CNNs to be applied to smaller datasets of several thousand images, reduces the amount of computational time required to fit the model, and provides boosted performance compared to training a new model from scratch \citep{yosinskiHowTransferableAre2014}.
-It has been successfully applied in automatic classification of medical images \citep{shinDeepConvolutionalNeural2016} as well as in various applications of shoe forensics \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017}.
-\mt{moved to end of para 1}
 
@@
 \paragraph{Annotated Training Data}
-\fix{The first sentence is a repeat from earlier in the paper, so we may want to move directly into the feature set we're using. Thoughts?}
-\mt{\sout{Class characteristics are characteristics which can be used to exclude shoes from a match at a crime scene, but cannot be used for individualized matching because they are shared by many shoes (e.g., all shoes of a specific brand and size).}} \mt{Existing research indicates that a} sufficiently well-defined set of features can be used to separate shoes into make and model categories \citep{grossVariabilitySignificanceClass2013}; the set of features used in that study included circle/oval, crepe, herringbone, hexagon, parallel lines, logo/lettering/numbering, perimeter lugs, star, and other.
+Existing research indicates that a sufficiently well-defined set of features can be used to separate shoes into make and model categories \citep{grossVariabilitySignificanceClass2013}; the set of features used in that study included circle/oval, crepe, herringbone, hexagon, parallel lines, logo/lettering/numbering, perimeter lugs, star, and other.
 After consulting with practitioners, we developed a set of categories suitable for automatic recognition by convolutional neural networks.
@@
 (2013) require contextual spatial information which is not preserved during labeling (for example, lugs are required to be on the perimeter of the shoe).
-\autoref{tab:class-char-examples} shows three examples of each class \mt{in our feature set}.
+\autoref{tab:class-char-examples} shows three examples of each class in our feature set.
 
@@
 The remaining two faces can be convex or straight, and the concave faces may have straight portions, so long as there is a concave region.
-%Using this definition, shapes such as butterflies are included as bowties.
 \item [Chevron] Chevron shapes include repeating parallel lines as well as individual ``v" shapes. They may be angular but can also be curved.
@@
 
-Thousands of outsole images were \mt{obtained} from online shoe retail sites and annotated using LabelMe, a tool for image annotation in computer vision problems \citep{labelme}.
+Thousands of outsole images were obtained from online shoe retail sites and annotated using LabelMe, a tool for image annotation in computer vision problems \citep{labelme}.
 After annotation, the minimum bounding rectangle of the region is identified and the image is cropped to that area; subsequently, the cropped image is scaled to 256 x 256 pixels.
@@
 
-\svp{This section header might need to move? Haven't decided yet...}
 \section{Results}
@@
 The model is able to identify the circle in all cases where a circle is truly present, regardless of whether the circle is part of the text or it is distinct; however, the circle prediction is also strong in three of the four images that do not contain a circle, which indicates that the model is having dificulty distinguishing the two categories.
-\svp{While the circle predictions are above the EER for letters such as C and S, these predictions are generally much lower than those where distinct circles or Os are seen, indicating that there are some indications that features are not as strong in the quasi-circle shapes, but that the model is not able to completely separate Os from C, S, and G shapes which are also very circular.} This confusion is complicated \mt{by} the frequency of images which contain both circles and text: text is present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("circle", dfunion$name)), 2)}\% of our labeled images containing circles, and circles are present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("text", dfunion$name)), 2)}\% of our images containing text.
+\svp{While the circle predictions are above the EER for letters such as C and S, these predictions are generally much lower than those where distinct circles or Os are seen, indicating that there are some indications that features are not as strong in the quasi-circle shapes, but that the model is not able to completely separate Os from C, S, and G shapes which are also very circular.} This confusion is complicated by the frequency of images which contain both circles and text: text is present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("circle", dfunion$name)), 2)}\% of our labeled images containing circles, and circles are present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("text", dfunion$name)), 2)}\% of our images containing text.
 In addition, the distinction between a circle and an ``o" comes from the context around the shape, which is much more difficult for the model to parse than simply detecting a shape.

### 2019-06-21 paper/paper.Rnw
Clean out addressed comments

--- 2019-06-20 paper/paper.Rnw
@@
 Many convolutional neural networks have been trained on a subset of images from an online image database called ImageNet \citep{dengImageNetLargeScaleHierarchical}.
-In particular, there is a standard set of images from ImageNet, consisting of 1.2 million images and spanning 1,000 catetories, that is commonly used in computer vision tasks; these images provide both a wide variety of images for training complex image-classification models and a standardized way to compare performance across different models.
-\mt{\sout{Images from ImageNet are hierarchically labeled, so an image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity.
-}}CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously.
+In particular, there is a standard set of images from ImageNet, consisting of 1.2 million images \mt{that span} 1,000 catetories, that is commonly used in computer vision tasks\mt{\sout{; these images provide both a wide variety of images for training complex image-classification models and a standardized way to compare performance across different models}}.
+CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously.
 As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks.
-\mt{Later in this paper, we will examine a selection of filters from a CNN trained on ImageNet that demonstrates the generalizability of pre-trained models.} \fix{Consider citing the figure number; e.g.
-Figure xxx shows a selection of filters from VGG16 which demonstrate the levels of detail in the initial layers of a model trained on generic data.
-The only issue is that this means figures are cited out of order...
-not sure whether that's kosher.} \mt{I think I used to have that, but then you said I hadn't introduced VGG16 yet so it didn't make sense here (which also totally makes sense).
-It's definitely a chicken and egg problem...}
-
-%\fix{This section repeats bits from the last two paragraphs. Consider reordering things so that there is less duplication? What's the best way to do that coherently?}
-%\mt{Note to Susan: Most of the following paragraph has been Frankensteined into the 'Transfer learning' paragraph. If it sounds okay there, we can delete this paragraph.}
-%Each layer in the model structure is comprised of sets of weights and connections that have been optimized during training. \mt{moved to para 1}
-%Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated \citep{oquabLearningTransferringMidlevel2014}.
-\mt{citation used in para 1}
-%In many cases, the entire model base is used with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with smaller, domain-specific sets of image data.
-\mt{redundant to above, not used}
-%Transfer learning allows CNNs to be applied to smaller datasets of several thousand images, reduces the amount of computational time required to fit the model, and provides boosted performance compared to training a new model from scratch \citep{yosinskiHowTransferableAre2014}.
-It has been successfully applied in automatic classification of medical images \citep{shinDeepConvolutionalNeural2016} as well as in various applications of shoe forensics \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017}.
-\mt{moved to end of para 1}
+In the next section, we will introduce the pre-trained network VGG16, pictured in \autoref{fig:VGG16-structure}, and examine a selection of its filters (shown in \autoref{vgg16layers}) that demonstrate the levels of detail in the initial layers of a model trained on generic data.
 
@@
 
-\fix{I don't really want to get into convolutional layers, but if we're going to mention them we probably need to say something about them. Thoughts?}
-\mt{I only want to give enough detail to briefly explain 1) how filters work as feature detectors and 2) why filters trained on not-shoes will work well on shoes.
-I think that point is made strongly by showing hierarchy in the blocks.
-I thought that the phrasing below gave just enough of a background to mention that filters are structured in layers, but that might raise more questions than it answers---maybe we could gloss over the layer-structure of the blocks and just mention they contain filters?}
-The base of VGG16 is made up of five blocks, which consist of two or three convolutional layers and one max pooling layer which aggregates the signal from the convolutional layers.
-Each convolutional layer is made up of hundreds of filters, such that each of the five convolutional blocks has between 128 and 1,536 filters.
-A filter is a set of numerical weights that, when applied to an image, quantifies the presence of the specific shapes and/or colors that the filter is trained to detect.
+\mt{The base of VGG16 is made up of five convolutional blocks, each of which contain between 128 and 1,536 filters.} A filter is a set of numerical weights that, when applied to an image, quantifies the presence of the specific shapes and/or colors that the filter is trained to detect.
 \autoref{fig:vgg16layers} shows a selection of filters from VGG16, with each row corresponding to one of the five convolutional blocks.
@@
 \paragraph{Annotated Training Data}
-\fix{The first sentence is a repeat from earlier in the paper, so we may want to move directly into the feature set we're using. Thoughts?}
-\mt{\sout{Class characteristics are characteristics which can be used to exclude shoes from a match at a crime scene, but cannot be used for individualized matching because they are shared by many shoes (e.g., all shoes of a specific brand and size).}} \mt{Existing research indicates that a} sufficiently well-defined set of features can be used to separate shoes into make and model categories \citep{grossVariabilitySignificanceClass2013}; the set of features used in that study included circle/oval, crepe, herringbone, hexagon, parallel lines, logo/lettering/numbering, perimeter lugs, star, and other.
+\mt{Existing research indicates that a} sufficiently well-defined set of features can be used to separate shoes into make and model categories \citep{grossVariabilitySignificanceClass2013}; the set of features used in that study included circle/oval, crepe, herringbone, hexagon, parallel lines, logo/lettering/numbering, perimeter lugs, star, and other.
 After consulting with practitioners, we developed a set of categories suitable for automatic recognition by convolutional neural networks.
@@
 (2013) require contextual spatial information which is not preserved during labeling (for example, lugs are required to be on the perimeter of the shoe).
-\autoref{tab:class-char-examples} shows three examples of each class \mt{in our feature set}.
+\autoref{tab:class-char-examples} shows three examples of each class \mt{in the developed feature set}.
 
@@
 \subsection{Model Accuracy}
-%
-% \mt{\textbf{OUTLINE OF RESULTS}
-% \begin{itemize}
-% \item ROC: shows model performance is generally good, gives broad comparison of class performances
-% \item Examples: show what predictions look like, the model isn't perfect, and motivate that some confusions are systematic between specific classes
-% \item Confusion Matrix: show patterns in misclassification, motivate diagnostics to explain the cause of common mistakes
-% \item Two main reasons for error: image contrast, and overlapping features between classes
-% \begin{itemize}
-% \item Low contrast: the model can't see features, which affects training and prediction
-% \item Overlapping features:
-% \begin{itemize}
-% \item Many classes share individual features (e.g., lines, angles, curves), so the model already has to integrate features and learn context (e.g., circle vs o)
-% \item Some classes are also much more variable (see brand plots), which means the model has to widen its definitions of certain shapes, which makes confusion between classes worse
-% \end{itemize}
-% \end{itemize}
-% \item Heatmaps: 1) verify that the model is using category-appropriate features to make decisions and 2) show examples of understandable mistakes
-% \end{itemize}
-% }
 
 \paragraph{Overall Accuracy}
-
 For each image in the test set, the ``true" labels are nine human-assigned labels of 0 or 1 corresponding to the presence or absence of each of the nine shape categories in the image.
@@
 [Expand on specific misclassifications more, after determining which examples to delete.]}
-\fix{We need to make sure that the x axis labels aren't crammed together...
-Overall the examples are good, but I think I might get rid of the all-white DC image, and possibly the one right below it as well.
-We can discuss on Monday.}
+
+\fix{DC moved to heatmap, green block text could still be removed}
+
 <<eers, include = F>>=
@@
 
-% \mt{Side note: I keep futzing with this section.
-I am going to forcibly move on and welcome Susan's feedback, and not allow myself to change any more of the consistency plots.} \svp{Good strategy.
-Biggest feedback is to change the text to actually match the image as it is currently, but I get why you stopped and I'm proud of you for noticing that it was a problem and finding a solution.
-:) Nice job.}
-\mt{As the above section is right now, the text should match the images...
-I had some issues with image sorting, because Windows will put 11 before 2 but not before 02, so maybe Linux has a different behavior that's resorting the images?
-I also cache them to speed up recompiling, so maybe something's up there?}
+% \fix{make sure text matches figure!}
 
@@
 @
-
-%\svp{Pros and cons of confusion matrix here? Maybe we just don't show the actual proportions on the graph? I'd like to discuss this more in person.}
 
@@
 \begin{itemize}
-\item An image with multiple true shapes that the model clearly predicts well, to prove that it has learned the features we think it should
+\item An image with multiple true shapes that the model clearly predicts well, to prove that it has learned the features we think it should (probably an UGG logo)
 \item (Seychelles circle/text confusion)
@@
 \end{itemize}}
-
-\mt{The captions on the old heatmaps are wrong since we switched models. Pls ignore :)}
-
-\mt{I added EER options to heatmaps, but there are still some not ideal bits (it makes the absence of border look strange), so we can decide how to improve it or whether to use it}
-
-<<setup-heatmap>>=
-imgs <- list.files(file.path(wd, "images", "heatmaps", "raw"), full.names = T)
-@
 
@@
 
-# # Turn all images in "heatmaps/raw" to top-3 heatmaps
-# sapply(imgs, function(x) {
-#   calc_heatmap(x, loaded_model) %>%
+# Turn all images in "heatmaps/raw" to top-4 heatmaps
+imgs <- list.files(file.path(wd, "images", "heatmaps", "raw"), full.names = T)
+sapply(imgs, function(x) {
+  calc_heatmap(x, loaded_model) %>%
+    prune_heatmap(top_n = 4, sort = "value") %>%
+    create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
+})
+
+# calc_heatmap(imgs[which(grepl("Ugg-grey", imgs))], loaded_model) %>%
 #   prune_heatmap(top_n = 3, sort = "value") %>%
-#     create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
-# })
-
-calc_heatmap(imgs[which(grepl("Ugg-grey", imgs))], loaded_model) %>%
-  prune_heatmap(top_n = 3, sort = "value") %>%
-  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F,
-                   eer = eers)
-
-calc_heatmap(imgs[which(grepl("text", imgs))], loaded_model) %>%
-  prune_heatmap(top_n = 3, sort = "value") %>%
-  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F,
-                   eer = eers)
-
-calc_heatmap(imgs[which(grepl("polygon", imgs))], loaded_model) %>%
-  prune_heatmap(top_n = 3, sort = "value") %>%
-  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F,
-                   eer = eers)
-
-calc_heatmap(imgs[which(grepl("corks", imgs))], loaded_model) %>%
-  prune_heatmap(top_n = 3, sort = "value") %>%
-  create_composite(save_file = T, outdir = h_imgdir, color_by_label = F,
-                   eer = eers)
+#   create_composite(save_file = T, outdir = h_imgdir, color_by_label = F,
+#                    eer = eers)
+#
+# calc_heatmap(imgs[which(grepl("text", imgs))], loaded_model) %>%
+#   prune_heatmap(top_n = 3, sort = "value") %>%
+#   create_composite(save_file = T, outdir = h_imgdir, color_by_label = F,
+#                    eer = eers)
+#
+# calc_heatmap(imgs[which(grepl("polygon", imgs))], loaded_model) %>%
+#   prune_heatmap(top_n = 3, sort = "value") %>%
+#   create_composite(save_file = T, outdir = h_imgdir, color_by_label = F,
+#                    eer = eers)
+#
+# calc_heatmap(imgs[which(grepl("corks", imgs))], loaded_model) %>%
+#   prune_heatmap(top_n = 3, sort = "value") %>%
+#   create_composite(save_file = T, outdir = h_imgdir, color_by_label = F,
+#                    eer = eers)
 #

### 2019-06-21 paper/paper.Rnw

@@
 
-%\fix{This paragraph may not be necessary, depending on the level of detail we need to get into.}
-%\mt{I think briefly explaining the pre-training data helps support the idea that the model can learn to classify like we do, but it can definitely be said more concisely.}
 Many convolutional neural networks have been trained on a subset of images from an online image database called ImageNet \citep{dengImageNetLargeScaleHierarchical}.
@@
 \paragraph{Annotated Training Data}
-\mt{Existing research indicates that a} sufficiently well-defined set of features can be used to separate shoes into make and model categories \citep{grossVariabilitySignificanceClass2013}; the set of features used in that study included circle/oval, crepe, herringbone, hexagon, parallel lines, logo/lettering/numbering, perimeter lugs, star, and other.
+Existing research indicates that a sufficiently well-defined set of features can be used to separate shoes into make and model categories \citep{grossVariabilitySignificanceClass2013}; the set of features used in that study included circle/oval, crepe, herringbone, hexagon, parallel lines, logo/lettering/numbering, perimeter lugs, star, and other.
 After consulting with practitioners, we developed a set of categories suitable for automatic recognition by convolutional neural networks.
@@
 (2013) require contextual spatial information which is not preserved during labeling (for example, lugs are required to be on the perimeter of the shoe).
-\autoref{tab:class-char-examples} shows three examples of each class \mt{in the developed feature set}.
+\autoref{tab:class-char-examples} shows three examples of each class in \mt{the modified} feature set.
 

### 2019-06-21 paper/paper.Rnw
Clean out more comments

@@
 Again, predictions are largely accurate, with most of the prominent shapes correctly identified above EER.
-\fix{These are mostly new examples from our last meeting because many of the previous images were in the training set...
-Grrrrr.....
-I have two or three too many, not sure about color contrast within images, etc., feel free to delete a few or we can discuss them.
-[Expand on specific misclassifications more, after determining which examples to delete.]}
+\fix{Expand on specific misclassifications more, after determining which examples to delete.}
 
@@
 \autoref{fig:circle-text-consistency} shows a number of images containing circles and/or text.
-\fix{this isn't accurate to what I'm seeing...} \fix{The top four rows show text contained within a separate, distinct circle shape, and the next five rows show circles represented as the letter ``o" in text.
+\fix{The top four rows show text contained within a separate, distinct circle shape, and the next five rows show circles represented as the letter ``o" in text.
 The remaining images contain only one of either text or circles, as well as other possible shape categories.} In particular, the first two of these images contain text that use the capital letter ``G", the next two images contain no circular shapes, and the final \fix{N} images contain pure circles that do not appear to belong to text.
 The model is able to identify the circle in all cases where a circle is truly present, regardless of whether the circle is part of the text or it is distinct; however, the circle prediction is also strong in three of the four images that do not contain a circle, which indicates that the model is having dificulty distinguishing the two categories.
-\svp{While the circle predictions are above the EER for letters such as C and S, these predictions are generally much lower than those where distinct circles or Os are seen, indicating that there are some indications that features are not as strong in the quasi-circle shapes, but that the model is not able to completely separate Os from C, S, and G shapes which are also very circular.} This confusion is complicated by the frequency of images which contain both circles and text: text is present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("circle", dfunion$name)), 2)}\% of our labeled images containing circles, and circles are present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("text", dfunion$name)), 2)}\% of our images containing text.
+While the circle predictions are above the EER for letters such as C and S, these predictions are generally much lower than those where distinct circles or Os are seen, indicating that there are some indications that features are not as strong in the quasi-circle shapes, but that the model is not able to completely separate Os from C, S, and G shapes which are also very circular.
+This confusion is complicated by the frequency of images which contain both circles and text: text is present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("circle", dfunion$name)), 2)}\% of our labeled images containing circles, and circles are present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("text", dfunion$name)), 2)}\% of our images containing text.
 In addition, the distinction between a circle and an ``o" comes from the context around the shape, which is much more difficult for the model to parse than simply detecting a shape.
@@
 Many of the misclassifications in these examples, however, are not just seen in individual images, but rather exist systematically between certain classes.
-% Nicely phrased!!!
-I couldn't have done it better myself.
 A confusion matrix is a type of visualization that summarizes the relationships between the labels assigned to an image and the labels predicted by the model across all test data, and is a helpful diagnostic tool to understand which classes are systematically misclassified.
 The confusion matrix for the fitted model is presented in \autoref{fig:confusion-matrix}.
-%mt{Each value in the matrix represents the probability of assigning the label associated with that value's corresponding row, conditional upon the image truly belonging to the category associated with the value's column.
-Thus, }
 The values along the diagonal of the matrix represent the proportion of true positives captured within each category.
@@
 
-%The confusion between text and circles is a natural consequence of the overlap between geometric shapes and components of written text; it is not indicative of a lack of fit as much as a weakness in the classification system used to define the model.
-
 \paragraph{Feature Overlap}
-%\paragraph{Brand Investigations}
 We have defined our categories through consultation with practitioners and modifications of previous research, but as a result, there are a number of features (such as lines, angles, and curves) which are not unique to only one category.
@@
 \begin{itemize}
-\item a triangle may be made up of three acute angles, but a quad can also contain three acute angles.
+\item a triangle may be made up of three acute angles, but a quadrilateral can also contain three acute angles.
 \item the only distinction between an ``o" and a circle is whether there is more text around it.
@@
 @
-
-% \mt{Note to self: I feel like I want the heatmaps concentrated in one spot, and it would feel like a strange interjection to explain them earlier.
-However, I mostly see including them as a way to support the claims we've been making throughout the paper, so I could definitely imagine an argument for including them earlier for specific examples so that saying ``the model sees..." is less speculatory...}
 

### 2019-06-22 paper/code/Generate_Model_Images.R
Clean up embarrassing code

@@ -325,11 +325,9 @@ create_composite <- function(heatmap_data, save_file = F, outdir = ".",
 
   if(!is.null(eer)) {
     above_eer <- heatmap_data$predictions > eer[heatmap_data$classes]
-  } else{
+  } else {
     above_eer <- rep(NA, n_classes)
   }
-
-
 
   for (j in 1:n_classes) {
 
@@ -353,11 +351,6 @@ create_composite <- function(heatmap_data, save_file = F, outdir = ".",
                                width = 14, height = 14, bg = NA, col = pal_col)
 
     if (j %in% heatmap_data$successful_heatmap) {
-      if(is.na(above_eer[j])) {
-        col_j <- "white"
-      } else {
-        col_j <- ifelse(above_eer[j], "navyblue", "gray90")
-      }
 
       # Create blank image with correct bg color
       filelist[j] <- image_blank(300, 300, color = bg_col) %>%
@@ -367,8 +360,13 @@ create_composite <- function(heatmap_data, save_file = F, outdir = ".",
         # add image
         image_composite(image, offset = "+22+6") %>%
         # add overlay
-        image_composite(overlay, offset = "+22+6") %>%
-        image_border(color = col_j, geometry = "2x2")
+        image_composite(overlay, offset = "+22+6")
+
+      if(!is.na(above_eer[j])) {
+        filelist[j] <- filelist[j] %>%
+          image_border(color = ifelse(above_eer[j], "navyblue", "gray90"),
+                       geometry = "2x2")
+      }
 
 
     } else {

### 2019-06-24 paper/paper.Rnw
Take a stab at adding methods to introduction (with more questions than answers...)

@@
 
-In this paper, we discuss an alternate approach that uses an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images.
+In this paper, we discuss an alternate approach \mt{using the base of a pre-trained convolutional neural network and training a} an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images.
 Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain.
@@
 
+%\mt{By labeling a set of images within a feature space similar to that used by examiners, then utilizing the generalizability of pre-trained CNN feature detectors to train a model to predict the presence of these features in outsole images, }
+
 \section{Materials and Methods}
 \fix{We may want to mention some of the methods in the introduction for continuity sake...}
+\mt{I'm not sure if adding a sentence to the above paragraph accomplished this goal.
+Do you think I need to write a new transition paragraph instead?
+I tried doing that (see commented out stuff), but wasn't sure how to avoid redundancy with the methods section...}
+
 Ultimately, most pattern evidence is stored in photographic form; as a result, we must use image analysis methods to create numerical features from the original visual record.
@@
 
-<<generate-heatmaps, echo = F, eval = T>>=
+<<generate-heatmaps, echo = F, eval = F>>=
 # My computer has troubles with the transparency of the overlay...
@@
     prune_heatmap(top_n = 4, sort = "value") %>%
-    create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
+    create_composite(save_file = T, outdir = h_imgdir,
+                     color_by_label = F, eer = eers)
 })
@@
 
-heatmaps <- list.files(h_imgdir, pattern = "^heatmap", full.names = T)
 @
 
-<<current-heatmaps, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.cap = "Correct or reasonably incorrect heatmaps", echo = F, message = F, warning = F>>=
+<<current-heatmaps, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.cap = "Correct or reasonably incorrect heatmaps", echo = F, message = F, warning = F, fig.pos = "!p">>=
+
+h_imgdir <- file.path(imgdir, "heatmaps", "processed")
+heatmaps <- list.files(h_imgdir, pattern = "^heatmap", full.names = T)
+
 knitr::include_graphics(heatmaps[grepl("Ugg-grey", heatmaps)])
@@
 
-sapply(heatmaps, knitr::include_graphics)
+#sapply(heatmaps, knitr::include_graphics)
 @

### 2019-06-24 paper/paper.Rnw
Attempt at conclusion

@@
 In these heatmaps, we can confirm that multiple classes have common features, resulting in misclassifications.
-\fix{Add EER border to heatmaps}
 
@@
 \svp{Rework this to address 3 main conclusions - 1: NNs work for automatic feature ID, 2: Different levels of feature variability within classes, and 3: correcting for contrast is important.}
+
+% Restate goal
 The goal of this research was to develop a method to automatically identify geometric class characteristics of shoe outsoles.
@@
 
-Our model was developed using the convolutional base of VGG16 and a newly trained classifier/head to identify the defined set of geometric features.
+% Briefly summarize methods
+% Overview of results: model generally performs well
+The model was developed using the \mt{pre-trained} convolutional base of VGG16 and a newly trained classifier/head to output predicted probabilities within the defined set of geometric features.
 After training, the model performs well on the data provided.
@@
 
-\mt{
-\begin{itemize}
-\item Many model errors are a result of overlapping geometric features between classes.
-The variability within certain classes exacerbates the difficulty in defining categories, leading to frequent confusion between circle/text and triangle/quad/polygon \svp{This is at least in part due to the features used by practitioners, and partially due to our generalization/operational definition.}
-\item Model predictions also indicate that contrast correction is important, and model should be retrained on cc'd images
-\item Mention cutting off conv block? \svp{Sure, just do it in friendly terms}
-\end{itemize}
-}
+% Discussion of weaknesses
+\fix{Although model predictions are overall quite good, there are a number of \fix{issues}.
+Firstly, the model has substantial difficulty predicting images with low contrast because image features are obscured; exploring methods of color correction, such as histogram normalization, may prove useful for eliminating the effect of contrast on predictions.
+
+A larger difficulty is the inherent overlap between features of common geometric shapes, and high variability of those shapes on outsoles.
+Overlapping features between shapes makes it more difficult for the model to distinguish between geometrically similar shapes, such as chevrons and triangles.
+When the physical shapes present within a category are highly variable, this further widens the sets of features associated with individual categories, leading to even more model confusion.
+These issues of overlap and variabailty are at least in part due to the features used by practitioners, and partially due to our generalization/operational definition.}
+
+\mt{What is the balance between comprehensive and redundant? I don't think I've found it... I'll try again when not exhausted}
+
+\mt{Final paragraph will suggest redefining shapes (yuck) and cutting off the final conv block to see whether those features are easier for the model to sort between categories.}
 

### 2019-06-24 paper/paper.Rnw
Consolidate conclusion

@@
 
-% Restate goal
+% Restate goal and briefly summarize methods
 The goal of this research was to develop a method to automatically identify geometric class characteristics of shoe outsoles.
+A set of geometric class characteristics was defined, based on feature sets currently used by examiners, to both broadly classify a large variety of shoes and to narrow down similarity into a manageable number and type of features for further use in a shoeprint analysis.
+Thousands of outsole images were obtained from online shoe retailers and labeled according to their geometric shapes.
+A model, the \mt{pre-trained} convolutional base of VGG16 and a newly trained classifier/head, was then trained on a portion of these labeled images to output predicted probabilities within the defined set of geometric features.
+After training, the model performs well on the data provided.
+In general, the model is able to identify many well-defined geometric shapes in the images.
+
+%The goal of this research was to develop a method to automatically identify geometric class characteristics of shoe outsoles.
 Such a method takes the difficult problem of quantifying outsole features in a given population and lays the groundwork for a tractable solution.
-A set of geometric class characteristics was defined, based on feature sets currently used by examiners, to both broadly classify a large variety of shoes and to narrow down similarity into a manageable number and type of features for further use in a shoeprint analysis.
-The final set of features used in this research achieves both goals quite well, and has potential for much finer-grained analysis when recognized features are combined with spatial information and relationships.
+\mt{Are we alluding to the scanner in this paper, or should I change that to focus on automating what examiners are doing now?} A set of geometric class characteristics was defined, based on feature sets currently used by examiners, to both broadly classify a large variety of shoes and to narrow down similarity into a manageable number and type of features for further use in a shoeprint analysis.
+\fix{I think our conclusions of overlap make this statement too rosy: The final set of features used in this research achieves both goals quite well, and has potential for much finer-grained analysis when recognized features are combined with spatial information and relationships.}
 
@@
 % Overview of results: model generally performs well
-The model was developed using the \mt{pre-trained} convolutional base of VGG16 and a newly trained classifier/head to output predicted probabilities within the defined set of geometric features.
+%The model was developed using the \mt{pre-trained} convolutional base of VGG16 and a newly trained classifier/head to output predicted probabilities within the defined set of geometric features.
 After training, the model performs well on the data provided.

### 2019-06-24 paper/paper.Rnw
Another stab at concluding

@@
 
-A larger difficulty is the inherent overlap between features of common geometric shapes, and high variability of those shapes on outsoles.
-Overlapping features between shapes makes it more difficult for the model to distinguish between geometrically similar shapes, such as chevrons and triangles.
-When the physical shapes present within a category are highly variable, this further widens the sets of features associated with individual categories, leading to even more model confusion.
-These issues of overlap and variabailty are at least in part due to the features used by practitioners, and partially due to our generalization/operational definition.}
-
-\mt{What is the balance between comprehensive and redundant? I don't think I've found it... I'll try again when not exhausted}
-
-\mt{Final paragraph will suggest redefining shapes (yuck) and cutting off the final conv block to see whether those features are easier for the model to sort between categories.}
+A larger difficulty is the inherent overlap between features of common geometric shapes; since CNNs make predictions via bottom-up feature integration, it is more difficult for the model to distinguish between geometrically similar shapes, such as chevrons and triangles.
+This issue is exacerbated by high variability of those shapes on outsoles, which requires the model to form a flexible understanding of  each shape category, but makes distinction even more difficult.
+These issues of overlap and variabailty are at least in part due to the features used by practitioners, and partially due to the operational definitions of each class used in this research.
+While a \fix{full} remedy may not be possible, there may be some modifications to operational class definitions that we can make to optimize labels for model prediction.}
 

### 2019-06-24 paper/paper.Rnw
minor edits

--- 2019-06-24 paper/paper.Rnw
@@
 <<setup, include = F>>=
-knitr::opts_chunk$set(echo = F, dev = "pdf", dpi = 1000, message = F, warning = F, fig.align = "center")
+knitr::opts_chunk$set(echo = F, dev = "pdf", dpi = 1000, message = F, warning = F, fig.align = "center", fig.path = "figure/")
 
@@
 
-In this paper, we discuss an alternate approach \mt{using the base of a pre-trained convolutional neural network and training a} an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images.
+In this paper, we discuss an alternate approach that uses an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images.
 Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain.
@@
 Do you think I need to write a new transition paragraph instead?
-I tried doing that (see commented out stuff), but wasn't sure how to avoid redundancy with the methods section...}
+I tried doing that (see commented out stuff), but wasn't sure how to avoid redundancy with the methods section...} \svp{I was talking about the methods above (e.g.
+fractal decomp, scale invariant feature recognition, etc., not methods we used)}
 
@@
 Neural networks are composed of sets of layers; the early layers, which are called the model base, contain feature detectors, and the final layers make up the classifier, or model head, which connects meaningful features to classification labels (see \autoref{fig:VGG16-structure} for an example of the structure of a CNN).
-\mt{Each of these layers is comprised of sets of weights and connections, or parameters, that are optimized during training.} A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power.
+Each of these layers is \svp{composed} of sets of weights and \svp{connections that} are optimized during training.
+A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power.
 Assembling sufficient data to train a network from scratch is a gargantuan task; the process of \emph{transfer learning} is a natural solution to this common problem.
 Transfer learning leverages the modularity of neural networks, that is, that the model base that detects features can be separated from the classifier which produces predictions.
-Transfer learning approaches use the base of a model trained on more general image data and train a new model head or classifier using a smaller set of domain-specific labeled images \mt{\citep{oquabLearningTransferringMidlevel2014}, allowing CNNs to be applied to smaller datasets of several thousand images, reduces the amount of computational time required to fit the model, and provides boosted performance compared to training a new model from scratch \citep{yosinskiHowTransferableAre2014}.
-It has been successfully applied in automatic classification of medical images \citep{shinDeepConvolutionalNeural2016} as well as in various applications of shoe forensics \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017}.}
+Transfer learning approaches use the base of a model trained on more general image data and train a new model head or classifier using a smaller set of domain-specific labeled images\citep{oquabLearningTransferringMidlevel2014}.
+\svp{This allows} CNNs to be applied to smaller datasets of several thousand images, reduces the amount of computational time required to fit the model, and provides boosted performance compared to training a new model from scratch \citep{yosinskiHowTransferableAre2014}.
+Transfer learning has been successfully applied in automatic classification of medical images \citep{shinDeepConvolutionalNeural2016} as well as in various applications of shoe forensics \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017}.
 
 Many convolutional neural networks have been trained on a subset of images from an online image database called ImageNet \citep{dengImageNetLargeScaleHierarchical}.
-In particular, there is a standard set of images from ImageNet, consisting of 1.2 million images \mt{that span} 1,000 catetories, that is commonly used in computer vision tasks\mt{\sout{; these images provide both a wide variety of images for training complex image-classification models and a standardized way to compare performance across different models}}.
+In particular, there is a standard set of images from ImageNet, consisting of 1.2 million images \mt{that span} 1,000 catetories, that is commonly used in computer vision tasks.
 CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously.
 As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks.
-In the next section, we will introduce the pre-trained network VGG16, pictured in \autoref{fig:VGG16-structure}, and examine a selection of its filters (shown in \autoref{vgg16layers}) that demonstrate the levels of detail in the initial layers of a model trained on generic data.
+In the next section, we will introduce the pre-trained network VGG16, pictured in \autoref{fig:VGG16-structure}, and examine a selection of its filters (shown in \autoref{fig:vgg16layers}) that demonstrate the levels of detail in the initial layers of a model trained on generic data.
 
@@
 \citep{krizhevskyImageNetClassificationDeep2012}.
-In contrast to other popular \mt{pre-trained} networks, like ResNet, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy.
+In contrast to other popular pre-trained networks, \svp{such as AlexNet or ResNet}, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy.
 The simplicity of this structure provides the ability to peer into the inner workings of the network for diagnostic purposes, providing a distinct advantage over more complicated network structures with slightly higher accuracy ratings.
 VGG16 is a common choice for transfer learning because of this structural simplicity; it has been used for detection of text in natural images \citep{zhongDeepTextUnifiedFramework2016}, medical imaging classification \citep{oquabLearningTransferringMidlevel2014}, classification of weld defects \citep{liuWeldDefectImages2018}, and many other domain-specific image recognition tasks that are more specific than the ImageNet data on which it was trained.
-\autoref{fig:VGG16-structure} shows the architecture of VGG16; with transfer learning, the VGG16 model head is replaced with a trained model head specific to recognition of shoeprint class characteristics.
+\autoref{fig:VGG16-structure} shows the architecture of VGG16; with transfer learning, the VGG16 model head is replaced with a model head \svp{trained specifically to recognize} shoeprint class characteristics.
 
-\mt{The base of VGG16 is made up of five convolutional blocks, each of which contain between 128 and 1,536 filters.} A filter is a set of numerical weights that, when applied to an image, quantifies the presence of the specific shapes and/or colors that the filter is trained to detect.
-\autoref{fig:vgg16layers} shows a selection of filters from VGG16, with each row corresponding to one of the five convolutional blocks.
+The base of VGG16 is made up of five convolutional blocks, each of which contain between 128 and 1,536 filters.
+A filter is a set of numerical weights that, when applied to an image, quantifies the presence of the specific shapes and/or colors that the filter is trained to detect.
+\autoref{fig:vgg16layers} shows a selection of \svp{images which maximally activate specific} filters from VGG16, % Note - they aren't actually the filters themselves, these are generated images which maximize activation of the filter in question.
+with each row corresponding to one of the five convolutional blocks.
 Filters in earlier convolutional blocks detect simple features, such as colors, lines, corners, and blobs, while later filters detect more complex combinations of features.
@@
 We can test these hypotheses by examining what the model ``sees" when making its predictions, using heatmaps, or class activation maps (CAMs).
-\fix{Heatmaps or Class Activation Maps} are a visual diagnostic tool that highlights the areas of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}.
+Heatmaps, \svp{also known as} Class Activation Maps\svp{,} are a visual diagnostic tool that highlights the areas of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}.
 Using these graphics, we can identify which locations in an image contribute most to the model's predictions.
@@
 #
-# calc_heatmap(imgs[which(grepl("polygon", imgs))], loaded_model) %>%
+# calc_heatmap(imgs[which(grepl("polygon_recolor", imgs))], loaded_model) %>%
 #   prune_heatmap(top_n = 3, sort = "value") %>%
@@
 
-<<current-heatmaps, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.cap = "Correct or reasonably incorrect heatmaps", echo = F, message = F, warning = F, fig.pos = "!p">>=
+<<current-heatmaps, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.cap = c("A section of an Ugg shoe containing the logo.
+The model correctly identifies triangles and text (and references the correct portion of the image in both cases).
+The model identifies a circle as well; while it could be argued that the triangles form an implicit circle, the heatmap demonstrates that the circular portion of the image is the 'G', which would not be a closed figure labeled as a circle.
+The model's prediction of the presence of a circle is understandable, but incorrect.", "In this image of a Seychelles shoe, the model identifies text and circle.
+The prediction of text is correct; the prediction of circle is not - there are very round shapes, but they do not form a closed figure that could be identified as a circle.", "polygon", "corks"), echo = F, message = F, warning = F, fig.pos = "!p">>=
 
@@
 knitr::include_graphics(heatmaps[grepl("text", heatmaps)])
-knitr::include_graphics(heatmaps[grepl("polygon", heatmaps)])
+knitr::include_graphics(heatmaps[grepl("polygon_recolor", heatmaps)])
 knitr::include_graphics(heatmaps[grepl("corks", heatmaps)])
@@
 % Discussion of weaknesses
-\fix{Although model predictions are overall quite good, there are a number of \fix{issues}.
-Firstly, the model has substantial difficulty predicting images with low contrast because image features are obscured; exploring methods of color correction, such as histogram normalization, may prove useful for eliminating the effect of contrast on predictions.
+\fix{Although model predictions are} \svp{generally accurate}, there are a number of \svp{systematic misclassifications}.
+The model has substantial difficulty predicting images with low contrast because image features are obscured.
+Exploring methods of color correction, such as histogram equalization, may prove useful for eliminating the effect of contrast on predictions.
 
-A larger difficulty is the inherent overlap between features of common geometric shapes, and high variability of those shapes on outsoles.
+\svp{Another source of systematic misclassifications stems from the} inherent overlap between features of common geometric shapes, \svp{in combination with} high variability of those shapes on outsoles.
 Overlapping features between shapes makes it more difficult for the model to distinguish between geometrically similar shapes, such as chevrons and triangles.
 When the physical shapes present within a category are highly variable, this further widens the sets of features associated with individual categories, leading to even more model confusion.
-These issues of overlap and variabailty are at least in part due to the features used by practitioners, and partially due to our generalization/operational definition.}
+These issues of overlap and variability are at least in part due to \svp{our decision to use} the features used by practitioners \svp{directly}; \svp{some of this may also result from our operational definitions of various shapes, which did not take into account the model's capabilities and initial training}.
 
-\mt{What is the balance between comprehensive and redundant? I don't think I've found it... I'll try again when not exhausted}
+% \mt{What is the balance between comprehensive and redundant?
+I don't think I've found it...
+I'll try again when not exhausted} I'll let you know if I ever find it, but this looks pretty great - I've made a few tiny tweaks to phrasing, but the content is exactly what I was hoping for :)
 
-\mt{Final paragraph will suggest redefining shapes (yuck) and cutting off the final conv block to see whether those features are easier for the model to sort between categories.}
+\mt{Final paragraph will suggest redefining shapes (yuck) and cutting off the final conv block to see whether those features are easier for the model to sort between categories.} \svp{Sounds good!}
 

### 2019-06-24 paper/paper.Rnw
comment the merge issues out

@@
 
+% <<<<<<< HEAD
 \svp{Another source of systematic misclassifications stems from the} inherent overlap between features of common geometric shapes, \svp{in combination with} high variability of those shapes on outsoles.
@@
 
-% \mt{What is the balance between comprehensive and redundant?
-I don't think I've found it...
-I'll try again when not exhausted} I'll let you know if I ever find it, but this looks pretty great - I've made a few tiny tweaks to phrasing, but the content is exactly what I was hoping for :)
-
-\mt{Final paragraph will suggest redefining shapes (yuck) and cutting off the final conv block to see whether those features are easier for the model to sort between categories.} \svp{Sounds good!}
+% =======
+A larger difficulty is the inherent overlap between features of common geometric shapes; since CNNs make predictions via bottom-up feature integration, it is more difficult for the model to distinguish between geometrically similar shapes, such as chevrons and triangles.
+This issue is exacerbated by high variability of those shapes on outsoles, which requires the model to form a flexible understanding of  each shape category, but makes distinction even more difficult.
+These issues of overlap and variabailty are at least in part due to the features used by practitioners, and partially due to the operational definitions of each class used in this research.
+While a \fix{full} remedy may not be possible, there may be some modifications to operational class definitions that we can make to optimize labels for model prediction.
+% >>>>>>> 30ea1b70f71998a523620fbe35230f47b7dc61fd
 

### 2019-06-24 paper/paper.Rnw
Fix methods mixup, clean up existing conclusion

--- 2019-06-24 paper/paper.Rnw
@@
 
-%\mt{By labeling a set of images within a feature space similar to that used by examiners, then utilizing the generalizability of pre-trained CNN feature detectors to train a model to predict the presence of these features in outsole images, }
-
 \section{Materials and Methods}
-\fix{We may want to mention some of the methods in the introduction for continuity sake...}
-\mt{I'm not sure if adding a sentence to the above paragraph accomplished this goal.
-Do you think I need to write a new transition paragraph instead?
-I tried doing that (see commented out stuff), but wasn't sure how to avoid redundancy with the methods section...} \svp{I was talking about the methods above (e.g.
-fractal decomp, scale invariant feature recognition, etc., not methods we used)}
-
+%\fix{We may want to mention some of the methods in the introduction for continuity sake...}
+%\svp{I was talking about the methods above (e.g. fractal decomp, scale invariant feature recognition, etc., not methods we used)}
 Ultimately, most pattern evidence is stored in photographic form; as a result, we must use image analysis methods to create numerical features from the original visual record.
-There are a number of methods that may be employed to identify shapes and features in an image, such as Fourier transforms, Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}.
+There are a number of methods that may be employed to identify shapes and features in an image, such as Fourier\mt{-Mellin} transforms, \mt{fractal decomposition, scale-invariant feature recognition}, Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}.
 While these methods are useful in identifying these specific features at a low level, they can be computationally intensive and only identify features on a very small scale; as a result, they cannot reliably identify large geometric shapes like those that may be found in a high resolution outsole image.
@@
 Neural networks are composed of sets of layers; the early layers, which are called the model base, contain feature detectors, and the final layers make up the classifier, or model head, which connects meaningful features to classification labels (see \autoref{fig:VGG16-structure} for an example of the structure of a CNN).
-Each of these layers is \svp{composed} of sets of weights and \svp{connections that} are optimized during training.
+Each of these layers is composed of sets of weights and connections that are optimized during training.
 A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power.
@@
 Transfer learning approaches use the base of a model trained on more general image data and train a new model head or classifier using a smaller set of domain-specific labeled images\citep{oquabLearningTransferringMidlevel2014}.
-\svp{This allows} CNNs to be applied to smaller datasets of several thousand images, reduces the amount of computational time required to fit the model, and provides boosted performance compared to training a new model from scratch \citep{yosinskiHowTransferableAre2014}.
+This allows CNNs to be applied to smaller datasets of several thousand images, reduces the amount of computational time required to fit the model, and provides boosted performance compared to training a new model from scratch \citep{yosinskiHowTransferableAre2014}.
 Transfer learning has been successfully applied in automatic classification of medical images \citep{shinDeepConvolutionalNeural2016} as well as in various applications of shoe forensics \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017}.
@@
 \citep{krizhevskyImageNetClassificationDeep2012}.
-In contrast to other popular pre-trained networks, \svp{such as AlexNet or ResNet}, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy.
+In contrast to other popular pre-trained networks, such as AlexNet or ResNet, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy.
 The simplicity of this structure provides the ability to peer into the inner workings of the network for diagnostic purposes, providing a distinct advantage over more complicated network structures with slightly higher accuracy ratings.
 VGG16 is a common choice for transfer learning because of this structural simplicity; it has been used for detection of text in natural images \citep{zhongDeepTextUnifiedFramework2016}, medical imaging classification \citep{oquabLearningTransferringMidlevel2014}, classification of weld defects \citep{liuWeldDefectImages2018}, and many other domain-specific image recognition tasks that are more specific than the ImageNet data on which it was trained.
-\autoref{fig:VGG16-structure} shows the architecture of VGG16; with transfer learning, the VGG16 model head is replaced with a model head \svp{trained specifically to recognize} shoeprint class characteristics.
+\autoref{fig:VGG16-structure} shows the architecture of VGG16; with transfer learning, the VGG16 model head is replaced with a model head trained specifically to recognize shoeprint class characteristics.
 
@@
 A filter is a set of numerical weights that, when applied to an image, quantifies the presence of the specific shapes and/or colors that the filter is trained to detect.
-\autoref{fig:vgg16layers} shows a selection of \svp{images which maximally activate specific} filters from VGG16, % Note - they aren't actually the filters themselves, these are generated images which maximize activation of the filter in question.
-with each row corresponding to one of the five convolutional blocks.
+\autoref{fig:vgg16layers} shows a selection of images which maximally activate specific} filters from VGG16, with each row corresponding to one of the five convolutional blocks.
 Filters in earlier convolutional blocks detect simple features, such as colors, lines, corners, and blobs, while later filters detect more complex combinations of features.
@@
 
-<<exemplars, fig.cap = "Examples containing a single shape", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 8, fig.height = 12>>=
+<<exemplars, fig.cap = "Examples containing a single shape", fig.pos = "p!", out.width = "100%", cache = F, fig.width = 8, fig.height = 12>>=
 single_exemplars <- list.files(file.path(wd, "images", "consistency",
@@
 We have proposed several hypotheses to explain why the model's predictions are not entirely consistent with manual labels on the images.
-We can test these hypotheses by examining what the model ``sees" when making its predictions, using heatmaps, or class activation maps (CAMs).
-Heatmaps, \svp{also known as} Class Activation Maps\svp{,} are a visual diagnostic tool that highlights the areas of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}.
+We can test these hypotheses by examining what the model ``sees" when making its predictions \mt{\sout{using heatmaps, or class activation maps (CAMs)}}.
+Heatmaps, also known as Class Activation Maps \mt{(CAMs)}, are a visual diagnostic tool that highlights the areas of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}.
 Using these graphics, we can identify which locations in an image contribute most to the model's predictions.
-In \fix{the figures in this subsection,} heatmaps are shown for a selection of three classes, which are either contained in the image or have high output probabilities, alongside the original image.
+In figures \fix{X-Y}, heatmaps are shown for a selection of \fix{three} classes, which are either contained in the image or have high output probabilities, alongside the original image.
 In these heatmaps, we can confirm that multiple classes have common features, resulting in misclassifications.
@@
 
-% Restate goal
+% Restate goal and briefly summarize methods
 The goal of this research was to develop a method to automatically identify geometric class characteristics of shoe outsoles.
-Such a method takes the difficult problem of quantifying outsole features in a given population and lays the groundwork for a tractable solution.
 A set of geometric class characteristics was defined, based on feature sets currently used by examiners, to both broadly classify a large variety of shoes and to narrow down similarity into a manageable number and type of features for further use in a shoeprint analysis.
-The final set of features used in this research achieves both goals quite well, and has potential for much finer-grained analysis when recognized features are combined with spatial information and relationships.
-
-% Briefly summarize methods
-% Overview of results: model generally performs well
-The model was developed using the \mt{pre-trained} convolutional base of VGG16 and a newly trained classifier/head to output predicted probabilities within the defined set of geometric features.
+Thousands of outsole images were obtained from online shoe retailers and labeled according to their geometric shapes.
+The pre-trained convolutional base of VGG16 was then used, with a new classifier trained on a portion of these labeled images, to output predicted probabilities for test images within the defined set of geometric features.
 After training, the model performs well on the data provided.
@@
 % Discussion of weaknesses
-\fix{Although model predictions are} \svp{generally accurate}, there are a number of \svp{systematic misclassifications}.
+Although model predictions are generally accurate, there are a number of systematic misclassifications.
 The model has substantial difficulty predicting images with low contrast because image features are obscured.
@@
 
-\svp{Another source of systematic misclassifications stems from the} inherent overlap between features of common geometric shapes, \svp{in combination with} high variability of those shapes on outsoles.
-Overlapping features between shapes makes it more difficult for the model to distinguish between geometrically similar shapes, such as chevrons and triangles.
-When the physical shapes present within a category are highly variable, this further widens the sets of features associated with individual categories, leading to even more model confusion.
-These issues of overlap and variability are at least in part due to \svp{our decision to use} the features used by practitioners \svp{directly}; \svp{some of this may also result from our operational definitions of various shapes, which did not take into account the model's capabilities and initial training}.
-
-% \mt{What is the balance between comprehensive and redundant?
-I don't think I've found it...
-I'll try again when not exhausted} I'll let you know if I ever find it, but this looks pretty great - I've made a few tiny tweaks to phrasing, but the content is exactly what I was hoping for :)
-
-\mt{Final paragraph will suggest redefining shapes (yuck) and cutting off the final conv block to see whether those features are easier for the model to sort between categories.} \svp{Sounds good!}
+Another source of systematic misclassifications stems from the inherent overlap between features of common geometric shapes; since CNNs make predictions via bottom-up feature integration, it is more difficult for the model to distinguish between geometrically similar shapes, such as chevrons and triangles.
+This issue is exacerbated by high variability of those shapes on outsoles, which requires the model to form a flexible understanding of each shape category, making distinction between similar shapes even more difficult.
+These issues of overlap and variability are at least in part due to our decision to use the features used by practitioners directly; some of this may also result from our operational definitions of various shapes, which did not take into account the model's capabilities and initial training.
+While a \fix{full} remedy may not be possible, there may be some modifications to operational class definitions that we can make to optimize labels for model prediction.
 

### 2019-06-24 paper/paper.Rnw
Finish draft of conclusion

@@
 \fix{DC moved to heatmap, green block text could still be removed}
+\fix{Write/check figure captions}
 
@@
 \section{Conclusions}
-\svp{Rework this to address 3 main conclusions - 1: NNs work for automatic feature ID, 2: Different levels of feature variability within classes, and 3: correcting for contrast is important.}
+%\svp{Rework this to address 3 main conclusions - 1: NNs work for automatic feature ID, 2: Different levels of feature variability within classes, and 3: correcting for contrast is important.}
 
@@
 
-%\section{Future Work}
-%\svp{Need to show image contrast stuff to use this paragraph.}
-%Although CoNNOR performs well in its current state, there are still a number of ways to potentially improve prediction accuracy.
-For one, image contrast  plays a large role in how well the model classifies the geometric shapes present.
-Thus, exploring methods of color correction, such as histogram normalization, may prove useful for eliminating the effect of contrast on predictions.
-Geometric features are also relatively simple with respect to the features that are being detected by the final convolutional block of VGG16, as in \autoref{fig:vgg16layers}; it is quite possible that prediction accuracy could improve by directly classifying the features that are output by the fourth block, rather than using the full convolutional base.
+\fix{Although the model is not perfect in its current form, as a first pass tool it is performing well and improvements seem promising.
+That most of the misclassifications are systematic shows that the model is able to learn features and apply rules consistently.
+There are possible modifications to the labeling scheme that may make model predictions align more closely to human-assigned labels, but model consistency is a much higher good than ``accuracy" (in the sense of agreeing with human-assigned labels).
+Examiners themselves may vary in how to label ambiguous features (e.g., the half-rounded quadrilaterals in \fix{the dc heatmap}), so a consistent classification method provides a more unified definition of certain geometric shapes and facilitates comparison of outsoles with similar geometric features.
+All in all, this model shows that CNNS are a useful tool for classifying a large number of outsoles within a common classification scheme and can ultimately help streamline the current process of sorting and comparing shoes by class characteristics.}
+
+\mt{I think the conclusion currently says everything I want it to (cutting off a conv block doesn't seem to help make the point). Definitely can use some language edits.}
 

### 2019-06-25 paper/paper.Rnw
language edits in conclusion

@@
 Exploring methods of color correction, such as histogram equalization, may prove useful for eliminating the effect of contrast on predictions.
+\svp{Initial attempts at implementing histogram equalization suggest that the overall accuracy improvement is modest; the model fit to images which have been equalized had approximately 92.5\% validation accuracy, an improvement of 1.5\% over the default model described in this paper.}
 
-Another source of systematic misclassifications stems from the inherent overlap between features of common geometric shapes; since CNNs make predictions via bottom-up feature integration, it is more difficult for the model to distinguish between geometrically similar shapes, such as chevrons and triangles.
-This issue is exacerbated by high variability of those shapes on outsoles, which requires the model to form a flexible understanding of each shape category, making distinction between similar shapes even more difficult.
-These issues of overlap and variability are at least in part due to our decision to use the features used by practitioners directly; some of this may also result from our operational definitions of various shapes, which did not take into account the model's capabilities and initial training.
-While a \fix{full} remedy may not be possible, there may be some modifications to operational class definitions that we can make to optimize labels for model prediction.
+Another source of systematic misclassifications stems from the inherent overlap between features of common geometric shapes\svp{.
+}CNNs make predictions via bottom-up feature integration; \svp{as a result,} it is more difficult for the model to distinguish between geometrically similar shapes, such as chevrons and triangles.
+This issue is exacerbated by high variability of some shapes \svp{commonly found in outsole patterns}, which requires the model to form a flexible understanding of each shape category \svp{and makes} distinction between similar shapes even more difficult.
+These issues of overlap and variability are at least in part due to our decision to use the features used by practitioners directly; some of this may also result from our operational definitions of various shapes, which \svp{were created before we had a complete understanding of the practical limitations of CNNs.\sout{did not take into account the model's capabilities and initial training.}} While a \fix{full} remedy may not be possible, there may be some modifications to operational class definitions that we can make to optimize labels for model prediction.
+\svp{In addition, we will continue to improve the labeling of shapes in our database with the goal of expanding the breadth and quantity of labeled images available for model training.}
 
-\fix{Although the model is not perfect in its current form, as a first pass tool it is performing well and improvements seem promising.
+\svp{Although there are many areas for potential improvement in future iterations of this model, we have demonstrated the effectiveness of transfer learning to provide an initial label for shoe outsole patterns.
+As} most of the misclassifications are systematic, it is clear that the model is able to learn features and apply rules consistently.
+\svp{We expect that future modifications to the labeling criteria and data processing pipeline will improve the model's accuracy, but the 91\% validation accuracy} \fix{really should be looking at test set accuracy, but I don't know that number off the top of my head...} \svp{is promising considering that we do not have estimates of the consistency of examiners' estimates on ambiguous features.}  \svp{Taken together, these factors } show that \svp{Convolutional Neural Networks} are a useful tool for classifying a large number of outsoles within a common classification scheme and can ultimately help streamline the current process of sorting and comparing shoes by class characteristics.
+
+% \fix{Although the model is not perfect in its current form, as a first pass tool it is performing well and improvements seem promising.
 That most of the misclassifications are systematic shows that the model is able to learn features and apply rules consistently.
@@
 All in all, this model shows that CNNS are a useful tool for classifying a large number of outsoles within a common classification scheme and can ultimately help streamline the current process of sorting and comparing shoes by class characteristics.}
-
-\mt{I think the conclusion currently says everything I want it to (cutting off a conv block doesn't seem to help make the point). Definitely can use some language edits.}
+%
+% \mt{I think the conclusion currently says everything I want it to (cutting off a conv block doesn't seem to help make the point). Definitely can use some language edits.}
 

### 2019-06-25 paper/paper.Rnw
Fix issues with model path that were causing issues for me. I don't *think* this will effect anything else.

@@
 
-if(grepl("!School", model_path)) {
+if (grepl("!School", model_path)) {
   newest_model <- get_newest(dir = model_path, pattern = "weights.h5")
@@
 load(list.files(model_dir, pattern = "-history.Rdata", full.names = T)[1])
-load(file.path(get_newest()$path, get_newest(pattern = "\\d.Rdata")$base_file))
+load(file.path(newest_model$path,
+               paste0(newest_model$start_date, "_",
+                      newest_model$prefix, ".Rdata")))
 

### 2019-06-25 paper/paper.Rnw
Remove extra }

@@
 A filter is a set of numerical weights that, when applied to an image, quantifies the presence of the specific shapes and/or colors that the filter is trained to detect.
-\autoref{fig:vgg16layers} shows a selection of images which maximally activate specific} filters from VGG16, with each row corresponding to one of the five convolutional blocks.
+\autoref{fig:vgg16layers} shows a selection of images which maximally activate specific filters from VGG16, with each row corresponding to one of the five convolutional blocks.
 Filters in earlier convolutional blocks detect simple features, such as colors, lines, corners, and blobs, while later filters detect more complex combinations of features.

### 2019-06-25 paper/paper.Rnw
All captions except heatmaps

@@
   newest_model <- get_newest(dir = model_path, pattern = "weights.h5")
+  newest_model$prefix = "vgg16_onehotaug_9class_256"
+  newest_model$start_date = "2019-05-14_19_45_22"
 } else {
@@
 \section{Materials and Methods}
-%\fix{We may want to mention some of the methods in the introduction for continuity sake...}
-%\svp{I was talking about the methods above (e.g. fractal decomp, scale invariant feature recognition, etc., not methods we used)}
+%\fix{We may want to mention some of the methods in the introduction for continuity sake...
 Ultimately, most pattern evidence is stored in photographic form; as a result, we must use image analysis methods to create numerical features from the original visual record.
@@
 \caption[VGG16 structure]{VGG16 consists of five convolutional blocks that make up the model base.
-Each convolutional block contains an increasing number of increasingly complex features.
+Each convolutional block contains an increasing number of increasingly complex \mt{feature-detecting filters}.
 After the convolutional blocks, the fully connected layers of the model head are used to make global connections between separate features.}\label{fig:VGG16-structure}
@@
 
-<<vgg16layers, fig.width = 7, fig.height = 5.75, fig.cap = "A selection of filters from the convolutional layers of VGG16.">>=
+<<vgg16layers, fig.width = 7, fig.height = 5.75, fig.cap = "{Visualizing convolutional filters requires generating images which maximally activate specific filters.
+This figure contains images that maximally activate a selection of filters from the convolutional blocks of VGG16.
+Filters from later blocks detect more complex features than those in earlier blocks.
+Much like human vision, a hierarchical structure of feature detection with increasing complexity allows CNN ``vision\" trained on a general task to be applied to new domains.}">>=
 
@@
 
-<<overall-roc, fig.width = 5, fig.height = 5, out.width = ".5\\textwidth", fig.cap = "ROC curve showing overall model performance.", fig.scap = "Overall model performance ROC curve.">>=
+<<overall-roc, fig.width = 5, fig.height = 5, out.width = ".5\\textwidth", fig.cap = "ROC curve showing overall model performance.
+\\mt{The curve indicates that model accuracy is generally high.}", fig.scap = "Overall model performance ROC curve.">>=
 library(pROC)
@@
 AUC is area under the curve, a measure of overall model performance.
-Equal error rates are marked, indicating the position at which there is equal probability of a false positive or false negative error.", fig.scap = "Class-by-class ROC curves.">>=
+Equal error rates (EER) are marked, indicating the position at which there is equal probability of a false positive or false negative error.", fig.scap = "Class-by-class ROC curves.">>=
 aucs <- plot_onehot_roc(preds, test_labs, str_to_title(classes))
@@
 The first rows of \autoref{fig:exemplars} are images which contain only one of the nine classes.
-Most of the predictions for these images are promising; there are some false-positive labels that indicate confusions between circle/text and polygon/quad, but all of the true shapes are corectly identified by the model above EER.
+Most of the predictions for these images are promising; there are some false-positive labels that indicate confusions between circle/text and polygon/quad, but all of the true shapes are corectly identified by the model above their respective EER.
 The next image belongs to the ``other" category, as it does not contain any of the nine classes; the model correctly assigns near-zero probabilities to each class.
@@
 Again, predictions are largely accurate, with most of the prominent shapes correctly identified above EER.
-\fix{Expand on specific misclassifications more, after determining which examples to delete.}
-
-\fix{DC moved to heatmap, green block text could still be removed}
-\fix{Write/check figure captions}
+\mt{The most notable misclassifications in these images are the false positive predictions of triangle in the image with green text and lines in the Adidas logo.
+Although the Adidas logo is technically made up of adjacent quadrilaterals, the prediction of a triangle is reasonable when considering the proximity of those quadrilaterals and the shape's features as a whole, and the relatively low predicted probability suggests that the model recognizes that this shape is ambiguous.
+All other predictions of these images are reasonably correct given our operational definition of class categories.}
 
@@
 
-<<exemplars, fig.cap = "Examples containing a single shape", fig.pos = "p!", out.width = "100%", cache = F, fig.width = 8, fig.height = 12>>=
+<<exemplars, fig.cap = "\\mt{Model predictions of images containing a single shape, no shape, or multiple shapes across all categories.
+Although there are some confusions (such as between circle and text in the single class images, and the false positive prediction of triangle in the image with green text), predictions are generally reasonable and consistent with the human-assigned labels of these images.}", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 8, fig.height = 12>>=
 single_exemplars <- list.files(file.path(wd, "images", "consistency",
@@
 \autoref{fig:circle-text-consistency} shows a number of images containing circles and/or text.
-\fix{The top four rows show text contained within a separate, distinct circle shape, and the next five rows show circles represented as the letter ``o" in text.
-The remaining images contain only one of either text or circles, as well as other possible shape categories.} In particular, the first two of these images contain text that use the capital letter ``G", the next two images contain no circular shapes, and the final \fix{N} images contain pure circles that do not appear to belong to text.
+The top four rows show text contained within a separate, distinct circle shape, and the next five rows show circles represented as the letter ``o" in text.
+The remaining images contain only one of either text or circles, as well as other possible shape categories.
+In particular, the first two of these images contain text that use the capital letter ``G", the next two images contain no circular shapes, and the final 3 images contain pure circles that do not appear to belong to text.
 The model is able to identify the circle in all cases where a circle is truly present, regardless of whether the circle is part of the text or it is distinct; however, the circle prediction is also strong in three of the four images that do not contain a circle, which indicates that the model is having dificulty distinguishing the two categories.
-While the circle predictions are above the EER for letters such as C and S, these predictions are generally much lower than those where distinct circles or Os are seen, indicating that there are some indications that features are not as strong in the quasi-circle shapes, but that the model is not able to completely separate Os from C, S, and G shapes which are also very circular.
+While the circle predictions are \mt{often} above the EER for letters such as G and S, these predictions are generally much lower than those where distinct circles or Os are seen, indicating that there are some indications that features are not as strong in the quasi-circle shapes, but that the model is not able to completely separate Os from C, S, and G shapes which are also very circular.
 This confusion is complicated by the frequency of images which contain both circles and text: text is present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("circle", dfunion$name)), 2)}\% of our labeled images containing circles, and circles are present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("text", dfunion$name)), 2)}\% of our images containing text.
@@
 
-% \fix{make sure text matches figure!}
-
-<<circle-text-consistency, fig.cap = "Text with circles or circle-like shapes", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 8, fig.height = 12>>=
+<<circle-text-consistency, fig.cap = "\\mt{Model predictions of images containing circles and/or text.
+The model correctly identifies circles in all images where circles are present, but also predicts moderate probabilities of circles in many images where text is present with curved letters, such as ``G\".
+Text is correctly identified in most images, with the exception of the false negative prediction of the low-contrast Teva image and the false positive prediction of concentric circles.
+In general, model predictions are reasonable but suggest a few sources of systematic misclassification, which are discussed in a later section.}", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 8, fig.height = 12>>=
 circle_text <- list.files(file.path(wd, "images", "consistency", "circle-text"), full.names = T)
@@
 
-<<low-contrast-consistency, fig.cap = "Images with low contrast", fig.pos = "p!", out.width = "100%", cache = F>>=
+<<low-contrast-consistency, fig.cap = "\\mt{Model predictions for images with low contrast.
+While these images do truly contain a variety of shapes, the diffuse probabilities assigned by the model across all categories indicate that features are obscured by low contrast, which propogates through model training and prediction.}", fig.pos = "p!", out.width = "100%", cache = T>>=
 lowcontrast <- list.files(file.path(wd, "images", "consistency", "low-contrast2"), full.names = T)
@@
 knitr::include_graphics(heatmaps[grepl("Ugg-grey", heatmaps)])
-knitr::include_graphics(heatmaps[grepl("text", heatmaps)])
+knitr::include_graphics(heatmaps[grepl("text-seychelles", heatmaps)])
 knitr::include_graphics(heatmaps[grepl("polygon_recolor", heatmaps)])
@@
 Exploring methods of color correction, such as histogram equalization, may prove useful for eliminating the effect of contrast on predictions.
-\svp{Initial attempts at implementing histogram equalization suggest that the overall accuracy improvement is modest; the model fit to images which have been equalized had approximately 92.5\% validation accuracy, an improvement of 1.5\% over the default model described in this paper.}
+Initial attempts at implementing histogram equalization suggest that the overall accuracy improvement is modest; the model fit to images which have been equalized had approximately 92.5\% validation accuracy, an improvement of 1.5\% over the default model described in this paper.
 
-Another source of systematic misclassifications stems from the inherent overlap between features of common geometric shapes\svp{.
-}CNNs make predictions via bottom-up feature integration; \svp{as a result,} it is more difficult for the model to distinguish between geometrically similar shapes, such as chevrons and triangles.
-This issue is exacerbated by high variability of some shapes \svp{commonly found in outsole patterns}, which requires the model to form a flexible understanding of each shape category \svp{and makes} distinction between similar shapes even more difficult.
-These issues of overlap and variability are at least in part due to our decision to use the features used by practitioners directly; some of this may also result from our operational definitions of various shapes, which \svp{were created before we had a complete understanding of the practical limitations of CNNs.\sout{did not take into account the model's capabilities and initial training.}} While a \fix{full} remedy may not be possible, there may be some modifications to operational class definitions that we can make to optimize labels for model prediction.
-\svp{In addition, we will continue to improve the labeling of shapes in our database with the goal of expanding the breadth and quantity of labeled images available for model training.}
+Another source of systematic misclassifications stems from the inherent overlap between features of common geometric shapes.
+CNNs make predictions via bottom-up feature integration; as a result, it is more difficult for the model to distinguish between geometrically similar shapes, such as chevrons and triangles.
+This issue is exacerbated by high variability of some shapes commonly found in outsole patterns, which requires the model to form a flexible understanding of each shape category and makes distinction between similar shapes even more difficult.
+These issues of overlap and variability are at least in part due to our decision to use the features used by practitioners directly; some of this may also result from our operational definitions of various shapes, which were created before we had a complete understanding of the practical limitations of CNNs.
+While a complete remedy may not be possible, there may be some modifications to operational class definitions that we can make to optimize labels for model prediction.
+In addition, we will continue to improve the labeling of shapes in our database with the goal of expanding the breadth and quantity of labeled images available for model training.
 
-\svp{Although there are many areas for potential improvement in future iterations of this model, we have demonstrated the effectiveness of transfer learning to provide an initial label for shoe outsole patterns.
-As} most of the misclassifications are systematic, it is clear that the model is able to learn features and apply rules consistently.
-\svp{We expect that future modifications to the labeling criteria and data processing pipeline will improve the model's accuracy, but the 91\% validation accuracy} \fix{really should be looking at test set accuracy, but I don't know that number off the top of my head...} \svp{is promising considering that we do not have estimates of the consistency of examiners' estimates on ambiguous features.}  \svp{Taken together, these factors } show that \svp{Convolutional Neural Networks} are a useful tool for classifying a large number of outsoles within a common classification scheme and can ultimately help streamline the current process of sorting and comparing shoes by class characteristics.
+Although there are many areas for potential improvement in future iterations of this model, we have demonstrated the effectiveness of transfer learning to provide an initial label for shoe outsole patterns.
+As most of the misclassifications are systematic, it is clear that the model is able to learn features and apply rules consistently.
+We expect that future modifications to the labeling criteria and data processing pipeline will improve the model's accuracy, but the 91\% validation accuracy \fix{really should be looking at test set accuracy, but I don't know that number off the top of my head...}\mt{See chunk below} is promising considering that we do not have estimates of the consistency of examiners' estimates on ambiguous features.
+Taken together, these factors show that Convolutional Neural Networks are a useful tool for classifying a large number of outsoles within a common classification scheme and can ultimately help streamline the current process of sorting and comparing shoes by class characteristics.
 
-% \fix{Although the model is not perfect in its current form, as a first pass tool it is performing well and improvements seem promising.
-That most of the misclassifications are systematic shows that the model is able to learn features and apply rules consistently.
-There are possible modifications to the labeling scheme that may make model predictions align more closely to human-assigned labels, but model consistency is a much higher good than ``accuracy" (in the sense of agreeing with human-assigned labels).
-Examiners themselves may vary in how to label ambiguous features (e.g., the half-rounded quadrilaterals in \fix{the dc heatmap}), so a consistent classification method provides a more unified definition of certain geometric shapes and facilitates comparison of outsoles with similar geometric features.
-All in all, this model shows that CNNS are a useful tool for classifying a large number of outsoles within a common classification scheme and can ultimately help streamline the current process of sorting and comparing shoes by class characteristics.}
-%
-% \mt{I think the conclusion currently says everything I want it to (cutting off a conv block doesn't seem to help make the point). Definitely can use some language edits.}
+<<test-accuracy, echo = F>>=
+# This is my stab at calculating test accuracy.
+# .786 doesn't seem too promising :(
+
+disc_preds <- sweep(preds, 2, eers) > 0 #predicted values minus EER, column wise
+test_accuracy <- sum(disc_preds == test_labs) / (length(test_labs))
+@
 

### 2019-06-25 paper/code/Generate_Model_Images.R
Update EER border width in pred_prob_plot

@@ -108,7 +108,8 @@ pred_prob_plot <- function(img_path, model, classes = default_classes,
   img_preds_df <- img_preds_mat %>%
     as.data.frame() %>%
     set_colnames(default_classes) %>%
-    set_rownames(purrr::map_chr(img_preds, "path") %>% basename() %>% str_remove("\\.jpg")) %>%
+    set_rownames(purrr::map_chr(img_preds, "path") %>% basename() %>%
+                   str_remove("\\.jpg")) %>%
     mutate(path = purrr::map_chr(img_preds, "path"))
 
   img_preds_df <- if (sort) {
@@ -158,7 +159,7 @@ pred_prob_plot <- function(img_path, model, classes = default_classes,
     geom_tile(aes(y = idx, x = as.numeric(factor(class)),
                   fill = value, color = yes,
                   width = w, height = h),
-              data = img_preds_df_long) +
+              data = img_preds_df_long, size = .4) +
     geom_text(aes(x = as.numeric(factor(class)), y = idx,
                   label = sprintf("%0.2f", value)), data = img_preds_df_long,
               hjust = 0.5, vjust = 0.5) +

### 2019-06-26 paper/code/Generate_Model_Images.R
Make all labels title case, wrap code around 80 chars

@@ -23,14 +23,16 @@ default_classes <- c(
 # --- Functions ----------------------------------------------------------------
 
 # Get most recent model
-get_newest <- function(dir = file.path(modeldir, "TrainedModels"), pattern = "fullimage.rdata") {
+get_newest <- function(dir = file.path(modeldir, "TrainedModels"),
+                       pattern = "fullimage.rdata") {
   dirlist <- list.dirs(path = dir) %>% rev()
   newest_file <- NULL
   startidx <- 1:pmin(length(dirlist) - 1, 10)
   while (is.null(newest_file) & max(startidx) <= length(dirlist)) {
     newest_file <- dirlist %>%
       magrittr::extract(startidx) %>%
-      sapply(function(x) list.files(x, pattern = pattern, recursive = F, full.names = T)) %>%
+      sapply(function(x) list.files(x, pattern = pattern, recursive = F,
+                                    full.names = T)) %>%
       unlist() %>%
       as.character()
 
@@ -48,7 +50,8 @@ get_newest <- function(dir = file.path(modeldir, "TrainedModels"), pattern = "fu
     str_remove(start_date) %>%
     str_remove("^_")
   process_dir <- basename(path)
-  return(list(path = path, base_file = base_file, prefix = prefix, start_date = start_date, process_dir = process_dir))
+  return(list(path = path, base_file = base_file, prefix = prefix,
+              start_date = start_date, process_dir = process_dir))
 }
 
 set_weights <- function(model_wts_file) {
@@ -78,7 +81,7 @@ set_weights <- function(model_wts_file) {
 
 
 # --- # Model predictions # ----------------------------------------------------
-predict_new <- function(img_path, model, classes = default_classes) {
+predict_new <- function(img_path, model, classes = str_to_title(default_classes)) {
   purrr::map(img_path, function(x) {
     img <- jpeg::readJPEG(x)
     dim(img) <- c(1, dim(img))
@@ -92,9 +95,9 @@ predict_new <- function(img_path, model, classes = default_classes) {
   })
 }
 
-pred_prob_plot <- function(img_path, model, classes = default_classes,
+pred_prob_plot <- function(img_path, model, classes = str_to_title(default_classes),
                            sort = T, eer = NULL) {
-  if(!is.null(eer)) {
+  if (!is.null(eer)) {
[…]

### 2019-06-26 paper/paper.Rnw
Edit authors - get rid of duplicate affiliation markings
accepting changes, tweaking figures/captions

@@
 % or include affiliations in footnotes:
-\author[1]{Miranda Tilton\fnref{isu}}
-\ead{tiltonm@iastate.edu}
-
-\author[1]{Susan Vanderplas\corref{cor1}\fnref{isu}}
+\author[1]{Miranda Tilton and Susan Vanderplas\corref{cor1}\fnref{isu}}
 \cortext[cor1]{Corresponding author}
@@
 Ultimately, most pattern evidence is stored in photographic form; as a result, we must use image analysis methods to create numerical features from the original visual record.
-There are a number of methods that may be employed to identify shapes and features in an image, such as Fourier\mt{-Mellin} transforms, \mt{fractal decomposition, scale-invariant feature recognition}, Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}.
-While these methods are useful in identifying these specific features at a low level, they can be computationally intensive and only identify features on a very small scale; as a result, they cannot reliably identify large geometric shapes like those that may be found in a high resolution outsole image.
+There are a number of methods that may be employed to identify shapes and features in an image, such as Fourier-Mellin transforms, Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}.
+While these methods are useful in identifying these specific features at a low level, they only identify features on a very small scale; as a result, they cannot reliably identify large geometric shapes like those that may be found in a high resolution outsole image.
 
@@
 <<overall-roc, fig.width = 5, fig.height = 5, out.width = ".5\\textwidth", fig.cap = "ROC curve showing overall model performance.
-\\mt{The curve indicates that model accuracy is generally high.}", fig.scap = "Overall model performance ROC curve.">>=
+\\svp{Model accuracy is generally high, as the ROC curve is significantly above the identity line $y=x$.}", fig.scap = "Overall model performance ROC curve.">>=
 library(pROC)
@@
 Again, predictions are largely accurate, with most of the prominent shapes correctly identified above EER.
-\mt{The most notable misclassifications in these images are the false positive predictions of triangle in the image with green text and lines in the Adidas logo.
+The most notable misclassifications in these images are the false positive predictions of triangle in the image with green text and lines in the Adidas logo.
 Although the Adidas logo is technically made up of adjacent quadrilaterals, the prediction of a triangle is reasonable when considering the proximity of those quadrilaterals and the shape's features as a whole, and the relatively low predicted probability suggests that the model recognizes that this shape is ambiguous.
-All other predictions of these images are reasonably correct given our operational definition of class categories.}
+All other predictions \svp{for} these images are \svp{accurate} given our operational definition of class categories.
 
@@
 
-<<exemplars, fig.cap = "\\mt{Model predictions of images containing a single shape, no shape, or multiple shapes across all categories.
-Although there are some confusions (such as between circle and text in the single class images, and the false positive prediction of triangle in the image with green text), predictions are generally reasonable and consistent with the human-assigned labels of these images.}", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 8, fig.height = 12>>=
+<<exemplars, fig.cap = "Model predictions of images containing \\svp{(from the top)} a single shape, no shape, or multiple shapes across all categories.
+\\svp{Predicted probabilities greater than the equal error rate threshold are outlined in black.} Although there are some \\svp{misclassifications}, predictions are generally consistent with the human-assigned labels of these images.", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 11/1.25, fig.height = 16/1.25>>=
 single_exemplars <- list.files(file.path(wd, "images", "consistency",
@@
 
-<<circle-text-consistency, fig.cap = "\\mt{Model predictions of images containing circles and/or text.
+<<circle-text-consistency, fig.cap = "Model predictions of images containing circles and/or text.
 The model correctly identifies circles in all images where circles are present, but also predicts moderate probabilities of circles in many images where text is present with curved letters, such as ``G\".
-Text is correctly identified in most images, with the exception of the false negative prediction of the low-contrast Teva image and the false positive prediction of concentric circles.
-In general, model predictions are reasonable but suggest a few sources of systematic misclassification, which are discussed in a later section.}", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 8, fig.height = 12>>=
+Text is correctly identified in most images, with the exception of the false negative prediction of the low-contrast \\svp{text within a hexagon} and the false positive prediction of concentric circles.", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 11/1.25, fig.height = 16/1.25>>=
 circle_text <- list.files(file.path(wd, "images", "consistency", "circle-text"), full.names = T)
@@
 <<confusion-matrix, fig.width = 9, fig.height = 8, out.width = "\\textwidth", fig.cap = "Confusion matrix, showing on the diagonal the correct classification rate and on the off-diagonal, classification errors.
-Note that in multi-label images, correct off-diagonal labels have been excluded from the calculation of false positives.", fig.scap = "Confusion matrix, with correct and incorrect model classifications.", dpi = 600>>=
+Note that in multi-label images, correct off-diagonal labels have been excluded from the calculation of false positives.
+\\svp{Quadrilaterals are generally over-predicted across true labels; polygons and triangles are under-predicted.
+Circles and text are also often confused.}", fig.scap = "Confusion matrix, with correct and incorrect model classifications.", dpi = 600>>=
 get_confusion_matrix(predictions = preds, classes = classes,
@@
   ggcorrplot(., hc.order = F, outline.col = "white", lab = T) +
+  theme_bw() +
   scale_fill_gradient("Classification\nRate", low = "white",
                       high = "cornflowerblue", limits = c(0, 1)) +
-  scale_x_discrete("Image Label") + scale_y_discrete("Prediction") +
+  scale_x_discrete("Image Label", labels = str_to_title(classes)) + scale_y_discrete("Prediction") +
   theme(axis.title.x = element_text(size = 14),
-        axis.title.y = element_text(size = 14, angle = 90, vjust = 1)) +
+        axis.title.y = element_text(size = 14, angle = 90, vjust = 1),
+        axis.text.y = element_text(angle = 90, hjust = .5)) +
   ggtitle("CoNNOR Multi-Class Confusion Matrix: Test Set Performance") +
-  theme_bw() +
   theme(panel.grid.major = element_line(color = "grey50"),
@@
 As a result, the features in low contrast images are not easily resolved by the filters; this is propogated through both model training and prediction of new images.
-In other words, the model 'learns' that when there are few meaningful features with which to determine which shapes are present, the best strategy for minimizing loss is to predict diffuse probabilities across many classes, as seen in \autoref{fig:low-contrast-consistency}.
+In other words, the model `learns' that when there are few meaningful features with which to determine which shapes are present, the best strategy for minimizing loss is to predict diffuse probabilities across many classes, as seen in \autoref{fig:low-contrast-consistency}.
 
@@
 
-<<current-heatmaps, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.cap = c("A section of an Ugg shoe containing the logo.
+<<current-heatmaps, eval = F, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.cap = c("A section of an Ugg shoe containing the logo.
 The model correctly identifies triangles and text (and references the correct portion of the image in both cases).
 The model identifies a circle as well; while it could be argued that the triangles form an implicit circle, the heatmap demonstrates that the circular portion of the image is the 'G', which would not be a closed figure labeled as a circle.
-The model's prediction of the presence of a circle is understandable, but incorrect.", "In this image of a Seychelles shoe, the model identifies text and circle.
-The prediction of text is correct; the prediction of circle is not - there are very round shapes, but they do not form a closed figure that could be identified as a circle.", "polygon", "corks"), echo = F, message = F, warning = F, fig.pos = "!p">>=
+The model's prediction of the presence of a circle is understandable, but incorrect.\\label{fig:ugg-heatmap}", "In this image of a Seychelles shoe, the model identifies text and circle.
+The prediction of text is correct; the prediction of circle is not - there are very round shapes, but they do not form a closed figure that could be identified as a circle.\\label{fig:seychelles-heatmap}", "polygon", "corks"), echo = F, message = F, warning = F, fig.pos = "!p">>=
 

### 2019-06-26 paper/code/Generate_Model_Images.R
Add optional comparison column for images

@@ -96,7 +96,7 @@ predict_new <- function(img_path, model, classes = str_to_title(default_classes)
 }
 
 pred_prob_plot <- function(img_path, model, classes = str_to_title(default_classes),
-                           sort = T, eer = NULL) {
+                           sort = T, eer = NULL, img_path2 = NULL) {
   if (!is.null(eer)) {
     assertthat::assert_that(is.numeric(eer))
     assertthat::assert_that(length(eer) == length(classes))
@@ -114,6 +114,11 @@ pred_prob_plot <- function(img_path, model, classes = str_to_title(default_class
     set_rownames(purrr::map_chr(img_preds, "path") %>% basename() %>%
                    str_remove("\\.jpg")) %>%
     mutate(path = purrr::map_chr(img_preds, "path"))
+
+  if (!is.null(img_path2)) {
+    assertthat::assert_that(length(img_path) == length(img_path2))
+    img_preds_df$path2 <- img_path2
+  }
 
   img_preds_df <- if (sort) {
     img_preds_df %>%
@@ -148,6 +153,23 @@ pred_prob_plot <- function(img_path, model, classes = str_to_title(default_class
     ~annotation_custom(grid::rasterGrob(readJPEG(.x), interpolate = T),
                        ymin = .y - .5, ymax = .y + .5,
                        xmin = -0.5, xmax = 0.5))
+  xlims <- -0.5 + c(0,length(classes))
+  xlabs <- classes
+  xbreaks <- 1:9
+
+  if (!is.null(img_path2)) {
+    features_xraster2 <- purrr::map2(
+      img_preds_df$path2, img_preds_df$idx,
+      ~annotation_custom(grid::rasterGrob(as.array(im_equalize(readJPEG(.x)),
+                                                   dim = c(256, 256, 3)),
+                                          interpolate = T),
+                         ymin = .y - .5, ymax = .y + .5,
+                         xmin = length(classes) + .5, xmax = length(classes) + 1.5))
+    xlims <- -0.5 + c(0,length(classes) + 2)
+    xlabs <- c("Original\nImage", xlabs, "Contrast\nCorrected\nImage")
+    xbreaks <- 0:(length(classes) + 1)
+    features_xraster <- list(features_xraster, features_xraster2)
+  }
 
   if (sum(eer) == length(classes)) {
     cols <- c("No" = "black", "Yes" = "black")
[…]

### 2019-06-26 paper/paper.Rnw
Add code to equalize images, extra column to pred_prob_plot

@@
 source(file.path(codedir, "Generate_Model_Images.R"))
+source(file.path(codedir, "equalize.R"))
 @
@@
 
-<<low-contrast-consistency, fig.cap = "\\mt{Model predictions for images with low contrast.
-While these images do truly contain a variety of shapes, the diffuse probabilities assigned by the model across all categories indicate that features are obscured by low contrast, which propogates through model training and prediction.}", fig.pos = "p!", out.width = "100%", cache = T>>=
+<<low-contrast-consistency, fig.cap = "Model predictions for images with low contrast.
+While these images contain a variety of shapes \\svp{(as shown in the color-corrected column on the right)}, the diffuse probabilities assigned by the model across all categories indicate that features are obscured by low contrast, which propogates through model training and prediction.", fig.pos = "p!", out.width = "100%", fig.width = 10/1.2, fig.height = 9/1.2, cache = T>>=
 lowcontrast <- list.files(file.path(wd, "images", "consistency", "low-contrast2"), full.names = T)
-pred_prob_plot(lowcontrast, loaded_model, sort = F, eer = eers)
+
+pred_prob_plot(lowcontrast, loaded_model, sort = F, eer = eers,
+               img_path2 = lowcontrast) +
+  theme(legend.position = "bottom")
 @

### 2019-06-26 paper/code/equalize.R

@@ -1,0 +1,87 @@
+
+
+equalize <- function(x) {
+  # Stolen from EBImage and modified
+  if (max(x) > 1) { # Assume LAB
+    range <- c(0, 100)
+  } else {
+    range <- c(0, 1)
+  }
+
+  levels <- 1024
+  r <- range(x)
+  if (r[1L] == r[2L]) {
+    warning("Image histogram is single-valued and cannot be equalized.")
+    x
+  } else {
+    if (!is.numeric(range) || length(range) != 2L) {
+      stop("'range' must be a numeric vector of length 2.")
+    }
+
+    levels <- as.integer(levels)
+    if (is.na(levels) || levels < 1L) {
+      stop("Levels must be at least equal 1.")
+    }
+    breaks <- seq(range[1L], range[2L], length.out = levels + 1L)
+    d <- dim(x)
+    n <- d[1L] * d[2L]
+
+    # equalize each frame separately
+    .equalize <- function(y) {
+      h <- hist.default(y, breaks = breaks, plot = FALSE)
+      cdf <- cumsum(h$counts)
+      cdf_min <- min(cdf[cdf > 0])
+
+      equalized <- ((cdf - cdf_min) / (prod(dim(y)) - cdf_min) * (range[2L] - range[1L])) + range[1L]
+      bins <- round((y - range[1L]) / (range[2L] - range[1L]) * (levels - 1L)) + 1L
+
+      res <- equalized[bins]
+    }
+
+    ld <- length(d[d != 1])
+    res <- if (ld > 2L) {
+      apply(x, 3L:ld, .equalize)
+    } else {
+      .equalize(x)
+    }
+
+    imager::as.cimg(res, dim = d)
+  }
+}
+
+im_equalize <- function(im, plot_res = F) {
+  reconv <- FALSE
+
+  if (!imager::is.cimg(im)) {
+    reconv <- TRUE
+    im <- imager::as.cimg(im) %>% imager::imrotate(angle = 90) %>% mirror("x")
+  }
+
+  tmp <- im %>%
+    imager::RGBtoHSL() %>%
+    imager::imsplit("c")
+
+  total_var <- purrr::map_dbl(imager::imsplit(im, "c"), var) %>% sum()
+  if (var(tmp[[3]]) > 0.0001 & total_var > 0.001) {
+    tmp[[3]] <- equalize(tmp[[3]])
+  }
+  res <- tmp %>%
[…]

### 2019-06-26 paper/paper.Rnw
comments on accuracy, restructuring titles, slight changes to figure width/height and captions

@@
 
-\paragraph{Model Predictions---Examples}
+\paragraph{Class Accuracy}
 The ROC curves provide only a general summary of model accuracy; to learn more about how specific images are classified, we can view images from the test set along with the nine class predictions assigned by the model.
@@
 <<exemplars, fig.cap = "Model predictions of images containing \\svp{(from the top)} a single shape, no shape, or multiple shapes across all categories.
-\\svp{Predicted probabilities greater than the equal error rate threshold are outlined in black.} Although there are some \\svp{misclassifications}, predictions are generally consistent with the human-assigned labels of these images.", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 11/1.25, fig.height = 16/1.25>>=
+\\svp{Predicted probabilities greater than the equal error rate threshold are outlined in black.} Although there are some \\svp{misclassifications}, predictions are generally consistent with the human-assigned labels of these images.", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 11.5/1.25, fig.height = 16/1.25>>=
 single_exemplars <- list.files(file.path(wd, "images", "consistency",
@@
 The model correctly identifies circles in all images where circles are present, but also predicts moderate probabilities of circles in many images where text is present with curved letters, such as ``G\".
-Text is correctly identified in most images, with the exception of the false negative prediction of the low-contrast \\svp{text within a hexagon} and the false positive prediction of concentric circles.", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 11/1.25, fig.height = 16/1.25>>=
+Text is correctly identified in most images, with the exception of the false negative prediction of the low-contrast \\svp{text within a hexagon} and the false positive prediction of concentric circles.", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 11.5/1.25, fig.height = 16/1.25>>=
 circle_text <- list.files(file.path(wd, "images", "consistency", "circle-text"), full.names = T)
@@
 
-\paragraph{Confusion Matrix}
+% \paragraph{Confusion Matrix}
 Figures \ref{fig:exemplars} and \ref{fig:circle-text-consistency} show predictions for a number of individual images.
@@
         panel.grid.minor = element_line(color = "grey60")) +
-  theme(plot.margin = grid::unit(c(0,0,0,0), "mm"),
+  theme(plot.margin = grid::unit(c(1,0,0,0), "mm"),
         plot.background = element_rect(fill = "transparent", color = NA),
@@
 Exploring methods of color correction, such as histogram equalization, may prove useful for eliminating the effect of contrast on predictions.
-Initial attempts at implementing histogram equalization suggest that the overall accuracy improvement is modest; the model fit to images which have been equalized had approximately 92.5\% validation accuracy, an improvement of 1.5\% over the default model described in this paper.
+Initial implementations of histogram equalization suggest that the overall accuracy improvement is modest; the model fit to images which have been equalized had approximately 92.5\% validation accuracy, an improvement of 1.5\% over the default model described in this paper.
 
@@
 # .786 doesn't seem too promising :(
+# That's with thresholding. If you use absolute error, we get an accuracy rate of 0.86, which is pretty good. (the contrast corrected model is 0.88, so much more reasonable)
+# Or if we compare to your calculations, we go from .786 -> .811. All in all, we're doing ok. We can definitely just report validation accuracy, though...
+# Interestingly enough, if you use a 0.5 threshold, you get an accuracy rate of .85...
+EER isn't getting us that much.
+The accuracy rate on the contrast-fixed model is exactly the same under that constraint, which suggests that there's some difference in the output probabilities, but not *that* much of a difference - enough to make a small difference when we're using EER.
+abs_err <- sum(abs(preds - test_labs))/length(test_labs)
 

### 2019-06-26 paper/paper.Rnw
Finish heatmap selection, captions, and in-text description (the latter is currently commented out)

@@
 Many convolutional neural networks have been trained on a subset of images from an online image database called ImageNet \citep{dengImageNetLargeScaleHierarchical}.
-In particular, there is a standard set of images from ImageNet, consisting of 1.2 million images \mt{that span} 1,000 catetories, that is commonly used in computer vision tasks.
+In particular, there is a standard set of images from ImageNet, consisting of 1.2 million images that span 1,000 catetories, that is commonly used in computer vision tasks.
 CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously.
@@
 
-<<vgg16layers, fig.width = 7, fig.height = 5.75, fig.cap = "{Visualizing convolutional filters requires generating images which maximally activate specific filters.
+<<vgg16layers, fig.width = 7, fig.height = 5.75, fig.cap = "Visualizing convolutional filters requires generating images which maximally activate specific filters.
 This figure contains images that maximally activate a selection of filters from the convolutional blocks of VGG16.
 Filters from later blocks detect more complex features than those in earlier blocks.
-Much like human vision, a hierarchical structure of feature detection with increasing complexity allows CNN ``vision\" trained on a general task to be applied to new domains.}">>=
+Much like human vision, a hierarchical structure of feature detection with increasing complexity allows CNN ``vision\" trained on a general task to be applied to new domains.">>=
 
@@
 (2013) require contextual spatial information which is not preserved during labeling (for example, lugs are required to be on the perimeter of the shoe).
-\autoref{tab:class-char-examples} shows three examples of each class in \mt{the modified} feature set.
+\autoref{tab:class-char-examples} shows three examples of each class in the modified feature set.
 
@@
 <<overall-roc, fig.width = 5, fig.height = 5, out.width = ".5\\textwidth", fig.cap = "ROC curve showing overall model performance.
-\\svp{Model accuracy is generally high, as the ROC curve is significantly above the identity line $y=x$.}", fig.scap = "Overall model performance ROC curve.">>=
+Model accuracy is generally high, as the ROC curve is significantly above the identity line $y=x$.", fig.scap = "Overall model performance ROC curve.">>=
 library(pROC)
@@
 Although the Adidas logo is technically made up of adjacent quadrilaterals, the prediction of a triangle is reasonable when considering the proximity of those quadrilaterals and the shape's features as a whole, and the relatively low predicted probability suggests that the model recognizes that this shape is ambiguous.
-All other predictions \svp{for} these images are \svp{accurate} given our operational definition of class categories.
+All other predictions for these images are accurate given our operational definition of class categories.
 
@@
 
-<<exemplars, fig.cap = "Model predictions of images containing \\svp{(from the top)} a single shape, no shape, or multiple shapes across all categories.
-\\svp{Predicted probabilities greater than the equal error rate threshold are outlined in black.} Although there are some \\svp{misclassifications}, predictions are generally consistent with the human-assigned labels of these images.", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 11.5/1.25, fig.height = 16/1.25>>=
+<<exemplars, fig.cap = "Model predictions of images containing (from the top) a single shape, no shape, or multiple shapes across all categories.
+Predicted probabilities greater than the equal error rate threshold are outlined in blue.
+Although there are some misclassifications, predictions are generally consistent with the human-assigned labels of these images.", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 11.5/1.25, fig.height = 16/1.25>>=
 single_exemplars <- list.files(file.path(wd, "images", "consistency",
@@
 The model is able to identify the circle in all cases where a circle is truly present, regardless of whether the circle is part of the text or it is distinct; however, the circle prediction is also strong in three of the four images that do not contain a circle, which indicates that the model is having dificulty distinguishing the two categories.
-While the circle predictions are \mt{often} above the EER for letters such as G and S, these predictions are generally much lower than those where distinct circles or Os are seen, indicating that there are some indications that features are not as strong in the quasi-circle shapes, but that the model is not able to completely separate Os from C, S, and G shapes which are also very circular.
+While the circle predictions are often above the EER for letters such as G and S, these predictions are generally much lower than those where distinct circles or Os are seen, indicating that there are some indications that features are not as strong in the quasi-circle shapes, but that the model is not able to completely separate Os from C, S, and G shapes which are also very circular.
 This confusion is complicated by the frequency of images which contain both circles and text: text is present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("circle", dfunion$name)), 2)}\% of our labeled images containing circles, and circles are present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("text", dfunion$name)), 2)}\% of our images containing text.
@@
 The model correctly identifies circles in all images where circles are present, but also predicts moderate probabilities of circles in many images where text is present with curved letters, such as ``G\".
-Text is correctly identified in most images, with the exception of the false negative prediction of the low-contrast \\svp{text within a hexagon} and the false positive prediction of concentric circles.", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 11.5/1.25, fig.height = 16/1.25>>=
+Text is correctly identified in most images, with the exception of the false negative prediction of the low-contrast text within a hexagon and the false positive prediction of concentric circles.", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 11.5/1.25, fig.height = 16/1.25>>=
 circle_text <- list.files(file.path(wd, "images", "consistency", "circle-text"), full.names = T)
@@
 Note that in multi-label images, correct off-diagonal labels have been excluded from the calculation of false positives.
-\\svp{Quadrilaterals are generally over-predicted across true labels; polygons and triangles are under-predicted.
-Circles and text are also often confused.}", fig.scap = "Confusion matrix, with correct and incorrect model classifications.", dpi = 600>>=
+Quadrilaterals are generally over-predicted across true labels; polygons and triangles are under-predicted.
+Circles and text are also often confused.", fig.scap = "Confusion matrix, with correct and incorrect model classifications.", dpi = 600>>=
 get_confusion_matrix(predictions = preds, classes = classes,
@@
 <<low-contrast-consistency, fig.cap = "Model predictions for images with low contrast.
-While these images contain a variety of shapes \\svp{(as shown in the color-corrected column on the right)}, the diffuse probabilities assigned by the model across all categories indicate that features are obscured by low contrast, which propogates through model training and prediction.", fig.pos = "p!", out.width = "100%", fig.width = 10/1.2, fig.height = 9/1.2, cache = T>>=
+While these images contain a variety of shapes (as shown in the color-corrected column on the right), the diffuse probabilities assigned by the model across all categories indicate that features are obscured by low contrast, which propogates through model training and prediction.", fig.pos = "p!", out.width = "100%", fig.width = 10/1.2, fig.height = 9/1.2, cache = T>>=
 lowcontrast <- list.files(file.path(wd, "images", "consistency", "low-contrast2"), full.names = T)
@@
 We have proposed several hypotheses to explain why the model's predictions are not entirely consistent with manual labels on the images.
-We can test these hypotheses by examining what the model ``sees" when making its predictions \mt{\sout{using heatmaps, or class activation maps (CAMs)}}.
-Heatmaps, also known as Class Activation Maps \mt{(CAMs)}, are a visual diagnostic tool that highlights the areas of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}.
+We can test these hypotheses by examining what the model ``sees" when making its predictions.
+Heatmaps, also known as Class Activation Maps (CAMs), are a visual diagnostic tool that highlights the areas of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}.
 Using these graphics, we can identify which locations in an image contribute most to the model's predictions.
-In figures \fix{X-Y}, heatmaps are shown for a selection of \fix{three} classes, which are either contained in the image or have high output probabilities, alongside the original image.
-In these heatmaps, we can confirm that multiple classes have common features, resulting in misclassifications.
+\mt{Figures \ref{fig:ugg-heatmap}-\ref{fig:dc-heatmap} show a number of images alongside heatmaps for the four classes with the highest output probabilities for each image.
+These heatmaps can confirm that a number of misclassifications are a result of overlapping features between classes, and that the model is using reasonable features to make predictions.}
 
-\fix{Choose three (or four) heatmaps:
-\begin{itemize}
-\item An image with multiple true shapes that the model clearly predicts well, to prove that it has learned the features we think it should (probably an UGG logo)
-\item (Seychelles circle/text confusion)
-\item An image with high quad/tri/poly confusion, but the model uses appropriate features to make educated guesses
-\item An image with no true classes but would yield false positives (e.g., corks, puzzle pieces) to show that there will always be weird shapes that our classification can't be expected to cover, but our model will still behave reasonably enough when encountering these cases.
-\end{itemize}}
+\mt{The captions are pretty detailed.
+Do I have to say more?} %\autoref{fig:ugg-heatmap} shows predictions and heatmaps for the logo from an UGG shoe.
+The model correctly identifies triangles and text, and references appropriate regions of the image.
+The model identifies a circle as well; while it could be argued that the triangles form an implicit circle, the heatmap demonstrates that the circular portion of the image is the 'G', which would not be a closed figure labeled as a circle.
+The model's prediction of the presence of a circle is understandable, but incorrect.
+In \autoref{fig:seychelles-heatmap}, the model assigns high probabilities to text and circle.
+The prediction of text is correct, but the prediction of circle is not---there are very round shapes, but they do not form a closed figure that could be identified as a circle.
+\autoref{fig:dc-heatmap} shows a DC shoe with repeating instances of a shape that is a combination of a circle and quadrilateral, but does not technically fall into either of those categories by our definition.
+The model assigns high probabilities of circle and quadrilateral and indicates that the proper image features are being used for this classification.
+The prediction of polygon is incorrect but reasonable considering that many small hexagons are indistinguishable from circles.
+
+% \fix{Choose three (or four) heatmaps:
+% \begin{itemize}
+% \item An image with multiple true shapes that the model clearly predicts well, to prove that it has learned the features we think it should (probably an UGG logo)
+% \item (Seychelles circle/text confusion)
+% \item \sout{An image with high quad/tri/poly confusion, but the model uses appropriate features to make educated guesses}
+% \item An image with no true classes but would yield false positives (e.g., corks, puzzle pieces) to show that there will always be weird shapes that our classification can't be expected to cover, but our model will still behave reasonably enough when encountering these cases.
+% \end{itemize}}
 
@@
 sapply(imgs, function(x) {
+  #cat(x);
   calc_heatmap(x, loaded_model) %>%
@@
 
-<<current-heatmaps, eval = F, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.cap = c("A section of an Ugg shoe containing the logo.
+<<current-heatmaps, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.cap = c("A section of an UGG shoe containing the logo.
 The model correctly identifies triangles and text (and references the correct portion of the image in both cases).
@@
 The model's prediction of the presence of a circle is understandable, but incorrect.\\label{fig:ugg-heatmap}", "In this image of a Seychelles shoe, the model identifies text and circle.
-The prediction of text is correct; the prediction of circle is not - there are very round shapes, but they do not form a closed figure that could be identified as a circle.\\label{fig:seychelles-heatmap}", "polygon", "corks"), echo = F, message = F, warning = F, fig.pos = "!p">>=
+The prediction of text is correct; the prediction of circle is not---there are very round shapes, but they do not form a closed figure that could be identified as a circle.\\label{fig:seychelles-heatmap}", "This image from a DC shoe contains repeating instances of a shape that appears to be a combination of a circle and quadrilateral, but does not technically fall into any of the nine categories used in our labeling scheme.
+The model assigns high probabilities of circle and quadrilateral and indicates that the proper image features are being used for this classification.
+The prediction of polygon is incorrect but reasonable considering that many small hexagons are indistinguishable from circles.\\label{fig:dc-heatmap}"), echo = F, fig.pos = "!p">>=
 
@@
 knitr::include_graphics(heatmaps[grepl("text-seychelles", heatmaps)])
-knitr::include_graphics(heatmaps[grepl("polygon_recolor", heatmaps)])
-knitr::include_graphics(heatmaps[grepl("corks", heatmaps)])
+#knitr::include_graphics(heatmaps[grepl("polygon_recolor", heatmaps)])
+knitr::include_graphics(heatmaps[grepl("dc", heatmaps)])
 
@@
 \section{Conclusions}
-%\svp{Rework this to address 3 main conclusions - 1: NNs work for automatic feature ID, 2: Different levels of feature variability within classes, and 3: correcting for contrast is important.}
-
 % Restate goal and briefly summarize methods
@@
 As most of the misclassifications are systematic, it is clear that the model is able to learn features and apply rules consistently.
-We expect that future modifications to the labeling criteria and data processing pipeline will improve the model's accuracy, but the 91\% validation accuracy \fix{really should be looking at test set accuracy, but I don't know that number off the top of my head...}\mt{See chunk below} is promising considering that we do not have estimates of the consistency of examiners' estimates on ambiguous features.
+We expect that future modifications to the labeling criteria and data processing pipeline will improve the model's accuracy, but the 91\% validation accuracy is promising considering that we do not have estimates of the consistency of examiners' estimates on ambiguous features.
 Taken together, these factors show that Convolutional Neural Networks are a useful tool for classifying a large number of outsoles within a common classification scheme and can ultimately help streamline the current process of sorting and comparing shoes by class characteristics.

### 2019-06-27 paper/paper.Rnw
response

@@
 \mt{The captions are pretty detailed.
-Do I have to say more?} %\autoref{fig:ugg-heatmap} shows predictions and heatmaps for the logo from an UGG shoe.
+Do I have to say more?} \svp{Yes, but make it a bit more general - reiterate that there are overlapping features and ambiguous shapes; the heatmaps demonstrate that the model is cuing into accurate features, but that these features don't always combine in the ways necessary to result in a human-assigned label.}
+%\autoref{fig:ugg-heatmap} shows predictions and heatmaps for the logo from an UGG shoe.
 The model correctly identifies triangles and text, and references appropriate regions of the image.

### 2019-06-28 paper/paper.Rnw
Finish text for heatmaps

--- 2019-06-26 paper/paper.Rnw
@@
 Using these graphics, we can identify which locations in an image contribute most to the model's predictions.
+
 \mt{Figures \ref{fig:ugg-heatmap}-\ref{fig:dc-heatmap} show a number of images alongside heatmaps for the four classes with the highest output probabilities for each image.
-These heatmaps can confirm that a number of misclassifications are a result of overlapping features between classes, and that the model is using reasonable features to make predictions.}
-
-\mt{The captions are pretty detailed.
-Do I have to say more?} %\autoref{fig:ugg-heatmap} shows predictions and heatmaps for the logo from an UGG shoe.
+\autoref{fig:ugg-heatmap} shows predictions and heatmaps for the logo from an UGG shoe.
 The model correctly identifies triangles and text, and references appropriate regions of the image.
-The model identifies a circle as well; while it could be argued that the triangles form an implicit circle, the heatmap demonstrates that the circular portion of the image is the 'G', which would not be a closed figure labeled as a circle.
-The model's prediction of the presence of a circle is understandable, but incorrect.
-In \autoref{fig:seychelles-heatmap}, the model assigns high probabilities to text and circle.
-The prediction of text is correct, but the prediction of circle is not---there are very round shapes, but they do not form a closed figure that could be identified as a circle.
+The prediction of a circle is shown to be associated with the letter 'G'; although this prediction is technically incorrect, it indicates that the model has learned to associate curved shapes with circles.
+This association is again seen in the high predicted probability of circle in \autoref{fig:seychelles-heatmap}.
+In this case, the prediction of text is again correct,, but the prediction of circle is not---there are very round shapes, but they do not form a closed figure that could be identified as a circle.
 \autoref{fig:dc-heatmap} shows a DC shoe with repeating instances of a shape that is a combination of a circle and quadrilateral, but does not technically fall into either of those categories by our definition.
 The model assigns high probabilities of circle and quadrilateral and indicates that the proper image features are being used for this classification.
-The prediction of polygon is incorrect but reasonable considering that many small hexagons are indistinguishable from circles.
-
-% \fix{Choose three (or four) heatmaps:
-% \begin{itemize}
-% \item An image with multiple true shapes that the model clearly predicts well, to prove that it has learned the features we think it should (probably an UGG logo)
-% \item (Seychelles circle/text confusion)
-% \item \sout{An image with high quad/tri/poly confusion, but the model uses appropriate features to make educated guesses}
-% \item An image with no true classes but would yield false positives (e.g., corks, puzzle pieces) to show that there will always be weird shapes that our classification can't be expected to cover, but our model will still behave reasonably enough when encountering these cases.
-% \end{itemize}}
+Additionally, the prediction of polygon is incorrect but reasonable considering that many small hexagons are indistinguishable from circles.
+These heatmaps can confirm that a number of misclassifications are a result of overlapping features between classes, and that the model is using reasonable features to make predictions.}
 

### 2019-06-28 paper/paper.Rnw
Add abstract

@@
 \begin{abstract}
-This template helps you to create a properly formatted \LaTeX\ manuscript.
+\mt{Current practices to assess similarity in forensic shoeprint analysis typically require examiners to manually identify and label features in images of a shoe prints or outsoles.
+This research aims to automate such feature identification and labeling within a feature set similar to that already used by forensic examiners.
+A set of geometric features was developed based on common geometric shapes, such as circles and quadrilaterals.
+A new classifier was then trained for the pretrained convolutional neural network base of VGG16 to create a model to automatically classify portions of outsole images into the new geometric scheme of outsole class characteristics.
+In general, the model performs well on images with unambiguous shapes and moderate color contrast.
+Although systematic misclassifications arise from the inherent overlap of features in geometric shapes, model diagnostics show that such misclassifications are generally consistent and based on appropriate features
+%, indicating that additional label modifications may remedy these confusions
+%, indicating reasonable model performance
+.
+Additional improvements may be realized by modifying the labeling scheme to reduce feature overlap and by preprocessing the images to improve contrast.
+This new approach represents a significant improvement to the current manual classification of footwear tread patterns, facilitating new modes of data collection and automatic processing that will expand the data available for assessment of footwear class characteristic frequency in the population.}
 \end{abstract}

### 2019-06-28 paper/paper.Rnw
modified abstract

--- 2019-06-28 paper/paper.Rnw
@@
 \begin{abstract}
-This template helps you to create a properly formatted \LaTeX\ manuscript.
+Currently, to assess similarity in forensic shoeprint analysis, examiners must manually identify and label features in images of shoe prints or outsoles.
+In this paper, we discuss an algorithm for automatically labeling shoe outsole images using the same feature space examiners use: a set of features based on geometric shapes such as circles, stars, and quadrilaterals.
+The automatic labeling algorithm performs well on images with unambiguous shapes and moderate color contrast, though some systematic errors arise from the physical similarities of geometric shapes.
+Additional improvements may be realized by modifying the labeling scheme to reduce feature overlap and by preprocessing the images to improve contrast.
+This new approach represents a significant improvement to the current manual classification of footwear tread patterns, and could be integrated into current matching databases to reduce the amount of time examiners spend on data entry.
 \end{abstract}

### 2019-07-01 paper/paper.Rnw
Fix typos

@@
 Many convolutional neural networks have been trained on a subset of images from an online image database called ImageNet \citep{dengImageNetLargeScaleHierarchical}.
-In particular, there is a standard set of images from ImageNet, consisting of 1.2 million images that span 1,000 catetories, that is commonly used in computer vision tasks.
+In particular, there is a standard set of images from ImageNet, consisting of 1.2 million images that span 1,000 categories, that is commonly used in computer vision tasks.
 CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously.
@@
 \caption[VGG16 structure]{VGG16 consists of five convolutional blocks that make up the model base.
-Each convolutional block contains an increasing number of increasingly complex \mt{feature-detecting filters}.
+Each convolutional block contains an increasing number of increasingly complex feature-detecting filters.
 After the convolutional blocks, the fully connected layers of the model head are used to make global connections between separate features.}\label{fig:VGG16-structure}
@@
 Although the filters of VGG16 have been trained and optimized on the ImageNet data, the large number of filters span a wide variety of shapes and features, many of which are useful when examining tread patterns.
-Just as humans can detect the features of an unfamiliar object (consituent parts, edges, textures, and corners), a CNN trained on a wide set of objects can be used to classify unrelated images, such as shoe soles.
+Just as humans can detect the features of an unfamiliar object (constituent parts, edges, textures, and corners), a CNN trained on a wide set of objects can be used to classify unrelated images, such as shoe soles.
 
@@
 Thus, to determine the accuracy of the model predictions, we must select a threshold for each label to discretize the model predictions.
-If the threshold is too low, moderate probabilites will be more likely to exceed the threshold, which increases the true positive and false positive predictions in the test data.
+If the threshold is too low, moderate probabilities will be more likely to exceed the threshold, which increases the true positive and false positive predictions in the test data.
 Conversely, if the threshold is too high, moderate probabilities will more often fail to meet the threshold, which increases the number of true and false negative predictions.
@@
 \autoref{fig:overall-roc} shows the ROC curve for our model across all classes, and \autoref{fig:class-roc} shows the curve for each class.
-ROC curves can be summerized by the Area Under the Curve (AUC), which quantifies the overall accuracy of the model.
+ROC curves can be summarized by the Area Under the Curve (AUC), which quantifies the overall accuracy of the model.
 Perfect prediction would be indicated by a right angle along the upper left corner of the plot, with a corresponding AUC of 1, and a diagonal line with an AUC of 0.5 would indicate that the classification method performs no better than random chance.
@@
 The saturation of blue color corresponds to the strength of the predicted probability, and the navy blue border around an image indicates that the predicted probability is above the EER threshold for the given class.
-Note that the EER is not the same for all classes; in other words, the same probablity value assigned across different categories may yield different conclusions about whether the model predicts the shape is present or absent in the image.
+Note that the EER is not the same for all classes; in other words, the same probability value assigned across different categories may yield different conclusions about whether the model predicts the shape is present or absent in the image.
 
 The first rows of \autoref{fig:exemplars} are images which contain only one of the nine classes.
-Most of the predictions for these images are promising; there are some false-positive labels that indicate confusions between circle/text and polygon/quad, but all of the true shapes are corectly identified by the model above their respective EER.
+Most of the predictions for these images are promising; there are some false-positive labels that indicate confusions between circle/text and polygon/quad, but all of the true shapes are correctly identified by the model above their respective EER.
 The next image belongs to the ``other" category, as it does not contain any of the nine classes; the model correctly assigns near-zero probabilities to each class.
@@
 In particular, the first two of these images contain text that use the capital letter ``G", the next two images contain no circular shapes, and the final 3 images contain pure circles that do not appear to belong to text.
-The model is able to identify the circle in all cases where a circle is truly present, regardless of whether the circle is part of the text or it is distinct; however, the circle prediction is also strong in three of the four images that do not contain a circle, which indicates that the model is having dificulty distinguishing the two categories.
+The model is able to identify the circle in all cases where a circle is truly present, regardless of whether the circle is part of the text or it is distinct; however, the circle prediction is also strong in three of the four images that do not contain a circle, which indicates that the model is having difficulty distinguishing the two categories.
 While the circle predictions are often above the EER for letters such as G and S, these predictions are generally much lower than those where distinct circles or Os are seen, indicating that there are some indications that features are not as strong in the quasi-circle shapes, but that the model is not able to completely separate Os from C, S, and G shapes which are also very circular.
@@
 Similarly, polygons and triangles produce a moderate number of false positive predictions in other categories, as evidenced by the vertical bands for these categories.
-It is also apparent that circles and text are commonly confused, which supports our earlier obsesrvartions from \autoref{fig:circle-text-consistency} that these two classes are difficult for the model to distinguish, in part because circles are relatively common components of latin script characters.
+It is also apparent that circles and text are commonly confused, which supports our earlier observations from \autoref{fig:circle-text-consistency} that these two classes are difficult for the model to distinguish, in part because circles are relatively common components of Latin script characters.
 
@@
 Features from outsoles that are primarily black or white in color are typically not as prominent as features from images with more varied color schemes.
-As a result, the features in low contrast images are not easily resolved by the filters; this is propogated through both model training and prediction of new images.
+As a result, the features in low contrast images are not easily resolved by the filters; this is propagated through both model training and prediction of new images.
 In other words, the model `learns' that when there are few meaningful features with which to determine which shapes are present, the best strategy for minimizing loss is to predict diffuse probabilities across many classes, as seen in \autoref{fig:low-contrast-consistency}.
@@
 
-\mt{Figures \ref{fig:ugg-heatmap}-\ref{fig:dc-heatmap} show a number of images alongside heatmaps for the four classes with the highest output probabilities for each image.
+Figures \ref{fig:ugg-heatmap}-\ref{fig:dc-heatmap} show a number of images alongside heatmaps for the four classes with the highest output probabilities for each image.
 \autoref{fig:ugg-heatmap} shows predictions and heatmaps for the logo from an UGG shoe.
 The model correctly identifies triangles and text, and references appropriate regions of the image.
-The prediction of a circle is shown to be associated with the letter 'G'; although this prediction is technically incorrect, it indicates that the model has learned to associate curved shapes with circles.
+The prediction of a circle is shown to be associated with the letter ``G"; although this prediction is technically incorrect, it indicates that the model has learned to associate curved shapes with circles.
 This association is again seen in the high predicted probability of circle in \autoref{fig:seychelles-heatmap}.
@@
 Additionally, the prediction of polygon is incorrect but reasonable considering that many small hexagons are indistinguishable from circles.
-These heatmaps can confirm that a number of misclassifications are a result of overlapping features between classes, and that the model is using reasonable features to make predictions.}
+These heatmaps can confirm that a number of misclassifications are a result of overlapping features between classes, and that the model is using reasonable features to make predictions.
 
@@
 The model correctly identifies triangles and text (and references the correct portion of the image in both cases).
-The model identifies a circle as well; while it could be argued that the triangles form an implicit circle, the heatmap demonstrates that the circular portion of the image is the 'G', which would not be a closed figure labeled as a circle.
+The model identifies a circle as well; while it could be argued that the triangles form an implicit circle, the heatmap demonstrates that the circular portion of the image is the ``G\", which would not be a closed figure labeled as a circle.
 The model's prediction of the presence of a circle is understandable, but incorrect.\\label{fig:ugg-heatmap}", "In this image of a Seychelles shoe, the model identifies text and circle.

### 2019-07-01 paper/paper.Rnw
Fix NIST attribution

--- 2019-06-28 paper/paper.Rnw
@@
 \title{A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}}
-\tnotetext[t1]{This document is the results of the research
-project funded ....}
+\tnotetext[t1]{This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement #70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.}
 

### 2019-07-01 paper/paper.Rnw
fix pound sign

--- 2019-07-01 paper/paper.Rnw
@@
 \title{A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}}
-\tnotetext[t1]{This document is the results of the research
-project funded ....}
+\tnotetext[t1]{This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.}
 

### 2019-07-01 paper/paper.Rnw
More tweaks to header/author statements

@@
 %% Group authors per affiliation:
-% \author{Miranda Tilton and Susan Vanderplas\fnref{csafe}}
-% \address{Iowa State University Statistics Department}
-% \fntext[csafe]{195 Durham Center, 613 Morill Rd., Ames, IA 50011}
+\author{Miranda Tilton and Susan Vanderplas\fnref{csafe}}
+\address{Center for Statistical Applications in Forensic Evidence, Iowa State University}
+\fntext[csafe]{195 Durham Center, 613 Morill Rd., Ames, IA 50011.}
 
 % or include affiliations in footnotes:
-\author[1]{Miranda Tilton and Susan Vanderplas\corref{cor1}\fnref{isu}}
-\cortext[cor1]{Corresponding author}
-\ead{srvander@iastate.edu}
-
-\address[1]{195 Durham Center, 613 Morill Rd, Ames, IA 50011}
-\fntext[isu]{Center for Statistical Applications in Forensic Evidence, Iowa State University}
+% \author[1]{Miranda Tilton and Susan Vanderplas\corref{cor1}\fnref{isu}}
+% \cortext[cor1]{Corresponding author}
+% \ead{srvander@iastate.edu}
+%
+% \address[1]{195 Durham Center, 613 Morill Rd, Ames, IA 50011}
+% \fntext[isu]{Center for Statistical Applications in Forensic Evidence, Iowa State University}
 

### 2019-07-01 paper/paper.Rnw
Fix title of CSAFE (Statistics *and* Applications in...)
Separate keywords with semicolons

@@
 \author{Miranda Tilton and Susan Vanderplas\fnref{csafe}}
-\address{Center for Statistical Applications in Forensic Evidence, Iowa State University}
+\address{Center for Statistics and Applications in Forensic Evidence, Iowa State University}
 \fntext[csafe]{195 Durham Center, 613 Morill Rd., Ames, IA 50011.}
@@
 % \address[1]{195 Durham Center, 613 Morill Rd, Ames, IA 50011}
-% \fntext[isu]{Center for Statistical Applications in Forensic Evidence, Iowa State University}
+% \fntext[isu]{Center for Statistics and Applications in Forensic Evidence, Iowa State University}
 
@@
 \begin{keyword}
-footwear, class characteristics, computer vision, neural networks
+Footwear; class characteristics; computer vision; neural networks
 \end{keyword}

### 2019-07-01 paper/paper.Rnw
doubleblind mode

--- 2019-07-01 paper/paper.Rnw
@@
-\documentclass[review]{elsarticle}
+\documentclass[doubleblind,review]{elsarticle}
 
@@
 %% Group authors per affiliation:
-\author{Miranda Tilton and Susan Vanderplas\fnref{csafe}}
-\address{Center for Statistical Applications in Forensic Evidence, Iowa State University}
-\fntext[csafe]{195 Durham Center, 613 Morill Rd., Ames, IA 50011.}
+\author{Miranda Tilton and Susan Vanderplas}
+\address{Center for Statistical Applications in Forensic Evidence, Iowa State University. 195 Durham Center, 613 Morill Rd., Ames, IA 50011}
 

### 2019-07-01 paper/paper.Rnw
fix csafe acronym

--- 2019-07-01 paper/paper.Rnw
@@
 %% Group authors per affiliation:
-\author{Miranda Tilton and Susan Vanderplas\fnref{csafe}}
-\address{Center for Statistics and Applications in Forensic Evidence, Iowa State University}
-\fntext[csafe]{195 Durham Center, 613 Morill Rd., Ames, IA 50011.}
+\author{Miranda Tilton and Susan Vanderplas}
+\address{Center for Statistics and Applications in Forensic Evidence, Iowa State University. 195 Durham Center, 613 Morill Rd., Ames, IA 50011}
 

### 2019-07-01 paper/coverletter.tex
cover letter

\documentclass[11pt]{article}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{float}
% \setlength{\textheight}{26cm}
% \setlength{\textwidth}{18cm}
\topmargin0in

\oddsidemargin-1cm
\evensidemargin0in
\textheight9in
\textwidth7in
\setlength{\parindent}{0em}

\pagestyle{fancyplain}
\addtolength{\headheight}{\baselineskip}
% \addtolength{\headheight}{43pt}
\rhead{}
\cfoot{}
\markright{{\hfill\large \sc Center for Statistics and Applications in Forensic Evidence}\\{\hfill\sc Iowa State University} \\{\hfill 195 Durham Center, 613 Morill Rd., Ames, IA 50011}}

\begin{document}
C. Jackowski \\
Institute of Forensic Medicine\vspace{.5em}\\
C. Cattaneo \\
Università degli Studi di Milano \vspace{2em}

\today\\\vspace{1em}

{\bf Re: Submission to Forensic Science International}\\\vspace{2em}

Dear Professor Jackowski and Professor Cattaneo, \vspace{1em}

Please find attached a manuscript, ``A Convolutional Neural Network for Outsole Recognition" intended for submission to FSI. \vspace{1em}

The manuscript describes a computer vision algorithm which automates the recognition of geometric shapes in images of shoe outsoles. This algorithm uses a scheme similar to that used by footwear forensic analysts, unlike other convolutional neural network based analyses in footwear forensics. We describe the model's structure, capabilities, and overall performance; we also explore situations where the model's automatic recognition process does not succeed and provide some possible solutions for these issues going forward. We believe this work lays the foundation to bridge the gap between theoretical applications of computer vision in footwear forensics and work that is done in the crime lab. \vspace{1em}

We hope that you find the manuscript acceptable for publication in FSI. If there is any additional information we can provide, please feel free to contact us directly. \vspace{0.25in}

Sincerely,\\
\begin{minipage}{.25\linewidth}
\vspace*{.5cm}
\includegraphics[width=3in,keepaspectratio=TRUE]{signature}
\vspace*{.5cm}
\end{minipage}

Miranda Tilton and Susan Vanderplas\\
Center for Statistics and Applications in Forensic Evidence\\
Iowa State University

\end{document}

### 2020-03-13 paper/paper - revision.Rnw
Very rough draft of new introduction, in the order I want but using crude sentences (for now)

--- 2019-07-01 paper/paper.Rnw
@@
 \section{Introduction}
-Most forensic examinations of footwear pattern evidence are limited to class characteristic based matches; that is, matches based on the positioning of geometric features within the tread pattern.
+
+NEW OUTLINE
+
+In the US, real forensic analysis of shoe print evidence is varied and difficult.
+It is often impossible to conduct an individualizing analysis when a print is partial or low quality.
+Beyond the question of identification, it is more common (and arguably more tractable) to consider class characteristics.
+At the class level,  one question of interest \mt{motivate this more?} is calculating random match probability: Given a questioned shoe print, how many shoes in the relevant population have a similar tread pattern?
+We consider this question as distinct from make and model determination, since many shoes of different makes and models can still have very similar tread patterns (e.g., knockoffs, same models over different years, etc)
+
+From a statistical perspective, random match probability could be calculated by sampling many shoes from a given population and counting the number of shoes in the sample that are "similar" to the given print.
+In practice, one potential method of sampling would be a low-profile scanner placed in a high traffic area.
+Such a scanner could take pictures of shoes walking past, and those images could then be used to create a database of shoes relevant to the local area.
+\mt{Are cameras more robust/weatherproof than pressure scanners?
+Why images instead of EverOS impressions?} \mt{What is the best way to say that such a scanner already exists, and/or that the grant is approved?} Assuming that such a database of images is available, the task becomes assessing the similarity of prints.
+This is done by identifying relevant features in the scans and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.
+
+CNNS have become the gold standard of image recognition.
+\mt{salvage the other intro to explain why simpler methods aren't sufficient.} There is precedent for using machine learning to identify tread patterns \mt{hunt through papers to find relevant references}
+
+%Most forensic examinations of footwear pattern evidence are limited to class characteristic based matches; that is, matches based on the positioning of geometric features within the tread pattern.
 Some of these features may include brand logos and size markings, while others are more generic geometric shapes, such as circles, triangles, stars, and chevrons.
@@
 
-In the wider field of computer vision, attention has turned to more robust methods for image recognition, such as convolutional neural networks (CNNs), which are capable of achieving near-human accuracy under even degraded image conditions \citep{krizhevskyImageNetClassificationDeep2012}.
+%In the wider field of computer vision, attention has turned to more robust methods for image recognition, such as convolutional neural networks (CNNs), which are capable of achieving near-human accuracy under even degraded image conditions \citep{krizhevskyImageNetClassificationDeep2012}.
 These networks are designed to mimic the process of human vision, and typically involve application of sets of filters, some of which mimic the filters used in the low-level techniques used in early automatic classification attempts.
@@
 
-In this paper, we discuss an alternate approach that uses an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images.
+%In this paper, we discuss an alternate approach that uses an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images.
 Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain.

### 2020-03-18 paper/paper - revision.Rnw
Flesh out the new outline

@@
 \begin{abstract}
+\mt{Update framing}
 Currently, to assess similarity in forensic shoeprint analysis, examiners must manually identify and label features in images of shoe prints or outsoles.
@@
 
-In the US, real forensic analysis of shoe print evidence is varied and difficult.
-It is often impossible to conduct an individualizing analysis when a print is partial or low quality.
-Beyond the question of identification, it is more common (and arguably more tractable) to consider class characteristics.
-At the class level,  one question of interest \mt{motivate this more?} is calculating random match probability: Given a questioned shoe print, how many shoes in the relevant population have a similar tread pattern?
-We consider this question as distinct from make and model determination, since many shoes of different makes and models can still have very similar tread patterns (e.g., knockoffs, same models over different years, etc)
+In the United States, forensic analysis of shoe print evidence in criminal cases is varied and difficult.
+The process of collecting print evidence from a crime scene is not standardized, and it is often impossible to conduct an individualizing analysis when a print is partial or of low quality.
+In cases where individualizing characteristics are not considered (estimated at 95\% of cases in the United States according to some experts \footnote{Leslie Hammer, presentation to CSAFE on March 5, 2018}), it is useful to be able to assess the random match probability.
+In other words, given a questioned shoe print, what is the probability that a randomly selected shoe in the relevant population has a similar tread pattern?
+Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns (e.g., knockoffs, same models over different years, etc)
 
-From a statistical perspective, random match probability could be calculated by sampling many shoes from a given population and counting the number of shoes in the sample that are "similar" to the given print.
-In practice, one potential method of sampling would be a low-profile scanner placed in a high traffic area.
+From a statistical perspective, random match probability could be calculated by sampling a large number of shoes from a given population and counting the number of shoes in the sample whose tread patterns are similar to the given print.
+In practice, one potential method of sampling could be a low-profile scanner placed in a high traffic area.
 Such a scanner could take pictures of shoes walking past, and those images could then be used to create a database of shoes relevant to the local area.
 \mt{Are cameras more robust/weatherproof than pressure scanners?
-Why images instead of EverOS impressions?} \mt{What is the best way to say that such a scanner already exists, and/or that the grant is approved?} Assuming that such a database of images is available, the task becomes assessing the similarity of prints.
-This is done by identifying relevant features in the scans and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.
+Why images instead of EverOS impressions?} \mt{(Note: a grant proposal for such a scanner has been approved by NIJ, and there currently exists a working prototype.)} Assuming that such a database of images is available, the task becomes assessing the similarity of the shoe tread patterns in the images.
+This is done by identifying a set of relevant features and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity measures based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.
 

### 2020-03-18 paper/paper - revision.Rnw
Minor wording changes

@@
 
-NEW OUTLINE
-
 In the United States, forensic analysis of shoe print evidence in criminal cases is varied and difficult.
 The process of collecting print evidence from a crime scene is not standardized, and it is often impossible to conduct an individualizing analysis when a print is partial or of low quality.
-In cases where individualizing characteristics are not considered (estimated at 95\% of cases in the United States according to some experts \footnote{Leslie Hammer, presentation to CSAFE on March 5, 2018}), it is useful to be able to assess the random match probability.
+In fact, some experts estimate that 95\% of cases in the United States do not consider individualizing characteristics of shoeprint evidence \footnote{Leslie Hammer, presentation to CSAFE on March 5, 2018}).
+Even so, in these cases it still may be useful, and much more tractable, to assess the random match probability of the footwear evidence.
 In other words, given a questioned shoe print, what is the probability that a randomly selected shoe in the relevant population has a similar tread pattern?
-Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns (e.g., knockoffs, same models over different years, etc)
+Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns (e.g., knockoffs, same models over different years, etc).
 

### 2020-03-27 paper/paper - revision.Rnw
Flesh out introduction, make formal to-do list

--- 2020-03-18 paper/paper - revision.Rnw
@@
 
-NEW OUTLINE
+\mt{General paper to-do:}
+\mt{1. Find a paper to cite the 95\% statistic}
+\mt{Option: Shoe marks are most common at a crime scene (Rida et al., 2019)}
+\mt{2. A zotero reference is missing details, remember to go find/fix it}
+\mt{3. Cite papers with manual data collection}
+\mt{4. Susan will send pictures of the prototype scanner from the grant proposal}
+\mt{5. Argue in methods that we know real image quality will be lower but we needed to start somewhere}
+\mt{6. Our goal is to narrow down what type of model works and assess the model(s) limitations}
+\mt{7.
+Add to conclusion: Overall, this concept seems to work, but we need to modify the classificiation scheme.
+Thus, this work is publishable, because it needs work and community input.
+Need more research into human comphrehensible and yet useful classification schemes for this purpose.
+We need better class scheme, important and promising, getting this right matters}
+\mt{8. Update abstract to reflect new framing of paper}
+\mt{}
 
@@
 The process of collecting print evidence from a crime scene is not standardized, and it is often impossible to conduct an individualizing analysis when a print is partial or of low quality.
-In cases where individualizing characteristics are not considered (estimated at 95\% of cases in the United States according to some experts \footnote{Leslie Hammer, presentation to CSAFE on March 5, 2018}), it is useful to be able to assess the random match probability.
+In fact, some experts estimate that 95\% of cases in the United States do not consider individualizing characteristics of shoeprint evidence \fix{(cite)}.
+\mt{(see list above)}
+%\footnote{Leslie Hammer, presentation to CSAFE on March 5, 2018}).
+Even so, in these cases it still may be \mt{both possible and useful} to assess the random match probability of the footwear evidence \mt{at the level of class characteristics}.
 In other words, given a questioned shoe print, what is the probability that a randomly selected shoe in the relevant population has a similar tread pattern?
-Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns (e.g., knockoffs, same models over different years, etc)
+Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns (e.g., knockoffs, similar models over different years, etc).
 
-From a statistical perspective, random match probability could be calculated by sampling a large number of shoes from a given population and counting the number of shoes in the sample whose tread patterns are similar to the given print.
+From a statistical perspective, random match probability could theoretically be calculated by sampling a large number of shoes from a given population and counting the number of shoes in the sample whose tread patterns are similar to the given print.
 In practice, one potential method of sampling could be a low-profile scanner placed in a high traffic area.
-Such a scanner could take pictures of shoes walking past, and those images could then be used to create a database of shoes relevant to the local area.
-\mt{Are cameras more robust/weatherproof than pressure scanners?
-Why images instead of EverOS impressions?} \mt{(Note: a grant proposal for such a scanner has been approved by NIJ, and there currently exists a working prototype.)} Assuming that such a database of images is available, the task becomes assessing the similarity of the shoe tread patterns in the images.
-This is done by identifying a set of relevant features and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity measures based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.
+Such a scanner could take \mt{impressions or} pictures of shoes walking past, and those scans could then be used to create a database of shoes relevant to the local area.
 
-CNNS have become the gold standard of image recognition.
-\mt{salvage the other intro to explain why simpler methods aren't sufficient.} There is precedent for using machine learning to identify tread patterns \mt{hunt through papers to find relevant references}
+However, studies by \fix{authors (yr)} and \fix{authors (yr)} demonstrate that even \fix{[a couple hundred]} shoe prints constitute an unweildy amount of data requiring \fix{[a large amount of data processing]}.
+Thus, in parallel with the development of such a scanner, it is necessary to develop a way to process large amounts of data automatically.
+\mt{The primary focus of this paper is to develop one such method for large-scale automatic data processing; in this case, "processing" refers to assessing} the similarity of the shoe tread patterns in the collected data.
+%\fix{Cut this?: This is done by identifying a set of relevant features and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity measures based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.}
+
+Although footwear impressions are more commonly used in practice for footwear analysis than outsole images, early prototypes of a footwear scanner have suggested that cameras are more robust than pressure scanners for placement in public locations that may experience inclement weather or other non-controlled conditions.
+Thus, \fix{this is an image recognition problem}.
+
+% From original intro:
+Several attempts have been made to automatically identify features in both prints and shoe tread images, with low-level image analysis methods including fractal decomposition \citep{alexanderAutomaticClassificationRecognition1999}, scale-invariant feature recognition \citep{pavlouAutomaticExtractionClassification2006}, Fourier-Mellin transformation \citep{guehamAutomaticRecognitionShoeprints2008}, and other classical image analysis methods.
+Some of these low-level methods perform relatively well in good conditions \citep{luostarinenMeasuringAccuracyAutomatic2014a} but have degraded performance under conditions which are suboptimal, including those commonly found at crime scenes.
+
+% From original intro:
+In the wider field of computer vision, attention has turned to more robust methods for image recognition, such as convolutional neural networks (CNNs), which are capable of achieving near-human accuracy under even degraded image conditions \citep{krizhevskyImageNetClassificationDeep2012}.
+These networks are designed to mimic the process of human vision, and typically involve application of sets of filters, some of which mimic the filters used in the low-level techniques used in early automatic classification attempts.
+CNNs have been applied in footwear forensics, making use of more general neural networks optimized to detect objects found in natural scenes, such as trees, animals, buildings, and cars; the pre-trained network's features are then used to compare shoes or prints to determine how well they match \citep{kongCrossDomainImageMatching2019a,kongCrossDomainForensicShoeprint2017,zhangAdaptingConvolutionalNeural2017}.
+Neural network methods appear to be more successful than traditional image analysis techniques, but the features used for matching are not generally informative for humans, who typically classify shoes by patterns and spatial relationships \citep{davisIntelligenceApproachFootwear1981,grossVariabilitySignificanceClass2013}.
+
+There is precedent for using machine learning to identify tread patterns \mt{hunt through papers to find relevant references}
+
+% ORIGINAL INTRO:
 

### 2020-03-27 paper/paper - revision.Rnw
Edits to introduction, ready for Susan's comments

@@
 \mt{3. Cite papers with manual data collection}
+\mt{3.5. Incorporate recommended citations from reviewer 3}
 \mt{4. Susan will send pictures of the prototype scanner from the grant proposal}
@@
 
-From a statistical perspective, random match probability could theoretically be calculated by sampling a large number of shoes from a given population and counting the number of shoes in the sample whose tread patterns are similar to the given print.
+From a statistical perspective, random match probability could theoretically be calculated by sampling a large number of shoes from a given population and calculating the proportion of shoes in the sample whose tread patterns are similar to the given print.
 In practice, one potential method of sampling could be a low-profile scanner placed in a high traffic area.
@@
 Thus, in parallel with the development of such a scanner, it is necessary to develop a way to process large amounts of data automatically.
-\mt{The primary focus of this paper is to develop one such method for large-scale automatic data processing; in this case, "processing" refers to assessing} the similarity of the shoe tread patterns in the collected data.
-%\fix{Cut this?: This is done by identifying a set of relevant features and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity measures based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.}
+\mt{The primary focus of this paper is the development of one such method for large-scale automatic data processing, where "data processing" refers to assessing} the similarity of the shoe tread patterns in the collected data.
 
 Although footwear impressions are more commonly used in practice for footwear analysis than outsole images, early prototypes of a footwear scanner have suggested that cameras are more robust than pressure scanners for placement in public locations that may experience inclement weather or other non-controlled conditions.
-Thus, \fix{this is an image recognition problem}.
+Thus, \fix{we move forward by considering how to assess similarity of tread patterns in images, in particular.} More specifically, this assessment is done by identifying a set of relevant features and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity measures based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.
 
@@
 
-There is precedent for using machine learning to identify tread patterns \mt{hunt through papers to find relevant references}
+% From original intro:
+In this paper, we discuss an alternate approach that uses an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images.
+Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain.
+\fix{Ultimately, we can use the output probabilities to quantify similarity between tread patterns for the ultimate goal of calculating random match probability for a given tread pattern.}
 

### 2020-03-27 paper/paper - revision.Rnw
Update beginning of methods and defend choice of online retailer images

@@
 \section{Materials and Methods}
-%\fix{We may want to mention some of the methods in the introduction for continuity sake...
-Ultimately, most pattern evidence is stored in photographic form; as a result, we must use image analysis methods to create numerical features from the original visual record.
+
+Ultimately, \mt{images collected from a footwear scanner (as described in the previous section) would be} stored in photographic form; as a result, we must use image analysis methods to create numerical features from the original visual record.
 There are a number of methods that may be employed to identify shapes and features in an image, such as Fourier-Mellin transforms, Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}.
@@
 To date, \Sexpr{unique(ann_df$base_image) %>% length()} shoes have been labeled, yielding \Sexpr{nrow(dfunion)} multi-label images.
+
+\fix{Rough idea: A reasonable criticism of this approach is that images from online retailers are of much higher quality than would be expected from the sidewalk scanner proposed earlier in this paper.
+This critique is especially fair considering that CNNs, and neural networks in general, have been shown to perform poorly on test data when it does not resemble the training data (find citation).
+However, this research is so new that we need to start somewhere.
+Once this work shows that CNNs can sufficiently distinguish outsole patterns/features, we can justify the effort to modify this approach for real/degraded images.}
 

### 2020-03-27 paper/paper - revision.Rnw
Move to-do to Trello

@@
 \section{Introduction}
-
-\mt{General paper to-do:}
-\mt{1. Find a paper to cite the 95\% statistic}
-\mt{Option: Shoe marks are most common at a crime scene (Rida et al., 2019)}
-\mt{2. A zotero reference is missing details, remember to go find/fix it}
-\mt{3. Cite papers with manual data collection}
-\mt{3.5. Incorporate recommended citations from reviewer 3}
-\mt{4. Susan will send pictures of the prototype scanner from the grant proposal}
-\mt{5. Argue in methods that we know real image quality will be lower but we needed to start somewhere}
-\mt{6. Our goal is to narrow down what type of model works and assess the model(s) limitations}
-\mt{7.
-Add to conclusion: Overall, this concept seems to work, but we need to modify the classificiation scheme.
-Thus, this work is publishable, because it needs work and community input.
-Need more research into human comphrehensible and yet useful classification schemes for this purpose.
-We need better class scheme, important and promising, getting this right matters}
-\mt{8. Update abstract to reflect new framing of paper}
-\mt{}
 

### 2020-03-30 paper/paper-revision.Rnw
Update abstract

--- 2020-03-27 paper/paper - revision.Rnw
@@
 \begin{abstract}
-\mt{Update framing}
-Currently, to assess similarity in forensic shoeprint analysis, examiners must manually identify and label features in images of shoe prints or outsoles.
+To assess random match probability at the level of class characteristics, one could imagine sampling shoes from a local population and calculating the proportion of shoes "similar" to a given print.
+One implemention of such a procedure could involve 1) a scanner to gather images of shoes and 2) software to automatically label images within a feature space and assess the images for similarity \fix{within} those features.
 In this paper, we discuss an algorithm for automatically labeling shoe outsole images using the same feature space examiners use: a set of features based on geometric shapes such as circles, stars, and quadrilaterals.

### 2020-03-30 paper/paper-revision.Rnw
Update deprecated code and add meeting notes (messy, to be fixed)

@@
 
-library(tidyverse)
 library(magrittr)
 library(tidyverse)
+library(tidyr)
+nest <- nest_legacy
+unnest <- unnest_legacy
 library(ggplot2)
@@
 
+\fix{We are working to dev a scanner using photo technology.
+It takes pics like Zappos.
+While images are degraded in some ways, the fundamental reqs are sim.
+We want to dev the methodology for this before we get actual data.}
+
+\fix{manual collection citations: presentation by Sandy Parent, benedict et al, benedict cites another, the one that says college students and criminals are basically the same}
+
 % From original intro:
@@
 \fix{Ultimately, we can use the output probabilities to quantify similarity between tread patterns for the ultimate goal of calculating random match probability for a given tread pattern.}
+
+\mt{subsection: motivation (w/ pics of scanner)}
 
@@
 However, this research is so new that we need to start somewhere.
+We need to work the kinks out of recognizing characteristics that examiners use.
+We can do that on real data once the scanner has been built and once we know how well our classification works.
+AKA find the pitfalls for clean images rather than introduce too many points of failure at once.
 Once this work shows that CNNs can sufficiently distinguish outsole patterns/features, we can justify the effort to modify this approach for real/degraded images.}
@@
   ungroup() %>%
-  tidyr::nest(c(str, value)) %>%
+  tidyr::nest_legacy(c(str, value)) %>%
   mutate(n_true = purrr::map_int(data, ~sum(.$value)),

### 2020-03-30 paper/code/Generate_Model_Images.R

@@ -560,7 +560,7 @@ plot_onehot_roc <- function(preds, labels, classes = str_to_title(default_classe
   cv <- 1:ncol(preds) %>% set_names(classes)
   tmp <- purrr::map(cv, ~pROC::roc(labels[,.], preds[,.]))
 
-  roc_data <- purrr::map_df(tmp, ~data_frame(tpr = .$sensitivities,
+  roc_data <- purrr::map_df(tmp, ~tibble(tpr = .$sensitivities,
                                              fpr = 1 - .$specificities,
                                              thresholds = .$thresholds,
                                              auc = .$auc[1]), .id = "class") %>%

### 2020-03-30 paper/paper-revision.Rnw
Add citation for prevalence of class characteristic analysis

@@
 The process of collecting print evidence from a crime scene is not standardized, and it is often impossible to conduct an individualizing analysis when a print is partial or of low quality.
-In fact, some experts estimate that 95\% of cases in the United States do not consider individualizing characteristics of shoeprint evidence \fix{(cite)}.
-\mt{(see list above)}
+In fact, some experts estimate that the majority of cases in the United States do not consider individualizing characteristics of shoeprint evidence \citep{smith_2011}.
 %\footnote{Leslie Hammer, presentation to CSAFE on March 5, 2018}).

### 2020-03-30 paper/paper-revision.Rnw
Incorporate scanner context

@@
 Although footwear impressions are more commonly used in practice for footwear analysis than outsole images, early prototypes of a footwear scanner have suggested that cameras are more robust than pressure scanners for placement in public locations that may experience inclement weather or other non-controlled conditions.
-Thus, \fix{we move forward by considering how to assess similarity of tread patterns in images, in particular.} More specifically, this assessment is done by identifying a set of relevant features and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity measures based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.
-
-\fix{We are working to dev a scanner using photo technology.
-It takes pics like Zappos.
-While images are degraded in some ways, the fundamental reqs are sim.
-We want to dev the methodology for this before we get actual data.}
-
-\fix{manual collection citations: presentation by Sandy Parent, benedict et al, benedict cites another, the one that says college students and criminals are basically the same}
+\mt{Thus, we are working to develop a scanner using photo technology} \fix{(reference/cite image of the scanner)}.
+\mt{This scanner would take pictures of shoe outsoles, much like images that are currently found on the websites of online shoe retailers \fix{(reference/cite test image of  an outsole on the scanner)}.
+Although the images from the scanner are more degraded in some ways than those found from online retailers, the fundamental requirements to analyze such data would be similar.} Therefore, \fix{we move forward by considering how to assess similarity of tread patterns in images of outsoles, in particular.} More specifically, this assessment is done by identifying a set of relevant features and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity measures based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.
 

### 2020-04-03 paper/paper-revision.Rnw
Add citations, clean up fix tags

@@
 To assess random match probability at the level of class characteristics, one could imagine sampling shoes from a local population and calculating the proportion of shoes "similar" to a given print.
-One implemention of such a procedure could involve 1) a scanner to gather images of shoes and 2) software to automatically label images within a feature space and assess the images for similarity \fix{within} those features.
-In this paper, we discuss an algorithm for automatically labeling shoe outsole images using the same feature space examiners use: a set of features based on geometric shapes such as circles, stars, and quadrilaterals.
+Our proposed implemention of such a procedure involves 1) a scanner to gather images of shoes and 2) software to automatically label images within a feature space and assess the images for similarity among those features.
+In this paper, we discuss an algorithm for automatically labeling shoe outsole images using a feature space similar to what examiners use: a set of features based on geometric shapes such as circles, stars, and quadrilaterals.
 The automatic labeling algorithm performs well on images with unambiguous shapes and moderate color contrast, though some systematic errors arise from the physical similarities of geometric shapes.
@@
 In the United States, forensic analysis of shoe print evidence in criminal cases is varied and difficult.
-The process of collecting print evidence from a crime scene is not standardized, and it is often impossible to conduct an individualizing analysis when a print is partial or of low quality.
-In fact, some experts estimate that the majority of cases in the United States do not consider individualizing characteristics of shoeprint evidence \citep{smith_2011}.
+The process of collecting print evidence from a crime scene is not standardized, and it is often impossible to conduct an individualizing analysis (i.e., one based on randomly acquired characteristics) when a print is partial or of low quality.
+In fact, the majority of footwear examinations in the United States reach conclusions regarding the correspondence of class characteristics rather than of individualizing characteristics \citep{smith_2011}.
 %\footnote{Leslie Hammer, presentation to CSAFE on March 5, 2018}).
-Even so, in these cases it still may be \mt{both possible and useful} to assess the random match probability of the footwear evidence \mt{at the level of class characteristics}.
-In other words, given a questioned shoe print, what is the probability that a randomly selected shoe in the relevant population has a similar tread pattern?
+In cases where individualizing characteristics are not used, it still may be both possible and useful to assess the random match probability of the footwear evidence at the level of class characteristic; in other words, given a questioned shoe print, what is the probability that a randomly selected shoe in the relevant population has a similar tread pattern?
 Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns (e.g., knockoffs, similar models over different years, etc).
@@
 
-However, studies by \fix{authors (yr)} and \fix{authors (yr)} demonstrate that even \fix{[a couple hundred]} shoe prints constitute an unweildy amount of data requiring \fix{[a large amount of data processing]}.
+However, studies by \citet{parent_sandy_significance_nodate} and \citet{benedict_geographical_2014} demonstrate that even \fix{[a couple hundred]} shoe prints constitute an unweildy amount of data requiring \fix{[a large amount of data processing]}.
 Thus, in parallel with the development of such a scanner, it is necessary to develop a way to process large amounts of data automatically.

### 2020-04-03 paper/paper-revision.Rnw
Add scanner images to paper (as figure) and small text edits

@@
 In fact, the majority of footwear examinations in the United States reach conclusions regarding the correspondence of class characteristics rather than of individualizing characteristics \citep{smith_2011}.
-%\footnote{Leslie Hammer, presentation to CSAFE on March 5, 2018}).
-In cases where individualizing characteristics are not used, it still may be both possible and useful to assess the random match probability of the footwear evidence at the level of class characteristic; in other words, given a questioned shoe print, what is the probability that a randomly selected shoe in the relevant population has a similar tread pattern?
+In the cases where individualizing characteristics are not used, it may be both possible and useful to assess the random match probability of the footwear evidence at the level of class characteristics; in other words, given a questioned shoe print, what is the probability that a randomly selected shoe in the relevant population has a similar tread pattern?
 Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns (e.g., knockoffs, similar models over different years, etc).
 
-From a statistical perspective, random match probability could theoretically be calculated by sampling a large number of shoes from a given population and calculating the proportion of shoes in the sample whose tread patterns are similar to the given print.
+From a statistical perspective, random match probability could theoretically be calculated by sampling a large number of shoes from a given population and calculating the observed proportion of shoes in the sample whose tread patterns are similar to the given print.
 In practice, one potential method of sampling could be a low-profile scanner placed in a high traffic area.
@@
 
-However, studies by \citet{parent_sandy_significance_nodate} and \citet{benedict_geographical_2014} demonstrate that even \fix{[a couple hundred]} shoe prints constitute an unweildy amount of data requiring \fix{[a large amount of data processing]}.
+However, studies by \citep{parent_sandy_significance_nodate} and \citep{benedict_geographical_2014} demonstrate that even a couple hundred shoe prints constitute an unweildy amount of data that then requires \fix{[a large amount of data processing]}.
 Thus, in parallel with the development of such a scanner, it is necessary to develop a way to process large amounts of data automatically.
@@
 Although footwear impressions are more commonly used in practice for footwear analysis than outsole images, early prototypes of a footwear scanner have suggested that cameras are more robust than pressure scanners for placement in public locations that may experience inclement weather or other non-controlled conditions.
-\mt{Thus, we are working to develop a scanner using photo technology} \fix{(reference/cite image of the scanner)}.
-\mt{This scanner would take pictures of shoe outsoles, much like images that are currently found on the websites of online shoe retailers \fix{(reference/cite test image of  an outsole on the scanner)}.
+\mt{Thus, we are working to develop a scanner using photo technology.
+This scanner would take pictures of shoe outsoles, much like images that are currently found on the websites of online shoe retailers \autoref{fig:prototype-scanner}.
 Although the images from the scanner are more degraded in some ways than those found from online retailers, the fundamental requirements to analyze such data would be similar.} Therefore, \fix{we move forward by considering how to assess similarity of tread patterns in images of outsoles, in particular.} More specifically, this assessment is done by identifying a set of relevant features and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity measures based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.
+
+\fix{How to make the scanner images a figure instead of a table?}
+
+\begin{figure}
+\centering
+\setlength\tabcolsep{1mm}
+\begin{tabular}{cc}
+\raisebox{-.5\height}{\includegraphics[width=.45\textwidth]{images/prototype/20200402_151639.jpg}} &
+\raisebox{-.5\height}{\includegraphics[width=.45\textwidth]{images/prototype/GOPR1231-blur-enhance_square.jpg}} \vspace{1mm}\\
+
+\end{tabular}
+\caption[Images of prototype scanner.]{Images of the prototype scanner for collecting a large number of outsole images.}\label{fig:prototype-scanner}
+\end{figure}
 

### 2020-04-09 paper/paper-revision.Rnw
Finish abstract, changes throughout intro and methods

@@
 To assess random match probability at the level of class characteristics, one could imagine sampling shoes from a local population and calculating the proportion of shoes "similar" to a given print.
-Our proposed implemention of such a procedure involves 1) a scanner to gather images of shoes and 2) software to automatically label images within a feature space and assess the images for similarity among those features.
-In this paper, we discuss an algorithm for automatically labeling shoe outsole images using a feature space similar to what examiners use: a set of features based on geometric shapes such as circles, stars, and quadrilaterals.
+Our proposed implemention of such a procedure involves 1) a scanner to gather images of shoes from a local population and 2) software to automatically label images within a feature space and assess the images for similarity among those features.
+In this paper, we discuss an algorithm for automatically labeling shoe outsole images using a classification scheme similar to what examiners use: a set of features based on geometric shapes such as circles, stars, and quadrilaterals.
 The automatic labeling algorithm performs well on images with unambiguous shapes and moderate color contrast, though some systematic errors arise from the physical similarities of geometric shapes.
 Additional improvements may be realized by modifying the labeling scheme to reduce feature overlap and by preprocessing the images to improve contrast.
-This new approach represents a significant improvement to the current manual classification of footwear tread patterns, and could be integrated into current matching databases to reduce the amount of time examiners spend on data entry.
+This new approach represents a significant step toward making tractable the computation of random match probabilities of tread patterns, and could potentially be modified to automate other types of footwear examinations based on class characteristics.
 \end{abstract}
@@
 
-\section{Introduction}
+\section{Introduction}\label{sec:Introduction}
 
-In the United States, forensic analysis of shoe print evidence in criminal cases is varied and difficult.
+In the United States, forensic analysis of shoe print evidence is varied and difficult.
 The process of collecting print evidence from a crime scene is not standardized, and it is often impossible to conduct an individualizing analysis (i.e., one based on randomly acquired characteristics) when a print is partial or of low quality.
@@
 In practice, one potential method of sampling could be a low-profile scanner placed in a high traffic area.
-Such a scanner could take \mt{impressions or} pictures of shoes walking past, and those scans could then be used to create a database of shoes relevant to the local area.
+Such a scanner could take impressions or pictures of shoes walking past, and those scans could then be used to create a database of shoes relevant to the local area.
 
-However, studies by \citep{parent_sandy_significance_nodate} and \citep{benedict_geographical_2014} demonstrate that even a couple hundred shoe prints constitute an unweildy amount of data that then requires \fix{[a large amount of data processing]}.
-Thus, in parallel with the development of such a scanner, it is necessary to develop a way to process large amounts of data automatically.
-\mt{The primary focus of this paper is the development of one such method for large-scale automatic data processing, where "data processing" refers to assessing} the similarity of the shoe tread patterns in the collected data.
+However, a number of footwear studies (such as those by \citep{parent_sandy_significance_nodate} and \citep{benedict_geographical_2014}) have demonstrated that even a couple hundred shoe prints constitute an unweildy amount of data that require a large amount of time and effort to process and analyze.
+Thus, in parallel with the development of the aforementioned scanner, it is necessary to develop a way to process large amounts of data automatically.
+The primary focus of this paper is the development of one such method for large-scale automatic data processing, where "data processing" refers to assessing the similarity of the shoe tread patterns in the collected data.
 
-Although footwear impressions are more commonly used in practice for footwear analysis than outsole images, early prototypes of a footwear scanner have suggested that cameras are more robust than pressure scanners for placement in public locations that may experience inclement weather or other non-controlled conditions.
-\mt{Thus, we are working to develop a scanner using photo technology.
-This scanner would take pictures of shoe outsoles, much like images that are currently found on the websites of online shoe retailers \autoref{fig:prototype-scanner}.
-Although the images from the scanner are more degraded in some ways than those found from online retailers, the fundamental requirements to analyze such data would be similar.} Therefore, \fix{we move forward by considering how to assess similarity of tread patterns in images of outsoles, in particular.} More specifically, this assessment is done by identifying a set of relevant features and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity measures based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.
-
-\fix{How to make the scanner images a figure instead of a table?}
+Although footwear impressions are more commonly used in practice for footwear analysis than outsole images, our early prototypes of a footwear scanner have suggested that cameras are more robust than pressure scanners for placement in public locations, which may experience inclement weather or other non-controlled conditions.
+Thus, we are working to develop a scanner that uses cameras to take pictures of shoe outsoles, much like images that are currently found on the websites of online shoe retailers \autoref{fig:prototype-scanner}.
+Although the images from the scanner are more degraded in some ways than images found from online retailers, the fundamental requirements to analyze such data would be similar.
+Therefore, \fix{we move forward by considering how to assess similarity of tread patterns in images of outsoles, in particular.} More specifically, this assessment is done by identifying a set of relevant features and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity measures based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.
 
@@
 
-\fix{Rough idea: A reasonable criticism of this approach is that images from online retailers are of much higher quality than would be expected from the sidewalk scanner proposed earlier in this paper.
-This critique is especially fair considering that CNNs, and neural networks in general, have been shown to perform poorly on test data when it does not resemble the training data (find citation).
-However, this research is so new that we need to start somewhere.
-We need to work the kinks out of recognizing characteristics that examiners use.
-We can do that on real data once the scanner has been built and once we know how well our classification works.
-AKA find the pitfalls for clean images rather than introduce too many points of failure at once.
-Once this work shows that CNNs can sufficiently distinguish outsole patterns/features, we can justify the effort to modify this approach for real/degraded images.}
+\mt{A reasonable critique of this approach is that images collected from online retailers are of much higher quality than would be expected from the sidewalk scanner proposed in \autoref{sec:Introduction}, and thus the CNN described in this paper is not immediately ready to process real degraded images.
+This concern is especially fair when considering that machine learning methods tend to perform poorly when applied to test data that do not resemble the training data.
+However, the proposed automation methods for this approach are quite novel in the domain of footwear examination, and a fully functioning scanner does not yet exist to collect large amounts of degraded images.
+Thus, the goal of this work is to evaluate the proposed classification scheme and determine whether CNNS can sufficiently distinguish features found on outsoles.
+Addressing these questions with clean retailer images provides both justification and guidance for modifying this approach for real degraded images.}
 

### 2020-04-10 paper/paper-revision.Rnw
Finish polishing intro

@@
 Thus, in parallel with the development of the aforementioned scanner, it is necessary to develop a way to process large amounts of data automatically.
-The primary focus of this paper is the development of one such method for large-scale automatic data processing, where "data processing" refers to assessing the similarity of the shoe tread patterns in the collected data.
+The primary focus of this paper is the development of one such method for large-scale automatic data processing.
 
@@
 Thus, we are working to develop a scanner that uses cameras to take pictures of shoe outsoles, much like images that are currently found on the websites of online shoe retailers \autoref{fig:prototype-scanner}.
-Although the images from the scanner are more degraded in some ways than images found from online retailers, the fundamental requirements to analyze such data would be similar.
-Therefore, \fix{we move forward by considering how to assess similarity of tread patterns in images of outsoles, in particular.} More specifically, this assessment is done by identifying a set of relevant features and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity measures based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.
+Despite the fact that images from the scanner would be more degraded in some ways than images found from online retailers, the fundamental requirements to analyze such data would be similar.
+Thus, retailer images provide a reasonable basis for the development of a method to assess similarity in tread images.
+More specifically, this assessment is done by identifying a set of relevant features and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity measures based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.
 
@@
 \end{figure}
+
+\subsection{Image Analysis}
 
@@
 Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain.
-\fix{Ultimately, we can use the output probabilities to quantify similarity between tread patterns for the ultimate goal of calculating random match probability for a given tread pattern.}
-
-\mt{subsection: motivation (w/ pics of scanner)}
+\fix{Ultimately, the output probabilities can be used to quantify similarity between tread patterns for the ultimate goal of calculating random match probability for a given tread pattern.}
 

### 2020-04-10 paper/paper-revision.Rnw
Push edits with compiled doc (and working references)

@@
 Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns (e.g., knockoffs, similar models over different years, etc).
+\fix{[This is an important point.
+Should I make a bigger deal out of it throughout the paper?]}
 
@@
 
-However, a number of footwear studies (such as those by \citep{parent_sandy_significance_nodate} and \citep{benedict_geographical_2014}) have demonstrated that even a couple hundred shoe prints constitute an unweildy amount of data that require a large amount of time and effort to process and analyze.
+A number of footwear studies (such as those by \citep{parent_sandy_significance_nodate} and \citep{benedict_geographical_2014}) have demonstrated that even a couple hundred shoe prints constitute an unweildy amount of data, which then require a large amount of time and effort to process and analyze.
 Thus, in parallel with the development of the aforementioned scanner, it is necessary to develop a way to process large amounts of data automatically.
@@
 Thus, we are working to develop a scanner that uses cameras to take pictures of shoe outsoles, much like images that are currently found on the websites of online shoe retailers \autoref{fig:prototype-scanner}.
-Despite the fact that images from the scanner would be more degraded in some ways than images found from online retailers, the fundamental requirements to analyze such data would be similar.
-Thus, retailer images provide a reasonable basis for the development of a method to assess similarity in tread images.
+Despite the fact that images from the scanner would be more degraded in some ways than images found from online retailers, the fundamental requirements to analyze such data would be similar, so retailer images provide a reasonable basis for the development of a method to assess similarity in tread images.
 More specifically, this assessment is done by identifying a set of relevant features and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity measures based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.
@@
 This concern is especially fair when considering that machine learning methods tend to perform poorly when applied to test data that do not resemble the training data.
-However, the proposed automation methods for this approach are quite novel in the domain of footwear examination, and a fully functioning scanner does not yet exist to collect large amounts of degraded images.
-Thus, the goal of this work is to evaluate the proposed classification scheme and determine whether CNNS can sufficiently distinguish features found on outsoles.
-Addressing these questions with clean retailer images provides both justification and guidance for modifying this approach for real degraded images.}
+However, the proposed automation methods for this approach are quite novel within the domain of footwear examination, and a fully functioning scanner does not yet exist to collect large amounts of degraded images.
+Thus, the goal of this work is to evaluate the proposed classification scheme and determine whether CNNs can sufficiently distinguish features found on outsoles.
+Addressing these questions with clean retailer images provides both justification for further developing the scanner and guidance for modifying the model to process real degraded images.}
 

### 2020-04-10 paper/paper-revision.Rnw
Added \mt{} tags to all changes since original version and corresponding PDF

@@
 \begin{abstract}
-To assess random match probability at the level of class characteristics, one could imagine sampling shoes from a local population and calculating the proportion of shoes "similar" to a given print.
-Our proposed implemention of such a procedure involves 1) a scanner to gather images of shoes from a local population and 2) software to automatically label images within a feature space and assess the images for similarity among those features.
-In this paper, we discuss an algorithm for automatically labeling shoe outsole images using a classification scheme similar to what examiners use: a set of features based on geometric shapes such as circles, stars, and quadrilaterals.
+\mt{To assess random match probability at the level of class characteristics, one could imagine sampling shoes from a local population and calculating the proportion of shoes "similar" to a given print.
+Our proposed implemention of such a procedure involves 1) a scanner to gather images of shoes from a local population and 2) software to automatically label images within a feature space and assess the images for similarity among those features.} In this paper, we discuss an algorithm for automatically labeling shoe outsole images using a \mt{classification scheme} similar to what examiners use: a set of features based on geometric shapes such as circles, stars, and quadrilaterals.
 The automatic labeling algorithm performs well on images with unambiguous shapes and moderate color contrast, though some systematic errors arise from the physical similarities of geometric shapes.
 Additional improvements may be realized by modifying the labeling scheme to reduce feature overlap and by preprocessing the images to improve contrast.
-This new approach represents a significant step toward making tractable the computation of random match probabilities of tread patterns, and could potentially be modified to automate other types of footwear examinations based on class characteristics.
+\mt{This new approach represents a significant step toward making tractable the computation of random match probabilities of tread patterns, and could potentially be modified to automate other types of footwear examinations based on class characteristics.}
 \end{abstract}
@@
 \begin{keyword}
-Footwear; class characteristics; computer vision; neural networks
+Footwear; class characteristics; \mt{random match probability}; computer vision; neural networks
 \end{keyword}
@@
 
-In the United States, forensic analysis of shoe print evidence is varied and difficult.
+\mt{In the United States, forensic analysis of shoe print evidence is varied and difficult.
 The process of collecting print evidence from a crime scene is not standardized, and it is often impossible to conduct an individualizing analysis (i.e., one based on randomly acquired characteristics) when a print is partial or of low quality.
@@
 In the cases where individualizing characteristics are not used, it may be both possible and useful to assess the random match probability of the footwear evidence at the level of class characteristics; in other words, given a questioned shoe print, what is the probability that a randomly selected shoe in the relevant population has a similar tread pattern?
-Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns (e.g., knockoffs, similar models over different years, etc).
-\fix{[This is an important point.
+Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns (e.g., knockoffs, similar models over different years, etc).} \fix{[This is an important point.
 Should I make a bigger deal out of it throughout the paper?]}
 
-From a statistical perspective, random match probability could theoretically be calculated by sampling a large number of shoes from a given population and calculating the observed proportion of shoes in the sample whose tread patterns are similar to the given print.
+\mt{From a statistical perspective, random match probability could theoretically be calculated by sampling a large number of shoes from a given population and calculating the observed proportion of shoes in the sample whose tread patterns are similar to the given print.
 In practice, one potential method of sampling could be a low-profile scanner placed in a high traffic area.
-Such a scanner could take impressions or pictures of shoes walking past, and those scans could then be used to create a database of shoes relevant to the local area.
+Such a scanner could take impressions or pictures of shoes walking past, and those scans could then be used to create a database of shoes relevant to the local area.}
 
-A number of footwear studies (such as those by \citep{parent_sandy_significance_nodate} and \citep{benedict_geographical_2014}) have demonstrated that even a couple hundred shoe prints constitute an unweildy amount of data, which then require a large amount of time and effort to process and analyze.
+\mt{A number of footwear studies (such as those by \citep{parent_sandy_significance_nodate} and \citep{benedict_geographical_2014}) have demonstrated that even a couple hundred shoe prints constitute an unweildy amount of data, which then require a large amount of time and effort to process and analyze.
 Thus, in parallel with the development of the aforementioned scanner, it is necessary to develop a way to process large amounts of data automatically.
-The primary focus of this paper is the development of one such method for large-scale automatic data processing.
+The primary focus of this paper is the development of one such method for large-scale automatic data processing.}
 
-Although footwear impressions are more commonly used in practice for footwear analysis than outsole images, our early prototypes of a footwear scanner have suggested that cameras are more robust than pressure scanners for placement in public locations, which may experience inclement weather or other non-controlled conditions.
+\mt{Although footwear impressions are more commonly used in practice for footwear analysis than outsole images, our early prototypes of a footwear scanner have suggested that cameras are more robust than pressure scanners for placement in public locations, which may experience inclement weather or other non-controlled conditions.
 Thus, we are working to develop a scanner that uses cameras to take pictures of shoe outsoles, much like images that are currently found on the websites of online shoe retailers \autoref{fig:prototype-scanner}.
 Despite the fact that images from the scanner would be more degraded in some ways than images found from online retailers, the fundamental requirements to analyze such data would be similar, so retailer images provide a reasonable basis for the development of a method to assess similarity in tread images.
-More specifically, this assessment is done by identifying a set of relevant features and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity measures based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.
+More specifically, this assessment is done by identifying a set of relevant features and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity measures based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.}
 
@@
 Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain.
-\fix{Ultimately, the output probabilities can be used to quantify similarity between tread patterns for the ultimate goal of calculating random match probability for a given tread pattern.}
-
-% ORIGINAL INTRO:
-
-%Most forensic examinations of footwear pattern evidence are limited to class characteristic based matches; that is, matches based on the positioning of geometric features within the tread pattern.
-Some of these features may include brand logos and size markings, while others are more generic geometric shapes, such as circles, triangles, stars, and chevrons.
-Class characteristics can be used for elimination, but are not sufficient for identification, as the characteristics are shared by all shoes with the same size, make, and model \citep{grossVariabilitySignificanceClass2013,bodziakfootwear2000}.
-Several attempts have been made to automatically identify features in both prints and shoe tread images, with low-level image analysis methods including fractal decomposition \citep{alexanderAutomaticClassificationRecognition1999}, scale-invariant feature recognition \citep{pavlouAutomaticExtractionClassification2006}, Fourier-Mellin transformation \citep{guehamAutomaticRecognitionShoeprints2008}, and other classical image analysis methods.
-Some of these low-level methods perform relatively well in good conditions \citep{luostarinenMeasuringAccuracyAutomatic2014a} but have degraded performance under conditions which are suboptimal, including those commonly found at crime scenes.
-
-%In the wider field of computer vision, attention has turned to more robust methods for image recognition, such as convolutional neural networks (CNNs), which are capable of achieving near-human accuracy under even degraded image conditions \citep{krizhevskyImageNetClassificationDeep2012}.
-These networks are designed to mimic the process of human vision, and typically involve application of sets of filters, some of which mimic the filters used in the low-level techniques used in early automatic classification attempts.
-CNNs have been applied in footwear forensics, making use of more general neural networks optimized to detect objects found in natural scenes, such as trees, animals, buildings, and cars; the pre-trained network's features are then used to compare shoes or prints to determine how well they match \citep{kongCrossDomainImageMatching2019a,kongCrossDomainForensicShoeprint2017,zhangAdaptingConvolutionalNeural2017}.
-Neural network methods appear to be more successful than traditional image analysis techniques, but the features used for matching are not generally informative for humans, who typically classify shoes by patterns and spatial relationships \citep{davisIntelligenceApproachFootwear1981,grossVariabilitySignificanceClass2013}.
-
-%In this paper, we discuss an alternate approach that uses an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images.
-Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain.
-In addition, much of the existing software to assess geometric similarity, such as SoleMate\footnote{\url{http://www.fosterfreeman.com/trace-evidence/357-sicar-6-solemate-2.html}}, requires examiners to identify and label features manually; automating classification within the currently used feature space could reduce or eliminate the need for such manual classification.
+\mt{These output probabilities can later be used to quantify similarity between tread patterns for the ultimate goal of calculating random match probability for a given tread pattern.}
 
@@
 
-Ultimately, \mt{images collected from a footwear scanner (as described in the previous section) would be} stored in photographic form; as a result, we must use image analysis methods to create numerical features from the original visual record.
+Images collected from a footwear scanner (as described in the previous section) would be stored in photographic form; as a result, we must use image analysis methods to create numerical features from the original visual record.
 There are a number of methods that may be employed to identify shapes and features in an image, such as Fourier-Mellin transforms, Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}.
@@
 
-We have proposed several hypotheses to explain why the model's predictions are not entirely consistent with manual labels on the images.
-We can test these hypotheses by examining what the model ``sees" when making its predictions.
+We have proposed several hypotheses \mt{to explain the inconsistencies between the manual labels and the model's predictions on the images.} We can test these hypotheses by examining what the model ``sees" when making its predictions.
 Heatmaps, also known as Class Activation Maps (CAMs), are a visual diagnostic tool that highlights the areas of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}.
@@
 % Restate goal and briefly summarize methods
-The goal of this research was to develop a method to automatically identify geometric class characteristics of shoe outsoles.
+
+\mt{The goal of this research was to develop a method to automatically identify geometric class characteristics of shoe outsoles.
+This goal is motivated by the desire to compute the random match probability of specific tread patterns, which could be achieved by applying this automatic image processing method to outsole images collected from a footwear scanner in a local population.}
+
 A set of geometric class characteristics was defined, based on feature sets currently used by examiners, to both broadly classify a large variety of shoes and to narrow down similarity into a manageable number and type of features for further use in a shoeprint analysis.
@@
 
-Although there are many areas for potential improvement in future iterations of this model, we have demonstrated the effectiveness of transfer learning to provide an initial label for shoe outsole patterns.
+Although there are many areas for potential improvement in future iterations of this model, we have demonstrated the effectiveness of using transfer learning to provide an initial label for shoe outsole patterns.
 As most of the misclassifications are systematic, it is clear that the model is able to learn features and apply rules consistently.
-We expect that future modifications to the labeling criteria and data processing pipeline will improve the model's accuracy, but the 91\% validation accuracy is promising considering that we do not have estimates of the consistency of examiners' estimates on ambiguous features.
-Taken together, these factors show that Convolutional Neural Networks are a useful tool for classifying a large number of outsoles within a common classification scheme and can ultimately help streamline the current process of sorting and comparing shoes by class characteristics.
+We expect that the proposed modifications to the labeling criteria and data processing pipeline will improve the model's accuracy, but the 91\% validation accuracy is promising considering that we do not have estimates of the consistency of examiners' estimates on ambiguous features.
+Taken together, these factors show that Convolutional Neural Networks are a useful tool for classifying a large number of outsoles within a common classification scheme, \mt{although further research is needed into designing human comphrehensible and yet useful classification schemes for this purpose.
+This work thus provides promising evidence that a well-trained model within an appropriate classification scheme, in conjunction with images collected from a footwear scanner, can be used to estimate the random match probability of footwear tread patterns.}
 

### 2020-04-14 paper/paper-revision.Rnw
Slowly working my way through this...

@@
 \begin{abstract}
-\mt{To assess random match probability at the level of class characteristics, one could imagine sampling shoes from a local population and calculating the proportion of shoes "similar" to a given print.
-Our proposed implemention of such a procedure involves 1) a scanner to gather images of shoes from a local population and 2) software to automatically label images within a feature space and assess the images for similarity among those features.} In this paper, we discuss an algorithm for automatically labeling shoe outsole images using a \mt{classification scheme} similar to what examiners use: a set of features based on geometric shapes such as circles, stars, and quadrilaterals.
-The automatic labeling algorithm performs well on images with unambiguous shapes and moderate color contrast, though some systematic errors arise from the physical similarities of geometric shapes.
+\svp{Quantitative assessment of class characteristic frequency is a difficult problem, in part because there is currently no way to collect and efficiently process and categorize the quantity of data which would be required for such assessments.
+The human labor required to both collect and systematically classify images collected from the population in sufficient quantity to make such assessments has been prohibitive.
+% I'm not thrilled with this bit, but I do think we need to drive the point home a bit harder as to why we have to automate both parts of this process
+In this paper, we discuss a computational method which complements a passive, unmanned scanning device that is currently under development.
+The automatic class characteristic recognition algorithm described in this paper has been developed with photographs of outsole images, and identifies high-level features} such as circles, stars, and quadrilaterals.
+\svp{These features are designed to mimic high-level features used by examiners, to facilitate automatic processing of features without resorting to black-box feature sets commonly used in neural networks.
+The algorithm}  performs well on images with unambiguous shapes and moderate color contrast, though some systematic errors arise from the physical similarities of geometric shapes.
 Additional improvements may be realized by modifying the labeling scheme to reduce feature overlap and by preprocessing the images to improve contrast.
-\mt{This new approach represents a significant step toward making tractable the computation of random match probabilities of tread patterns, and could potentially be modified to automate other types of footwear examinations based on class characteristics.}
+This new approach represents a significant step toward \svp{enabling the collection and application of population-level class characteristic data to support quantitative assessments of the value of class-characteristic based similarity assessments.} % We do NOT want to even hint at automating away anyone's job.
 \end{abstract}
@@
 
-\mt{In the United States, forensic analysis of shoe print evidence is varied and difficult.
-The process of collecting print evidence from a crime scene is not standardized, and it is often impossible to conduct an individualizing analysis (i.e., one based on randomly acquired characteristics) when a print is partial or of low quality.
-In fact, the majority of footwear examinations in the United States reach conclusions regarding the correspondence of class characteristics rather than of individualizing characteristics \citep{smith_2011}.
-In the cases where individualizing characteristics are not used, it may be both possible and useful to assess the random match probability of the footwear evidence at the level of class characteristics; in other words, given a questioned shoe print, what is the probability that a randomly selected shoe in the relevant population has a similar tread pattern?
-Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns (e.g., knockoffs, similar models over different years, etc).} \fix{[This is an important point.
-Should I make a bigger deal out of it throughout the paper?]}
+% \mt{In the United States, forensic analysis of shoe print evidence is varied and difficult.}% this is very imprecise
+\svp{Collection and use of footwear evidence varies heavily across jurisdictions within the United States, and, though it is more common for such analyses to be used in Europe, the U.K., and Israel, many of the same fundamental issues remain.}
+\svp{It is often difficult to recover footwear evidence with sufficient detail to identify individualizing characteristics, either because of degradation or because the medium did not record the individualizing characteristics in sufficient detail to examine.
+It is common for prints to be only partially recorded or recovered, which further compounds the difficulty of performing these types of analysis.}
+\svp{As a result, }
+% \mt{The process of collecting print evidence from a crime scene is not standardized, and it is often impossible to conduct an individualizing analysis (i.e., one based on randomly acquired characteristics) when a print is partial or of low quality.
+In fact,
+\mt{ the majority of footwear examinations in the United States reach conclusions regarding the correspondence of class characteristics rather than of individualizing characteristics \citep{smith_2011}.
+\svp{This tendency is also reflected in the SWGTREAD range of conclusions scale, which has two separate categories to indicate degrees of class characteristic based association out of a range of 7 conclusions.}
+\svp{Currently, the guidelines also caution against any statistical assessment because accurate data and models to support statistical calculations do not exist.}
+\svp{If it were possible to collect a sufficiently large quantity of footwear impression data from a local population, it would be helpful to be able to calculate the frequency of shoes with similar class characteristics in the population in order to characterize the determinative value of a particular piece of evidence.}
+% In the cases where individualizing characteristics are not used, it may be both possible and useful to assess the random match probability of the footwear evidence at the level of class characteristics; in other words, given a questioned shoe print, what is the probability that a randomly selected shoe in the relevant population has a similar tread pattern?
+\mt{Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns} \svp{on the class characteristic level.} %\fix{[This is an important point.
+Should I make a bigger deal out of it throughout the paper?]}% if it fits in, that's a good idea.
 

### 2020-04-14 paper/paper-revision.Rnw
Remove FSI specific stuff

@@
-\documentclass[doubleblind,review]{elsarticle}
+\documentclass{article}
 
@@
 
-% Allow strikethrough text
-\usepackage[normalem]{ulem}
-
-\journal{Forensic Science International}
-
+\usepackage[normalem]{ulem} % Allow strikethrough text
+\usepackage{float}
+\usepackage{booktabs}
+\usepackage{fullpage}
+\usepackage{graphicx}
+\usepackage{amsmath,amssymb}
+\usepackage{natbib}
 %%%%%%%%%%%%%%%%%%%%%%%
@@
 %% `Elsevier LaTeX' style
-\bibliographystyle{elsarticle-num}
+% \bibliographystyle{elsarticle-num}
+\bibliographystyle{plain}
 %%%%%%%%%%%%%%%%%%%%%%%
@@
 %%% Commenting %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage[table]{xcolor}
+% \usepackage[table,dvipsnames]{xcolor}
 \newcommand{\mt}[1]{{\color{magenta} #1}}
@@
 knitr::opts_chunk$set(echo = F, dev = "pdf", dpi = 1000, message = F, warning = F, fig.align = "center", fig.path = "figure/")
+options(kableExtra.latex.load_packages = FALSE)
+library(kableExtra)
+knit_hooks$set(document = function(x) {sub('\\usepackage[]{color}', '\\usepackage[table]{xcolor}', x, fixed = TRUE)})
 
@@
 
-\begin{document}
-
-\begin{frontmatter}
-
-\title{A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}}
-\tnotetext[t1]{This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.}
+\title{A Convolutional Neural Network for Outsole Recognition\footnote{This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.}}
 
@@
 \author{Miranda Tilton and Susan Vanderplas}
-\address{Center for Statistics and Applications in Forensic Evidence, Iowa State University. 195 Durham Center, 613 Morill Rd., Ames, IA 50011}
+% \address{Center for Statistics and Applications in Forensic Evidence, Iowa State University. 195 Durham Center, 613 Morill Rd., Ames, IA 50011}
 
-% or include affiliations in footnotes:
-% \author[1]{Miranda Tilton and Susan Vanderplas\corref{cor1}\fnref{isu}}
-% \cortext[cor1]{Corresponding author}
-% \ead{srvander@iastate.edu}
-%
-% \address[1]{195 Durham Center, 613 Morill Rd, Ames, IA 50011}
-% \fntext[isu]{Center for Statistics and Applications in Forensic Evidence, Iowa State University}
-
+\begin{document}
+\maketitle
 \begin{abstract}
@@
 \end{abstract}
+%
+% \begin{keyword}
+% Footwear; class characteristics; \mt{random match probability}; computer vision; neural networks
+% \end{keyword}
 
-\begin{keyword}
-Footwear; class characteristics; \mt{random match probability}; computer vision; neural networks
-\end{keyword}
-
-\end{frontmatter}
+% \end{frontmatter}
 
@@
 In fact,
-\mt{ the majority of footwear examinations in the United States reach conclusions regarding the correspondence of class characteristics rather than of individualizing characteristics \citep{smith_2011}.
+the majority of footwear examinations in the United States reach conclusions regarding the correspondence of class characteristics rather than of individualizing characteristics \citep{smith_2011}.
 \svp{This tendency is also reflected in the SWGTREAD range of conclusions scale, which has two separate categories to indicate degrees of class characteristic based association out of a range of 7 conclusions.}
@@
 % In the cases where individualizing characteristics are not used, it may be both possible and useful to assess the random match probability of the footwear evidence at the level of class characteristics; in other words, given a questioned shoe print, what is the probability that a randomly selected shoe in the relevant population has a similar tread pattern?
-\mt{Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns} \svp{on the class characteristic level.} %\fix{[This is an important point.
+Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns \svp{on the class characteristic level.} %\fix{[This is an important point.
 Should I make a bigger deal out of it throughout the paper?]}% if it fits in, that's a good idea.
 
-\mt{From a statistical perspective, random match probability could theoretically be calculated by sampling a large number of shoes from a given population and calculating the observed proportion of shoes in the sample whose tread patterns are similar to the given print.
-In practice, one potential method of sampling could be a low-profile scanner placed in a high traffic area.
-Such a scanner could take impressions or pictures of shoes walking past, and those scans could then be used to create a database of shoes relevant to the local area.}
+% \mt{From a statistical perspective,
+\svp{Statistically, } the \svp{coincidental (or random)} match probability is estimated by sampling a large number of shoes from a given population and calculating the observed proportion of shoes in the sample whose tread patterns are similar to the given print.
+
+% In practice, one potential method of sampling could be a low-profile scanner placed in a high traffic area.
+Such a scanner could take impressions or pictures of shoes walking past, and those scans could then be used to create a database of shoes relevant to the local area.
 

### 2020-04-14 paper/paper-revision.Rnw
finished corrections (this round).

@@
 \usepackage{amsmath,amssymb}
+
 \usepackage{natbib}
-%%%%%%%%%%%%%%%%%%%%%%%
-%% Elsevier bibliography styles
-%%%%%%%%%%%%%%%%%%%%%%%
-%% To change the style, put a % in front of the second line of the current style and
-%% remove the % from the second line of the style you would like to use.
-%%%%%%%%%%%%%%%%%%%%%%%
+\bibliographystyle{plain}
 
-%% Numbered
-%\bibliographystyle{model1-num-names}
-
-%% Numbered without titles
-%\bibliographystyle{model1a-num-names}
-
-%% Harvard
-%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
-
-%% Vancouver numbered
-%\usepackage{numcompress}\bibliographystyle{model3-num-names}
-
-%% Vancouver name/year
-%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
-
-%% APA style
-%\bibliographystyle{model5-names}\biboptions{authoryear}
-
-%% AMA style
-%\usepackage{numcompress}\bibliographystyle{model6-num-names}
-
-%% `Elsevier LaTeX' style
-% \bibliographystyle{elsarticle-num}
-\bibliographystyle{plain}
 %%%%%%%%%%%%%%%%%%%%%%%
@@
 the majority of footwear examinations in the United States reach conclusions regarding the correspondence of class characteristics rather than of individualizing characteristics \citep{smith_2011}.
-\svp{This tendency is also reflected in the SWGTREAD range of conclusions scale, which has two separate categories to indicate degrees of class characteristic based association out of a range of 7 conclusions.}
+\svp{This tendency is also reflected in the SWGTREAD range of conclusions scale, which has two separate categories to indicate degrees of class characteristic based association out of a range of 7 conclusions.}\fix{add citation}
 \svp{Currently, the guidelines also caution against any statistical assessment because accurate data and models to support statistical calculations do not exist.}
@@
 \svp{Statistically, } the \svp{coincidental (or random)} match probability is estimated by sampling a large number of shoes from a given population and calculating the observed proportion of shoes in the sample whose tread patterns are similar to the given print.
-
-% In practice, one potential method of sampling could be a low-profile scanner placed in a high traffic area.
-Such a scanner could take impressions or pictures of shoes walking past, and those scans could then be used to create a database of shoes relevant to the local area.
-
-\mt{A number of footwear studies (such as those by \citep{parent_sandy_significance_nodate} and \citep{benedict_geographical_2014}) have demonstrated that even a couple hundred shoe prints constitute an unweildy amount of data, which then require a large amount of time and effort to process and analyze.
-Thus, in parallel with the development of the aforementioned scanner, it is necessary to develop a way to process large amounts of data automatically.
-The primary focus of this paper is the development of one such method for large-scale automatic data processing.}
-
-\mt{Although footwear impressions are more commonly used in practice for footwear analysis than outsole images, our early prototypes of a footwear scanner have suggested that cameras are more robust than pressure scanners for placement in public locations, which may experience inclement weather or other non-controlled conditions.
-Thus, we are working to develop a scanner that uses cameras to take pictures of shoe outsoles, much like images that are currently found on the websites of online shoe retailers \autoref{fig:prototype-scanner}.
-Despite the fact that images from the scanner would be more degraded in some ways than images found from online retailers, the fundamental requirements to analyze such data would be similar, so retailer images provide a reasonable basis for the development of a method to assess similarity in tread images.
-More specifically, this assessment is done by identifying a set of relevant features and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity measures based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.}
+The primary obstacle to coincidental match probability calculation, then, is one of data collection and annotation.
+\svp{The (limited)} number of \svp{population class-characteristic} studies \citep{parent_sandy_significance_nodate,benedict_geographical_2014} \svp{demonstrate the effort required to collect, process, and analyze } even a \svp{moderate amount of data manually}.
+\svp{However, technological innovations over the past decade have automated previously manual tasks such as license plate reading, tracking individuals on video feeds, and other complex tasks.}
+\svp{In these comparable situations, two separate but related technological innovations were required: inexpensive and automatic data collection (e.g.
+via CCTV or video cameras) hardware, and automatic, accurate data processing software.}
+\svp{We are in the process of developing a low-profile scanner which, when embedded in a pedestrian area, passively collects images of shoe treads (a prototype is shown in \autoref{fig:prototype-scanner}).} \svp{In parallel, however, it is necessary to develop software which can process the data from a similarly oriented scanner to generate features which are used to make class characteristic comparisons.} \svp{Through the combination of hardware and software, we hope to enable the collection of population-level footwear data which will facilitate quantitative assessment of class characteristic frequency to support examiner evaluation.}
 
@@
 % From original intro:
-In this paper, we discuss an alternate approach that uses an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images.
-Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain.
-\mt{These output probabilities can later be used to quantify similarity between tread patterns for the ultimate goal of calculating random match probability for a given tread pattern.}
+
+Although footwear impressions are more commonly used in practice for footwear analysis than outsole images, our early prototypes of a footwear scanner \svp{suggest} that cameras are more robust than \svp{impression-based systems (e.g.
+EverOS) for use in outdoor settings.} \svp{As a result, in this paper describing an initial automated software system for class characteristic recognition, we use outsole images (such as those found on online retail sites), rather than impression-based training data.} \svp{While we recognize that images from the scanner are likely to show some degradation compared to} images found from online retailers, the fundamental features of the \svp{underlying machine learning models are similar.} \svp{In this paper, we present a proof-of-concept model designed to identify class characteristics probabilistically, using labels similar to those used by examiners.} \svp{We discuss the strengths and limitations of both the classification scheme and the neural-network based model, with important implications for future attempts to develop automatic high-level class characteristic recognition systems.}
 
@@
 
-\mt{A reasonable critique of this approach is that images collected from online retailers are of much higher quality than would be expected from the sidewalk scanner proposed in \autoref{sec:Introduction}, and thus the CNN described in this paper is not immediately ready to process real degraded images.
-This concern is especially fair when considering that machine learning methods tend to perform poorly when applied to test data that do not resemble the training data.
-However, the proposed automation methods for this approach are quite novel within the domain of footwear examination, and a fully functioning scanner does not yet exist to collect large amounts of degraded images.
-Thus, the goal of this work is to evaluate the proposed classification scheme and determine whether CNNs can sufficiently distinguish features found on outsoles.
-Addressing these questions with clean retailer images provides both justification for further developing the scanner and guidance for modifying the model to process real degraded images.}
+A reasonable critique of this approach is that images collected from online retailers are of much higher quality than would be expected from the sidewalk scanner proposed in \autoref{sec:Introduction}, and thus the CNN described in this paper is not immediately ready to process real degraded images.
+\svp{This consideration is justified, as} machine learning methods tend to perform poorly when applied to test data that do not resemble the training data.
+However, the \svp{use of CNNs to predict class characteristics using the same categories used by human examiners is novel and introduces some interesting complications.} \svp{As a full-featured automatic scanner is still under development, obtaining representative degraded images is not possible (because the scanner's optics are still in flux).} The goal of this work is to evaluate the proposed classification scheme and determine whether CNNs can sufficiently distinguish class characteristic features found on outsoles.
+Addressing these questions with clean retailer images provides \svp{justification for additional algorithm development as well as continuing investment in optimization of the scanner design}
 
@@
 
-We have proposed several hypotheses \mt{to explain the inconsistencies between the manual labels and the model's predictions on the images.} We can test these hypotheses by examining what the model ``sees" when making its predictions.
+We have proposed several hypotheses to explain the inconsistencies between the manual labels and the model's predictions on the images.
+We can test these hypotheses by examining what the model ``sees" when making its predictions.
 Heatmaps, also known as Class Activation Maps (CAMs), are a visual diagnostic tool that highlights the areas of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}.
@@
 
-\mt{The goal of this research was to develop a method to automatically identify geometric class characteristics of shoe outsoles.
-This goal is motivated by the desire to compute the random match probability of specific tread patterns, which could be achieved by applying this automatic image processing method to outsole images collected from a footwear scanner in a local population.}
-
-A set of geometric class characteristics was defined, based on feature sets currently used by examiners, to both broadly classify a large variety of shoes and to narrow down similarity into a manageable number and type of features for further use in a shoeprint analysis.
-Thousands of outsole images were obtained from online shoe retailers and labeled according to their geometric shapes.
-The pre-trained convolutional base of VGG16 was then used, with a new classifier trained on a portion of these labeled images, to output predicted probabilities for test images within the defined set of geometric features.
-After training, the model performs well on the data provided.
-In general, the model is able to identify many well-defined geometric shapes in the images.
+\svp{In this paper, we have presented} a method \svp{for automatic identification of} geometric class characteristics  \svp{in outsole tread patterns}.
+A set of geometric class characteristics, defined based on feature sets currently used by examiners, were used to classify a large variety of tread patterns.
+Labeling these images provided the opportunity to select a manageable number of features and define criteria for labeling these features on newly collected footwear images.
+Using the convolutional base of VGG16, a pre-trained neural network, we created a new classifier optimized to work with outsole images.
+This classifier returns the probability of each geometric feature for new outsole images.
 
 % Discussion of weaknesses
-Although model predictions are generally accurate, there are a number of systematic misclassifications.
+Although model predictions are generally accurate, there are a number of systematic misclassifications \svp{and weaknesses} which can be addressed in future iterations of the modeling process.
 The model has substantial difficulty predicting images with low contrast because image features are obscured.

### 2020-04-17 paper/paper-revision.Rnw
Accept edits, minor additional changes

@@
 \begin{abstract}
-\svp{Quantitative assessment of class characteristic frequency is a difficult problem, in part because there is currently no way to collect and efficiently process and categorize the quantity of data which would be required for such assessments.
-The human labor required to both collect and systematically classify images collected from the population in sufficient quantity to make such assessments has been prohibitive.
-% I'm not thrilled with this bit, but I do think we need to drive the point home a bit harder as to why we have to automate both parts of this process
+Quantitative assessment of class characteristic frequency is a difficult problem, in part because there is currently no way to collect and efficiently process the quantity of data that would be required for such assessments without prohibitive amounts of human labor.
 In this paper, we discuss a computational method which complements a passive, unmanned scanning device that is currently under development.
-The automatic class characteristic recognition algorithm described in this paper has been developed with photographs of outsole images, and identifies high-level features} such as circles, stars, and quadrilaterals.
-\svp{These features are designed to mimic high-level features used by examiners, to facilitate automatic processing of features without resorting to black-box feature sets commonly used in neural networks.
-The algorithm}  performs well on images with unambiguous shapes and moderate color contrast, though some systematic errors arise from the physical similarities of geometric shapes.
+The automatic class characteristic recognition algorithm described in this paper has been developed with photographs of outsole images, and identifies high-level \mt{geometric} features such as circles, stars, and quadrilaterals.
+These features are designed to mimic high-level features used by examiners to facilitate automatic processing of features without resorting to black-box feature sets commonly used in neural networks.
+The algorithm performs well on images with unambiguous shapes and moderate color contrast, though some systematic errors arise from the physical similarities of geometric shapes.
 Additional improvements may be realized by modifying the labeling scheme to reduce feature overlap and by preprocessing the images to improve contrast.
-This new approach represents a significant step toward \svp{enabling the collection and application of population-level class characteristic data to support quantitative assessments of the value of class-characteristic based similarity assessments.} % We do NOT want to even hint at automating away anyone's job.
+This new approach represents a significant step toward enabling the collection and application of population-level class characteristic data to support quantitative assessments of the value of class-characteristic based similarity assessments.
 \end{abstract}
@@
 
-% \mt{In the United States, forensic analysis of shoe print evidence is varied and difficult.}% this is very imprecise
-\svp{Collection and use of footwear evidence varies heavily across jurisdictions within the United States, and, though it is more common for such analyses to be used in Europe, the U.K., and Israel, many of the same fundamental issues remain.}
-\svp{It is often difficult to recover footwear evidence with sufficient detail to identify individualizing characteristics, either because of degradation or because the medium did not record the individualizing characteristics in sufficient detail to examine.
-It is common for prints to be only partially recorded or recovered, which further compounds the difficulty of performing these types of analysis.}
-\svp{As a result, }
-% \mt{The process of collecting print evidence from a crime scene is not standardized, and it is often impossible to conduct an individualizing analysis (i.e., one based on randomly acquired characteristics) when a print is partial or of low quality.
-In fact,
-the majority of footwear examinations in the United States reach conclusions regarding the correspondence of class characteristics rather than of individualizing characteristics \citep{smith_2011}.
-\svp{This tendency is also reflected in the SWGTREAD range of conclusions scale, which has two separate categories to indicate degrees of class characteristic based association out of a range of 7 conclusions.}\fix{add citation}
-\svp{Currently, the guidelines also caution against any statistical assessment because accurate data and models to support statistical calculations do not exist.}
-\svp{If it were possible to collect a sufficiently large quantity of footwear impression data from a local population, it would be helpful to be able to calculate the frequency of shoes with similar class characteristics in the population in order to characterize the determinative value of a particular piece of evidence.}
-% In the cases where individualizing characteristics are not used, it may be both possible and useful to assess the random match probability of the footwear evidence at the level of class characteristics; in other words, given a questioned shoe print, what is the probability that a randomly selected shoe in the relevant population has a similar tread pattern?
-Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns \svp{on the class characteristic level.} %\fix{[This is an important point.
+Collection and use of footwear evidence varies heavily across jurisdictions within the United States, and, though it is more common for such analyses to be used in Europe, the U.K., and Israel, many of the same fundamental issues remain.
+It is often difficult to recover footwear evidence with sufficient detail to identify individualizing characteristics, either because of degradation or because the medium did not record the individualizing characteristics in sufficient detail to examine.
+It is also common for prints to be only partially recorded or recovered, which further compounds the difficulty of performing these types of analysis.
+As a result, the majority of footwear examinations in the United States reach conclusions regarding the correspondence of class characteristics rather than of individualizing characteristics \citep{smith_2011}.
+This tendency is also reflected in the SWGTREAD range of conclusions scale, which has two separate categories to indicate degrees of class characteristic based association out of a range of 7 conclusions.\fix{add citation} Currently, the guidelines also caution against any statistical assessment because accurate data and models to support statistical calculations do not exist.
+\mt{However,} if it were possible to collect a sufficiently large quantity of footwear impression data from a local population, it would be helpful to be able to calculate the frequency of shoes \mt{in the population that have similar class characteristics to a given shoe or print in order to characterize the determinative value of that} particular piece of evidence.
+Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns on the class characteristic level.
+%\fix{[This is an important point.
 Should I make a bigger deal out of it throughout the paper?]}% if it fits in, that's a good idea.
 
-% \mt{From a statistical perspective,
-\svp{Statistically, } the \svp{coincidental (or random)} match probability is estimated by sampling a large number of shoes from a given population and calculating the observed proportion of shoes in the sample whose tread patterns are similar to the given print.
+Statistically, the coincidental (or random) match probability is estimated by sampling a large number of shoes from a given population and calculating the observed proportion of shoes in the sample whose tread patterns are similar to the given print.
 The primary obstacle to coincidental match probability calculation, then, is one of data collection and annotation.
-\svp{The (limited)} number of \svp{population class-characteristic} studies \citep{parent_sandy_significance_nodate,benedict_geographical_2014} \svp{demonstrate the effort required to collect, process, and analyze } even a \svp{moderate amount of data manually}.
-\svp{However, technological innovations over the past decade have automated previously manual tasks such as license plate reading, tracking individuals on video feeds, and other complex tasks.}
-\svp{In these comparable situations, two separate but related technological innovations were required: inexpensive and automatic data collection (e.g.
-via CCTV or video cameras) hardware, and automatic, accurate data processing software.}
-\svp{We are in the process of developing a low-profile scanner which, when embedded in a pedestrian area, passively collects images of shoe treads (a prototype is shown in \autoref{fig:prototype-scanner}).} \svp{In parallel, however, it is necessary to develop software which can process the data from a similarly oriented scanner to generate features which are used to make class characteristic comparisons.} \svp{Through the combination of hardware and software, we hope to enable the collection of population-level footwear data which will facilitate quantitative assessment of class characteristic frequency to support examiner evaluation.}
+The (limited) number of population class-characteristic studies \citep{parent_sandy_significance_nodate,benedict_geographical_2014} demonstrate the effort required to collect, process, and analyze even a moderate amount of data manually.
+However, technological innovations over the past decade have automated previously manual tasks such as license plate reading, tracking individuals on video feeds, and other complex tasks.
+In these comparable situations, two separate but related technological innovations were required: inexpensive and automatic data collection (e.g.
+via CCTV or video cameras) hardware, and automatic, accurate data processing software.
+We are in the process of developing a low-profile scanner which, when embedded in a pedestrian area, passively collects images of shoe treads (a prototype is shown in \autoref{fig:prototype-scanner}).
+In parallel, however, it is necessary to develop software which can process the data from a similarly oriented scanner to generate features which are used to make class characteristic comparisons.
+Through the combination of hardware and software, we hope to enable the collection of population-level footwear data which will facilitate quantitative assessment of class characteristic frequency to support examiner evaluation.
 
@@
 
-% From original intro:
-
-Although footwear impressions are more commonly used in practice for footwear analysis than outsole images, our early prototypes of a footwear scanner \svp{suggest} that cameras are more robust than \svp{impression-based systems (e.g.
-EverOS) for use in outdoor settings.} \svp{As a result, in this paper describing an initial automated software system for class characteristic recognition, we use outsole images (such as those found on online retail sites), rather than impression-based training data.} \svp{While we recognize that images from the scanner are likely to show some degradation compared to} images found from online retailers, the fundamental features of the \svp{underlying machine learning models are similar.} \svp{In this paper, we present a proof-of-concept model designed to identify class characteristics probabilistically, using labels similar to those used by examiners.} \svp{We discuss the strengths and limitations of both the classification scheme and the neural-network based model, with important implications for future attempts to develop automatic high-level class characteristic recognition systems.}
+Although footwear impressions are more commonly used in practice for footwear analysis than outsole images, our early prototypes of a footwear scanner suggest that cameras are more robust than impression-based systems (e.g.
+EverOS) for use in outdoor settings.
+As a result, in this paper describing an initial automated software system for class characteristic recognition, we use outsole images (such as those found on online retail sites), rather than impression-based training data.
+While we recognize that images from the scanner are likely to show some degradation compared to images found from online retailers, the fundamental features of the underlying machine learning models are similar.
+In this paper, we present a proof-of-concept model designed to identify class characteristics probabilistically, using labels similar to those used by examiners.
+We discuss the strengths and limitations of both the classification scheme and the neural-network based model, with important implications for future attempts to develop automatic high-level class characteristic recognition systems.
 
@@
 A reasonable critique of this approach is that images collected from online retailers are of much higher quality than would be expected from the sidewalk scanner proposed in \autoref{sec:Introduction}, and thus the CNN described in this paper is not immediately ready to process real degraded images.
-\svp{This consideration is justified, as} machine learning methods tend to perform poorly when applied to test data that do not resemble the training data.
-However, the \svp{use of CNNs to predict class characteristics using the same categories used by human examiners is novel and introduces some interesting complications.} \svp{As a full-featured automatic scanner is still under development, obtaining representative degraded images is not possible (because the scanner's optics are still in flux).} The goal of this work is to evaluate the proposed classification scheme and determine whether CNNs can sufficiently distinguish class characteristic features found on outsoles.
-Addressing these questions with clean retailer images provides \svp{justification for additional algorithm development as well as continuing investment in optimization of the scanner design}
+This consideration is justified, as machine learning methods tend to perform poorly when applied to test data that do not resemble the training data.
+However, the use of CNNs to predict class characteristics using the same categories used by human examiners is novel and introduces some interesting complications.
+As a full-featured automatic scanner is still under development, obtaining representative degraded images is not possible (because the scanner's optics are still in flux).
+The goal of this work is to evaluate the proposed classification scheme and determine whether CNNs can sufficiently distinguish class characteristic features found on outsoles.
+Addressing these questions with clean retailer images provides justification for additional algorithm development as well as continuing investment in optimization of the scanner design.
 
@@
 
-\svp{In this paper, we have presented} a method \svp{for automatic identification of} geometric class characteristics  \svp{in outsole tread patterns}.
-A set of geometric class characteristics, defined based on feature sets currently used by examiners, were used to classify a large variety of tread patterns.
+In this paper, we have presented a method for automatic identification of geometric class characteristics in outsole tread patterns.
+A set of geometric class characteristics were defined based on feature sets currently used by examiners and used to classify a large variety of tread patterns.
 Labeling these images provided the opportunity to select a manageable number of features and define criteria for labeling these features on newly collected footwear images.
@@
 % Discussion of weaknesses
-Although model predictions are generally accurate, there are a number of systematic misclassifications \svp{and weaknesses} which can be addressed in future iterations of the modeling process.
+Although model predictions are generally accurate, there are a number of systematic misclassifications and weaknesses which can be addressed in future iterations of the modeling process.
 The model has substantial difficulty predicting images with low contrast because image features are obscured.

### 2020-04-17 paper/paper-revision.Rnw
A few introduction tweaks, add SWGTREAD citation

@@
 In this paper, we discuss a computational method which complements a passive, unmanned scanning device that is currently under development.
-The automatic class characteristic recognition algorithm described in this paper has been developed with photographs of outsole images, and identifies high-level \mt{geometric} features such as circles, stars, and quadrilaterals.
+The automatic class characteristic recognition algorithm described in this paper has been developed with photographs of outsole images, and identifies high-level geometric features such as circles, stars, and quadrilaterals.
 These features are designed to mimic high-level features used by examiners to facilitate automatic processing of features without resorting to black-box feature sets commonly used in neural networks.
@@
 As a result, the majority of footwear examinations in the United States reach conclusions regarding the correspondence of class characteristics rather than of individualizing characteristics \citep{smith_2011}.
-This tendency is also reflected in the SWGTREAD range of conclusions scale, which has two separate categories to indicate degrees of class characteristic based association out of a range of 7 conclusions.\fix{add citation} Currently, the guidelines also caution against any statistical assessment because accurate data and models to support statistical calculations do not exist.
-\mt{However,} if it were possible to collect a sufficiently large quantity of footwear impression data from a local population, it would be helpful to be able to calculate the frequency of shoes \mt{in the population that have similar class characteristics to a given shoe or print in order to characterize the determinative value of that} particular piece of evidence.
+This tendency is also reflected in the SWGTREAD range of conclusions scale, which has two separate categories to indicate degrees of class characteristic based association out of a range of 7 conclusions\citep{SWGTREAD2013}.
+Currently, the guidelines also caution against any statistical assessment because accurate data and models to support statistical calculations do not exist.
+However, if it were possible to collect a sufficiently large quantity of footwear impression data from a local population, it would be helpful to be able to calculate the frequency of shoes in the population that have similar class characteristics to a given shoe or print in order to characterize the determinative value of that particular piece of evidence.
 Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns on the class characteristic level.
-%\fix{[This is an important point.
-Should I make a bigger deal out of it throughout the paper?]}% if it fits in, that's a good idea.
 
@@
 However, technological innovations over the past decade have automated previously manual tasks such as license plate reading, tracking individuals on video feeds, and other complex tasks.
-In these comparable situations, two separate but related technological innovations were required: inexpensive and automatic data collection (e.g.
-via CCTV or video cameras) hardware, and automatic, accurate data processing software.
+In these comparable situations, two separate but related technological innovations were required: inexpensive and automatic data collection hardware (e.g.
+via CCTV or video cameras), and automatic, accurate data processing software.
 We are in the process of developing a low-profile scanner which, when embedded in a pedestrian area, passively collects images of shoe treads (a prototype is shown in \autoref{fig:prototype-scanner}).
@@
 EverOS) for use in outdoor settings.
-As a result, in this paper describing an initial automated software system for class characteristic recognition, we use outsole images (such as those found on online retail sites), rather than impression-based training data.
+As a result, in this paper describing an initial automated software system for class characteristic recognition, we use outsole images (such as those found on online retail sites) rather than impression-based training data.
 While we recognize that images from the scanner are likely to show some degradation compared to images found from online retailers, the fundamental features of the underlying machine learning models are similar.

### 2020-04-23 paper/paper.Rnw
fix refs
Rename files to play nicely with auto-submission system

--- 2019-07-01 paper/paper.Rnw
@@
 
-% Allow strikethrough text
-\usepackage[normalem]{ulem}
+\usepackage[normalem]{ulem} % Allow strikethrough text
+\usepackage{float}
+\usepackage{booktabs}
+\usepackage{fullpage}
+\usepackage{graphicx}
+\usepackage{amsmath,amssymb}
 
-\journal{Forensic Science International}
+\usepackage{natbib}
+\bibliographystyle{plain}
 
-%%%%%%%%%%%%%%%%%%%%%%%
-%% Elsevier bibliography styles
-%%%%%%%%%%%%%%%%%%%%%%%
-%% To change the style, put a % in front of the second line of the current style and
-%% remove the % from the second line of the style you would like to use.
-%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Numbered
-%\bibliographystyle{model1-num-names}
-
-%% Numbered without titles
-%\bibliographystyle{model1a-num-names}
-
-%% Harvard
-%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
-
-%% Vancouver numbered
-%\usepackage{numcompress}\bibliographystyle{model3-num-names}
-
-%% Vancouver name/year
-%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
-
-%% APA style
-%\bibliographystyle{model5-names}\biboptions{authoryear}
-
-%% AMA style
-%\usepackage{numcompress}\bibliographystyle{model6-num-names}
-
-%% `Elsevier LaTeX' style
-\bibliographystyle{elsarticle-num}
 %%%%%%%%%%%%%%%%%%%%%%%
@@
 knitr::opts_chunk$set(echo = F, dev = "pdf", dpi = 1000, message = F, warning = F, fig.align = "center", fig.path = "figure/")
+options(kableExtra.latex.load_packages = FALSE)
+library(kableExtra)
+knit_hooks$set(document = function(x) {sub('\\usepackage[]{color}', '\\usepackage[table]{xcolor}', x, fixed = TRUE)})
 
-library(tidyverse)
 library(magrittr)
 library(tidyverse)
+library(tidyr)
+nest <- nest_legacy
+unnest <- unnest_legacy
 library(ggplot2)
@@
 \author{Miranda Tilton and Susan Vanderplas}
-\address{Center for Statistical Applications in Forensic Evidence, Iowa State University. 195 Durham Center, 613 Morill Rd., Ames, IA 50011}
+% \address{Center for Statistics and Applications in Forensic Evidence, Iowa State University. 195 Durham Center, 613 Morill Rd., Ames, IA 50011}
 
-% or include affiliations in footnotes:
-% \author[1]{Miranda Tilton and Susan Vanderplas\corref{cor1}\fnref{isu}}
-% \cortext[cor1]{Corresponding author}
-% \ead{srvander@iastate.edu}
+\begin{document}
+\maketitle
+\begin{abstract}
+Quantitative assessment of class characteristic frequency is a difficult problem, in part because there is currently no way to collect and efficiently process the quantity of data that would be required for such assessments without prohibitive amounts of human labor.
+In this paper, we discuss a computational method which complements a passive, unmanned scanning device that is currently under development.
+The automatic class characteristic recognition algorithm described in this paper has been developed with photographs of outsole images, and identifies high-level geometric features such as circles, stars, and quadrilaterals.
+These features are designed to mimic high-level features used by examiners to facilitate automatic processing of features without resorting to black-box feature sets commonly used in neural networks.
+The algorithm performs well on images with unambiguous shapes and moderate color contrast, though some systematic errors arise from the physical similarities of geometric shapes.
+Additional improvements may be realized by modifying the labeling scheme to reduce feature overlap and by preprocessing the images to improve contrast.
+This new approach represents a significant step toward enabling the collection and application of population-level class characteristic data to support quantitative assessments of the value of class-characteristic based similarity assessments.
+\end{abstract}
 %
-% \address[1]{195 Durham Center, 613 Morill Rd, Ames, IA 50011}
-% \fntext[isu]{Center for Statistical Applications in Forensic Evidence, Iowa State University}
+% \begin{keyword}
+% Footwear; class characteristics; \mt{random match probability}; computer vision; neural networks
+% \end{keyword}
 
-\begin{abstract}
-Currently, to assess similarity in forensic shoeprint analysis, examiners must manually identify and label features in images of shoe prints or outsoles.
-In this paper, we discuss an algorithm for automatically labeling shoe outsole images using the same feature space examiners use: a set of features based on geometric shapes such as circles, stars, and quadrilaterals.
-The automatic labeling algorithm performs well on images with unambiguous shapes and moderate color contrast, though some systematic errors arise from the physical similarities of geometric shapes.
-Additional improvements may be realized by modifying the labeling scheme to reduce feature overlap and by preprocessing the images to improve contrast.
-This new approach represents a significant improvement to the current manual classification of footwear tread patterns, and could be integrated into current matching databases to reduce the amount of time examiners spend on data entry.
-\end{abstract}
-
-\begin{keyword}
-footwear, class characteristics, computer vision, neural networks
-\end{keyword}
-
-\end{frontmatter}
+% \end{frontmatter}
 
@@
 
-\section{Introduction}
-Most forensic examinations of footwear pattern evidence are limited to class characteristic based matches; that is, matches based on the positioning of geometric features within the tread pattern.
-Some of these features may include brand logos and size markings, while others are more generic geometric shapes, such as circles, triangles, stars, and chevrons.
-Class characteristics can be used for elimination, but are not sufficient for identification, as the characteristics are shared by all shoes with the same size, make, and model \citep{grossVariabilitySignificanceClass2013,bodziakfootwear2000}.
+\section{Introduction}\label{sec:Introduction}
+
+Collection and use of footwear evidence varies heavily across jurisdictions within the United States, and, though it is more common for such analyses to be used in Europe, the U.K., and Israel, many of the same fundamental issues remain.
+It is often difficult to recover footwear evidence with sufficient detail to identify individualizing characteristics, either because of degradation or because the medium did not record the individualizing characteristics in sufficient detail to examine.
+It is also common for prints to be only partially recorded or recovered, which further compounds the difficulty of performing these types of analysis.
+As a result, the majority of footwear examinations in the United States reach conclusions regarding the correspondence of class characteristics rather than of individualizing characteristics \citep{smith_2011}.
+This tendency is also reflected in the SWGTREAD range of conclusions scale, which has two separate categories to indicate degrees of class characteristic based association out of a range of 7 conclusions\citep{SWGTREAD2013}.
+Currently, the guidelines also caution against any statistical assessment because accurate data and models to support statistical calculations do not exist.
+However, if it were possible to collect a sufficiently large quantity of footwear impression data from a local population, it would be helpful to be able to calculate the frequency of shoes in the population that have similar class characteristics to a given shoe or print in order to characterize the determinative value of that particular piece of evidence.
+Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns on the class characteristic level.
+
+Statistically, the coincidental (or random) match probability is estimated by sampling a large number of shoes from a given population and calculating the observed proportion of shoes in the sample whose tread patterns are similar to the given print.
+The primary obstacle to coincidental match probability calculation, then, is one of data collection and annotation.
+The (limited) number of population class-characteristic studies \citep{parent_sandy_significance_nodate,benedict_geographical_2014} demonstrate the effort required to collect, process, and analyze even a moderate amount of data manually.
+However, technological innovations over the past decade have automated previously manual tasks such as license plate reading, tracking individuals on video feeds, and other complex tasks.
+In these comparable situations, two separate but related technological innovations were required: inexpensive and automatic data collection hardware (e.g.
+via CCTV or video cameras), and automatic, accurate data processing software.
+We are in the process of developing a low-profile scanner which, when embedded in a pedestrian area, passively collects images of shoe treads (a prototype is shown in \autoref{fig:prototype-scanner}).
+In parallel, however, it is necessary to develop software which can process the data from a similarly oriented scanner to generate features which are used to make class characteristic comparisons.
+Through the combination of hardware and software, we hope to enable the collection of population-level footwear data which will facilitate quantitative assessment of class characteristic frequency to support examiner evaluation.
+
+\begin{figure}
+\centering
+\setlength\tabcolsep{1mm}
+\begin{tabular}{cc}
+\raisebox{-.5\height}{\includegraphics[width=.45\textwidth]{images/prototype/20200402_151639.jpg}} &
+\raisebox{-.5\height}{\includegraphics[width=.45\textwidth]{images/prototype/GOPR1231-blur-enhance_square.jpg}} \vspace{1mm}\\
+
+\end{tabular}
+\caption[Images of prototype scanner.]{Images of the prototype scanner for collecting a large number of outsole images.}\label{fig:prototype-scanner}
+\end{figure}
+
+\subsection{Image Analysis}
+
+% From original intro:
 Several attempts have been made to automatically identify features in both prints and shoe tread images, with low-level image analysis methods including fractal decomposition \citep{alexanderAutomaticClassificationRecognition1999}, scale-invariant feature recognition \citep{pavlouAutomaticExtractionClassification2006}, Fourier-Mellin transformation \citep{guehamAutomaticRecognitionShoeprints2008}, and other classical image analysis methods.
@@
 
+% From original intro:
 In the wider field of computer vision, attention has turned to more robust methods for image recognition, such as convolutional neural networks (CNNs), which are capable of achieving near-human accuracy under even degraded image conditions \citep{krizhevskyImageNetClassificationDeep2012}.
@@
 
-In this paper, we discuss an alternate approach that uses an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images.
-Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain.
-In addition, much of the existing software to assess geometric similarity, such as SoleMate\footnote{\url{http://www.fosterfreeman.com/trace-evidence/357-sicar-6-solemate-2.html}}, requires examiners to identify and label features manually; automating classification within the currently used feature space could reduce or eliminate the need for such manual classification.
+Although footwear impressions are more commonly used in practice for footwear analysis than outsole images, our early prototypes of a footwear scanner suggest that cameras are more robust than impression-based systems (e.g.
+EverOS) for use in outdoor settings.
+As a result, in this paper describing an initial automated software system for class characteristic recognition, we use outsole images (such as those found on online retail sites) rather than impression-based training data.
+While we recognize that images from the scanner are likely to show some degradation compared to images found from online retailers, the fundamental features of the underlying machine learning models are similar.
+In this paper, we present a proof-of-concept model designed to identify class characteristics probabilistically, using labels similar to those used by examiners.
+We discuss the strengths and limitations of both the classification scheme and the neural-network based model, with important implications for future attempts to develop automatic high-level class characteristic recognition systems.
 
 \section{Materials and Methods}
-%\fix{We may want to mention some of the methods in the introduction for continuity sake...
-Ultimately, most pattern evidence is stored in photographic form; as a result, we must use image analysis methods to create numerical features from the original visual record.
+
+Images collected from a footwear scanner (as described in the previous section) would be stored in photographic form; as a result, we must use image analysis methods to create numerical features from the original visual record.
 There are a number of methods that may be employed to identify shapes and features in an image, such as Fourier-Mellin transforms, Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}.
@@
 
+A reasonable critique of this approach is that images collected from online retailers are of much higher quality than would be expected from the sidewalk scanner proposed in \autoref{sec:Introduction}, and thus the CNN described in this paper is not immediately ready to process real degraded images.
+This consideration is justified, as machine learning methods tend to perform poorly when applied to test data that do not resemble the training data.
+However, the use of CNNs to predict class characteristics using the same categories used by human examiners is novel and introduces some interesting complications.
+As a full-featured automatic scanner is still under development, obtaining representative degraded images is not possible (because the scanner's optics are still in flux).
+The goal of this work is to evaluate the proposed classification scheme and determine whether CNNs can sufficiently distinguish class characteristic features found on outsoles.
+Addressing these questions with clean retailer images provides justification for additional algorithm development as well as continuing investment in optimization of the scanner design.
+
 \begin{figure}[hbt]
@@
 \centering
-\includegraphics[width=.1\textwidth]{images/augmentation/bowtie(R)-1-ugg-sienna-enamel-blue_product_8726365_color_120557.jpg}\hfill
-\includegraphics[width=.1\textwidth]{images/augmentation/aug_bowtie(R)-1-ugg-sienna-enamel-blue_product_8726365_color_120557_0_8344.jpg}\hfill\hfill\hfill
-\includegraphics[width=.1\textwidth]{images/augmentation/quad-2-nike-kids-downshifter-7-infant-toddler-gunsmoke-sunset-pulse-atmosphere-grey_product_8800875_color_720973.jpg}\hfill
-\includegraphics[width=.1\textwidth]{images/augmentation/aug_quad-2-nike-kids-downshifter-7-infant-toddler-gunsmoke-sunset-pulse-atmosphere-grey_product_8800875_color_720973_0_7533.jpg}\hfill\hfill\hfill
-\includegraphics[width=.1\textwidth]{images/augmentation/quad-2-puma-suede-classic-x-mac-three-port-royale-port-royale_product_9123838_color_251426.jpg}\hfill
-\includegraphics[width=.1\textwidth]{images/augmentation/aug_quad-2-puma-suede-classic-x-mac-three-port-royale-port-royale_product_9123838_color_251426_0_5689.jpg}\hfill\hfill\hfill
-\includegraphics[width=.1\textwidth]{images/augmentation/text-1-bernie-mev-kids-catwalk-little-kid-big-kid-multi-camo_product_9084647_color_87089.jpg}\hfill
-\includegraphics[width=.1\textwidth]{images/augmentation/aug_text-1-bernie-mev-kids-catwalk-little-kid-big-kid-multi-camo_product_9084647_color_87089_0_1891.jpg}
+\includegraphics[width=.1\textwidth]{images/augmentation/bowtie_orig.jpg}\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/aug_bowtie.jpg}\hfill\hfill\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/quad-2_orig.jpg}\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/aug_quad-2.jpg}\hfill\hfill\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/quad_orig.jpg}\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/aug_quad.jpg}\hfill\hfill\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/text_orig.jpg}\hfill
+\includegraphics[width=.1\textwidth]{images/augmentation/aug_text.jpg}
 \caption[Original and augmented images.]{Four sets of original (left) and augmented (right) labeled images.}\label{fig:augmented}
@@
 
-We have proposed several hypotheses to explain why the model's predictions are not entirely consistent with manual labels on the images.
+We have proposed several hypotheses to explain the inconsistencies between the manual labels and the model's predictions on the images.
 We can test these hypotheses by examining what the model ``sees" when making its predictions.
@@
 % Restate goal and briefly summarize methods
-The goal of this research was to develop a method to automatically identify geometric class characteristics of shoe outsoles.
-A set of geometric class characteristics was defined, based on feature sets currently used by examiners, to both broadly classify a large variety of shoes and to narrow down similarity into a manageable number and type of features for further use in a shoeprint analysis.
-Thousands of outsole images were obtained from online shoe retailers and labeled according to their geometric shapes.
-The pre-trained convolutional base of VGG16 was then used, with a new classifier trained on a portion of these labeled images, to output predicted probabilities for test images within the defined set of geometric features.
-After training, the model performs well on the data provided.
-In general, the model is able to identify many well-defined geometric shapes in the images.
+
+In this paper, we have presented a method for automatic identification of geometric class characteristics in outsole tread patterns.
+A set of geometric class characteristics were defined based on feature sets currently used by examiners and used to classify a large variety of tread patterns.
+Labeling these images provided the opportunity to select a manageable number of features and define criteria for labeling these features on newly collected footwear images.
+Using the convolutional base of VGG16, a pre-trained neural network, we created a new classifier optimized to work with outsole images.
+This classifier returns the probability of each geometric feature for new outsole images.
 
 % Discussion of weaknesses
-Although model predictions are generally accurate, there are a number of systematic misclassifications.
+Although model predictions are generally accurate, there are a number of systematic misclassifications and weaknesses which can be addressed in future iterations of the modeling process.
 The model has substantial difficulty predicting images with low contrast because image features are obscured.
@@
 
-Although there are many areas for potential improvement in future iterations of this model, we have demonstrated the effectiveness of transfer learning to provide an initial label for shoe outsole patterns.
+Although there are many areas for potential improvement in future iterations of this model, we have demonstrated the effectiveness of using transfer learning to provide an initial label for shoe outsole patterns.
 As most of the misclassifications are systematic, it is clear that the model is able to learn features and apply rules consistently.
-We expect that future modifications to the labeling criteria and data processing pipeline will improve the model's accuracy, but the 91\% validation accuracy is promising considering that we do not have estimates of the consistency of examiners' estimates on ambiguous features.
-Taken together, these factors show that Convolutional Neural Networks are a useful tool for classifying a large number of outsoles within a common classification scheme and can ultimately help streamline the current process of sorting and comparing shoes by class characteristics.
+We expect that the proposed modifications to the labeling criteria and data processing pipeline will improve the model's accuracy, but the 91\% validation accuracy is promising considering that we do not have estimates of the consistency of examiners' estimates on ambiguous features.
+Taken together, these factors show that Convolutional Neural Networks are a useful tool for classifying a large number of outsoles within a common classification scheme, although further research is needed into designing human comphrehensible classification schemes for this purpose.
+This work thus provides the necessary proof-of-concept that a well-trained model within an appropriate classification scheme, in conjunction with images collected from a footwear scanner, can be used to estimate the coincidental match probability of footwear tread patterns.
 

### 2020-04-23 paper/coverletter.tex
Update cover letter and other submission files for SA&DM

@@
 \begin{document}
-C. Jackowski \\
-Institute of Forensic Medicine\vspace{.5em}\\
-C. Cattaneo \\
-Università degli Studi di Milano \vspace{2em}
+Jia Li \\
+Stanford University
+
+\vspace{2em}
 
@@
 
-{\bf Re: Submission to Forensic Science International}\\\vspace{2em}
+{\bf Re: Submission to Statistical Analysis and Data Mining}\\\vspace{2em}
 
-Dear Professor Jackowski and Professor Cattaneo, \vspace{1em}
+Dear Dr. Li, \vspace{1em}
 
-Please find attached a manuscript, ``A Convolutional Neural Network for Outsole Recognition" intended for submission to FSI. \vspace{1em}
+Please find attached a manuscript, ``A Convolutional Neural Network for Outsole Recognition" intended for submission to Statistical Analysis and Data Mining. \vspace{1em}
 
@@
 This algorithm uses a scheme similar to that used by footwear forensic analysts, unlike other convolutional neural network based analyses in footwear forensics.
-We describe the model's structure, capabilities, and overall performance; we also explore situations where the model's automatic recognition process does not succeed and provide some possible solutions for these issues going forward.
-We believe this work lays the foundation to bridge the gap between theoretical applications of computer vision in footwear forensics and work that is done in the crime lab.
+My coauthor, Dr.
+Susan VanderPlas, and I describe the model's structure, capabilities, and overall performance; we also explore situations where the model's automatic recognition process does not succeed and provide some possible solutions for these issues going forward.
+We believe this work lays the foundation for assessing footwear evidence quantitatively, which will improve the utility of forensic evidence and make the justice system's evaluation of evidence more fair.
 \vspace{1em}
 
-We hope that you find the manuscript acceptable for publication in FSI. If there is any additional information we can provide, please feel free to contact us directly. \vspace{0.25in}
+We hope that you find the manuscript acceptable for publication in Statistical Analysis and Data Mining.
+If there is any additional information we can provide, please feel free to contact me directly.
+\vspace{0.25in}
 
@@
 
-Miranda Tilton and Susan Vanderplas\\
+Miranda Tilton\\
 Center for Statistics and Applications in Forensic Evidence\\

### 2020-04-24 paper/paper.tex
bbl file and tex file

--- 2020-04-23 paper/paper.Rnw
@@
-\documentclass{article}
+\documentclass{article}\usepackage[]{graphicx}\usepackage[table]{xcolor}
+% maxwidth is the original width if it is less than linewidth
+% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 
@@
 
-<<setup, include = F>>=
-knitr::opts_chunk$set(echo = F, dev = "pdf", dpi = 1000, message = F, warning = F, fig.align = "center", fig.path = "figure/")
-options(kableExtra.latex.load_packages = FALSE)
-library(kableExtra)
-knit_hooks$set(document = function(x) {sub('\\usepackage[]{color}', '\\usepackage[table]{xcolor}', x, fixed = TRUE)})
-
-library(magrittr)
-library(tidyverse)
-library(tidyr)
-nest <- nest_legacy
-unnest <- unnest_legacy
-library(ggplot2)
-library(keras)
-
-mytheme <- theme_bw()
-theme_set(mytheme)
-@
-
-<<results_setup, include = F>>=
-if ("paper" %in% list.files()) {
-  wd <- file.path(getwd(), "paper")
-} else {
-  wd <- getwd()
-}
-
-codedir <- file.path(wd, "code")
-modeldir <- file.path(wd, "model")
-imgdir <- file.path(wd, "images")
-
-source(file.path(codedir, "Generate_Model_Images.R"))
-source(file.path(codedir, "equalize.R"))
-@
-
-<<labeled-data-setup, include = F>>=
-model_path <- file.path(modeldir, "TrainedModels")
-
-if (grepl("!School", model_path)) {
-  newest_model <- get_newest(dir = model_path, pattern = "weights.h5")
-  newest_model$prefix = "vgg16_onehotaug_9class_256"
-  newest_model$start_date = "2019-05-14_19_45_22"
-} else {
-  newest_model <- list(path = file.path(model_path, "20190514-155453"),
-      base_file = "2019-05-14_19:45:22_vgg16_onehotaug_9class_256-weights.h5",
-      prefix = "vgg16_onehotaug_9class_256", start_date = "2019-05-14_19:45:22",
-      process_dir = "20190514-155453")
-}
-
-newest_data_file <- file.path(modeldir, "RProcessedImages",
-                              newest_model$process_dir, "cropped_photos.Rdata")
-
-load(newest_data_file)
-source(file.path(codedir, "count_images.R"))
-@
-
-<<load-keras-model, include = F, eval = T, echo = F>>=
-model_dir <- newest_model$path
-load(list.files(model_dir, pattern = "-history.Rdata", full.names = T)[1])
-load(file.path(newest_model$path,
-               paste0(newest_model$start_date, "_",
-                      newest_model$prefix, ".Rdata")))
-
-model_wts_file <- file.path(newest_model$path, newest_model$base_file)
-loaded_model <- set_weights(model_wts_file)
-@
-
 \title{A Convolutional Neural Network for Outsole Recognition\footnote{This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.}}
@@
 % \address{Center for Statistics and Applications in Forensic Evidence, Iowa State University. 195 Durham Center, 613 Morill Rd., Ames, IA 50011}
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 
-<<vgg16layers, fig.width = 7, fig.height = 5.75, fig.cap = "Visualizing convolutional filters requires generating images which maximally activate specific filters.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+
+{\centering \includegraphics[width=\maxwidth]{/home/susan/Projects/Work/2019-CoNNORFSI/paper/images/labeled_filters}
+
+}
+
+\caption[Visualizing convolutional filters requires generating images which maximally activate specific filters]{Visualizing convolutional filters requires generating images which maximally activate specific filters.
 This figure contains images that maximally activate a selection of filters from the convolutional blocks of VGG16.
 Filters from later blocks detect more complex features than those in earlier blocks.
-Much like human vision, a hierarchical structure of feature detection with increasing complexity allows CNN ``vision\" trained on a general task to be applied to new domains.">>=
+Much like human vision, a hierarchical structure of feature detection with increasing complexity allows CNN ``vision" trained on a general task to be applied to new domains.}\label{fig:vgg16layers}
+\end{figure}
 
-if(!file.exists(file.path(wd, "images", "unlabeled_filters.png"))) {
-  imgs <- list.files(file.path(wd, "images", "vgg16layers"),
-                     recursive = T, full.names = T)
-  pngs <- lapply(imgs, function(x) png::readPNG(x)[46:1596, 32:1582, ])
-  grobs <- lapply(pngs, grid::rasterGrob)
-  filters <- gridExtra::arrangeGrob(grobs = grobs, nrow = 5)
-  ggsave(filename = file.path(wd, "images", "unlabeled_filters.png"),
-         plot = filters, device = "png", width = 6, height = 5.3)
-}
-
-if(!file.exists(file.path(wd, "images", "labeled_filters.png"))) {
-  library(magick)
-  b <- image_blank(width = 300, height = 1590)
-  f <- image_read(file.path(wd, "images", "unlabeled_filters.png"))
-  f_l <- image_append(c(b, f)) %>%
-    image_annotate("Block 1:", location = "+60+139",
-                   color = "black", size = 60) %>%
-    image_annotate("Block 2:", location = "+60+457",
-                   color = "black", size = 60) %>%
-    image_annotate("Block 3:", location = "+60+775",
-                   color = "black", size = 60) %>%
-    image_annotate("Block 4:", location = "+60+1093",
-                   color = "black", size = 60) %>%
-    image_annotate("Block 5:", location = "+60+1411",
-                   color = "black", size = 60) %>%
-    image_scale(geometry = "600x")
-  image_write(image = f_l, path = file.path(wd, "images", "labeled_filters.png"),
-              format = "png")
-}
-
-knitr::include_graphics(file.path(wd, "images", "labeled_filters.png"))
-@
+\end{knitrout}
 
@@
 During this process, aspect ratio is not preserved, though efforts are made to label regions which are relatively square to minimize the effect of this distortion.
-To date, \Sexpr{unique(ann_df$base_image) %>% length()} shoes have been labeled, yielding \Sexpr{nrow(dfunion)} multi-label images.
+To date, 4571 shoes have been labeled, yielding 28081 multi-label images.
 
@@
 
-<<class-characteristic-barchart, fig.width = 6, fig.height = 4, dpi = 300, out.width = "80%", fig.cap = "Distribution of classes in all labeled images.
-Quadrilaterals, lines, circles, text, and chevrons are relatively common; stars, polygons, and bowties are relatively rare.", fig.scap = "Distribution of classes in all labeled images.">>=
-annotated_imgs %>%
-  mutate(label_type = str_replace(label_type, "multi", "multiple")) %>%
-  mutate(labels = str_to_title(labels)) %>%
-  mutate(labels = factor(labels, levels = c("Quad", "Line", "Text", "Circle",
-                                            "Chevron", "Triangle", "Polygon",
-                                            "Star", "Bowtie", "Other"))) %>%
-  ggplot() +
-  geom_bar(aes(x = labels, fill = label_type), color = "black") +
-  scale_fill_manual("Labels", values = c("single" = "#6ba2b9",
-                                         "multiple" = "#2e5597")) +
-  coord_flip() +
-  ylab("# Labeled Images") +
-  xlab("")  +
-  ggtitle("Current Class Distribution (All Labeled Images)") +
-  mytheme +
-  theme(legend.position = c(1, 1),
-        legend.justification = c(1.03, 1.03),
-        legend.background = element_rect(fill = "white"))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+
+{\centering \includegraphics[width=0.8\linewidth]{figure/class-characteristic-barchart-1}
+
+}
+
+\caption[Distribution of classes in all labeled images.]{Distribution of classes in all labeled images.
+Quadrilaterals, lines, circles, text, and chevrons are relatively common; stars, polygons, and bowties are relatively rare.}\label{fig:class-characteristic-barchart}
+\end{figure}
+
+\end{knitrout}
 
@@
 
-The \Sexpr{nrow(dfunion)} images were split such that 60\% were used for training. Since the categories do not exist in equal proportion in the labeled data,
+The 28081 images were split such that 60\% were used for training. Since the categories do not exist in equal proportion in the labeled data,
 the training data were weighted by proportion during the training process to ensure that the model can identify both rare and common geometric shapes.
@@
 
-<<training-accuracy, fig.width = 7.5, fig.height = 5, out.width = ".75\\textwidth", fig.cap = "Training and validation accuracy and loss for each epoch of the fitting process.
-Training and validation accuracy reach 89.5\\% around epoch 9.
-After that point, validation loss remains the same and training loss decreases slightly, while validation accuracy increases more slowly than training accuracy.", fig.scap = "Training and validation accuracy and loss during each epoch.", fig.pos = 'h'>>=
-data.frame(history$metrics) %>%
-  mutate(epoch = 1:n()) %>%
-  gather(key = "measure", value = "value", -epoch) %>%
-  mutate(Type = ifelse(str_detect(measure, "val"), "Validation", "Training"),
-         measure = ifelse(str_detect(measure, "acc"), "Accuracy", "Loss")) %>%
-  # bind_rows(tibble(epoch = NA, value =  .6, measure = "Accuracy", Type = "Validation")) %>%
-  # bind_rows(tibble(epoch = NA, value =  .33, measure = "Loss", Type = "Validation")) %>%
-  ggplot(aes(x = epoch, y = value, color = Type)) +
-  geom_point() +
-  geom_smooth(se = F) +
-  facet_grid(measure~., scales = "free_y", switch = "both") +
-  theme_bw() +
-  scale_y_continuous("") +
-  scale_x_continuous("Epoch") +
-  ggtitle("CoNNOR Training Performance") + mytheme +
-  theme(axis.title.y = element_blank(), legend.position = c(1, .5),
-        legend.justification = c(1.03, -0.05),
-        legend.background = element_rect(fill = "white"))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[h]
+
+{\centering \includegraphics[width=.75\textwidth]{figure/training-accuracy-1}
+
+}
+
+\caption[Training and validation accuracy and loss during each epoch.]{Training and validation accuracy and loss for each epoch of the fitting process.
+Training and validation accuracy reach 89.5\% around epoch 9.
+After that point, validation loss remains the same and training loss decreases slightly, while validation accuracy increases more slowly than training accuracy.}\label{fig:training-accuracy}
+\end{figure}
+
+\end{knitrout}
 
@@
 
-<<overall-roc, fig.width = 5, fig.height = 5, out.width = ".5\\textwidth", fig.cap = "ROC curve showing overall model performance.
-Model accuracy is generally high, as the ROC curve is significantly above the identity line $y=x$.", fig.scap = "Overall model performance ROC curve.">>=
-library(pROC)
-pred_df <- as_tibble(preds) %>% gather(key = feature, value = value)
-test_labs_df <- as_tibble(test_labs) %>% gather(key = feature, value = value)
-whole_model_roc <- roc(test_labs_df$value, pred_df$value)
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
 
-whole_model_roc_df <- tibble(tpr = whole_model_roc$sensitivities,
-                             fpr = 1 - whole_model_roc$specificities,
-                             thresholds = whole_model_roc$thresholds,
-                             auc = whole_model_roc$auc[1]) %>%
-  nest(tpr, fpr, thresholds, .key =  "roc_plot") %>%
-  mutate(eer = purrr::map(roc_plot, eer))
-ggplot() +
-  geom_line(aes(x = fpr, y = tpr), data = unnest(whole_model_roc_df, roc_plot),
-            size = 1.25) +
-  geom_label(aes(x = 1, y = .07, label = sprintf("AUC: %0.2f", auc)), hjust = 1,
-             vjust = -0.2, data = whole_model_roc_df) +
-  geom_point(aes(x = fpr, y = tpr, color = "Equal Error Rate"),
-             data = unnest(whole_model_roc_df, eer), size = 2) +
-  scale_color_manual("", values = "black") +
-  scale_x_continuous("False Positive Rate", breaks = c(0, .25, .5, .75, 1),
-                     labels = c("0.0", "", "0.5", "", "1.0")) +
-  scale_y_continuous("True Positive Rate", breaks = c(0, .25, .5, .75, 1),
-                     labels = c("0.0", "", "0.5", "", "1.0")) +
-  ggtitle("CoNNOR Test Set Performance (All Classes)") +
-  coord_fixed() + mytheme +
-  theme(legend.position = c(1, 0), legend.justification = c(1.01, -0.01),
-        legend.title = element_blank(),
-        legend.background = element_rect(fill = "white"))
-@
+{\centering \includegraphics[width=.5\textwidth]{figure/overall-roc-1}
 
-<<class-roc, fig.width = 8, fig.height = 6, out.width = "\\textwidth", fig.cap = "Class-by-class ROC curves.
+}
+
+\caption[Overall model performance ROC curve.]{ROC curve showing overall model performance.
+Model accuracy is generally high, as the ROC curve is significantly above the identity line $y=x$.}\label{fig:overall-roc}
+\end{figure}
+
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/class-roc-1}
+
+}
+
+\caption[Class-by-class ROC curves.]{Class-by-class ROC curves.
 AUC is area under the curve, a measure of overall model performance.
-Equal error rates (EER) are marked, indicating the position at which there is equal probability of a false positive or false negative error.", fig.scap = "Class-by-class ROC curves.">>=
-aucs <- plot_onehot_roc(preds, test_labs, str_to_title(classes))
-thresholds <- purrr::map_dbl(aucs$data$eer, ~.$thresholds)
-aucs$data$thresholds <- thresholds
+Equal error rates (EER) are marked, indicating the position at which there is equal probability of a false positive or false negative error.}\label{fig:class-roc}
+\end{figure}
 
-ggplot() +
-  geom_line(aes(x = fpr, y = tpr), data = unnest(aucs$data, roc_plot), size = 1.25) +
-  geom_label(aes(x = 1, y = 0, label = sprintf("AUC: %0.2f\nEER: %0.2f", auc, thresholds)), hjust = 1, vjust = -0.02, data = aucs$data) +
-  geom_point(aes(x = fpr, y = tpr, color = "Equal Error\nRate (EER)"), data = unnest(aucs$data, eer), size = 2.5) +
-  scale_color_manual("", values = "black") +
-  facet_wrap(~class) +
-  scale_x_continuous("False Positive Rate", breaks = c(0, .25, .5, .75, 1), labels = c("0.0", "", "0.5", "", "1.0")) +
-  scale_y_continuous("True Positive Rate", breaks = c(0, .25, .5, .75, 1), labels = c("0.0", "", "0.5", "", "1.0")) +
-  ggtitle("CoNNOR Test Set Performance") +
-  facet_wrap(~class, nrow = 2) +
-  coord_fixed() +
-  theme(legend.position = c(1, 0), legend.justification = c(1, 0))
-@
+\end{knitrout}
 
-The full model has an AUC of \Sexpr{round(whole_model_roc_df$auc,2)}, and the AUC for individual classes ranges from \Sexpr{round(min(aucs$data$auc),2)} (for \Sexpr{sprintf("%s",classes[which.min(aucs$data$auc)])}) to \Sexpr{round(max(aucs$data$auc),2)}  (for \Sexpr{sprintf("%s",classes[which.max(aucs$data$auc)])}).
+The full model has an AUC of 0.88, and the AUC for individual classes ranges from 0.81 (for line) to 0.93  (for bowtie).
 While the class performances do vary slightly, each ROC curve indicates that the model performs significantly better than random chance.
@@
 
-<<eers, include = F>>=
-eers <- purrr::map_dbl(aucs$data$eer, function(x){x$thresholds}) %>%
-  setNames(classes)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[p!]
 
-<<exemplars, fig.cap = "Model predictions of images containing (from the top) a single shape, no shape, or multiple shapes across all categories.
+{\centering \includegraphics[width=1\linewidth]{figure/exemplars-1}
+
+}
+
+\caption[Model predictions of images containing (from the top) a single shape, no shape, or multiple shapes across all categories]{Model predictions of images containing (from the top) a single shape, no shape, or multiple shapes across all categories.
 Predicted probabilities greater than the equal error rate threshold are outlined in blue.
-Although there are some misclassifications, predictions are generally consistent with the human-assigned labels of these images.", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 11.5/1.25, fig.height = 16/1.25>>=
-single_exemplars <- list.files(file.path(wd, "images", "consistency",
-                                         "single-exemplar"), full.names = T)
-multi_exemplars <- list.files(file.path(wd, "images", "consistency",
-                                        "multi-exemplar"), full.names = T)
-pred_prob_plot(rev(c(single_exemplars, multi_exemplars)),
-               loaded_model, sort = F, eer = eers)
-@
+Although there are some misclassifications, predictions are generally consistent with the human-assigned labels of these images.}\label{fig:exemplars}
+\end{figure}
+
+\end{knitrout}
 
@@
 While the circle predictions are often above the EER for letters such as G and S, these predictions are generally much lower than those where distinct circles or Os are seen, indicating that there are some indications that features are not as strong in the quasi-circle shapes, but that the model is not able to completely separate Os from C, S, and G shapes which are also very circular.
-This confusion is complicated by the frequency of images which contain both circles and text: text is present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("circle", dfunion$name)), 2)}\% of our labeled images containing circles, and circles are present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("text", dfunion$name)), 2)}\% of our images containing text.
+This confusion is complicated by the frequency of images which contain both circles and text: text is present in 39\% of our labeled images containing circles, and circles are present in 31\% of our images containing text.
 In addition, the distinction between a circle and an ``o" comes from the context around the shape, which is much more difficult for the model to parse than simply detecting a shape.
@@
 
-<<circle-text-consistency, fig.cap = "Model predictions of images containing circles and/or text.
-The model correctly identifies circles in all images where circles are present, but also predicts moderate probabilities of circles in many images where text is present with curved letters, such as ``G\".
-Text is correctly identified in most images, with the exception of the false negative prediction of the low-contrast text within a hexagon and the false positive prediction of concentric circles.", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 11.5/1.25, fig.height = 16/1.25>>=
-circle_text <- list.files(file.path(wd, "images", "consistency", "circle-text"), full.names = T)
-pred_prob_plot(rev(circle_text), loaded_model, sort = F, eer = eers)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[p!]
+
+{\centering \includegraphics[width=1\linewidth]{figure/circle-text-consistency-1}
+
+}
+
+\caption[Model predictions of images containing circles and/or text]{Model predictions of images containing circles and/or text.
+The model correctly identifies circles in all images where circles are present, but also predicts moderate probabilities of circles in many images where text is present with curved letters, such as ``G".
+Text is correctly identified in most images, with the exception of the false negative prediction of the low-contrast text within a hexagon and the false positive prediction of concentric circles.}\label{fig:circle-text-consistency}
+\end{figure}
+
+\end{knitrout}
 
@@
 
-<<confusion-matrix, fig.width = 9, fig.height = 8, out.width = "\\textwidth", fig.cap = "Confusion matrix, showing on the diagonal the correct classification rate and on the off-diagonal, classification errors.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/confusion-matrix-1}
+
+}
+
+\caption[Confusion matrix, with correct and incorrect model classifications.]{Confusion matrix, showing on the diagonal the correct classification rate and on the off-diagonal, classification errors.
 Note that in multi-label images, correct off-diagonal labels have been excluded from the calculation of false positives.
 Quadrilaterals are generally over-predicted across true labels; polygons and triangles are under-predicted.
-Circles and text are also often confused.", fig.scap = "Confusion matrix, with correct and incorrect model classifications.", dpi = 600>>=
-get_confusion_matrix(predictions = preds, classes = classes,
-                     test_labels = test_labs, threshold = thresholds) %>%
-  set_names(str_to_title(classes)) %>%
-  ggcorrplot(., hc.order = F, outline.col = "white", lab = T) +
-  theme_bw() +
-  scale_fill_gradient("Classification\nRate", low = "white",
-                      high = "cornflowerblue", limits = c(0, 1)) +
-  scale_x_discrete("Image Label", labels = str_to_title(classes)) + scale_y_discrete("Prediction") +
-  theme(axis.title.x = element_text(size = 14),
-        axis.title.y = element_text(size = 14, angle = 90, vjust = 1),
-        axis.text.y = element_text(angle = 90, hjust = .5)) +
-  ggtitle("CoNNOR Multi-Class Confusion Matrix: Test Set Performance") +
-  theme(panel.grid.major = element_line(color = "grey50"),
-        panel.grid.minor = element_line(color = "grey60")) +
-  theme(plot.margin = grid::unit(c(1,0,0,0), "mm"),
-        plot.background = element_rect(fill = "transparent", color = NA),
-        plot.subtitle = element_blank(), plot.caption = element_blank(),
-        panel.spacing = unit(c(0, 0, 0, 0), "mm"))
-@
+Circles and text are also often confused.}\label{fig:confusion-matrix}
+\end{figure}
+
+\end{knitrout}
 
@@
 
-<<low-contrast-consistency, fig.cap = "Model predictions for images with low contrast.
-While these images contain a variety of shapes (as shown in the color-corrected column on the right), the diffuse probabilities assigned by the model across all categories indicate that features are obscured by low contrast, which propogates through model training and prediction.", fig.pos = "p!", out.width = "100%", fig.width = 10/1.2, fig.height = 9/1.2, cache = T>>=
-lowcontrast <- list.files(file.path(wd, "images", "consistency", "low-contrast2"), full.names = T)
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[p!]
 
-pred_prob_plot(lowcontrast, loaded_model, sort = F, eer = eers,
-               img_path2 = lowcontrast) +
-  theme(legend.position = "bottom")
-@
+{\centering \includegraphics[width=1\linewidth]{figure/low-contrast-consistency-1}
+
+}
+
+\caption[Model predictions for images with low contrast]{Model predictions for images with low contrast.
+While these images contain a variety of shapes (as shown in the color-corrected column on the right), the diffuse probabilities assigned by the model across all categories indicate that features are obscured by low contrast, which propogates through model training and prediction.}\label{fig:low-contrast-consistency}
+\end{figure}
+
+\end{knitrout}
 
@@
 
-<<brand-plots, out.width = "100%", fig.width = 7, fig.height = 5, fig.cap = "Proportions of top 10 shoe brands in our labeled dataset within each shape category, sorted by prevalence.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figure/brand-plots-1}
+
+}
+
+\caption[Proportions of top 10 shoe brands in our labeled dataset within each shape category, sorted by prevalence]{Proportions of top 10 shoe brands in our labeled dataset within each shape category, sorted by prevalence.
 Brand text is present when the brand makes up more than five percent of labeled shoes for the shape category.
 Some shapes, like bowtie and polygon, are relatively dominated by a small number of brands while other shapes, such as text and quad, are comprised of a large number of brands.
-Typically, model predictions are best for shape categories without much variability, which, in general, coincides with shapes dominated by a small number of brands.">>=
-img_names <- dfunion$filename %>% basename() %>% as.vector
+Typically, model predictions are best for shape categories without much variability, which, in general, coincides with shapes dominated by a small number of brands.}\label{fig:brand-plots}
+\end{figure}
 
-labs <- sapply(classes, grepl, x = str_extract(img_names, "^.*?-\\d?-"))
-str <- gsub("^.*?-\\d{0,2}-", "", img_names)
-
-# Extract first, first two, and first three words of string
-brand_1 <- str_extract(str, "^.*?-")
-brand_2 <- str_extract(str, "^.*?-.*?-")
-brand_3 <- str_extract(str, "^.*?-.*?-.*?-")
-
-# These are the beginning of brands that need an extra word extracted
-multi_brands <- c("the", "the-north", "polo", "polo-ralph", "5", "5-11",
-                  "1", "to", "dr", "la", "new", "old", "under",
-                  "steve", "bernie", "cole", "tory",
-                  "harley", "kristin", "eric",
-                  "spring", "chinese", "dirty")
-
-brand <- ifelse(brand_1 %in% paste0(multi_brands, "-"),
-                ifelse(brand_2 %in% paste0(multi_brands, "-"),
-                       brand_3, brand_2),
-                brand_1) %>%
-  gsub(. , pattern = "-", replacement = " ") %>%
-  trimws(., "right"); rm(brand_1, brand_2, brand_3, multi_brands)
-
-df_images <- cbind(tibble(str, brand), labs)
-
-df_shoes <- df_images %>%
-  group_by(str, brand) %>%
-  summarize_at(vars(bowtie:triangle), any)
-
-df_shoes_long <- df_shoes %>%
-  tidyr::gather(key = feature, value = value, -str, -brand) %>%
-  ungroup() %>%
-  tidyr::nest_legacy(c(str, value)) %>%
-  mutate(n_true = purrr::map_int(data, ~sum(.$value)),
-         n_shoes_brand = purrr::map_int(data, ~length(.$value)))
-
-brand_prop_by_feature <- df_shoes_long %>%
-  group_by(feature) %>%
-  mutate(prop = n_true/sum(n_true)) %>%
-  arrange(feature, desc(prop)) %>%
-  mutate(brand_cat = ifelse(row_number() <= 10, brand, "other")) %>%
-  group_by(feature) %>%
-  mutate(rank = pmin(row_number(), 11))
-
-brand_prop_by_feature_10 <- brand_prop_by_feature %>%
-  group_by(feature, brand_cat, rank) %>%
-  summarize(prop = sum(prop)) %>%
-  ungroup() %>%
-  group_by(feature) %>%
-  arrange(feature, rank) %>%
-  mutate(cprop = cumsum(prop),
-         cprop2 = lag(cprop, 1, default = 0),
-         mid = (cprop + cprop2)/2,
-         rank = ifelse(rank == 11, NA, rank))
-
-ggplot(data = brand_prop_by_feature_10) +
-  geom_bar(aes(x = feature, y = prop, fill = factor(rank)),
-           position = "stack", stat = "identity", color = "black") +
-  geom_text(aes(x = feature, y = 1-mid, label = brand_cat),
-            data = filter(brand_prop_by_feature_10, prop > .05), size = 2.5) +
-  scale_fill_brewer(palette = "Paired", guide = F, na.value = "grey") +
-  scale_y_continuous("Proportion") + scale_x_discrete("") +
-  ggtitle("Proportion of Labeled Features by Brand")
-@
+\end{knitrout}
 
@@
 
-<<generate-heatmaps, echo = F, eval = F>>=
-# My computer has troubles with the transparency of the overlay...
-# This code is functional, but highest quality will come from generating these
-# images on Bigfoot and loading them statically.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!p]
 
-h_imgdir <- file.path(imgdir, "heatmaps", "processed")
-if(!dir.exists(h_imgdir)) {dir.create(h_imgdir)}
+{\centering \includegraphics[width=.95\textwidth]{/home/susan/Projects/Work/2019-CoNNORFSI/paper/images/heatmaps/processed/heatmap-a7-Ugg-grey}
 
-# Turn all images in "heatmaps/raw" to top-4 heatmaps
-imgs <- list.files(file.path(wd, "images", "heatmaps", "raw"), full.names = T)
-sapply(imgs, function(x) {
-  #cat(x);
-  calc_heatmap(x, loaded_model) %>%
-    prune_heatmap(top_n = 4, sort = "value") %>%
-    create_composite(save_file = T, outdir = h_imgdir,
-                     color_by_label = F, eer = eers)
-})
+}
 
-# calc_heatmap(imgs[which(grepl("Ugg-grey", imgs))], loaded_model) %>%
-#   prune_heatmap(top_n = 3, sort = "value") %>%
-#   create_composite(save_file = T, outdir = h_imgdir, color_by_label = F,
-#                    eer = eers)
-#
-# calc_heatmap(imgs[which(grepl("text", imgs))], loaded_model) %>%
-#   prune_heatmap(top_n = 3, sort = "value") %>%
-#   create_composite(save_file = T, outdir = h_imgdir, color_by_label = F,
-#                    eer = eers)
-#
-# calc_heatmap(imgs[which(grepl("polygon_recolor", imgs))], loaded_model) %>%
-#   prune_heatmap(top_n = 3, sort = "value") %>%
-#   create_composite(save_file = T, outdir = h_imgdir, color_by_label = F,
-#                    eer = eers)
-#
-# calc_heatmap(imgs[which(grepl("corks", imgs))], loaded_model) %>%
-#   prune_heatmap(top_n = 3, sort = "value") %>%
-#   create_composite(save_file = T, outdir = h_imgdir, color_by_label = F,
-#                    eer = eers)
-#
-# calc_heatmap(imgs[which(grepl("chevron-circle", imgs))], loaded_model) %>%
-#   #prune_heatmap(c("chevron", "quad", "circle")) %>%
-#   create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
-#
-# calc_heatmap(imgs[which(grepl("triangle", imgs))], loaded_model) %>%
-#   #prune_heatmap(c("triangle", "quad", "chevron")) %>%
-#   create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
-#
-# calc_heatmap(imgs[which(grepl("star", imgs))], loaded_model) %>%
-#   #prune_heatmap(c("star", "bowtie", "polygon")) %>%
-#   create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
-#
-# calc_heatmap(imgs[which(grepl("other", imgs))], loaded_model) %>%
-#   #prune_heatmap(c("star", "chevron", "polygon")) %>%
-#   create_composite(save_file = T, outdir = h_imgdir, color_by_label = F)
+\caption{A section of an UGG shoe containing the logo.
+The model correctly identifies triangles and text (and references the correct portion of the image in both cases).
+The model identifies a circle as well; while it could be argued that the triangles form an implicit circle, the heatmap demonstrates that the circular portion of the image is the ``G", which would not be a closed figure labeled as a circle.
+The model's prediction of the presence of a circle is understandable, but incorrect.\label{fig:ugg-heatmap}}\label{fig:current-heatmaps1}
+\end{figure}
 
-@
+\begin{figure}[!p]
 
-<<current-heatmaps, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.cap = c("A section of an UGG shoe containing the logo.
-The model correctly identifies triangles and text (and references the correct portion of the image in both cases).
-The model identifies a circle as well; while it could be argued that the triangles form an implicit circle, the heatmap demonstrates that the circular portion of the image is the ``G\", which would not be a closed figure labeled as a circle.
-The model's prediction of the presence of a circle is understandable, but incorrect.\\label{fig:ugg-heatmap}", "In this image of a Seychelles shoe, the model identifies text and circle.
-The prediction of text is correct; the prediction of circle is not---there are very round shapes, but they do not form a closed figure that could be identified as a circle.\\label{fig:seychelles-heatmap}", "This image from a DC shoe contains repeating instances of a shape that appears to be a combination of a circle and quadrilateral, but does not technically fall into any of the nine categories used in our labeling scheme.
+{\centering \includegraphics[width=.95\textwidth]{/home/susan/Projects/Work/2019-CoNNORFSI/paper/images/heatmaps/processed/heatmap-text-seychelles}
+
+}
+
+\caption{In this image of a Seychelles shoe, the model identifies text and circle.
+The prediction of text is correct; the prediction of circle is not---there are very round shapes, but they do not form a closed figure that could be identified as a circle.\label{fig:seychelles-heatmap}}\label{fig:current-heatmaps2}
+\end{figure}
+
+\begin{figure}[!p]
+
+{\centering \includegraphics[width=.95\textwidth]{/home/susan/Projects/Work/2019-CoNNORFSI/paper/images/heatmaps/processed/heatmap-04-circle_quad-2-dc-vestrey-black-camo_product_9065558_color_12784}
+
+}
+
+\caption{This image from a DC shoe contains repeating instances of a shape that appears to be a combination of a circle and quadrilateral, but does not technically fall into any of the nine categories used in our labeling scheme.
 The model assigns high probabilities of circle and quadrilateral and indicates that the proper image features are being used for this classification.
-The prediction of polygon is incorrect but reasonable considering that many small hexagons are indistinguishable from circles.\\label{fig:dc-heatmap}"), echo = F, fig.pos = "!p">>=
+The prediction of polygon is incorrect but reasonable considering that many small hexagons are indistinguishable from circles.\label{fig:dc-heatmap}}\label{fig:current-heatmaps3}
+\end{figure}
 
-h_imgdir <- file.path(imgdir, "heatmaps", "processed")
-heatmaps <- list.files(h_imgdir, pattern = "^heatmap", full.names = T)
-
-knitr::include_graphics(heatmaps[grepl("Ugg-grey", heatmaps)])
-knitr::include_graphics(heatmaps[grepl("text-seychelles", heatmaps)])
-#knitr::include_graphics(heatmaps[grepl("polygon_recolor", heatmaps)])
-knitr::include_graphics(heatmaps[grepl("dc", heatmaps)])
-
-#sapply(heatmaps, knitr::include_graphics)
-@
-
-<<chevron-circle-heatmap, eval = F, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing chevrons and circles.", fig.cap = "Chevrons are identified by sharp corners and parallel lines, but the thick line between the two shape sections is falsely identified as a quadrilateral.", fig.pos = "!h">>=
-heatmaps <- list.files(file.path(wd, "images", "heatmaps", "static"),
-                       pattern = "^heatmap", full.names = T)
-knitr::include_graphics(heatmaps[grepl("chevron-circle", heatmaps)])
-@
-
-<<text-seychelles-heatmap, eval = F, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing text.", fig.cap = "This heatmap illustrates that the high probability of a circle is a result of the highly curled tails of the 'S' in the text.">>=
-knitr::include_graphics(heatmaps[grepl("text", heatmaps)])
-@
-
-<<corks-heatmap, eval = F, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing none of the target classes.", fig.cap = "This outsole shows repeated outlines of corks, which do not fall into any of the nine classes; the concave corners where the corks overlap are predicted to be stars, and the round edges lead to predictions of circles.\\svp{Not really...
-only 0.03 probability}">>=
-knitr::include_graphics(heatmaps[grepl("corks", heatmaps)])
-@
-
-<<puzzle-heatmap, eval = F, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing none of the target classes.", fig.cap = "This outsole shows repeated outlines of puzzle pieces, which do not fall into any of the nine classes.">>=
-knitr::include_graphics(heatmaps[grepl("other", heatmaps)])
-@
+\end{knitrout}
 
@@
 
-<<test-accuracy, echo = F>>=
-# This is my stab at calculating test accuracy.
-# .786 doesn't seem too promising :(
-# That's with thresholding. If you use absolute error, we get an accuracy rate of 0.86, which is pretty good. (the contrast corrected model is 0.88, so much more reasonable)
-# Or if we compare to your calculations, we go from .786 -> .811. All in all, we're doing ok. We can definitely just report validation accuracy, though...
-# Interestingly enough, if you use a 0.5 threshold, you get an accuracy rate of .85...
-EER isn't getting us that much.
-The accuracy rate on the contrast-fixed model is exactly the same under that constraint, which suggests that there's some difference in the output probabilities, but not *that* much of a difference - enough to make a small difference when we're using EER.
-abs_err <- sum(abs(preds - test_labs))/length(test_labs)
-
-disc_preds <- sweep(preds, 2, eers) > 0 #predicted values minus EER, column wise
-test_accuracy <- sum(disc_preds == test_labs) / (length(test_labs))
-@
-
 \bibliography{mybibfile}

### 2020-04-27 paper/paper.Rnw
Updates to try to debug compiler

@@
 As a result, the majority of footwear examinations in the United States reach conclusions regarding the correspondence of class characteristics rather than of individualizing characteristics \citep{smith_2011}.
-This tendency is also reflected in the SWGTREAD range of conclusions scale, which has two separate categories to indicate degrees of class characteristic based association out of a range of 7 conclusions\citep{SWGTREAD2013}.
+This tendency is also reflected in the SWGTREAD range of conclusions scale, which has two separate categories to indicate degrees of class characteristic based association out of a range of 7 conclusions \citep{SWGTREAD2013}.
 Currently, the guidelines also caution against any statistical assessment because accurate data and models to support statistical calculations do not exist.
@@
 Transfer learning leverages the modularity of neural networks, that is, that the model base that detects features can be separated from the classifier which produces predictions.
-Transfer learning approaches use the base of a model trained on more general image data and train a new model head or classifier using a smaller set of domain-specific labeled images\citep{oquabLearningTransferringMidlevel2014}.
+Transfer learning approaches use the base of a model trained on more general image data and train a new model head or classifier using a smaller set of domain-specific labeled images \citep{oquabLearningTransferringMidlevel2014}.
 This allows CNNs to be applied to smaller datasets of several thousand images, reduces the amount of computational time required to fit the model, and provides boosted performance compared to training a new model from scratch \citep{yosinskiHowTransferableAre2014}.

### 2020-04-27 paper/paper.tex
Updates to try to debug compiler

@@
 
-{\centering \includegraphics[width=\maxwidth]{/home/susan/Projects/Work/2019-CoNNORFSI/paper/images/labeled_filters}
+{\centering \includegraphics[width=\maxwidth]{images/labeled_filters}
 
@@
 
-{\centering \includegraphics[width=.95\textwidth]{/home/susan/Projects/Work/2019-CoNNORFSI/paper/images/heatmaps/processed/heatmap-a7-Ugg-grey}
+{\centering \includegraphics[width=.95\textwidth]{images/heatmaps/processed/heatmap-a7-Ugg-grey}
 
@@
 The model identifies a circle as well; while it could be argued that the triangles form an implicit circle, the heatmap demonstrates that the circular portion of the image is the ``G", which would not be a closed figure labeled as a circle.
-The model's prediction of the presence of a circle is understandable, but incorrect.\label{fig:ugg-heatmap}}\label{fig:current-heatmaps1}
+The model's prediction of the presence of a circle is understandable, but incorrect.\label{fig:ugg-heatmap}}\label{fig:current-heatmaps}
 \end{figure}
@@
 
-{\centering \includegraphics[width=.95\textwidth]{/home/susan/Projects/Work/2019-CoNNORFSI/paper/images/heatmaps/processed/heatmap-text-seychelles}
+{\centering \includegraphics[width=.95\textwidth]{images/heatmaps/processed/heatmap-text-seychelles}
 
@@
 \caption{In this image of a Seychelles shoe, the model identifies text and circle.
-The prediction of text is correct; the prediction of circle is not---there are very round shapes, but they do not form a closed figure that could be identified as a circle.\label{fig:seychelles-heatmap}}\label{fig:current-heatmaps2}
+The prediction of text is correct; the prediction of circle is not---there are very round shapes, but they do not form a closed figure that could be identified as a circle.\label{fig:seychelles-heatmap}}\label{fig:current-heatmaps}
 \end{figure}
@@
 
-{\centering \includegraphics[width=.95\textwidth]{/home/susan/Projects/Work/2019-CoNNORFSI/paper/images/heatmaps/processed/heatmap-04-circle_quad-2-dc-vestrey-black-camo_product_9065558_color_12784}
+{\centering \includegraphics[width=.95\textwidth]{images/heatmaps/processed/heatmap-04-circle_quad-2-dc-vestrey-black-camo_product_9065558_color_12784}
 
@@
 The model assigns high probabilities of circle and quadrilateral and indicates that the proper image features are being used for this classification.
-The prediction of polygon is incorrect but reasonable considering that many small hexagons are indistinguishable from circles.\label{fig:dc-heatmap}}\label{fig:current-heatmaps3}
+The prediction of polygon is incorrect but reasonable considering that many small hexagons are indistinguishable from circles.\label{fig:dc-heatmap}}\label{fig:current-heatmaps}
 \end{figure}
